### 2018-01-26 applications.tex
initial check in

Consider the MDP in Figure~\ref{figure:simple-mdp-mdp}. Suppose the
states $+$ and $-$ are neighbors. Let us compute $\{ s : M,
\neighbor{S}, s \models \dpriv{3}{0} (\F Y) \}$. First, the self
product $M^2$ is depicted in Figure~\ref{figure:simple-mdp-product}
(the states $++$ and $--$ and associated transitions are ignored
for clarity). Applying
Theorem~\ref{theorem:eventuallyC} with $\epsilon = 4$ and $\delta = 0$, 
consider the following linear programming problem:
\[
\begin{array}{rcrcrcrcrcl}
  \multicolumn{11}{c}{
  \min x_{++} + x_{+-} + x_{-+} + x_{--} + x_{YY} + x_{YN} + x_{NY} + x_{NN}
  }\\
  \multicolumn{9}{r}{ x_{YN} } & = & 1 \\
  \multicolumn{9}{r}{ x_{YY} } & = & -3 \\
  \multicolumn{9}{r}{ x_{NY} } & = & -4 \\
  \multicolumn{9}{r}{ x_{NN} } & = & 0 \\
  x_{++} &-&   .5625 x_{YY} &-& .1875 x_{YN} &-& .1875 x_{NY} &-&
               .0625 x_{NN} & \geq & 0\\
  x_{++} &-&   .64   x_{YY} &-& .16   x_{YN} &-& .16   x_{NY} &-&
               .04   x_{NN} & \geq & 0\\
  x_{+-} &-&   .1875 x_{YY} &-& .5625 x_{YN} &-& .0625 x_{NY} &-& 
               .1875 x_{NN} & \geq & 0\\
  x_{+-} &-&   .16   x_{YY} &-& .64   x_{YN} &-& .04   x_{NY} &-&
               .16   x_{NN} & \geq & 0\\
  x_{-+} &-&   .1875 x_{YY} &-& .0625 x_{YN} &-& .5625 x_{NY} &-& 
               .1875 x_{NN} & \geq & 0\\
  x_{-+} &-&   .16   x_{YY} &-& .04   x_{YN} &-& .64   x_{NY} &-&
               .16   x_{NN} & \geq & 0\\
  x_{--} &-&   .0625 x_{YY} &-& .1875 x_{YN} &-& .1875 x_{NY} &-&
               .5625 x_{NN} & \geq & 0\\
  x_{--} &-&   .04   x_{YY} &-& .16   x_{YN} &-& .16   x_{NY} &-&
               .64   x_{NN} & \geq & 0
\end{array}
\]
The optimum is attained at 
$(x_{++}, x_{+-}, x_{-+}, x_{--}) = (-2.25, 0, -2.75, -0.6)$. By
Theorem~\ref{theorem:eventuallyC}, we have
\[
\begin{array}{l}
\max_{\scheduler{S}}
\myPr{+}{M_{\scheduler{S}}}{\neighbor{S}}{\F Y} -
4 \myPr{+}{M_{\scheduler{S}}}{\neighbor{S}}{\F Y}
= -2.25 \leq 0\\
\max_{\scheduler{S}}
\myPr{+}{M_{\scheduler{S}}}{\neighbor{S}}{\F Y} -
4 \myPr{-}{M_{\scheduler{S}}}{\neighbor{S}}{\F Y}
= 0 \leq 0 \\
\max_{\scheduler{S}}
\myPr{-}{M_{\scheduler{S}}}{\neighbor{S}}{\F Y} -
4 \myPr{+}{M_{\scheduler{S}}}{\neighbor{S}}{\F Y}
= -2.75 \leq 0\\
\max_{\scheduler{S}}
\myPr{-}{M_{\scheduler{S}}}{\neighbor{S}}{\F Y} -
4 \myPr{-}{M_{\scheduler{S}}}{\neighbor{S}}{\F Y}
= -0.6 \leq 0
\end{array}
\]
Therefore $+, - \in \{ s : M, \neighbor{M}, s \models \dpriv{3}{0}
(\F Y) \}$. This is expected. Recall that low- and high-accuracy
mechanisms are $(3, 0)$- and $(4, 0)$-differentially private
respectively. Regardless of request types, the interactive mechanism
in Figure~\ref{figure:simple-mdp-mdp} can only achieve $(4,
0)$-differential privacy at best.

In stream
processing mechanisms, input data are retrieved and analyzed
continuously.

Continual observation is commonly found in on-line service
providers~\cite{DNPR:10:DPCO,DNPRY:10:PPSA}. 
Consider monitoring events $a$ and $b$ by streaming inputs. We would
like to compute the number of event types occurred in inputs.

For each event, a bit denotes whether it has occurred or not. To model
adjacent inputs, we use four actions symbols to denote events and its
replacement. For instance, the action $a[a]$ means the event $a$ is
observed without replacement; $b[a]$ means the event $b$ is observed
but replaced with $a$ in an adjacent input.

\begin{figure}
  \centering
  \begin{subfigure}{\columnwidth}
    \centering
      \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
      distance=2cm,node/.style={circle,draw}]
      \node[node] (I) at  ( 0,  0) { $I$ };
      \node[node] (00) at (-1.5,  1.5) { $00$ };
      \node[node] (01) at (-1.5, -1.5) { $01$ };
      \node[node] (10) at ( 1.5,  1.5) { $10$ };
      \node[node] (11) at ( 1.5, -1.5) { $11$ };

      \path
      (I) edge node [below=2] { $1/4$ } (00)
      (I) edge node [above=2] { $1/4$ } (01)
      (I) edge node [below=2] { $1/4$ } (10)
      (I) edge node [above=2] { $1/4$ } (11)

      (00) edge [loop above] node { $a[\uscore]$, .375 } (00)
      (00) edge [bend left=10] node { $a[\uscore]$, .625 } (10)
      (00) edge [loop left] node [above=1,rotate=90] { $b[\uscore]$, .375 } (00)
      (00) edge [bend left=10] node [left=3,above=1,rotate=270] { $b[\uscore]$, .625 } (01)

      (10) edge [loop above] node { $a[\uscore]$, .625 } (00)
      (10) edge [bend left=10] node { $a[\uscore]$, .375 } (00)
      (10) edge [loop right] node [above=1,rotate=270] { $b[\uscore]$, .375 } (10)
      (10) edge [bend left=10] node [above=1,rotate=270] { $b[\uscore]$, .625 } (11)

      (11) edge [loop below] node { $a[\uscore]$, .625 } (11)
      (11) edge [bend left=10] node { $a[\uscore]$, .375 } (01)
      (11) edge [loop right] node [above=1,rotate=270] { $b[\uscore]$, .625 } (11)
      (11) edge [bend left=10] node [left=-3,above=1,rotate=90] { $b[\uscore]$, .375 } (10)

      (01) edge [loop below] node { $a[\uscore]$, .375 } (01)
      (01) edge [bend left=10] node { $a[\uscore]$, .675 } (11)
      (01) edge [loop left] node [above=1,rotate=90] { $b[\uscore]$, .625 } (01)
      (01) edge [bend left=10] node [left=-3,above=1,rotate=90] { $b[\uscore]$, .375 } (00)

      ;

      \node[node] (I') at  ( 5.5,    0) { $\underline{I}$ };
      \node[node] (00') at (   4,  1.5) { $\underline{00}$ };
      \node[node] (01') at (   4, -1.5) { $\underline{01}$ };
      \node[node] (10') at (   7,  1.5) { $\underline{10}$ };
      \node[node] (11') at (   7, -1.5) { $\underline{11}$ };

      \path
      (I') edge node [below=2] { $1/4$ } (00')
      (I') edge node [above=2] { $1/4$ } (01')
      (I') edge node [below=2] { $1/4$ } (10')
      (I') edge node [above=2] { $1/4$ } (11')

      (00') edge [loop above] node { $\uscore[a]$, .375 } (00')
      (00') edge [bend left=10] node { $\uscore[a]$, .625 } (10')
      (00') edge [loop left] node [above=1,rotate=90] { $\uscore[b]$, .375 } (00')
      (00') edge [bend left=10] node [left=3,above=1,rotate=270] { $\uscore[b]$, .625 } (01')

      (10') edge [loop above] node { $\uscore[a]$, .625 } (00')
      (10') edge [bend left=10] node { $\uscore[a]$, .375 } (00')
      (10') edge [loop right] node [above=1,rotate=270] { $\uscore[b]$, .375 } (10')
      (10') edge [bend left=10] node [above=1,rotate=270] { $\uscore[b]$, .625 } (11')

      (11') edge [loop below] node { $\uscore[a]$, .625 } (11')
      (11') edge [bend left=10] node { $\uscore[a]$, .375 } (01')
      (11') edge [loop right] node [above=1,rotate=270] { $\uscore[b]$, .625 } (11')
      (11') edge [bend left=10] node [left=-3,above=1,rotate=90] { $\uscore[b]$, .375 } (10')

      (01') edge [loop below] node { $\uscore[a]$, .375 } (01')
      (01') edge [bend left=10] node { $\uscore[a]$, .675 } (11')
      (01') edge [loop left] node [above=1,rotate=90] { $\uscore[b]$, .625 } (01')
      (01') edge [bend left=10] node [left=-3,above=1,rotate=90] { $\uscore[b]$, .375 } (00')

      ;
      \end{tikzpicture}
      \caption{Adjacent Inputs}
      \label{figure:adjacent-inputs}
    \end{subfigure}
    \begin{subfigure}{\columnwidth}
      \centering
      \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
      distance=2cm,node/.style={circle,draw}]
      \node[node] (0) at (-5,  0) { $0$ };
      \node[node] (1) at ( 0,  0) { $1$ };
      \node[node] (2) at ( 5,  0) { $2$ };

      \node[node] (00) at (-4.5, 1.5) { $00$ };
      \node[node] (01) at (-1.5, 2) { $01$ };
      \node[node] (10) at ( 1.5, 2) { $10$ };
      \node[node] (11) at ( 4.5, 1.5) { $11$ };

      \node[node] (00') at (-4.5, -1.5) { $\underline{00}$ };
      \node[node] (01') at (-1.5, -2) { $\underline{01}$ };
      \node[node] (10') at ( 1.5, -2) { $\underline{10}$ };
      \node[node] (11') at ( 4.5, -1.5) { $\underline{11}$ };

      \path
      (00) edge node [left] { $2/3$ } (0)
      (00) edge node [left=50,above] { $1/6$ } (1)
      (00) edge node [left=100,above=15] { $1/6$ } (2)

      (01) edge node [right=30,above=20] { $1/3$ } (0)
      (01) edge node [left=20,above=5] { $1/3$ } (1)
      (01) edge node [left=70,above=20] { $1/3$ } (2)

      (10) edge node [right=70,above=20] { $1/3$ } (0)
      (10) edge node [right=20,above=5] { $1/3$ } (1)
      (10) edge node [left=30,above=20] { $1/3$ } (2)

      (11) edge node [right=100,above=15] { $1/6$ } (0)
      (11) edge node [right=50,above] { $1/6$ } (1)
      (11) edge node [right] { $2/3$ } (2)

      (00') edge node [left] { $2/3$ } (0)
      (00') edge node [left=50,below] { $1/6$ } (1)
      (00') edge node [left=100,below=15] { $1/6$ } (2)

      (01') edge node [right=30,below=20] { $1/3$ } (0)
      (01') edge node [left=20,below=5] { $1/3$ } (1)
      (01') edge node [left=70,below=20] { $1/3$ } (2)

      (10') edge node [right=70,below=20] { $1/3$ } (0)
      (10') edge node [right=20,below=5] { $1/3$ } (1)
      (10') edge node [left=30,below=20] { $1/3$ } (2)

      (11') edge node [right=100,below=15] { $1/6$ } (0)
      (11') edge node [right=50,below] { $1/6$ } (1)
      (11') edge node [right] { $2/3$ } (2)
      ;

      \end{tikzpicture}
      \caption{Randomized Frequency}
      \label{figure:randomized-frequency}
    \end{subfigure}
  \caption{Frequency Estimator}
  \label{figure:frequency-estimator}
\end{figure}

\todo{verify this is $1$-DP}

### 2018-01-26 dpCTL.tex
initial check in

The logic $\dpCTL$ is designed to specify differentially private
mechanisms. We introduce the differentially private path quantifer
$\dpriv{\epsilon}{\delta}$ in $\dpCTL$. For any path formula
$\phi$, a state $s$ in a Markov chain $K$ satisfies
$\dpriv{\epsilon}{\delta} \phi$ if the probability of having $\phi$-paths 
from $s$ is close to the probabilities of 
having $\phi$-paths from its neighboring states. 
We give semantics for both Markov chains and Markov decision processes.

### 2018-01-26 examples.tex
initial check in

\begin{figure}
  \centering
  \begin{subfigure}{.48\columnwidth}
      \[
      \begin{array}{|c|c|c|}
        \hline
        &
        \multicolumn{2}{c|}{output}
        \\
        \hline
        state & Y & N \\
        \hline
        + & \frac{3}{4} = \frac{1}{2} \cdot \frac{1}{2} + \frac{1}{2} \cdot 1 
          & \frac{1}{4} = \frac{1}{2} \cdot \frac{1}{2} + \frac{1}{2} \cdot 0 
        \\
        \hline
        - & \frac{1}{4} = \frac{1}{2} \cdot \frac{1}{2} + \frac{1}{2} \cdot 0 
          & \frac{3}{4} = \frac{1}{2} \cdot \frac{1}{2} + \frac{1}{2} \cdot 1 
        \\
        \hline
      \end{array}
      \]
    \caption{Survey Mechanism}
    \label{figure:2-dp-table}
  \end{subfigure}  
  \hspace{.05\columnwidth}
  \begin{subfigure}{.40\columnwidth}
    \resizebox{\columnwidth}{!}{
    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
      distance=2cm,node/.style={circle,draw}]
      \node[node] (p) at ( 0,  .5) { $+$ };
      \node[node] (n) at ( 0, -.5) { $-$ };
      \node[node] (Y) at (-1.5,  0) { $Y$ };
      \node[node] (N) at ( 1.5,  0) { $N$ };

      \path
      (p) edge node [above] { $\frac{3}{4}$ } (Y)
      (p) edge node [above] { $\frac{1}{4}$ } (N)
      (n) edge node [below] { $\frac{1}{4}$ } (Y)
      (n) edge node [below] { $\frac{3}{4}$ } (N)
      ;
      \end{tikzpicture}
    }
    \caption{Corresponding Markov Chain}
    \label{figure:2-dp-mdp}
  \end{subfigure}
  \caption{Survey Mechanism with 2-Differential Privacy}
  \label{figure:2-dp}
\end{figure}

A simple way to design differentially private mechanisms is to add 
random noises. Consider the survey question: have you been diagnosed
with the disease $X$? In order to protect surveyees' privacy, each
answers the question as follows. The surveyee first flips a
coin. If it is tail, she answers the question truthfully. Otherwise, 
she randomly answers \textit{Yes} or \textit{No}
uniformly~\cite{DR:14:AFDP}. Figure~\ref{figure:2-dp} shows the survey
mechanism and its corresponding Markov chain. In the figure, the states
$+$ and $-$ denote the diagnoses are positive or negative
respectively; the states $Y$ and $N$ denote answers to the survey
question. The random noise boosts the probability of answering
\textit{Yes} or \textit{No} to at least $1/4$ regardless of
diagnoses. Inferences on individual diagnosis can be plausibly denied.

Let us analyze the survey mechanism briefly. The data universe
$\mathcal{X}$ is $\{ +, - \}$. The mechanism $M$ is a randomized
algorithm with inputs in $\mathcal{X}$ and outputs in $\{ Y, N
\}$. For any $x \in \mathcal{X}$, we have $1/4 \leq \Pr[M(x) = Y] \leq
3/4$. Hence for any neighboring $x, x' \in \mathcal{X}$, $\Pr[M(x) =
Y] \leq 3/4 = 3 \cdot 1/4 \leq 3 \Pr[M (x') =
Y]$. Similarly, $\Pr[M (x) = N] \leq 3 \Pr[M (x') = N]$ for any
neighboring $x, x' \in \mathcal{X}$. The survey mechanism is
$(3, 0)$-differentially private.

More sophisticated differentially private mechanisms are
available. Consider a query 
$f : \mathcal{X}^n \rightarrow \bbfZ^{\geq 0}$. Let $\alpha \in (0, 1)$. 
The \emph{$\alpha$-geometric mechanism} 
outputs $f(x) + Y$ where $Y$ is a random variable with the geometric
distribution~\cite{GRS:09:UUPM,GRS:12:UUPM} :
\[
\Pr[Y = y] = \frac{1 - \alpha}{1 + \alpha}\alpha^{|y|}
\textmd{ for } y \in \bbfZ.
\]
The $\alpha$-geometric mechanism is oblivious since it has the same
output distribution on any inputs $x, x'$ with $f (x) = f
(x')$. Moreover, it is also $(\alpha^{\Delta (f)}, 0)$-differentially
private for any query $f$ with sensitivity $\Delta (f)$. The range of
the mechanism 
however is $\bbfZ$. It may give nonsensical outputs such as
negative integers for queries with positive ranges.

The \emph{truncated $\alpha$-geometric mechanism over $\underline{n}$}
outputs $f (x) + Z$ where $Z$ is a random variable with the following
distribution: 
\[
\Pr[Z = z] =
\left\{
  \begin{array}{ll}
    0 & \textmd{ if } z < - f (x) \\
    \frac{\alpha^{f (x)}}{1 + \alpha} & \textmd{ if } z = -f (x)\\
    \frac{1 - \alpha}{1 + \alpha}\alpha^{|z|} & 
    \textmd{ if } -f (x) < z < n - f (x)\\
    \frac{\alpha^{n - f (x)}}{1 + \alpha} & \textmd{ if } z = n-f (x)\\
    0 & \textmd{ if } z > n - f (x)
  \end{array}
\right.
\]
Note the range of the truncated $\alpha$-geometric mechanism is
$\underline{n}$. The truncated $\alpha$-geometric mechanism is 
again oblivious; it is also $(\alpha^{\Delta (f)}, 0)$-differentially
private for any query $f$ with sensitivity $\Delta (f)$.
The truncated $1/2$-geometric mechanism over $\underline{5}$ is
given in Figure~\ref{figure:geometric-mechanism-table}.

\begin{figure}
  \centering
  \begin{subfigure}{.40\columnwidth}
    \[
    \begin{array}{|c|c|c|c|c|c|c|}
      \hline
      \textmd{input/output} & 0 & 1 & 2 & 3 & 4 & 5\\
      \hline
      0 & {2}/{3}  & {1}/{6} & {1}/{12} & {1}/{24}  & {1}/{48} & {1}/{48}\\
      \hline
      1 & {1}/{3}  & {1}/{3} & {1}/{6} & {1}/{12}  & {1}/{24} & {1}/{24}\\
      \hline
      2 & {1}/{6}  & {1}/{6} & {1}/{3} & {1}/{6}   & {1}/{12} & {1}/{12}\\
      \hline
      3 & {1}/{12} & {1}/{12} & {1}/{6} & {1}/{3}   & {1}/{6} & {1}/{6}\\
      \hline
      4 & {1}/{24} & {1}/{24} & {1}/{12} & {1}/{6}   & {1}/{3} & {1}/{3}\\
      \hline
      5 & {1}/{48} & {1}/{48} & {1}/{24} & {1}/{12} & {1}/{6} & {2}/{3}\\
      \hline
    \end{array}
    \]
    \caption{$1/2$-Geometric Mechanism}
    \label{figure:geometric-mechanism-table}
  \end{subfigure}
  \hspace{.08\columnwidth}
  \begin{subfigure}{.50\columnwidth}
    \centering
  \resizebox{.85\columnwidth}{!}{
    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
      distance=2cm,node/.style={circle,draw}]
      \node[node] (i0) at (0,  1) { $in_0$ };
      \node at (0, 0) { $\vdots$ };
      \node at (.8, -.8) { $\vdots$ };
      \node[node] (i5) at (0, -1) { $in_5$ };

      \node[node] (o0) at (4,  2.5) { $out_0$ };
      \node[node] (o1) at (4,  1.5) { $out_1$ };
      \node[node] (o2) at (4,  0.5) { $out_2$ };
      \node[node] (o3) at (4, -0.5) { $out_3$ };
      \node[node] (o4) at (4, -1.5) { $out_4$ };
      \node[node] (o5) at (4, -2.5) { $out_5$ };

      \path
      (i0) edge node [right=30,above=10] { $2/3$ } (o0)
      (i0) edge node [right=30,above=3] { $1/6$ } (o1)
      (i0) edge node [right=30,above=-3] { $1/12$ } (o2)
      (i0) edge node [right=30,below=-5] { $1/24$ } (o3)
      (i0) edge node [right=30,below] { $1/48$ } (o4)
      (i0) edge node [right=30,below=8] { $1/48$ } (o5)

      \hide{
      (i1) edge node [above] {  } (o0)
      (i1) edge node [above] {  } (o2)
      (i1) edge node [above] {  } (o3)
      (i1) edge node [above] {  } (o4)
      (i1) edge node [above] {  } (o5)

      (i2) edge node [above] {  } (o0)
      (i2) edge node [above] {  } (o2)
      (i2) edge node [above] {  } (o3)
      (i2) edge node [above] {  } (o4)
      (i2) edge node [above] {  } (o5)

      (i3) edge node [above] {  } (o0)
      (i3) edge node [above] {  } (o2)
      (i3) edge node [above] {  } (o3)
      (i3) edge node [above] {  } (o4)
      (i3) edge node [above] {  } (o5)

      (i4) edge node [above] {  } (o0)
      (i4) edge node [above] {  } (o2)
      (i4) edge node [above] {  } (o3)
      (i4) edge node [above] {  } (o4)
      (i4) edge node [above] {  } (o5)
      }

      (i5) edge node [right=10,above=16] { $1/48$ } (o0)
      \hide{
      (i5) edge node [above] {  } (o2)
      (i5) edge node [above] {  } (o3)
      (i5) edge node [above] {  } (o4)
      }
      (i5) edge node [right=15,below=8] { $2/3$ } (o5)
      ;
      \end{tikzpicture}
    }
    \caption{Markov Chain}
    \label{figure:geometric-mechanism-markov-chain}
  \end{subfigure}

  \caption{A Markov Chain for $1/2$-Geometric Mechanism}
  \label{figure:geometric-mechanism}
\end{figure}

Similar to the survey mechanism, it is 
straightfoward to model the truncated $1/2$-geometric
mechanisms as Markov chains. Let the state $in_k$ and $out_l$
denote the input value $k$ and output value $l$ respectively. Define
the state set $S = \{ in_k, out_k : k \in \underline{n} \}$.  
The probability transition $\wp (in_k, out_l)$ is the probability of the
output value $l$ on the input value $k$ as defined in the
mechanism. Figure~\ref{figure:geometric-mechanism-markov-chain} gives 
the Markov chain of the truncated
$1/2$-geometric mechanism over $\underline{5}$.

\begin{figure}
  \centering
  \begin{subfigure}{.40\columnwidth}
    \resizebox{\columnwidth}{!}{
    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
      distance=2cm,node/.style={circle,draw}]
      \node[node] (p) at ( 0,  1.5) { $+$ };
      \node[node] (q) at ( 0, -1.5) { $-$ };
      \node[node] (F) at (-1.5,  0) { $Y$ };
      \node[node] (T) at ( 1.5,  0) { $N$ };

      \path
      (p) edge [bend right=45] node [left] { $L, .75$ } (F)
      (p) edge [bend left=45] node [right] { $L, .25$ } (T)
      (p) edge [bend left=45] node [right=-12,above] { $H, .8$ } (F)
      (p) edge [bend right=45] node [right=12,above] { $H, .2$ } (T)

      (q) edge [bend left=45] node [left] { $L, .25$ } (F)
      (q) edge [bend right=45] node [right] { $L, .75$ } (T)
      (q) edge [bend right=45] node [left=12,below] { $H, .2$ } (F)
      (q) edge [bend left=45] node [right=12,below] { $H, .8$ } (T)

      (F) edge [loop left] node [below] { $-, 1$ } (F)
      (T) edge [loop right] node [below] { $-, 1$ } (T)
      ;
      \end{tikzpicture}
    }
    \caption{Markov Decision Process}
    \label{figure:simple-mdp-mdp}
  \end{subfigure}
  \begin{subfigure}{.58\columnwidth}
    \resizebox{\columnwidth}{!}{
    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
      distance=2cm,node/.style={circle,draw,inner sep=1pt}]
      \node[node] (pq) at (-2.1, 2.1) { $+-$ };
      \node[node] (qp) at ( 2.1,-2.1) { $-+$ };
      \node[node] (RR) at (-2.1,-2.1) { $YY$ };
      \node[node] (BB) at ( 2.1, 2.1) { $NN$ };
      \node[node] (RB) at (- .7,- .7) { $YN$ };
      \node[node] (BR) at (  .7,  .7) { $NY$ };

      \path
      (pq) edge [bend right=5] node [left=1,below=12,rotate=270] { $L, .1875$ } (RR)
      (pq) edge [bend left=5]  node [above=5,right=1] { $L, .1875$ } (BB)

      (pq) edge [bend left=5]  node [right=4,rotate=270] { $H, .16$ } (RR)
      (pq) edge [bend right=5] node [below=5,right=1] { $H, .16$ } (BB)

      (pq) edge [bend left=5]  node [left,below=10,rotate=295]  { $L, .5625$ } (RB)
      (pq) edge [bend right=5] node [left,below,rotate=330]  { $L, .0625$ } (BR)

      (pq) edge [bend right=5] node [right=2,above,rotate=300] { $H, .64$ } (RB)
      (pq) edge [bend left=5]  node [right,above,rotate=340] { $H, .04$ } (BR)

      (qp) edge [bend right=5] node [above=5,left=1] { $L, .1875$ } (RR)
      (qp) edge [bend left=5]  node [above=30,left=5,rotate=90] { $L, .1875$ } (BB)

      (qp) edge [bend left=5]  node [below=5,left=1] { $H, .16$ } (RR)
      (qp) edge [bend right=5] node [above,right=5,rotate=90] { $H, .16$ } (BB)

      (qp) edge [bend left=5]  node [left,below,rotate=330]  { $L, .0625$ } (RB)
      (qp) edge [bend right=5] node [left=3,below=3,rotate=295]  { $L, .5625$ } (BR)

      (qp) edge [bend right=5] node [left=5,above,rotate=330] { $H, .04$ } (RB)
      (qp) edge [bend left=5]  node [right=1,above=1,rotate=295] { $H, .64$ } (BR)

      (RR) edge [loop left] node [above] { $-, 1$ } (RR)
      (BB) edge [loop right] node [above] { $-, 1$ } (BB)
      (RB) edge [loop below] node [left] { $-, 1$ } (RB)
      (BR) edge [loop above] node [right] { $-, 1$ } (BR)
      ;
      \end{tikzpicture}
\hide{
    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
      distance=2cm,node/.style={circle,draw,inner sep=1pt}]
      \node[node] (pq) at (   0,   0) { $+-$ };
      \node[node] (RR) at (-  2,   0) { $YY$ };
      \node[node] (BB) at (   2,   0) { $NN$ };
      \node[node] (RB) at (   0, 1.5) { $YN$ };
      \node[node] (BR) at (   0,-1.5) { $NY$ };

      \path
      (pq) edge [bend right=15] node [above] { $L, .1875$ } (RR)
      (pq) edge [bend left=15]  node [above] { $L, .1875$ } (BB)

      (pq) edge [bend left=15]  node [below] { $H, .16$ } (RR)
      (pq) edge [bend right=15] node [below] { $H, .16$ } (BB)

      (pq) edge [bend left=15]  node [left]  { $L, .5625$ } (RB)
      (pq) edge [bend right=15] node [left]  { $L, .0625$ } (BR)

      (pq) edge [bend right=15] node [right] { $H, .64$ } (RB)
      (pq) edge [bend left=15]  node [right] { $H, .04$ } (BR)

      (RR) edge [loop below] node [below] { $-, 1$ } (RR)
      (BB) edge [loop above] node [above] { $-, 1$ } (BB)
      (RB) edge [loop left] node [left] { $-, 1$ } (RB)
      (BR) edge [loop right] node [right] { $-, 1$ } (BR)
      ;
      \end{tikzpicture}
}
    }
    \caption{Product Markov Decision Process}
    \label{figure:simple-mdp-product}
  \end{subfigure}  
  \caption{Markov Decision Process and its Self-Product}
  \label{figure:simple-mdp}
\end{figure}

Mechanisms are not necessarily closed randomized algorithms; they may
perform different compution on users' requests. 
We use Markov decision processes to model such 
interactive mechanisms. Specifically, external inputs are modeled by
actions. Behaviors associated with different inputs are modeled by
distributions associated with actions.

Consider again the survey mechanism. Suppose we would like to design
an interactive mechanism which adjusts random noises on surveyers'
requests. When the surveyer requests low-accuracy answers, the
surveyee uses the survey mechanism as before. When high-accuracy
answers are requested, the surveyee answers \textit{Yes} with
probability $4/5$ and \textit{No} with probability $1/5$ when she has
positive diagnosis. She answers \textit{Yes} with probability $1/5$
and \textit{No} with probability $4/5$ when she is not
diagnosed with the disease X. This gives an interactive mechanism
corresponding to the Markov decision process shown in
Figure~\ref{figure:simple-mdp-mdp}.

In the figure, the states $+$,
$-$, $Y$, and $N$ are interpreted as before. The actions $L$ and $H$
denote low- and high-accuracy requests respectively. Let $M_H$ denote
the Markov chain derived by high-accuracy requests.
Observe that $1/5 \leq \Pr[M_H(x) = Y] \leq 4/5$ for any $x
\in \mathcal{X}$. Hence $\Pr[M_H (x) = Y] \leq 4/5 = 4 \cdot 1/5 \leq
4 \Pr[M_H (x') = Y]$ for any neighboring $x, x' \in
\mathcal{X}$. Similarly, $\Pr[M_H (x) = N] \leq 4 \Pr[M_H (x') =
N]$. The high-accuracy mechanism is $(4,0)$-differentially private. 
The privacy guarantees vary from accuracy requests.

### 2018-01-26 main.tex
initial check in

\documentclass{llncs}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{xspace}
\pagestyle{plain}
\usepackage{algorithm,algpseudocode}
\usepackage{caption}
\usepackage{subcaption}
\captionsetup{compatibility=false}
\usepackage{mathtools}
\usepackage{tikz}
\usepackage{url}
\usetikzlibrary{arrows,calc}

\input{macros}

\title{Model Checking Differentially Private Properties}
\author{
Depeng Liu\inst{1}
\and
Bow-Yaw Wang\inst{2}
\and 
Lijun Zhang\inst{1}}
\institute{
Chinese Academy of Sciences 
\and
Academia Sinica
}

\begin{document}

\maketitle

\begin{abstract}
  We introduce the branching time temporal logic \dpCTL for specifying
  differential privacy. Its model checking algorithm is also developed. 
\end{abstract}

\section{Introduction}
\label{section:introduction}
\input{introduction}

This paper is organized as follows. After the introduction,
preliminaries are given in Section~\ref{section:preliminaries}. The
logic $\dpCTL$ is presented in
Section~\ref{section:dpCTL}. The model cheking algorithms are developed in
Section~\ref{section:model-checking-algorithms}. They are followed by
applications (Section~\ref{section:applications}. Finally,
Section~\ref{section:conclusions} concludes our presentation.

\section{Preliminaries}
\label{section:preliminaries}
\input{preliminaries}

\section{Examples}
\label{section:examples}
\input{examples}

\section{The Logic $\dpCTL$}
\label{section:dpCTL}
\input{dpCTL}

\subsection{Syntax}
\label{subsection:syntax}
\input{syntax}

\subsection{Semantics}
\label{subsection:semantics}
\input{semantics}

\section{Model Checking Algorithms}
\label{section:model-checking-algorithms}
\input{algorithms}

\section{Applications}
\label{section:applications}
\input{applications}

\section{Conclusions}
\label{section:conclusions}

\bibliographystyle{splncs03}
\bibliography{refs}

\end{document}

### 2018-01-26 preliminaries.tex
initial check in

Let $\bbfZ$ and $\bbfZ^{\geq 0}$ be the sets of integers and
non-negative integers respectively. For $n \geq 0$, $\underline{n} =
\{ 0, 1, \ldots, n \} \subseteq \bbfZ^{\geq 0}$.

We briefly review the definitions of differential privacy, Markov
chains, and Markov decision processes. For differential privacy, we
follow the non-standard \todo{why?} definition in~\cite{GRS:09:UUPM,GRS:12:UUPM}. Our
definitions of Markov chains and Markov decision processes are adopted
from~\cite{BK:08:PMC}.

\subsection{Differential Privacy}

We denote the data universe by $\mathcal{X}$; $x \in \mathcal{X}^n$ is
a database with $n$ rows from the data universe. Two databases $x$ and
$x'$ are \emph{neighbors} (denoted by $d(x, x') \leq 1$) if they are
identical except at most one row. A \emph{query} $f$ is a function
from $\mathcal{X}^n$ to its range $R$. The \emph{sensitivity} of the
query $f$ (written $\Delta (f)$) is $\max_{d(x, x') \leq 1} | f (x) -
f (x') |$. For instance, a \emph{counting} query counts the number
of rows with certain attributes (say, female). The sensitivity of a
counting query is $1$ since any neighbor can change the count by at
most one. We only consider queries with finite  ranges for simplicity. 
A \emph{data analysis mechanism} (or
\emph{mechanism} for brevity) $M_f$ for a query $f$
is a randomized algorithm with inputs in $\mathcal{X}^n$ and outputs
in $\tilde{R}$. 
A mechanism may not have the same output range as its query, that is,
$\tilde{R} \neq R$ in general. 
A mechanism $M_f$ for $f$ is \emph{oblivious} if 
$\Pr[M_f(x) = \tilde{r}] = \Pr[M_f(x') = \tilde{r}]$ for every
$r \in \tilde{R}$ when $f (x) = f (x')$. In words, outputs of an
oblivious mechanism depend on the query result $f (x)$ but not on the
input database $x$. The order of rows in a database, for instance, is
irrelevant to oblivious mechanisms. A database $x$ is
\emph{$(\epsilon, \delta)$-close} to $x'$ in a mechanism
$M_f$ if for every $\tilde{r} \in \tilde{R}$,
\[
\Pr[M_f (x) = \tilde{r}] \leq \epsilon \Pr[M_f (x') =
\tilde{r}] + \delta.\footnote{The standard definition requires 
$\Pr[M_f (x) = \tilde{r}] \leq e^\epsilon \Pr[M_f (x') =
\tilde{r}] + \delta$~\cite{DMNS:06:CNSPD,D:06:DP}. 
Since $\epsilon = e^{\ln \epsilon}$, our $(\epsilon, \delta)$-closedness is
equivalent to the standard $(\ln \epsilon,
\delta)$-closedness~\cite{GRS:09:UUPM,GRS:12:UUPM}.}  
\]
A mechanism $M_f$ is \emph{$(\epsilon, \delta)$-differentially
  private} % $(\epsilon, \delta \approx 0)$ 
if for every $x, x' \in \mathcal{X}^n$ with $d(x, x') \leq 1$,
$x$ is $(\epsilon, \delta)$-close to $x'$ in $M_f$.

The parameters $\epsilon$ and $\delta$ quantify probabilistically
similar behaviors; 
the smaller they are, the behaviors are more similar.
Informally, a differentially private mechamism has probabilistically
similar behabiors on neighboring databases. It will have similar
output distributions when a row is added, deleted, or replaced by
another in a given 
database. Consider, for instance, a database contains the row
corresponding to an individual. A differentially private mechanism
will have a similar output distribution when the individual is not in
the database. Regardless of the presence of the individual, such
mechanisms will behave similarly. Privacy of the individual is thus
preserved by differentially private mechanisms. \todo{not clear}

\subsection{Markov Chains and Markov Decision Processes}

Let $\AP$ be the set of \emph{atomic propositions}.
A \emph{(finite) discrete-time Markov chain} $K = (S, \wp, L)$ consists
of a non-empty finite set $S$ of \emph{states}, a \emph{transition
  probability function} $\wp : S \times S \rightarrow [0, 1]$ with
$\sum_{t \in S} \wp(s, t) = 1$ for every $s \in S$, and
a \emph{labeling function} $L : S \rightarrow 2^{\AP}$. A \emph{path}
in $K$ is an infinite sequence $\pi = \pi_0 \pi_1 \cdots \pi_n \cdots$
of states with $\wp (\pi_i, \pi_{i+1}) > 0$ for all $i \geq 0$. We write
$\pi[j]$ for the suffix $\pi_j \pi_{j+1} \cdots$. % Thus $\pi[0] = \pi$.

Let $\Act$ denote the set of \emph{actions}.
A \emph{(finite) Markov decision process} $M = (S, \Act, \wp, L)$ consists
of a \emph{transition probability function} $\wp : S \times \Act
\times S \rightarrow [0, 1]$ with $\sum_{t \in S} \wp(s, \alpha, t)
\in \{ 0, 1 \}$ for every $s \in S$ and $\alpha \in \Act$ and $S$,
$L$ as for Markov chains. Define
$\Act(s) = \{ \alpha : \sum_{t \in S} \wp (s, \alpha, t) = 1
\}$. That is, $\Act(s)$ is the set of \emph{enabled} actions in state
$s$. A Markov devision process is \emph{reactive} if $\Act(s) = \Act$
for every $s \in S$.
A \emph{path} $\pi$ in $M$ is an infinite sequence $\pi_0 \alpha_1
\pi_1 \cdots \pi_n \alpha_{n+1} \cdots$ with 
$\wp(\pi_i, \alpha_{i+1}, \pi_{i+1}) > 0$ for all $i \geq 0$. 
Similarly, we write $\pi[j]$ for the suffix $\pi_j \alpha_{j+1}
\pi_{j+1} \cdots$ of $\pi$. 
\hide{
For any measurable set $\Pi$ of paths, we
write $\Pr[\Pi]$ for the \emph{probability measure} of $\Pi$.
}

Let $M = (S, \Act, \wp, L)$ be a Markov decision process. A
\emph{scheduler} for $M$ is a function $\scheduler{S} : S^+
\rightarrow \Act$ with $\scheduler{S}(s_0s_1\cdots s_n) \in
\Act(s_n)$. A scheduler is \emph{position-independent} if
$\scheduler{S}(\sigma) = \scheduler{S}(\sigma')$ when $|\sigma| =
|\sigma'|$. 
A path $\pi =
\pi_0 \alpha_1 \pi_1 \cdots \pi_n \alpha_{n+1} \cdots$ is an
\emph{$\scheduler{S}$-path} if $\alpha_{i+1} = \scheduler{S}(\pi_0
\pi_1 \cdots \pi_i)$ for all $i \geq 0$. 
Note that a Markov decision process with a
scheduler $\scheduler{S}$ induces a Markov chain $M_{\scheduler{S}} =
(S^+, \wp_{\scheduler{S}}, L')$ where $L' (\sigma s) = L (s)$, 
$\wp_{\scheduler{S}} (\sigma s, \sigma s t) = \wp (s,
\scheduler{S}(\sigma s), t)$ for $\sigma \in S^*$ and $s, t \in S$.
\hide{
Subsequently, we write
$\Pr_{\scheduler{S}} (\Pi)$ for the \emph{probability measure} of the
measurable path set $\Pi$ on $M$ given the scheduler $\scheduler{S}$. 
}

\hide{
A \emph{reward function} on an MDP $M = (S, \wp, I, L)$ is a function
$r : S \rightarrow \bbfR$. The \emph{reward} for $r (\pi)$ a finite
path $\pi$ in $M$ is $\sum_{i=0}^n s_i$ where $\pi = s_0 \alpha_1 s_1
\cdots s_n$. 
}

### 2018-01-26 semantics.tex
initial check in

Given a finite discrete-time Markov chain $K = (S, \wp, L)$, a
\emph{neighborhood relation on $S$} $\neighbor{S} \subseteq S \times S$
is a reflexive and symmetric relation on $S$. We will write $s
\neighbor{S} t$ when $(s, t) \in \neighbor{S}$. If $s \neighbor{S} t$,
we say $s$ and $t$ are \emph{neighbors} or $t$ is a \emph{neighbor} of
$s$. Define the satisfaction relation $K, \neighbor{S}, s
\models \Phi$ as follows.
\begin{eqnarray*}
  K, \neighbor{S}, s \models \top\\
  K, \neighbor{S}, s \models p
  & \textmd{ if } &
  p \in L(s)\\
  K, \neighbor{S}, s \models \neg \Phi
  & \textmd{ if } &
  K, \neighbor{S}, s \not\models \Phi\\
  K, \neighbor{S}, s \models \Phi_0 \wedge \Phi_1
  & \textmd{ if } &
  K, \neighbor{S}, s \models \Phi_0 \textmd{ and }
  K, \neighbor{S}, s \models \Phi_1\\
  K, \neighbor{S}, s \models \PJ{J} \phi
  & \textmd{ if } &
  \Pr [\{ \pi : K, \neighbor{S}, \pi \models \phi \textmd{ with }
                    \pi_0 = s\}] \in J \\
  K, \neighbor{S}, s \models \dpriv{\epsilon}{\delta} \phi
  & \textmd{ if } & 
  \textmd{for every } t \textmd{ with }  s \neighbor{S} t,
      p(s) \leq \epsilon p(t) + \delta \textmd{ and } 
      p(t) \leq \\
  & & \epsilon p(s)  + \delta  \textmd{ where } p(x) = \Pr [\{
      \pi : K, \neighbor{S}, \pi \models \phi \textmd{ with}\\
  & &  \pi_0 = x \}]
\end{eqnarray*}

Moreover, the relation $K, \neighbor{S}, \pi \models \phi$ is defined as
follows.
\begin{eqnarray*}
  K, \neighbor{S}, \pi \models \X \Phi
  & \textmd{ if } &
  K, \neighbor{S}, \pi[1] \models \Phi\\
  K, \neighbor{S}, \pi \models \Phi \until \Psi 
  & \textmd{ if } &
  \textmd{there is a } j \geq 0 \textmd{ such that }
  K, \neighbor{S}, \pi[j] \models \Psi \textmd{ and } \\
  & & K, \neighbor{S}, \pi[k] \models \Phi
      \textmd{ for every } 0 \leq k < j
\hide{
  K, \neighbor{S}, \pi \models \Phi \buntil{n} \Psi 
  & \textmd{ if } &
  \textmd{there is a } 0 \leq j \leq n \textmd{ such that }
  K, \neighbor{S}, \pi[j] \models \Psi \textmd{ and } \\
  & & K, \neighbor{S}, \pi[k] \models \Phi
      \textmd{ for every } 0 \leq k < j
}
\end{eqnarray*}

To simplify our notation, for any Markov chain $K$, neighborhood
relation $N$ on $S$, a state $s \in S$, and a path formula $\phi$,
define
\[
\myPr{s}{K}{N}{\phi} =
\Pr [ \{ \pi : K, N, \pi \models \phi \textmd{ with } \pi_0 = s \} ].
\]
That is, $\myPr{s}{K}{N}{\phi}$ denotes the probability of paths
satisfying $\phi$ from $s$ on $K$ with $N$. The semantics of the
differentially private operator hence becomes
\begin{eqnarray*}
  K, \neighbor{S}, s \models \dpriv{\epsilon}{\delta} \phi
  & \textmd{ if } & 
  \textmd{for every } t \textmd{ with }  s \neighbor{S} t,
      \myPr{s}{K}{N}{\phi} \leq \epsilon \myPr{t}{K}{N}{\phi} + \delta 
      \textmd{ and }\\
  & &  \myPr{t}{K}{N}{\phi} \leq \epsilon \myPr{s}{K}{N}{\phi} + \delta
\end{eqnarray*}

Other than the differentially private operator, the semantics of
$\dpCTL$ is standard~\cite{BK:08:PMC}. 
To intuit the semantics of $\dpriv{\epsilon}{\delta} \phi$, 
recall that  $\myPr{s}{K}{N}{\phi}$ is the probability of having
paths satisfying $\phi$ from $s$. Thus, a state $s$ satisfies 
$\dpriv{\epsilon}{\delta} \phi$ if having paths satisfying $\phi$
from $s$ is $(\epsilon, \delta)$-close to having paths satisfying
$\phi$ from every neighbor of $s$. In other words, it is
probabilistically similar to observe paths satisfying $\phi$ from $s$
and from all neighbors of $s$.

Note that $\dpCTL$ generalizes differential privacy to infinite
behaviors naturally. Classical differential privacy discusses properties about
randomized terminating mechanisms. Sophisticated mechanisms for
streaming data have also been discussed in social networks and
internet of things. Such mechanisms inherently have infinite
computation. Their privacy guarantees are approximated by those of the 
prefixes of computation.
With \dpCTL, the subformula $\phi$ in $\dpriv{\epsilon}{\delta} \phi$
can be an arbitrary $\dpCTL$ path formula. One can specify
$(\epsilon, \delta)$-closedness over infinite behaviors naturally. For instance,
$\dpriv{\epsilon}{\delta} (\G \PJ{[1,1]} (\F \mathit{high}))$
specifies states having $(\epsilon, \delta)$-close 
infinitely often $\mathit{high}$ behaviors to its neighbors.

\noindent
\textit{Example.} 
Consider the Markov chain in
Figure~\ref{figure:geometric-mechanism-markov-chain}. Assume 
$L(out_i) = \{ out_i \}$ and $\Delta (f) = 1$. Then $in_k$ and $in_l$ are
neighbors iff $| k - l | \leq 1$. Define the $\dpCTL$ formula
$\psi = \dpriv{.5}{0} (\F out_0) \wedge \dpriv{.5}{0} (\F out_1) \wedge
\cdots \wedge \dpriv{.5}{0} (\F out_5)$. If the state $in_k$ satisfies
$\psi$ for $k = 0, \ldots, 5$, then the $.5$-geometric mechanism is
$(0.5, 0)$-differentially private.

The semantics over Markov chains generalizes to Markov decision
processes as well. Note that two states with different enabled actions
are trivially distinguishable; no privacy can be preserved. Without
loss of generality, we consider reactive Markov decision processes.
Let $M = (S, \Act, \wp, L)$ be a reactive Markov
decision process. Define the satisfaction
relation $M, \neighbor{S}, s \models \Phi$ as follows.
\begin{eqnarray*}
  M, \neighbor{S}, s \models \top\\
  M, \neighbor{S}, s \models p
  & \textmd{ if } &
  p \in L(s)\\
  M, \neighbor{S}, s \models \neg \Phi
  & \textmd{ if } &
  M, \neighbor{S}, s \not\models \Phi\\
  M, \neighbor{S}, s \models \Phi_0 \wedge \Phi_1
  & \textmd{ if } &
  M, \neighbor{S}, s \models \Phi_0 \textmd{ and }
  M, \neighbor{S}, s \models \Phi_1\\
  M, \neighbor{S}, s \models \PJ{J} \phi
  & \textmd{ if } &
  \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \in J  
  \textmd{ for every scheduler } \scheduler{S}\\
  M, \neighbor{S}, s \models \dpriv{\epsilon}{\delta} \phi
  & \textmd{ if } &
  \textmd{for all } t \textmd{ with } s \neighbor{S} t \textmd{ and
   position-independent scheduler } \\
  & & \scheduler{S}, 
   \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq \epsilon \cdot
   \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} + \delta 
   \textmd{ and }
   \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq \epsilon \cdot
  \\
  & &
   \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} + \delta
\end{eqnarray*}
Recall that $M_{\scheduler{S}}$ is but a Markov chain. The semantics
of $M_{\scheduler{S}}, \neighbor{S}, \pi \models \phi$ and hence the
probability $\myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}$ are
defined as in Markov chains. 
The semantics of $\dpCTL$ on reactive Markov decision processes
is again standard except the differentially private operator
$\dpriv{\epsilon}{\delta}$. For any path formula $\phi$,
$\dpriv{\epsilon}{\delta} \phi$ specifies states which are $(\epsilon,
\delta)$-close to all its neighbors in having paths satisfying $\phi$
for arbitrary position-independent schedulers. That is, no
position-independent scheduler can force any of neighbors to have
probabilistically distinguishable behaviors.

\noindent
\emph{Discussion.}
Another definition of the differentially private operator
might be: 
\begin{eqnarray*}
  M, \neighbor{S}, s \models \dpriv{\epsilon}{\delta}^{\mathit{bad}} \phi
  & \textmd{ if } &
  \textmd{for all } t \textmd{ with } s \neighbor{S} t \textmd{ and
  scheduler } \scheduler{S},
  \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq
  \epsilon \cdot\\
  && \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} + \delta \textmd{ and }
  \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq
  \epsilon \cdot \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} 
  + \delta
\end{eqnarray*}
A state satisfies $\dpriv{\epsilon}{\delta}^{\mathit{bad}} \phi$ if
no scheduler could differentiate the probabilities of paths satisfying
$\phi$ from neighbors. Such a definition would allow strategies to
take different actions from different states. It might need different
action sequences to distinguish neighbors. Two neighbors could still be
differentiated by a action sequence and lose their privacy. We hence
prefer the original definition.

\hide{
Since the scheduler attaining the maximally probable
behavior from a state may be different from the scheduler attaining
the maximally probable behavior from its neighbors, two
neighbors may still be distinguished by a scheduler. The
weaker definition does not preserve differential privacy. We hence
prefer the original definition.
}

### 2018-01-26 syntax.tex
initial check in

Consider
\begin{eqnarray*}
  \Phi & ::= & \top \ |\ p \ |\ \neg \Phi \ |\ \Phi \wedge \Phi \ |\ 
               \PJ{J} \phi \ |\ \dpriv{\epsilon}{\delta} \phi\\
  \phi & ::= & \X \Phi \ |\ \Phi \until \Phi
\end{eqnarray*}
A \emph{state} formula $\Phi$ is either $\top$, an atomic proposition
$p$, the negation of a sate formula, the conjunction of two state
formulae, a \emph{probabilistic} operator $\PJ{J}$ with $J$
an interval in $[0, 1]$ followed by a path formula, or a
\emph{differentially private} operator $\dpriv{\epsilon}{\delta}$
 with two non-negative real numbers $\epsilon$ and $\delta$ followed
 by a path formula. A 
\emph{path} formula $\phi$ is a \emph{next} operator ($\X$)
followed by a state formula, or an \emph{until} operator
($\until$) operator enclosed by two state formulae.

### 2018-01-26 semantics.tex
first try
added some properties

@@
 Moreover, the relation $K, \neighbor{S}, \pi \models \phi$ is defined as
-follows.
+for the standard LTL formulas. We only give the semantics for $\X$ and $\until$ as follows.
 \begin{eqnarray*}
-  K, \neighbor{S}, \pi \models \X \Phi
+  K, \neighbor{S}, \pi \models \X \phi
   & \textmd{ if } &
-  K, \neighbor{S}, \pi[1] \models \Phi\\
-  K, \neighbor{S}, \pi \models \Phi \until \Psi 
+  K, \neighbor{S}, \pi[1] \models \phi\\
+  K, \neighbor{S}, \pi \models \phi \until \psi
   & \textmd{ if } &
   \textmd{there is a } j \geq 0 \textmd{ such that }
-  K, \neighbor{S}, \pi[j] \models \Psi \textmd{ and } \\
-  & & K, \neighbor{S}, \pi[k] \models \Phi
+  K, \neighbor{S}, \pi[j] \models \psi \textmd{ and } \\
+  & & K, \neighbor{S}, \pi[k] \models \phi
       \textmd{ for every } 0 \leq k < j
@@
 
+\section{Specifying Properties in $\dpCTL$}
+In this section we describe what type of properties in the differential privacy context can be expressed using $\dpCTL$.
+
+\paragraph{Standard Differential Privacy.} 
+The formula $\dpriv{3}{0} (\X Y)$ holds in state $s$ if for  $t$ with $sNt$, the probability of satisfying $\X Y$ from $s$ and $t$ satisfy:
+$\myPr{s}{K}{N}{\X Y}\le 3\myPr{t}{K}{N}{\X Y}$.
+The formula $\dpriv{3}{0} (\X Y)\wedge \dpriv{3}{0} (\X N)$ thus specifies the differential privacy property for states $+$ and $-$ in Example~\ref{exa:survey}.
+
+Further, consider the Markov chain in
+Figure~\ref{figure:geometric-mechanism-markov-chain}. Assume
+$L(out_i) = \{ out_i \}$ and $\Delta (f) = 1$. Then $in_k$ and $in_l$ are
+neighbors iff $| k - l | \leq 1$\lz{put above sentence to the def of the model}. Define the $\dpCTL$ formula
+$\psi = \dpriv{.5}{0} (\F out_0) \wedge \dpriv{.5}{0} (\F out_1) \wedge
+\cdots \wedge \dpriv{.5}{0} (\F out_5)$\lz{I think we better use $\X$ operator here ;)}. If the state $in_k$ satisfies
+$\psi$ for $k = 0, \ldots, 5$, then the $.5$-geometric mechanism is
+$(0.5, 0)$-differentially private.
+
+\paragraph{Compositional Properties.}
+Compositional aspect is one of the building block for differentially private algorithms. According to the compositional theorem \cite{}, for $(\epsilon_1,0)$-differentially private algorithm $M_1$ and
+$(\epsilon_2,0)$-differentially private algorithm $M_2$, their combination $M$ defined by $M(x)=(M_1(x), M_2(x))$ is $(\epsilon_1\epsilon_2,0)$-differentially private.
+The degradation is rooted in the repeated computation.
+To illustrate this property, we consider the extended example below \lz{here example of the survey example, expanded twice: we should discuss the graph before it is drawn...}
+
+We consider the formula $\dpriv{9}{0} (\X (Y\wedge \X Y))$.
+A path satisfies $\X (Y\wedge \X Y)$ if the second state satisfies $Y$ and the third state satisfies $Y$ as well.
+We verify that this formula is satisfied  for state $+$ and $-$.
+Moreover, the bound $\epsilon=9$ is bound, since \lz{tbd once graph is given}.
+Finally, the formula $\wedge_{a_1,a_2}\dpriv{9}{0} (\X (a_1\wedge \X a_2))$ corresponds to the desired compositional property for the model, where $a_1,a_2$ range over all atomic propositions $\{Y,N\}$.
+
+We consider another simpler formula for comparision: $\dpriv{3}{0} (\X \X Y)$.
+In this case we donot have privacy lost, even though we are querying twice: the reason is that the output of the first query is not utilized at all.
+It is easy to verify that it is indeed satisfied by $+$ and $-$.
+
+\paragraph{Advanced Compositional Properties}
+
+One of the advantage of applying model checking is that we may get tighter bounds for compositional properties. Consider ...
+\lz{find a -- simpler -- formula such that the bound is not tight, maybe $\delta>0$ is needed}
+
+\paragraph{Infinite ...}
 Note that $\dpCTL$ generalizes differential privacy to infinite
@@
 $(\epsilon, \delta)$-closedness over infinite behaviors naturally. For instance,
-$\dpriv{\epsilon}{\delta} (\G \PJ{[1,1]} (\F \mathit{high}))$
+$\dpriv{\epsilon}{\delta} (\G \PJ{[1,1]} (\F \mathit{high}))$ \lz{now we can use $G F high$}
 specifies states having $(\epsilon, \delta)$-close
@@
 
-\noindent
-\textit{Example.} 
-Consider the Markov chain in
-Figure~\ref{figure:geometric-mechanism-markov-chain}. Assume 
-$L(out_i) = \{ out_i \}$ and $\Delta (f) = 1$. Then $in_k$ and $in_l$ are
-neighbors iff $| k - l | \leq 1$. Define the $\dpCTL$ formula
-$\psi = \dpriv{.5}{0} (\F out_0) \wedge \dpriv{.5}{0} (\F out_1) \wedge
-\cdots \wedge \dpriv{.5}{0} (\F out_5)$. If the state $in_k$ satisfies
-$\psi$ for $k = 0, \ldots, 5$, then the $.5$-geometric mechanism is
-$(0.5, 0)$-differentially private.
-
+\section{Markov Decision Processes}
+\lz{let put it as another section after Markov chains}
 The semantics over Markov chains generalizes to Markov decision

### 2018-01-26 semantics.tex
try

@@
 
+\section{Model Checking Algorithms}
+The model checking algorithms for $\dpCTL$ follow the classical
+algorithms for $\PCTL$ by computing the states satisfying
+sub state-formulae inductively~\cite{BK:08:PMC}. It hence suffices to
+consider the inductive step where the states satisfying the subformula
+$\dpriv{\epsilon}{\delta} (\phi)$ is computed.
+
+%We start with the algorithm for Markov chains.
+
+In the classical $\PCTL$ model checking algorithm for Markov chains,
+states satisfying the subformula $\PJ{J} \phi$ are obtained by
+computing $\myPr{s}{K}{\neighbor{S}}{\phi}$ for state $s \in S$.
+These probabilities can be computed by solving linear equations or
+through iterative approximations. We summarize it in the following
+theorem~\cite{BK:08:PMC}:
+
+\begin{lemma}
+  Let $K = (S, \wp, L)$ be a Markov chain, $s
+  \in S$, and $B, C \subseteq S$. The probabilities
+  $\myPr{s}{K}{\neighbor{S}}{\gX B}$ and
+  $\myPr{s}{K}{\neighbor{S}}{B \mho C}$
+\hide{
+  $\Pr[\{ \pi : K, \neighbor{K}, \pi \models B \buntil{n} C \textmd{
+    with } \pi_0 = s \}]$
+} are computable within time polynomial in
+  $|S|$.
+  \label{lemma:PJ-subroutines}
+\end{lemma}
+
+In Lemma~\ref{lemma:PJ-subroutines}, we abuse the notation slightly to
+admit path formulae of the form $\gX B$ (next $B$) and
+$B \guntil C$ ($B$ until $C$) with $B, C \subseteq S$ as
+in~\cite{BK:08:PMC}. They are interpreted by introducing new atomic
+propositions $B$ and $C$ for each $s \in B$ and $s \in C$
+respectively. We will also use graphical symbols $\gF C$ ($C$ in
+the future) and $\gG C$ (globally $C$) in such path formulae
+$\varphi$ to minimize confusion.
+
+In order to determine the set $\{ s : K, \neighbor{S}, s \models
+\dpriv{\epsilon}{\delta} \varphi \}$, our algorithm first computes
+the probabilities $p (s) = \myPr{s}{K}{\neighbor{S}}{\varphi}$ for every
+$s \in S$ (Algorithm~\ref{algorithm:sat-dpriv-dtmc}). For each $s \in S$,
+it then compares the probabilities $p (s)$ and $p (t)$ for every
+neighbor $t$ of $s$. If there is a neighbor $t$ such that $p (s)$ and
+$p (t)$ are not $(\epsilon, \delta)$-close, the state $s$
+is removed from the result. Hence
+Algorithm~\ref{algorithm:sat-dpriv-dtmc} returns all states which are
+$(\epsilon, \delta)$-close to their neighbors.
+The algorithm requires at most $O
+(|S|^2)$ additional steps. We hence have the following results:
+
+\begin{algorithm}
+  \begin{algorithmic}[1]
+    \Function{SAT}{$K$, $\neighbor{S}$, $\phi$}
+    \Match{$\phi$}
+    \Comment{by Lemma~\ref{lemma:PJ-subroutines}}
+    \Case{$\X \Psi$}
+      \State{$B \leftarrow \textmd{SAT} (K, \neighbor{S}, \Psi)$}
+      \State{$p(s) \leftarrow \myPr{s}{K}{\neighbor{S}}{\gX B}$
+        for every $s \in S$}
+    \EndCase
+    \Case{$\Psi \until \Psi'$}
+      \State{$B \leftarrow \textmd{SAT} (K, \neighbor{S}, \Psi)$}
+      \State{$C \leftarrow \textmd{SAT} (K, \neighbor{S}, \Psi')$}
+      \State{$p(s) \leftarrow \myPr{s}{K}{\neighbor{S}}{B \guntil C}$ for
+        every $s \in S$}
+    \EndCase
+    \EndMatch
+    \State{$R \leftarrow S$}
+    \For{$s \in S$}
+      \For{$t$ with $s \neighbor{K} t$}
+        \If{$p(s) \not\leq \epsilon p(t) + \delta$ or
+            $p(t) \not\leq \epsilon p(s) + \delta$}
+        {remove $s$ from $R$}
+        \EndIf
+      \EndFor
+    \EndFor
+
+    \State{\Return $R$}
+    \EndFunction
+  \end{algorithmic}
+  \caption{SAT($\phi$, $\neighbor{K}$)}
+  \label{algorithm:sat-dpriv-dtmc}
+\end{algorithm}
+
+\begin{proposition}
+  Let $K = (S, \wp, L)$ be a Markov chain.
+  $\{ s : K, \neighbor{K}, s \models \dpriv{\epsilon}{\delta} \phi \}$ is
+  computable within time polynomial in $|S|$.
+\end{proposition}
+
+\begin{corollary}
+  Let $K = (S, \wp, L)$ be a Markov chain and $\Phi$ a $\dpCTL$
+  formula. $\{ s : K, \neighbor{K}, s \models \Phi \}$ is
+  computable within time polynomial in $|S|$ and $|\Phi|$.
+\end{corollary}
+
 \section{Specifying Properties in $\dpCTL$}
@@
 \paragraph{Advanced Compositional Properties}
+One of the advantage of applying model checking is that we may get tighter bounds for compositional properties.
+Consider our survey example, and the property $\dpriv{1}{.5} (\X Y)$.
+Obviously, it holds in states $+$ and $-$.
+A careful check ensures that one cannot decrease the value of $\epsilon_1=1$ or $\delta_1=.5$ without increasing the other one.
 
-One of the advantage of applying model checking is that we may get tighter bounds for compositional properties. Consider ...
-\lz{find a -- simpler -- formula such that the bound is not tight, maybe $\delta>0$ is needed}
+Now we consider the formula $\dpriv{\epsilon_2}{\delta_2} (\X (Y\wedge \X Y))$ under the model in ???.
+Applying the compositional theorem\lz{Bow-Yaw, as we only refer to the book, I wonder whether using the notion $e^\epsilon$ is easier to explain?}, one can choose $\epsilon_2=\epsilon_1^2=1$ and $\delta_2=2\delta_1=1$.
+However, we can check easily that one can get better privacy parameter $(1,0.5)$ or $(0,\frac{9}{16})$.
 
@@
 \section{Markov Decision Processes}
-\lz{let put it as another section after Markov chains}
 The semantics over Markov chains generalizes to Markov decision
@@
 \end{eqnarray*}
+
+\lz{how about using "query sequence" instead of position-independent scheduler, and then argue that it makes sense for the resulting MDP model from differential privacy?
+This corresponds to a special scheduler that is applied on both neighboring states.
+}
 Recall that $M_{\scheduler{S}}$ is but a Markov chain. The semantics

### 2018-01-26 syntax.tex
first try
added some properties
try

@@
                \PJ{J} \phi \ |\ \dpriv{\epsilon}{\delta} \phi\\
-  \phi & ::= & \X \Phi \ |\ \Phi \until \Phi
+  \phi & ::= & \Phi \ |\ \neg\phi  \ |\  \phi\wedge\phi  \ |\  \gX \phi \ |\ \phi \until \phi
 \end{eqnarray*}
+\lz{Bow-Yaw: we use $\X$ or $\gX$?}
 A \emph{state} formula $\Phi$ is either $\top$, an atomic proposition

### 2018-01-26 applications.tex
motivate pan privacy

@@
 
-In stream
-processing mechanisms, input data are retrieved and analyzed
-continuously.
+Streaming processing gives another motivating example for Markov
+decision processes. In stream processing, input data are retrieved and
+analyzed continuously. Such continual observation is commonly found
+in social media, service providers, or even electronic surveillance
+systems. To protect privacy, differentially private mechanisms have
+been formulated and designed~\cite{DNPR:10:DPCO,DNPRY:10:PPSA}. 
+We model a discrete variant of the mechanisms by Markov decision processes.
 
-Continual observation is commonly found in on-line service
-providers~\cite{DNPR:10:DPCO,DNPRY:10:PPSA}. 
-Consider monitoring events $a$ and $b$ by streaming inputs. We would
-like to compute the number of event types occurred in inputs.
+For illustration, consider monitoring events of the types $a$ and $b$
+continuously. An observation is a sequence of event types. We would
+like to compute the number of event types in observations. Hence for
+observations $a \cdots a$ and $b \cdots b$ of arbitrary positive
+lengths, the query result is $1$; and the query result is $2$ for
+other non-empty observations.
+
+In order to formulate differential privacy, define two obsevations
+are \emph{neighbors} if one can be obtained from the other by replacing
+event types any number of times, that is, if they are of the same
+length. \bw{Later, we will consider a different definition of
+neighborhood by restricting the number of replacements.} Classical
+differential privacy requires the distributions of outcomes $0$,
+$1$, and $2$ are similar on neighboring observations. For streaming
+processing, observations are of indefinite lengths. Data analysis
+algorithms necessarily stay in their internal states indefinitely
+before outcomes are produced. Internal states however may reveal
+private information. When data analysis algorithms are in internal
+states, privacy may still be lost if an intrusion occurs. Classical
+differential privacy is not very useful during continual observation.
+
+In pan privacy, data analysis algorithms are required to respect
+privacy even in their internal states~\cite{DNPR:10:DPCO,DNPRY:10:PPSA}.
 

### 2018-01-26 applications.tex
reorganize presentation

--- 2018-01-26 applications.tex
@@
 
-In stream
-processing mechanisms, input data are retrieved and analyzed
-continuously.
+Streaming processing gives another motivating example for Markov
+decision processes. In stream processing, input data are retrieved and
+analyzed continuously. Such continual observation is commonly found
+in social media, service providers, or even electronic surveillance
+systems. To protect privacy, privacy-respecting mechanisms must be
+formulated and designed. Classical differential privacy however is
+insufficient in streaming processing. Note that observations are of
+indefinite lengths in such applications. All differentially private
+mechanism necessarily stay in their internal states indefinitely
+before outcomes are produced. Although differential privacy demands
+privacy-respecting outcomes, internal states may reveal private
+information in such mechanisms. When differentially private mechanisms
+are in their internal states, privacy is still lost if an intrusion
+occurs.
 
-Continual observation is commonly found in on-line service
-providers~\cite{DNPR:10:DPCO,DNPRY:10:PPSA}. 
-Consider monitoring events $a$ and $b$ by streaming inputs. We would
-like to compute the number of event types occurred in inputs.
+In pan privacy, data analysis mechanisms are required to respect
+privacy even in their internal states~\cite{DNPR:10:DPCO,DNPRY:10:PPSA}. 
+The idea is to show reachability probabilities are similar on
+neighboring observations. In the following, we model a discrete
+variant of the mechanisms by Markov decision processes. We will also
+specify pan privacy on our model using $\dpCTL$.
+
+Consider monitoring events of the types $a$ and $b$ continuously. An
+observation is a sequence of event types. Two obsevations are
+{neighbors} if one can be obtained from the other by replacing 
+event types any number of times, that is, if they are of the same
+length. \bw{Later, we will consider a different definition of
+neighborhood by restricting the number of replacements.} 
+We would like to compute the number of event types in
+observations. Hence for observations $a \cdots a$ and $b \cdots b$ of
+arbitrary positive lengths, the query result is $1$; and the query
+result is $2$ for other non-empty observations.
 

### 2018-01-27 dpCTL.tex
structures and some examples

@@
-The logic $\dpCTL$ is designed to specify differentially private
+The logic $\dpCTLstar$ is designed to specify differentially private
 mechanisms. We introduce the differentially private path quantifer
-$\dpriv{\epsilon}{\delta}$ in $\dpCTL$. For any path formula
+$\dpriv{\epsilon}{\delta}$ in $\dpCTLstar$. For any path formula
 $\phi$, a state $s$ in a Markov chain $K$ satisfies
@@
 having $\phi$-paths from its neighboring states.
-We give semantics for both Markov chains and Markov decision processes.
+%We give semantics for both Markov chains and Markov decision processes.

### 2018-01-27 examples.tex
first try
added some properties
structures and some examples

@@
+As discussed in previous section, a simple way to design differentially private mechanisms is to add
+random noises. 
+Thus, in this section, we discuss how to model differentially private mechanisms using Markov chains. We explain the main idea first with some examples.
+
+\paragraph{Survey Mechanisms.}
+
 \begin{figure}
@@
 
-A simple way to design differentially private mechanisms is to add 
-random noises. Consider the survey question: have you been diagnosed
+Consider the survey question: have you been diagnosed
 with the disease $X$? In order to protect surveyees' privacy, each
@@
 she randomly answers \textit{Yes} or \textit{No}
-uniformly~\cite{DR:14:AFDP}. Figure~\ref{figure:2-dp} shows the survey
+uniformly~\cite{DR:14:AFDP}.
+
+\begin{example}\label{exa:survey}
+  Figure~\ref{figure:2-dp} shows the survey
 mechanism and its corresponding Markov chain. In the figure, the states
@@
 respectively; the states $Y$ and $N$ denote answers to the survey
-question. The random noise boosts the probability of answering
+question. We have $AP=\{Y,N\}$, $L(Y)=\{Y\}$, $L(N)=\{N\}$, and $L(+)=L(-)=\emptyset$. \lz{draw it on the picture?}
+
+The random noise boosts the probability of answering
 \textit{Yes} or \textit{No} to at least $1/4$ regardless of
 diagnoses. Inferences on individual diagnosis can be plausibly denied.
+\end{example}
 
-Let us analyze the survey mechanism briefly. The data universe
+Let us recall briefly the analysis of the survey mechanism. The data universe
 $\mathcal{X}$ is $\{ +, - \}$. The mechanism $M$ is a randomized
@@
 
+\paragraph{Truncated $\alpha$-Geometric Mechanism.}
 More sophisticated differentially private mechanisms are
@@
 
+\subsection{Markov Chains for Differentially Private Mechanisms}
+Now we describe our main ideas for the modelling differentially private mechanisms. 
+A state $s$ stores at least the a database $x\in \mathcal{X}^n$, which is the data hold by a trusted \emph{curator}. A transition corresponds to the data analysis mechanisms.
+States are typically labelled with the corresponding outputs
+the curator generates.
+
+In the differentially privacy context, we differentiate between offline and online models. 
+In the offline, or non-interactive, model the curator releases the outputs only once and plays no further role.
+Thus the models we constructed above are offline models.
+On the contrary, in the online, or interactive, model the curator can permit further queries.
+In this setting it is obvious that queries do not change database themselves.
+To maintain differentially privacy, the curator however can decide to change data analysis mechanisms accordingly.
+For instance, after a certain output is disclosed or a certain number of queries, it may decide to disable further queries.
+
+Below we describe an online model.
+
 \begin{figure}
     \centering
-  \begin{subfigure}{.40\columnwidth}
+    \begin{subfigure}{.4\columnwidth}
 \resizebox{\columnwidth}{!}{
     \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
-      distance=2cm,node/.style={circle,draw}]
-      \node[node] (p) at ( 0,  1.5) { $+$ };
-      \node[node] (q) at ( 0, -1.5) { $-$ };
-      \node[node] (F) at (-1.5,  0) { $Y$ };
-      \node[node] (T) at ( 1.5,  0) { $N$ };
+    distance=1cm,node/.style={circle,draw}]
+    \node[node] (p) at ( -1.5,  0) { $+$ };
+    \node[node] (Y1) at ( 0, .8) { $Y_1$ };
+    \node[node] (N1) at ( 0, -.8) { $N_1$ };
+    \node[node] (Y2) at (1.5,  .8) { $Y_2$ };
+    \node[node] (N2) at ( 1.5, -.8) { $N_2$ };
 
     \path
-      (p) edge [bend right=45] node [left] { $L, .75$ } (F)
-      (p) edge [bend left=45] node [right] { $L, .25$ } (T)
-      (p) edge [bend left=45] node [right=-12,above] { $H, .8$ } (F)
-      (p) edge [bend right=45] node [right=12,above] { $H, .2$ } (T)
-
-      (q) edge [bend left=45] node [left] { $L, .25$ } (F)
-      (q) edge [bend right=45] node [right] { $L, .75$ } (T)
-      (q) edge [bend right=45] node [left=12,below] { $H, .2$ } (F)
-      (q) edge [bend left=45] node [right=12,below] { $H, .8$ } (T)
-
-      (F) edge [loop left] node [below] { $-, 1$ } (F)
-      (T) edge [loop right] node [below] { $-, 1$ } (T)
+    (p) edge node [above] { $\frac{3}{4}$ } (Y1)
+    (p) edge node [below] { $\frac{1}{4}$ } (N1)
+    (Y1) edge node [above] { $\frac{3}{4}$ } (Y2)
+    (Y1) edge node [above] { $\frac{1}{4}$ } (N2)
+    (N1) edge node [below] { $\frac{3}{4}$ } (Y2)
+    (N1) edge node [below] { $\frac{1}{4}$ } (N2)
     ;
@@
 }
-    \caption{Markov Decision Process}
-    \label{figure:simple-mdp-mdp}
+%        \caption{Markov Chain of }
+%        \label{figure:2-dp-mc1}
     \end{subfigure}
-  \begin{subfigure}{.58\columnwidth}
+    \hspace{.05\columnwidth}
+    \begin{subfigure}{.4\columnwidth}
         \resizebox{\columnwidth}{!}{
             \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
-      distance=2cm,node/.style={circle,draw,inner sep=1pt}]
-      \node[node] (pq) at (-2.1, 2.1) { $+-$ };
-      \node[node] (qp) at ( 2.1,-2.1) { $-+$ };
-      \node[node] (RR) at (-2.1,-2.1) { $YY$ };
-      \node[node] (BB) at ( 2.1, 2.1) { $NN$ };
-      \node[node] (RB) at (- .7,- .7) { $YN$ };
-      \node[node] (BR) at (  .7,  .7) { $NY$ };
+    distance=1cm,node/.style={circle,draw}]
+    \node[node] (n) at ( -1.5,  0) { $-$ };
+    \node[node] (Y1) at ( 0, .8) { $Y_1$ };
+    \node[node] (N1) at ( 0, -.8) { $N_1$ };
+    \node[node] (Y2) at (1.5,  .8) { $Y_2$ };
+    \node[node] (N2) at ( 1.5, -.8) { $N_2$ };
 
     \path
-      (pq) edge [bend right=5] node [left=1,below=12,rotate=270] { $L, .1875$ } (RR)
-      (pq) edge [bend left=5]  node [above=5,right=1] { $L, .1875$ } (BB)
-
-      (pq) edge [bend left=5]  node [right=4,rotate=270] { $H, .16$ } (RR)
-      (pq) edge [bend right=5] node [below=5,right=1] { $H, .16$ } (BB)
-
-      (pq) edge [bend left=5]  node [left,below=10,rotate=295]  { $L, .5625$ } (RB)
-      (pq) edge [bend right=5] node [left,below,rotate=330]  { $L, .0625$ } (BR)
-
-      (pq) edge [bend right=5] node [right=2,above,rotate=300] { $H, .64$ } (RB)
-      (pq) edge [bend left=5]  node [right,above,rotate=340] { $H, .04$ } (BR)
-
-      (qp) edge [bend right=5] node [above=5,left=1] { $L, .1875$ } (RR)
-      (qp) edge [bend left=5]  node [above=30,left=5,rotate=90] { $L, .1875$ } (BB)
-
-      (qp) edge [bend left=5]  node [below=5,left=1] { $H, .16$ } (RR)
-      (qp) edge [bend right=5] node [above,right=5,rotate=90] { $H, .16$ } (BB)
-
-      (qp) edge [bend left=5]  node [left,below,rotate=330]  { $L, .0625$ } (RB)
-      (qp) edge [bend right=5] node [left=3,below=3,rotate=295]  { $L, .5625$ } (BR)
-
-      (qp) edge [bend right=5] node [left=5,above,rotate=330] { $H, .04$ } (RB)
-      (qp) edge [bend left=5]  node [right=1,above=1,rotate=295] { $H, .64$ } (BR)
-
-      (RR) edge [loop left] node [above] { $-, 1$ } (RR)
-      (BB) edge [loop right] node [above] { $-, 1$ } (BB)
-      (RB) edge [loop below] node [left] { $-, 1$ } (RB)
-      (BR) edge [loop above] node [right] { $-, 1$ } (BR)
-      ;
-      \end{tikzpicture}
-\hide{
-    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
-      distance=2cm,node/.style={circle,draw,inner sep=1pt}]
-      \node[node] (pq) at (   0,   0) { $+-$ };
-      \node[node] (RR) at (-  2,   0) { $YY$ };
-      \node[node] (BB) at (   2,   0) { $NN$ };
-      \node[node] (RB) at (   0, 1.5) { $YN$ };
-      \node[node] (BR) at (   0,-1.5) { $NY$ };
-
-      \path
-      (pq) edge [bend right=15] node [above] { $L, .1875$ } (RR)
-      (pq) edge [bend left=15]  node [above] { $L, .1875$ } (BB)
-
-      (pq) edge [bend left=15]  node [below] { $H, .16$ } (RR)
-      (pq) edge [bend right=15] node [below] { $H, .16$ } (BB)
-
-      (pq) edge [bend left=15]  node [left]  { $L, .5625$ } (RB)
-      (pq) edge [bend right=15] node [left]  { $L, .0625$ } (BR)
-
-      (pq) edge [bend right=15] node [right] { $H, .64$ } (RB)
-      (pq) edge [bend left=15]  node [right] { $H, .04$ } (BR)
-
-      (RR) edge [loop below] node [below] { $-, 1$ } (RR)
-      (BB) edge [loop above] node [above] { $-, 1$ } (BB)
-      (RB) edge [loop left] node [left] { $-, 1$ } (RB)
-      (BR) edge [loop right] node [right] { $-, 1$ } (BR)
+    (n) edge node [above] { $\frac{1}{4}$ } (Y1)
+    (n) edge node [below] { $\frac{3}{4}$ } (N1)
+    (Y1) edge node [above] { $\frac{1}{4}$ } (Y2)
+    (Y1) edge node [above] { $\frac{3}{4}$ } (N2)
+    (N1) edge node [below] { $\frac{1}{4}$ } (Y2)
+    (N1) edge node [below] { $\frac{3}{4}$ } (N2)
     ;
@@
         }
-    }
-    \caption{Product Markov Decision Process}
-    \label{figure:simple-mdp-product}
+%        \caption{Markov Chain of}
+%        \label{figure:2-dp-mc2}
     \end{subfigure}
-  \caption{Markov Decision Process and its Self-Product}
-  \label{figure:simple-mdp}
+    \caption{Markov Chain of }
+    \label{figure:2-dp-mc}
 \end{figure}
+\begin{example}
+  First, assume that the curator allows at most two survey queries.
+The corresponding model is depicted in Figure~\ref{figure:2-dp-mc}, where states $Y_1,Y_2$ are labelled by $Y$, and $N_1,N_2$ are labelled by $N$.
+\end{example}
 
-Mechanisms are not necessarily closed randomized algorithms; they may
-perform different compution on users' requests. 
-We use Markov decision processes to model such 
-interactive mechanisms. Specifically, external inputs are modeled by
-actions. Behaviors associated with different inputs are modeled by
-distributions associated with actions.
-
-Consider again the survey mechanism. Suppose we would like to design
-an interactive mechanism which adjusts random noises on surveyers'
-requests. When the surveyer requests low-accuracy answers, the
-surveyee uses the survey mechanism as before. When high-accuracy
-answers are requested, the surveyee answers \textit{Yes} with
-probability $4/5$ and \textit{No} with probability $1/5$ when she has
-positive diagnosis. She answers \textit{Yes} with probability $1/5$
-and \textit{No} with probability $4/5$ when she is not
-diagnosed with the disease X. This gives an interactive mechanism
-corresponding to the Markov decision process shown in
-Figure~\ref{figure:simple-mdp-mdp}.
-
-In the figure, the states $+$,
-$-$, $Y$, and $N$ are interpreted as before. The actions $L$ and $H$
-denote low- and high-accuracy requests respectively. Let $M_H$ denote
-the Markov chain derived by high-accuracy requests.
-Observe that $1/5 \leq \Pr[M_H(x) = Y] \leq 4/5$ for any $x
-\in \mathcal{X}$. Hence $\Pr[M_H (x) = Y] \leq 4/5 = 4 \cdot 1/5 \leq
-4 \Pr[M_H (x') = Y]$ for any neighboring $x, x' \in
-\mathcal{X}$. Similarly, $\Pr[M_H (x) = N] \leq 4 \Pr[M_H (x') =
-N]$. The high-accuracy mechanism is $(4,0)$-differentially private. 
-The privacy guarantees vary from accuracy requests.
+\lz{tbd, the complex threshold example we discussed, by Depeng}

### 2018-01-27 mdp.tex
structures and some examples

\subsection{Markov Decision Processes}

\begin{figure}
  \centering
  \begin{subfigure}{.40\columnwidth}
    \resizebox{\columnwidth}{!}{
    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
      distance=2cm,node/.style={circle,draw}]
      \node[node] (p) at ( 0,  1.5) { $+$ };
      \node[node] (q) at ( 0, -1.5) { $-$ };
      \node[node] (F) at (-1.5,  0) { $Y$ };
      \node[node] (T) at ( 1.5,  0) { $N$ };

      \path
      (p) edge [bend right=45] node [left] { $L, .75$ } (F)
      (p) edge [bend left=45] node [right] { $L, .25$ } (T)
      (p) edge [bend left=45] node [right=-12,above] { $H, .8$ } (F)
      (p) edge [bend right=45] node [right=12,above] { $H, .2$ } (T)

      (q) edge [bend left=45] node [left] { $L, .25$ } (F)
      (q) edge [bend right=45] node [right] { $L, .75$ } (T)
      (q) edge [bend right=45] node [left=12,below] { $H, .2$ } (F)
      (q) edge [bend left=45] node [right=12,below] { $H, .8$ } (T)

      (F) edge [loop left] node [below] { $-, 1$ } (F)
      (T) edge [loop right] node [below] { $-, 1$ } (T)
      ;
      \end{tikzpicture}
    }
    \caption{Markov Decision Process}
    \label{figure:simple-mdp-mdp}
  \end{subfigure}
  \begin{subfigure}{.58\columnwidth}
    \resizebox{\columnwidth}{!}{
    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
      distance=2cm,node/.style={circle,draw,inner sep=1pt}]
      \node[node] (pq) at (-2.1, 2.1) { $+-$ };
      \node[node] (qp) at ( 2.1,-2.1) { $-+$ };
      \node[node] (RR) at (-2.1,-2.1) { $YY$ };
      \node[node] (BB) at ( 2.1, 2.1) { $NN$ };
      \node[node] (RB) at (- .7,- .7) { $YN$ };
      \node[node] (BR) at (  .7,  .7) { $NY$ };

      \path
      (pq) edge [bend right=5] node [left=1,below=12,rotate=270] { $L, .1875$ } (RR)
      (pq) edge [bend left=5]  node [above=5,right=1] { $L, .1875$ } (BB)

      (pq) edge [bend left=5]  node [right=4,rotate=270] { $H, .16$ } (RR)
      (pq) edge [bend right=5] node [below=5,right=1] { $H, .16$ } (BB)

      (pq) edge [bend left=5]  node [left,below=10,rotate=295]  { $L, .5625$ } (RB)
      (pq) edge [bend right=5] node [left,below,rotate=330]  { $L, .0625$ } (BR)

      (pq) edge [bend right=5] node [right=2,above,rotate=300] { $H, .64$ } (RB)
      (pq) edge [bend left=5]  node [right,above,rotate=340] { $H, .04$ } (BR)

      (qp) edge [bend right=5] node [above=5,left=1] { $L, .1875$ } (RR)
      (qp) edge [bend left=5]  node [above=30,left=5,rotate=90] { $L, .1875$ } (BB)

      (qp) edge [bend left=5]  node [below=5,left=1] { $H, .16$ } (RR)
      (qp) edge [bend right=5] node [above,right=5,rotate=90] { $H, .16$ } (BB)

      (qp) edge [bend left=5]  node [left,below,rotate=330]  { $L, .0625$ } (RB)
      (qp) edge [bend right=5] node [left=3,below=3,rotate=295]  { $L, .5625$ } (BR)

      (qp) edge [bend right=5] node [left=5,above,rotate=330] { $H, .04$ } (RB)
      (qp) edge [bend left=5]  node [right=1,above=1,rotate=295] { $H, .64$ } (BR)

      (RR) edge [loop left] node [above] { $-, 1$ } (RR)
      (BB) edge [loop right] node [above] { $-, 1$ } (BB)
      (RB) edge [loop below] node [left] { $-, 1$ } (RB)
      (BR) edge [loop above] node [right] { $-, 1$ } (BR)
      ;
      \end{tikzpicture}
\hide{
    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
      distance=2cm,node/.style={circle,draw,inner sep=1pt}]
      \node[node] (pq) at (   0,   0) { $+-$ };
      \node[node] (RR) at (-  2,   0) { $YY$ };
      \node[node] (BB) at (   2,   0) { $NN$ };
      \node[node] (RB) at (   0, 1.5) { $YN$ };
      \node[node] (BR) at (   0,-1.5) { $NY$ };

      \path
      (pq) edge [bend right=15] node [above] { $L, .1875$ } (RR)
      (pq) edge [bend left=15]  node [above] { $L, .1875$ } (BB)

      (pq) edge [bend left=15]  node [below] { $H, .16$ } (RR)
      (pq) edge [bend right=15] node [below] { $H, .16$ } (BB)

      (pq) edge [bend left=15]  node [left]  { $L, .5625$ } (RB)
      (pq) edge [bend right=15] node [left]  { $L, .0625$ } (BR)

      (pq) edge [bend right=15] node [right] { $H, .64$ } (RB)
      (pq) edge [bend left=15]  node [right] { $H, .04$ } (BR)

      (RR) edge [loop below] node [below] { $-, 1$ } (RR)
      (BB) edge [loop above] node [above] { $-, 1$ } (BB)
      (RB) edge [loop left] node [left] { $-, 1$ } (RB)
      (BR) edge [loop right] node [right] { $-, 1$ } (BR)
      ;
      \end{tikzpicture}
}
    }
    \caption{Product Markov Decision Process}
    \label{figure:simple-mdp-product}
  \end{subfigure}
  \caption{Markov Decision Process and its Self-Product}
  \label{figure:simple-mdp}
\end{figure}

Mechanisms are not necessarily closed randomized algorithms; they may
perform different compution on users' requests.
We use Markov decision processes to model such
interactive mechanisms. Specifically, external inputs are modeled by
actions. Behaviors associated with different inputs are modeled by
distributions associated with actions.

Consider again the survey mechanism. Suppose we would like to design
an interactive mechanism which adjusts random noises on surveyers'
requests. When the surveyer requests low-accuracy answers, the
surveyee uses the survey mechanism as before. When high-accuracy
answers are requested, the surveyee answers \textit{Yes} with
probability $4/5$ and \textit{No} with probability $1/5$ when she has
positive diagnosis. She answers \textit{Yes} with probability $1/5$
and \textit{No} with probability $4/5$ when she is not
diagnosed with the disease X. This gives an interactive mechanism
corresponding to the Markov decision process shown in
Figure~\ref{figure:simple-mdp-mdp}.

In the figure, the states $+$,
$-$, $Y$, and $N$ are interpreted as before. The actions $L$ and $H$
denote low- and high-accuracy requests respectively. Let $M_H$ denote
the Markov chain derived by high-accuracy requests.
Observe that $1/5 \leq \Pr[M_H(x) = Y] \leq 4/5$ for any $x
\in \mathcal{X}$. Hence $\Pr[M_H (x) = Y] \leq 4/5 = 4 \cdot 1/5 \leq
4 \Pr[M_H (x') = Y]$ for any neighboring $x, x' \in
\mathcal{X}$. Similarly, $\Pr[M_H (x) = N] \leq 4 \Pr[M_H (x') =
N]$. The high-accuracy mechanism is $(4,0)$-differentially private.
The privacy guarantees vary from accuracy requests.

\subsection{Semantics}
The semantics over Markov chains generalizes to Markov decision
processes as well. Note that two states with different enabled actions
are trivially distinguishable; no privacy can be preserved. Without
loss of generality, we consider reactive Markov decision processes.
Let $M = (S, \Act, \wp, L)$ be a reactive Markov
decision process. Define the satisfaction
relation $M, \neighbor{S}, s \models \Phi$ as follows.
\begin{eqnarray*}
  M, \neighbor{S}, s \models \top\\
  M, \neighbor{S}, s \models p
  & \textmd{ if } &
  p \in L(s)\\
  M, \neighbor{S}, s \models \neg \Phi
  & \textmd{ if } &
  M, \neighbor{S}, s \not\models \Phi\\
  M, \neighbor{S}, s \models \Phi_0 \wedge \Phi_1
  & \textmd{ if } &
  M, \neighbor{S}, s \models \Phi_0 \textmd{ and }
  M, \neighbor{S}, s \models \Phi_1\\
  M, \neighbor{S}, s \models \PJ{J} \phi
  & \textmd{ if } &
  \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \in J
  \textmd{ for every scheduler } \scheduler{S}\\
  M, \neighbor{S}, s \models \dpriv{\epsilon}{\delta} \phi
  & \textmd{ if } &
  \textmd{for all } t \textmd{ with } s \neighbor{S} t \textmd{ and
   position-independent scheduler } \\
  & & \scheduler{S},
   \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq \epsilon \cdot
   \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} + \delta
   \textmd{ and }
   \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq \epsilon \cdot
  \\
  & &
   \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} + \delta
\end{eqnarray*}

\lz{how about using "query sequence" instead of position-independent scheduler, and then argue that it makes sense for the resulting MDP model from differential privacy? This corresponds to a special scheduler that is applied on both neighboring states. }
Recall that $M_{\scheduler{S}}$ is but a Markov chain. The semantics
of $M_{\scheduler{S}}, \neighbor{S}, \pi \models \phi$ and hence the
probability $\myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}$ are
defined as in Markov chains.
The semantics of $\dpCTL$ on reactive Markov decision processes
is again standard except the differentially private operator
$\dpriv{\epsilon}{\delta}$. For any path formula $\phi$,
$\dpriv{\epsilon}{\delta} \phi$ specifies states which are $(\epsilon,
\delta)$-close to all its neighbors in having paths satisfying $\phi$
for arbitrary position-independent schedulers. That is, no
position-independent scheduler can force any of neighbors to have
probabilistically distinguishable behaviors.

\noindent
\emph{Discussion.}
Another definition of the differentially private operator
might be:
\begin{eqnarray*}
  M, \neighbor{S}, s \models \dpriv{\epsilon}{\delta}^{\mathit{bad}} \phi
  & \textmd{ if } &
  \textmd{for all } t \textmd{ with } s \neighbor{S} t \textmd{ and
  scheduler } \scheduler{S},
  \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq
  \epsilon \cdot\\
  && \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} + \delta \textmd{ and }
  \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq
  \epsilon \cdot \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}
  + \delta
\end{eqnarray*}
A state satisfies $\dpriv{\epsilon}{\delta}^{\mathit{bad}} \phi$ if
no scheduler could differentiate the probabilities of paths satisfying
$\phi$ from neighbors. Such a definition would allow strategies to
take different actions from different states. It might need different
action sequences to distinguish neighbors. Two neighbors could still be
differentiated by a action sequence and lose their privacy. We hence
prefer the original definition.

\hide{
Since the scheduler attaining the maximally probable
behavior from a state may be different from the scheduler attaining
the maximally probable behavior from its neighbors, two
neighbors may still be distinguished by a scheduler. The
weaker definition does not preserve differential privacy. We hence
prefer the original definition.
}

### 2018-01-27 semantics.tex
structures and some examples

@@
 Other than the differentially private operator, the semantics of
-$\dpCTL$ is standard~\cite{BK:08:PMC}.
+$\dpCTLstar$ is standard~\cite{BK:08:PMC}.
 To intuit the semantics of $\dpriv{\epsilon}{\delta} \phi$,
@@
 \section{Model Checking Algorithms}
-The model checking algorithms for $\dpCTL$ follow the classical
+We first describe the model checking algorithm for $\dpCTL$, which follows the classical
 algorithms for $\PCTL$ by computing the states satisfying
@@
 \end{corollary}
-
-\section{Specifying Properties in $\dpCTL$}
-In this section we describe what type of properties in the differential privacy context can be expressed using $\dpCTL$.
-
-\paragraph{Standard Differential Privacy.} 
-The formula $\dpriv{3}{0} (\X Y)$ holds in state $s$ if for  $t$ with $sNt$, the probability of satisfying $\X Y$ from $s$ and $t$ satisfy:
-$\myPr{s}{K}{N}{\X Y}\le 3\myPr{t}{K}{N}{\X Y}$.
-The formula $\dpriv{3}{0} (\X Y)\wedge \dpriv{3}{0} (\X N)$ thus specifies the differential privacy property for states $+$ and $-$ in Example~\ref{exa:survey}.
-
-Further, consider the Markov chain in
-Figure~\ref{figure:geometric-mechanism-markov-chain}. Assume
-$L(out_i) = \{ out_i \}$ and $\Delta (f) = 1$. Then $in_k$ and $in_l$ are
-neighbors iff $| k - l | \leq 1$\lz{put above sentence to the def of the model}. Define the $\dpCTL$ formula
-$\psi = \dpriv{.5}{0} (\F out_0) \wedge \dpriv{.5}{0} (\F out_1) \wedge
-\cdots \wedge \dpriv{.5}{0} (\F out_5)$\lz{I think we better use $\X$ operator here ;)}. If the state $in_k$ satisfies
-$\psi$ for $k = 0, \ldots, 5$, then the $.5$-geometric mechanism is
-$(0.5, 0)$-differentially private.
-
-\paragraph{Compositional Properties.}
-Compositional aspect is one of the building block for differentially private algorithms. According to the compositional theorem \cite{}, for $(\epsilon_1,0)$-differentially private algorithm $M_1$ and
-$(\epsilon_2,0)$-differentially private algorithm $M_2$, their combination $M$ defined by $M(x)=(M_1(x), M_2(x))$ is $(\epsilon_1\epsilon_2,0)$-differentially private.
-The degradation is rooted in the repeated computation.
-To illustrate this property, we consider the extended example below \lz{here example of the survey example, expanded twice: we should discuss the graph before it is drawn...}
-
-We consider the formula $\dpriv{9}{0} (\X (Y\wedge \X Y))$.
-A path satisfies $\X (Y\wedge \X Y)$ if the second state satisfies $Y$ and the third state satisfies $Y$ as well.
-We verify that this formula is satisfied  for state $+$ and $-$.
-Moreover, the bound $\epsilon=9$ is bound, since \lz{tbd once graph is given}.
-Finally, the formula $\wedge_{a_1,a_2}\dpriv{9}{0} (\X (a_1\wedge \X a_2))$ corresponds to the desired compositional property for the model, where $a_1,a_2$ range over all atomic propositions $\{Y,N\}$.
-
-We consider another simpler formula for comparision: $\dpriv{3}{0} (\X \X Y)$.
-In this case we donot have privacy lost, even though we are querying twice: the reason is that the output of the first query is not utilized at all.
-It is easy to verify that it is indeed satisfied by $+$ and $-$.
-
-\paragraph{Advanced Compositional Properties}
-One of the advantage of applying model checking is that we may get tighter bounds for compositional properties.
-Consider our survey example, and the property $\dpriv{1}{.5} (\X Y)$.
-Obviously, it holds in states $+$ and $-$.
-A careful check ensures that one cannot decrease the value of $\epsilon_1=1$ or $\delta_1=.5$ without increasing the other one.
-
-Now we consider the formula $\dpriv{\epsilon_2}{\delta_2} (\X (Y\wedge \X Y))$ under the model in ???.
-Applying the compositional theorem\lz{Bow-Yaw, as we only refer to the book, I wonder whether using the notion $e^\epsilon$ is easier to explain?}, one can choose $\epsilon_2=\epsilon_1^2=1$ and $\delta_2=2\delta_1=1$.
-However, we can check easily that one can get better privacy parameter $(1,0.5)$ or $(0,\frac{9}{16})$.
-
-\paragraph{Infinite ...}
-Note that $\dpCTL$ generalizes differential privacy to infinite
-behaviors naturally. Classical differential privacy discusses properties about
-randomized terminating mechanisms. Sophisticated mechanisms for
-streaming data have also been discussed in social networks and
-internet of things. Such mechanisms inherently have infinite
-computation. Their privacy guarantees are approximated by those of the
-prefixes of computation.
-With \dpCTL, the subformula $\phi$ in $\dpriv{\epsilon}{\delta} \phi$
-can be an arbitrary $\dpCTL$ path formula. One can specify
-$(\epsilon, \delta)$-closedness over infinite behaviors naturally. For instance,
-$\dpriv{\epsilon}{\delta} (\G \PJ{[1,1]} (\F \mathit{high}))$ \lz{now we can use $G F high$}
-specifies states having $(\epsilon, \delta)$-close
-infinitely often $\mathit{high}$ behaviors to its neighbors.
-
-\section{Markov Decision Processes}
-The semantics over Markov chains generalizes to Markov decision
-processes as well. Note that two states with different enabled actions
-are trivially distinguishable; no privacy can be preserved. Without
-loss of generality, we consider reactive Markov decision processes.
-Let $M = (S, \Act, \wp, L)$ be a reactive Markov
-decision process. Define the satisfaction
-relation $M, \neighbor{S}, s \models \Phi$ as follows.
-\begin{eqnarray*}
-  M, \neighbor{S}, s \models \top\\
-  M, \neighbor{S}, s \models p
-  & \textmd{ if } &
-  p \in L(s)\\
-  M, \neighbor{S}, s \models \neg \Phi
-  & \textmd{ if } &
-  M, \neighbor{S}, s \not\models \Phi\\
-  M, \neighbor{S}, s \models \Phi_0 \wedge \Phi_1
-  & \textmd{ if } &
-  M, \neighbor{S}, s \models \Phi_0 \textmd{ and }
-  M, \neighbor{S}, s \models \Phi_1\\
-  M, \neighbor{S}, s \models \PJ{J} \phi
-  & \textmd{ if } &
-  \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \in J
-  \textmd{ for every scheduler } \scheduler{S}\\
-  M, \neighbor{S}, s \models \dpriv{\epsilon}{\delta} \phi
-  & \textmd{ if } &
-  \textmd{for all } t \textmd{ with } s \neighbor{S} t \textmd{ and
-   position-independent scheduler } \\
-  & & \scheduler{S},
-   \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq \epsilon \cdot
-   \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} + \delta
-   \textmd{ and }
-   \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq \epsilon \cdot
-  \\
-  & &
-   \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} + \delta
-\end{eqnarray*}
-
-\lz{how about using "query sequence" instead of position-independent scheduler, and then argue that it makes sense for the resulting MDP model from differential privacy?
-This corresponds to a special scheduler that is applied on both neighboring states.
-}
-Recall that $M_{\scheduler{S}}$ is but a Markov chain. The semantics
-of $M_{\scheduler{S}}, \neighbor{S}, \pi \models \phi$ and hence the
-probability $\myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}$ are
-defined as in Markov chains.
-The semantics of $\dpCTL$ on reactive Markov decision processes
-is again standard except the differentially private operator
-$\dpriv{\epsilon}{\delta}$. For any path formula $\phi$,
-$\dpriv{\epsilon}{\delta} \phi$ specifies states which are $(\epsilon,
-\delta)$-close to all its neighbors in having paths satisfying $\phi$
-for arbitrary position-independent schedulers. That is, no
-position-independent scheduler can force any of neighbors to have
-probabilistically distinguishable behaviors.
-
-\noindent
-\emph{Discussion.}
-Another definition of the differentially private operator
-might be:
-\begin{eqnarray*}
-  M, \neighbor{S}, s \models \dpriv{\epsilon}{\delta}^{\mathit{bad}} \phi
-  & \textmd{ if } &
-  \textmd{for all } t \textmd{ with } s \neighbor{S} t \textmd{ and
-  scheduler } \scheduler{S},
-  \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq
-  \epsilon \cdot\\
-  && \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} + \delta \textmd{ and }
-  \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq
-  \epsilon \cdot \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}
-  + \delta
-\end{eqnarray*}
-A state satisfies $\dpriv{\epsilon}{\delta}^{\mathit{bad}} \phi$ if
-no scheduler could differentiate the probabilities of paths satisfying
-$\phi$ from neighbors. Such a definition would allow strategies to
-take different actions from different states. It might need different
-action sequences to distinguish neighbors. Two neighbors could still be
-differentiated by a action sequence and lose their privacy. We hence
-prefer the original definition.
-
-\hide{
-Since the scheduler attaining the maximally probable
-behavior from a state may be different from the scheduler attaining
-the maximally probable behavior from its neighbors, two
-neighbors may still be distinguished by a scheduler. The
-weaker definition does not preserve differential privacy. We hence
-prefer the original definition.
-}

### 2018-01-27 specifications.tex
structures and some examples

In this section we describe what type of properties in the differential privacy context can be expressed using $\dpCTL$.

\paragraph{Standard Differential Privacy.}
The formula $\dpriv{3}{0} (\X Y)$ holds in state $s$ if for  $t$ with $sNt$, the probability of satisfying $\X Y$ from $s$ and $t$ satisfy:
$\myPr{s}{K}{N}{\X Y}\le 3\myPr{t}{K}{N}{\X Y}$. The formula $\dpriv{3}{0} (\X Y)\wedge \dpriv{3}{0} (\X N)$ thus specifies the differential privacy property for states $+$ and $-$ in Example~\ref{exa:survey}.

Further, consider the Markov chain in
Figure~\ref{figure:geometric-mechanism-markov-chain}. Assume
$L(out_i) = \{ out_i \}$ and $\Delta (f) = 1$. Then $in_k$ and $in_l$ are
neighbors iff $| k - l | \leq 1$\lz{put above sentence to the def of the model}. Define the $\dpCTL$ formula
$\psi = \dpriv{.5}{0} (\F out_0) \wedge \dpriv{.5}{0} (\F out_1) \wedge
\cdots \wedge \dpriv{.5}{0} (\F out_5)$\lz{I think we better use $\X$ operator here ;)}. If the state $in_k$ satisfies
$\psi$ for $k = 0, \ldots, 5$, then the $.5$-geometric mechanism is
$(0.5, 0)$-differentially private.

\paragraph{Compositional Properties.}
Compositional aspect is one of the building block for differentially private algorithms. According to the compositional theorem \cite{}, for $(\epsilon_1,0)$-differentially private algorithm $M_1$ and
$(\epsilon_2,0)$-differentially private algorithm $M_2$, their combination $M$ defined by $M(x)=(M_1(x), M_2(x))$ is $(\epsilon_1\epsilon_2,0)$-differentially private. The degradation is rooted in the repeated computation. To illustrate this property, we consider the extended example below \lz{here example of the survey example, expanded twice: we should discuss the graph before it is drawn...}

We consider the formula $\dpriv{9}{0} (\X (Y\wedge \X Y))$. A path satisfies $\X (Y\wedge \X Y)$ if the second state satisfies $Y$ and the third state satisfies $Y$ as well. We verify that this formula is satisfied  for state $+$ and $-$. Moreover, the bound $\epsilon=9$ is bound, since \lz{tbd once graph is given}. Finally, the formula $\wedge_{a_1,a_2}\dpriv{9}{0} (\X (a_1\wedge \X a_2))$ corresponds to the desired compositional property for the model, where $a_1,a_2$ range over all atomic propositions $\{Y,N\}$.

We consider another simpler formula for comparision: $\dpriv{3}{0} (\X \X Y)$. In this case we donot have privacy lost, even though we are querying twice: the reason is that the output of the first query is not utilized at all. It is easy to verify that it is indeed satisfied by $+$ and $-$.

\paragraph{Tighter Privacy Bounds for Compositional Properties}
One of the advantage of applying model checking is that we may get tighter bounds for compositional properties. Consider our survey example, and the property $\dpriv{1}{.5} (\X Y)$. Obviously, it holds in states $+$ and $-$. A careful check ensures that one cannot decrease the value of $\epsilon_1=1$ or $\delta_1=.5$ without increasing the other one.

Now we consider the formula $\dpriv{\epsilon_2}{\delta_2} (\X (Y\wedge \X Y))$ under the model in ???. Applying the compositional theorem\lz{Bow-Yaw, as we only refer to the book, I wonder whether using the notion $e^\epsilon$ is easier to explain?}, one can choose $\epsilon_2=\epsilon_1^2=1$ and $\delta_2=2\delta_1=1$. However, we can check easily that one can get better privacy parameter $(1,0.5)$ or $(0,\frac{9}{16})$.

\paragraph{Infinite ...}
Note that $\dpCTL$ generalizes differential privacy to infinite
behaviors naturally. Classical differential privacy discusses properties about
randomized terminating mechanisms. Sophisticated mechanisms for
streaming data have also been discussed in social networks and
internet of things. Such mechanisms inherently have infinite
computation. Their privacy guarantees are approximated by those of the
prefixes of computation.
With \dpCTL, the subformula $\phi$ in $\dpriv{\epsilon}{\delta} \phi$
can be an arbitrary $\dpCTL$ path formula. One can specify
$(\epsilon, \delta)$-closedness over infinite behaviors naturally. For instance,
$\dpriv{\epsilon}{\delta} (\G \PJ{[1,1]} (\F \mathit{high}))$ \lz{now we can use $G F high$}
specifies states having $(\epsilon, \delta)$-close
infinitely often $\mathit{high}$ behaviors to its neighbors.

### 2018-01-27 syntax.tex
structures and some examples

--- 2018-01-26 syntax.tex
@@
                \PJ{J} \phi \ |\ \dpriv{\epsilon}{\delta} \phi\\
-  \phi & ::= & \X \Phi \ |\ \Phi \until \Phi
+  \phi & ::= & \Phi \ |\ \neg\phi  \ |\  \phi\wedge\phi  \ |\  \X \phi \ |\ \phi \until \phi
 \end{eqnarray*}
@@
  by a path formula. A
-\emph{path} formula $\phi$ is a \emph{next} operator ($\X$)
-followed by a state formula, or an \emph{until} operator
-($\until$) operator enclosed by two state formulae.
+\emph{path} formula $\phi$ is simply a LTL formula, with temporal operator \emph{next}  ($\X$)
+and  \emph{until} operator
+($\until$) operator enclosed by two path formulae.
+
+As in the classical setting, we consider the sublogic $\dpCTL$ by allowing only path formulae of the form $\X\Phi$ and $\Phi\until\Phi$.

### 2018-01-27 mdp.tex
more changes ;)

@@
 
+\paragraph{Model Checking Algorithm.}
+
+Our next task is to compute $\{ M, \neighbor{S}, s \models
+\dpriv{\epsilon}{\delta} \phi \}$ for a Markov decision process $M =
+(S, \Act, \wp, L)$. Recall the semantics of $\dpriv{\epsilon}{\delta}
+\phi$. Given $s, t$ with $s \neighbor{M} t$ and a path formula
+$\phi$, we need to decide whether
+$\myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq
+\epsilon \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} + \delta$
+for every scheduler $\scheduler{S}$.
+If the path formula only contain next operators, it is easy to deal with.
+For $\phi:=\X B$ with $B\subseteq S$ only one step behaviour needs to be considered, thus only the first action in the query sequence is needed to compute the maximum
+\begin{eqnarray}
+  \label{eqn:max-nextB}
+\max_{\scheduler{S}}
+\myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\gX B} -
+\epsilon \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\gX B}
+\end{eqnarray}
+with $s, t \in S$. It can be easily generalized to nested next operators:  one needs only to enumerate 
+all actions query sequences of length obtained by the number of nexted $\X$ operators.
+
+In general, the computation of the maximal probability for MDPs is non-trivial. \lz{we argue that it is difficult and leave as future work}
+
+Finally, we the property 
+$\wedge_{i\in N} \dpriv{\epsilon}{0}(\X^k \top) \wedge \dpriv{\epsilon}{0}(F \top)$ that is used to express the privacy for our threshold Markov chains.
+It is only a valid formula if we would allow infinite conjunctions.
+However, it is interesting to note the relation to language equivalence for probabilistic automata\lz{to be discussed}
+
 \hide{

### 2018-01-27 semantics.tex
more changes ;)

@@
 \end{corollary}
+
+The model checking algorithm for $\dpCTLstar$ can be treated as in the classical setting \cite{BK:08:PMC}, as all we need is to  compute the the probability
+  $\myPr{s}{K}{\neighbor{S}}{\phi}$ with general path formula $\phi$, one first construct a deterministic $\omega$-automaton $R$, such as deterministic Rabin automaton.
+Then, the probability reduces to a
+  reachability probability in the product Markov chain obtained from $K$ and $R$. There are more efficient algorithms without the product construction, see \cite{}. We skip the details.

### 2018-01-27 specifications.tex
more changes ;)

@@
 
-Further, consider the Markov chain in
-Figure~\ref{figure:geometric-mechanism-markov-chain}. Assume
-$L(out_i) = \{ out_i \}$ and $\Delta (f) = 1$. Then $in_k$ and $in_l$ are
-neighbors iff $| k - l | \leq 1$\lz{put above sentence to the def of the model}. Define the $\dpCTL$ formula
+ Define the $\dpCTL$ formula
 $\psi = \dpriv{.5}{0} (\F out_0) \wedge \dpriv{.5}{0} (\F out_1) \wedge
@@
 
-\paragraph{Tighter Privacy Bounds for Compositional Properties}
+\paragraph{Tighter Privacy Bounds for Compositional Properties.}
 One of the advantage of applying model checking is that we may get tighter bounds for compositional properties.
@@
 However, we can check easily that one can get better privacy parameter $(1,0.5)$ or $(0,\frac{9}{16})$.
+
+\paragraph{Reachability Privacy} 
+For our threshold \lz{to do} model, we have the desired privacy property $\wedge_{i\in N} \dpriv{\epsilon}{0}(\X^k \top) \wedge \dpriv{\epsilon}{0}(F \top)$.
+Here $\dpriv{\epsilon}{0}(F \top)$ is a reachability privacy: it states that related states should have similar probabilities of reaching $\top$.
 

### 2018-01-27 examples.tex
more changes ;)
threshold example

@@
 respectively; the states $Y$ and $N$ denote answers to the survey
-question. We have $AP=\{Y,N\}$, $L(Y)=\{Y\}$, $L(N)=\{N\}$, and $L(+)=L(-)=\emptyset$. \lz{draw it on the picture?}
+question. We have $AP=\{Y,N\}$, $L(Y)=\{Y\}$, $L(N)=\{N\}$, and $L(+)=L(-)=\emptyset$. States $+$ and $-$ are neighbors.\lz{draw it on the picture?}
 
@@
 straightfoward to model the truncated $1/2$-geometric
-mechanisms as Markov chains. Let the state $in_k$ and $out_l$
+mechanisms as Markov chains.
+
+\begin{example}
+Let the state $in_k$ and $out_l$
 denote the input value $k$ and output value $l$ respectively. Define
@@
 the Markov chain of the truncated
-$1/2$-geometric mechanism over $\underline{5}$.
+$1/2$-geometric mechanism over $\underline{5}$. Further, assume
+$L(out_i) = \{ out_i \}$ and $\Delta (f) = 1$. Then $in_k$ and $in_l$ are
+neighbors iff $| k - l | \leq 1$.
+\end{example}
 
 \subsection{Markov Chains for Differentially Private Mechanisms}
-Now we describe our main ideas for the modelling differentially private mechanisms. 
+Now we describe our main ideas for  modelling differentially private mechanisms.
 A state $s$ stores at least the a database $x\in \mathcal{X}^n$, which is the data hold by a trusted \emph{curator}. A transition corresponds to the data analysis mechanisms.
@@
 
-Below we describe an online model.
+\paragraph{Online Model.}
+Below we describe an online model adapted from \cite{}: it .....
 
-\begin{figure}
-    \centering
-    \begin{subfigure}{.4\columnwidth}
-\resizebox{\columnwidth}{!}{
-    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
-    distance=1cm,node/.style={circle,draw}]
-    \node[node] (p) at ( -1.5,  0) { $+$ };
-    \node[node] (Y1) at ( 0, .8) { $Y_1$ };
-    \node[node] (N1) at ( 0, -.8) { $N_1$ };
-    \node[node] (Y2) at (1.5,  .8) { $Y_2$ };
-    \node[node] (N2) at ( 1.5, -.8) { $N_2$ };
+We consider the database set $D = \{0,1,2,3,4,5\}$, a threshold
+$t \in \{0,1,2,3,4,5\}$, queries $\{f| f $ with sensitivity $\Delta (f) = 1 \}$.
+In order to protect privacy, a truncated $1/2-$geometric mechanisms over $\underline{5}$ is applied to each query and the threshold, and what we observe is the first time that a query result is greater than or equal to the threshold after interference.
+We are interested in the differential privacy property of the data in $D$, assuming that they are all neighbours.
+The modeling can be done in guarded command language and some fragments are shown below.\todo{the algorithm seems redundant}
 
-    \path
-    (p) edge node [above] { $\frac{3}{4}$ } (Y1)
-    (p) edge node [below] { $\frac{1}{4}$ } (N1)
-    (Y1) edge node [above] { $\frac{3}{4}$ } (Y2)
-    (Y1) edge node [above] { $\frac{1}{4}$ } (N2)
-    (N1) edge node [below] { $\frac{3}{4}$ } (Y2)
-    (N1) edge node [below] { $\frac{1}{4}$ } (N2)
-    ;
-    \end{tikzpicture}
-}
-%        \caption{Markov Chain of }
-%        \label{figure:2-dp-mc1}
-    \end{subfigure}
-    \hspace{.05\columnwidth}
-    \begin{subfigure}{.4\columnwidth}
-        \resizebox{\columnwidth}{!}{
-            \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
-    distance=1cm,node/.style={circle,draw}]
-    \node[node] (n) at ( -1.5,  0) { $-$ };
-    \node[node] (Y1) at ( 0, .8) { $Y_1$ };
-    \node[node] (N1) at ( 0, -.8) { $N_1$ };
-    \node[node] (Y2) at (1.5,  .8) { $Y_2$ };
-    \node[node] (N2) at ( 1.5, -.8) { $N_2$ };
+\begin{algorithm}
+    \begin{algorithmic}[1]
 
-    \path
-    (n) edge node [above] { $\frac{1}{4}$ } (Y1)
-    (n) edge node [below] { $\frac{3}{4}$ } (N1)
-    (Y1) edge node [above] { $\frac{1}{4}$ } (Y2)
-    (Y1) edge node [above] { $\frac{3}{4}$ } (N2)
-    (N1) edge node [below] { $\frac{1}{4}$ } (Y2)
-    (N1) edge node [below] { $\frac{3}{4}$ } (N2)
-    ;
-            \end{tikzpicture}
-        }
-%        \caption{Markov Chain of}
-%        \label{figure:2-dp-mc2}
-    \end{subfigure}
-    \caption{Markov Chain of }
-    \label{figure:2-dp-mc}
-\end{figure}
-\begin{example}
-  First, assume that the curator allows at most two survey queries.
-The corresponding model is depicted in Figure~\ref{figure:2-dp-mc}, where states $Y_1,Y_2$ are labelled by $Y$, and $N_1,N_2$ are labelled by $N$.
-\end{example}
+        \Function{OnlineModel}{$d$, $t$}
 
-\lz{tbd, the complex threshold example we discussed, by Depeng}
+        \State{$\mathbf{If}$  $t=0$ $\rightarrow   2/3 : (\hat{t} = 0) + 1/6 : (\hat{t} = 1) + 1/12 : (\hat{t} = 2)$ }
+        \State{$+ 1/24 : (\hat{t} = 3) + 1/48 : (\hat{t} = 4) + 1/48: (\hat{t} = 5);$}
+        \State{$| $ $ t=1\rightarrow   1/3 : (\hat{t} = 0) + 1/3 : (\hat{t} = 1) + 1/6 : (\hat{t} = 2)$}
+        \State{$+ 1/12 : (\hat{t} = 3) + 1/24 : (\hat{t} = 4) + 1/24: (\hat{t} = 5);$}
+        \State{$...$  $...$}
+        \Comment{Assignments of $\hat{t}$ by the geometric mechanism}
+        \State{$\mathbf{Fi}$}
+
+        \For{each query $f_i$}
+        \State{$\mathbf{If}$  $d=0$ $\rightarrow   2/3 : (\hat{f} = 0) + 1/6 : (\hat{f} = 1) + 1/12 : (\hat{f} = 2)$ }
+        \State{$+ 1/24 : (\hat{f} = 3) + 1/48 : (\hat{f} = 4) + 1/48: (\hat{f} = 5);$}
+        \State{$| $ $ d=1\rightarrow   1/3 : (\hat{f} = 0) + 1/3 : (\hat{f} = 1) + 1/6 : (\hat{f} = 2)$}
+        \State{$+ 1/12 : (\hat{f} = 3) + 1/24 : (\hat{f} = 4) + 1/24: (\hat{f} = 5);$}
+        \State{$...$  $...$}
+        \Comment{Assignments of $\hat{f}$ by the geometric mechanism}
+        \State{$\mathbf{Fi}$}
+        \State{$\mathbf{If}$ $\hat{f}\ge \hat{t}$ $a_i=\top$ \Return }
+        \State{$|$   $\hat{f} < \hat{t}$ $a_i=\bot$}
+        \State{$\mathbf{Fi}$}
+        \EndFor
+        \EndFunction
+
+    \end{algorithmic}
+    \caption{Input is a private database $d \in \{0,1,2,3,4,5\}$, an adaptively chosen stream of sensitivity 1 queries $f_1$, .
+.
+., a threshold $t \in \{0,1,2,3,4,5\}$.Output is a stream of answers $a_1$,...}
+    \label{algorithm:online-model}
+\end{algorithm}

### 2018-01-27 semantics.tex
threshold example

@@
 
-\section{Model Checking Algorithms}
+\subsection{Model Checking Algorithms}
 We first describe the model checking algorithm for $\dpCTL$, which follows the classical

### 2018-01-27 specifications.tex
threshold example

@@
 The degradation is rooted in the repeated computation.
-To illustrate this property, we consider the extended example below \lz{here example of the survey example, expanded twice: we should discuss the graph before it is drawn...}
+To illustrate this property, we consider the extended survey example below where the curator allows two consequent queries.
+The corresponding model is depicted in Figure~\ref{figure:2-dp-mc}, where states $Y_1,Y_2$ are labelled by $Y$, and $N_1,N_2$ are labelled by $N$.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}{.4\columnwidth}
+\resizebox{\columnwidth}{!}{
+    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
+    distance=1cm,node/.style={circle,draw}]
+    \node[node] (p) at ( -1.5,  0) { $+$ };
+    \node[node] (Y1) at ( 0, .8) { $Y_1$ };
+    \node[node] (N1) at ( 0, -.8) { $N_1$ };
+    \node[node] (Y2) at (1.5,  .8) { $Y_2$ };
+    \node[node] (N2) at ( 1.5, -.8) { $N_2$ };
+
+    \path
+    (p) edge node [above] { $\frac{3}{4}$ } (Y1)
+    (p) edge node [below] { $\frac{1}{4}$ } (N1)
+    (Y1) edge node [above] { $\frac{3}{4}$ } (Y2)
+    (Y1) edge node [above] { $\frac{1}{4}$ } (N2)
+    (N1) edge node [below] { $\frac{3}{4}$ } (Y2)
+    (N1) edge node [below] { $\frac{1}{4}$ } (N2)
+    ;
+    \end{tikzpicture}
+}
+%        \caption{Markov Chain of }
+%        \label{figure:2-dp-mc1}
+    \end{subfigure}
+    \hspace{.05\columnwidth}
+    \begin{subfigure}{.4\columnwidth}
+        \resizebox{\columnwidth}{!}{
+            \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
+    distance=1cm,node/.style={circle,draw}]
+    \node[node] (n) at ( -1.5,  0) { $-$ };
+    \node[node] (Y1) at ( 0, .8) { $Y_1$ };
+    \node[node] (N1) at ( 0, -.8) { $N_1$ };
+    \node[node] (Y2) at (1.5,  .8) { $Y_2$ };
+    \node[node] (N2) at ( 1.5, -.8) { $N_2$ };
+
+    \path
+    (n) edge node [above] { $\frac{1}{4}$ } (Y1)
+    (n) edge node [below] { $\frac{3}{4}$ } (N1)
+    (Y1) edge node [above] { $\frac{1}{4}$ } (Y2)
+    (Y1) edge node [above] { $\frac{3}{4}$ } (N2)
+    (N1) edge node [below] { $\frac{1}{4}$ } (Y2)
+    (N1) edge node [below] { $\frac{3}{4}$ } (N2)
+    ;
+            \end{tikzpicture}
+        }
+%        \caption{Markov Chain of}
+%        \label{figure:2-dp-mc2}
+    \end{subfigure}
+    \caption{Markov Chain of Double Surveys}
+    \label{figure:2-dp-mc}
+\end{figure}
 
@@
 We verify that this formula is satisfied  for state $+$ and $-$.
-Moreover, the bound $\epsilon=9$ is bound, since \lz{tbd once graph is given}.
+Moreover, the bound $\epsilon=9$ is bound, since the probability of satisfying $\X (Y\wedge \X Y)$ from states $+$ and $-$ are $9$ and $1$, respectively.
 Finally, the formula $\wedge_{a_1,a_2}\dpriv{9}{0} (\X (a_1\wedge \X a_2))$ corresponds to the desired compositional property for the model, where $a_1,a_2$ range over all atomic propositions $\{Y,N\}$.
@@
 
-Now we consider the formula $\dpriv{\epsilon_2}{\delta_2} (\X (Y\wedge \X Y))$ under the model in ???.
-Applying the compositional theorem\lz{Bow-Yaw, as we only refer to the book, I wonder whether using the notion $e^\epsilon$ is easier to explain?}, one can choose $\epsilon_2=\epsilon_1^2=1$ and $\delta_2=2\delta_1=1$.
-However, we can check easily that one can get better privacy parameter $(1,0.5)$ or $(0,\frac{9}{16})$.
+Now we consider the formula $\dpriv{\epsilon_2}{\delta_2} (\X (Y\wedge \X Y))$ under the model in Figure~\ref{figure:2-dp-mc}.
+Applying the compositional theorem, one can choose $\epsilon_2=\epsilon_1^2=1$ and $\delta_2=2\delta_1=1$.
+However, we can check easily that one gets better privacy parameter $(1,0.5)$ or $(0,\frac{9}{16})$.
 

### 2018-01-27 examples.tex
add background info about the "threshold" algorithm

@@
 \paragraph{Online Model.}
-Below we describe an online model adapted from \cite{}: it .....
+Below we describe an online model adapted from \cite{}: it is used in the scenario where a large amount of queries occur, but we only care for the queries which 
+are above a given threshold.
+Other queries below the threshold are considered "irrelevent" and might be discarded.
+In \cite{}, since the model is set for continuous database, the Laplace mechanism is applied to both queries and the threshold.
+In our model, the database is discrete and instead the geometric mechanism is applied.
 
-We consider the database set $D = \{0,1,2,3,4,5\}$, a threshold
-$t \in \{0,1,2,3,4,5\}$, queries $\{f| f $ with sensitivity $\Delta (f) = 1 \}$.
+We consider the database set $D = \{0,1,2,3,4,5\}$, a threshold $t \in \{0,1,2,3,4,5\}$, queries $\{f| f $ with sensitivity $\Delta (f) = 1 \}$.
 In order to protect privacy, a truncated $1/2-$geometric mechanisms over $\underline{5}$ is applied to each query and the threshold, and what we observe is the first time that a query result is greater than or equal to the threshold after interference.

### 2018-01-28 applications.tex
add frequency estimator

@@
 variant of the mechanisms by Markov decision processes. We will also
-specify pan privacy on our model using $\dpCTL$.
+specify pan privacy in our model using $\dpCTL$.
 
-Consider monitoring events of the types $a$ and $b$ continuously. An
-observation is a sequence of event types. Two obsevations are
-{neighbors} if one can be obtained from the other by replacing 
-event types any number of times, that is, if they are of the same
-length. \bw{Later, we will consider a different definition of
+Consider monitoring events of the types $a$ and $b$ continuously.
+A sequence of event types is said to be an observation. Two
+obsevations are {neighbors} if one can be obtained from the other 
+by replacing event types any number of times, that is, if they are of
+the same length. \bw{Later, we will consider a different definition of
 neighborhood by restricting the number of replacements.} 
-We would like to compute the number of event types in
-observations. Hence for observations $a \cdots a$ and $b \cdots b$ of
-arbitrary positive lengths, the query result is $1$; and the query
-result is $2$ for other non-empty observations.
+We would like to inquire the number of event types in
+observations. For instance, the query result result is $1$ 
+for observations $a \cdots a$ and $b \cdots b$ of arbitrary positive
+lengths; and it is $2$ for other non-empty observations.
+
+Based on the density estimation algorithm in \cite{DNPRY:10:PPSA}, we
+consider the following mechanism for event type streams. Our mechanism
+keeps a flag for each event type.
+
+\begin{algorithm}
+  \begin{algorithmic}[1]
+  \Procedure{FrequencyEstimator}{}
+    \State{$f_a \leftarrow 0, 1$ with probability 
+      $\frac{1}{2}, \frac{1}{2}$ respectively}
+    \Comment{Initialization}
+    \State{$f_b \leftarrow 0, 1$ with probability
+      $\frac{1}{2}, \frac{1}{2}$ respectively}
+    \Match{current event type $e$}
+    \Comment{Processing}
+        \Case{$a$}
+          \State{$f_a \leftarrow 0, 1$ with probability
+            $\frac{3}{8}, \frac{5}{8}$ respectively}
+        \EndCase
+        \Case{$b$}
+          \State{$f_b \leftarrow 0, 1$ with probability
+            $\frac{3}{8}, \frac{5}{8}$ respectively}
+        \EndCase
+    \EndMatch
+    \Match{$f_a + f_b$}
+    \Comment{Output}
+      \Case{0}
+        \State{$\mathit{out} \leftarrow 0, 1, 2$ with probability
+          $\frac{2}{3}, \frac{1}{6}, \frac{1}{6}$ respectively}
+      \EndCase
+      \Case{1}
+        \State{$\mathit{out} \leftarrow 0, 1, 2$ with probability
+          $\frac{1}{3}, \frac{1}{3}, \frac{1}{3}$ respectively}
+      \EndCase
+      \Case{2}
+        \State{$\mathit{out} \leftarrow 0, 1, 2$ with probability
+          $\frac{1}{6}, \frac{1}{6}, \frac{2}{3}$ respectively}
+      \EndCase
+    \EndMatch
+    \State{\Return $\mathit{out}$}
+  \EndProcedure
+  \end{algorithmic}
+  \caption{FrequencyEstimator}
+  \label{algorithm:frequency-estimator}
+\end{algorithm}
 

### 2018-01-28 applications.tex
add more descriptions

@@
 consider the following mechanism for event type streams. Our mechanism
-keeps a flag for each event type.
+keeps a flag for each event type. The flags are initialized to $0$ or
+$1$ uniformly. On observing an event type in the input stream, the
+flag for the event type is set to $0$ and $1$ with probability
+$\frac{3}{8}$ and $\frac{5}{8}$ respectively. The mechanism continues
+to update its flags until it is asked to return the number of observed
+event types.
 

### 2018-01-28 examples.tex
examples

@@
 
-In the differentially privacy context, we differentiate between offline and online models.
-In the offline, or non-interactive, model the curator releases the outputs only once and plays no further role.
-Thus the models we constructed above are offline models.
-On the contrary, in the online, or interactive, model the curator can permit further queries.
+In the differentially privacy context, we differentiate between \emph{offline} and \emph{online} models.
+In the offline (or \emph{non-interactive}) model the curator releases the outputs only once and plays no further role.
+Thus the models we constructed in previous section are offline models.
+On the contrary, in the online (or \emph{interactive}) model the curator can permit further queries.
 In this setting it is obvious that queries do not change database themselves.
@@
 
+\lz{use $\frac{2}{3}$, not $2/3$, and use $t'$, not $\hat{t}$}
 \begin{algorithm}

### 2018-01-28 semantics.tex
examples

@@
 
-The model checking algorithm for $\dpCTLstar$ can be treated as in the classical setting \cite{BK:08:PMC}, as all we need is to  compute the the probability
-  $\myPr{s}{K}{\neighbor{S}}{\phi}$ with general path formula $\phi$, one first construct a deterministic $\omega$-automaton $R$, such as deterministic Rabin automaton.
+The model checking algorithm for $\dpCTLstar$ can be treated as in the classical setting \cite{BK:08:PMC}, as all we need is to  compute the  probability
+  $\myPr{s}{K}{\neighbor{S}}{\phi}$ with general path formula $\phi$.
+For this purpose one first constructs a deterministic $\omega$-automaton $R$, such as deterministic Rabin automaton.
 Then, the probability reduces to a
-  reachability probability in the product Markov chain obtained from $K$ and $R$. There are more efficient algorithms without the product construction, see \cite{}. We skip the details.
+  reachability probability in the product Markov chain obtained from $K$ and $R$. There are more efficient algorithms without the product construction, see \cite{} for details.

### 2018-01-28 specifications.tex
examples

@@
-In this section we describe what type of properties in the differential privacy context can be expressed using $\dpCTL$.
+In this section we describe how properties in the differential privacy literature can be expressed using $\dpCTLstar$ formulae.
 
@@
 $\psi$ for $k = 0, \ldots, 5$, then the $.5$-geometric mechanism is
-$(0.5, 0)$-differentially private.
+$(.5, 0)$-differentially private.
 
 \paragraph{Compositional Properties.}
-Compositional aspect is one of the building block for differentially private algorithms. According to the compositional theorem \cite{}, for $(\epsilon_1,0)$-differentially private algorithm $M_1$ and
-$(\epsilon_2,0)$-differentially private algorithm $M_2$, their combination $M$ defined by $M(x)=(M_1(x), M_2(x))$ is $(\epsilon_1\epsilon_2,0)$-differentially private.
+Compositional aspect is one of the building block for differentially private algorithms.
+According to the compositional theorem \cite[Theorem 3.16]{DR:14:AFDP}, for $(\epsilon_1,\delta_1)$-differentially private algorithm $M_1$ and
+$(\epsilon_2,\delta_2)$-differentially private algorithm $M_2$, their combination $M$ defined by $M(x)=(M_1(x), M_2(x))$ is $(\epsilon_1\epsilon_2,\delta_1+\delta_2)$-differentially private.
 The degradation is rooted in the repeated computation.
@@
 We verify that this formula is satisfied  for state $+$ and $-$.
-Moreover, the bound $\epsilon=9$ is bound, since the probability of satisfying $\X (Y\wedge \X Y)$ from states $+$ and $-$ are $9$ and $1$, respectively.
+Moreover, the bound $\epsilon=9$ is tight, since the probability of satisfying $\X (Y\wedge \X Y)$ from states $+$ and $-$ are $\frac{9}{16}$ and $\frac{1}{16}$, respectively.
 Finally, the formula $\wedge_{a_1,a_2}\dpriv{9}{0} (\X (a_1\wedge \X a_2))$ corresponds to the desired compositional property for the model, where $a_1,a_2$ range over all atomic propositions $\{Y,N\}$.
 
-We consider another simpler formula for comparision: $\dpriv{3}{0} (\X \X Y)$.
+We consider two other formulae for comparision: 
+(a) Formula $\dpriv{3}{0} (\X \X Y)$.
 In this case we donot have privacy lost, even though we are querying twice: the reason is that the output of the first query is not utilized at all.
 It is easy to verify that it is indeed satisfied by $+$ and $-$.
+(b) Formula $\dpriv{3}{0} (\X (Y\wedge \dpriv{3}{0}(\X Y)))$.
+This is a nested $\dpCTL$ formula, where the inner state formula $\dpriv{3}{0}(\X Y)$ states the one step differential privacy.
+Observe the inner formula is satisfied by all neighbours.
+Thus, the outer formula also has no privacy lost.
 
@@
 
-\paragraph{Reachability Privacy}
-For our threshold \lz{to do} model, we have the desired privacy property $\wedge_{i\in N} \dpriv{\epsilon}{0}(\X^k \top) \wedge \dpriv{\epsilon}{0}(F \top)$.
-Here $\dpriv{\epsilon}{0}(F \top)$ is a reachability privacy: it states that related states should have similar probabilities of reaching $\top$.
+\paragraph{Reachability Privacy.}
+For our threshold \lz{to do} model, we have the differential privacy property $\wedge_{i=1}^{c} \dpriv{\epsilon}{0}(\X^k \top) \wedge \dpriv{\epsilon}{0}(F \top)$.
+Here $\dpriv{\epsilon}{0}(F \top)$ is a reachability privacy: it states that related states should have similar probabilities of eventually reaching $\top$.
 
-\paragraph{Infinite ...}
-Note that $\dpCTL$ generalizes differential privacy to infinite
+\paragraph{Infinite Differential Privacy.}
+Due to the power of linear temporal logic, our logic $\dpCTLstar$ generalizes differential privacy to infinite
 behaviors naturally. Classical differential privacy discusses properties about
@@
 prefixes of computation.
-With \dpCTL, the subformula $\phi$ in $\dpriv{\epsilon}{\delta} \phi$
-can be an arbitrary $\dpCTL$ path formula. One can specify
-$(\epsilon, \delta)$-closedness over infinite behaviors naturally. For instance,
-$\dpriv{\epsilon}{\delta} (\G \PJ{[1,1]} (\F \mathit{high}))$ \lz{now we can use $G F high$}
+With $\dpCTLstar$, the subformula $\phi$ in $\dpriv{\epsilon}{\delta} \phi$
+can be an arbitrary $\dpCTLstar$ path formula. One can specify
+$(\epsilon, \delta)$-closedness over infinite behaviors naturally.
+
+We consider the Bluetooth example from the PRISM benchmarks \cite{prism-bluetooth-page}. The model is a 
+Markov chain, and it ...
+
+For instance,
+$\dpriv{\epsilon}{\delta} (\G   \F \mathit{discovered})$ 
 specifies states having $(\epsilon, \delta)$-close
-infinitely often $\mathit{high}$ behaviors to its neighbors.
+infinitely often $\mathit{discovered}$ behaviors to its neighbors.

### 2018-01-28 applications.tex
add privacy analysis and specification

--- 2018-01-28 applications.tex
@@
-Consider the MDP in Figure~\ref{figure:simple-mdp-mdp}. Suppose the
-states $+$ and $-$ are neighbors. Let us compute $\{ s : M,
-\neighbor{S}, s \models \dpriv{3}{0} (\F Y) \}$. First, the self
-product $M^2$ is depicted in Figure~\ref{figure:simple-mdp-product}
-(the states $++$ and $--$ and associated transitions are ignored
-for clarity). Applying
-Theorem~\ref{theorem:eventuallyC} with $\epsilon = 4$ and $\delta = 0$, 
-consider the following linear programming problem:
-\[
-\begin{array}{rcrcrcrcrcl}
-  \multicolumn{11}{c}{
-  \min x_{++} + x_{+-} + x_{-+} + x_{--} + x_{YY} + x_{YN} + x_{NY} + x_{NN}
-  }\\
-  \multicolumn{9}{r}{ x_{YN} } & = & 1 \\
-  \multicolumn{9}{r}{ x_{YY} } & = & -3 \\
-  \multicolumn{9}{r}{ x_{NY} } & = & -4 \\
-  \multicolumn{9}{r}{ x_{NN} } & = & 0 \\
-  x_{++} &-&   .5625 x_{YY} &-& .1875 x_{YN} &-& .1875 x_{NY} &-&
-               .0625 x_{NN} & \geq & 0\\
-  x_{++} &-&   .64   x_{YY} &-& .16   x_{YN} &-& .16   x_{NY} &-&
-               .04   x_{NN} & \geq & 0\\
-  x_{+-} &-&   .1875 x_{YY} &-& .5625 x_{YN} &-& .0625 x_{NY} &-& 
-               .1875 x_{NN} & \geq & 0\\
-  x_{+-} &-&   .16   x_{YY} &-& .64   x_{YN} &-& .04   x_{NY} &-&
-               .16   x_{NN} & \geq & 0\\
-  x_{-+} &-&   .1875 x_{YY} &-& .0625 x_{YN} &-& .5625 x_{NY} &-& 
-               .1875 x_{NN} & \geq & 0\\
-  x_{-+} &-&   .16   x_{YY} &-& .04   x_{YN} &-& .64   x_{NY} &-&
-               .16   x_{NN} & \geq & 0\\
-  x_{--} &-&   .0625 x_{YY} &-& .1875 x_{YN} &-& .1875 x_{NY} &-&
-               .5625 x_{NN} & \geq & 0\\
-  x_{--} &-&   .04   x_{YY} &-& .16   x_{YN} &-& .16   x_{NY} &-&
-               .64   x_{NN} & \geq & 0
-\end{array}
-\]
-The optimum is attained at 
-$(x_{++}, x_{+-}, x_{-+}, x_{--}) = (-2.25, 0, -2.75, -0.6)$. By
-Theorem~\ref{theorem:eventuallyC}, we have
-\[
-\begin{array}{l}
-\max_{\scheduler{S}}
-\myPr{+}{M_{\scheduler{S}}}{\neighbor{S}}{\F Y} -
-4 \myPr{+}{M_{\scheduler{S}}}{\neighbor{S}}{\F Y}
-= -2.25 \leq 0\\
-\max_{\scheduler{S}}
-\myPr{+}{M_{\scheduler{S}}}{\neighbor{S}}{\F Y} -
-4 \myPr{-}{M_{\scheduler{S}}}{\neighbor{S}}{\F Y}
-= 0 \leq 0 \\
-\max_{\scheduler{S}}
-\myPr{-}{M_{\scheduler{S}}}{\neighbor{S}}{\F Y} -
-4 \myPr{+}{M_{\scheduler{S}}}{\neighbor{S}}{\F Y}
-= -2.75 \leq 0\\
-\max_{\scheduler{S}}
-\myPr{-}{M_{\scheduler{S}}}{\neighbor{S}}{\F Y} -
-4 \myPr{-}{M_{\scheduler{S}}}{\neighbor{S}}{\F Y}
-= -0.6 \leq 0
-\end{array}
-\]
-Therefore $+, - \in \{ s : M, \neighbor{M}, s \models \dpriv{3}{0}
-(\F Y) \}$. This is expected. Recall that low- and high-accuracy
-mechanisms are $(3, 0)$- and $(4, 0)$-differentially private
-respectively. Regardless of request types, the interactive mechanism
-in Figure~\ref{figure:simple-mdp-mdp} can only achieve $(4,
-0)$-differential privacy at best.
-
-Streaming processing gives another motivating example for Markov
+Streaming processing gives a motivating example for Markov
 decision processes. In stream processing, input data are retrieved and
@@
 consider the following mechanism for event type streams. Our mechanism
-keeps a flag for each event type.
+keeps a bit for each event type. The bits are initialized to $0$ or
+$1$ uniformly. On observing an event type in the input stream, the
+bit for the event type is set to $0$ and $1$ with probability
+$\frac{3}{8}$ and $\frac{5}{8}$ respectively. The mechanism continues
+to update its bits until it is asked to return the number of observed
+event types. To compute the number of observed event types, the
+mechanism adopts the truncated $1/2$-geometric mechanism on the sum of
+two bits for event types. If the sum is $0$, the mechanism outputs $0$
+with probability $\frac{2}{3}$ and $1, 2$ each with probability
+$\frac{1}{6}$. If the sum is $1$, one of $0, 1, 2$ is returned with
+the uniform distribution. If the sum is $2$, the mechanism outputs $2$
+with probability $\frac{2}{3}$ and $0, 1$ each with probability
+$\frac{1}{6}$ (Algorithm~\ref{algorithm:frequency-estimator}).
 
@@
 
-For each event, a bit denotes whether it has occurred or not. To model
-adjacent inputs, we use four actions symbols to denote events and its
-replacement. For instance, the action $a[a]$ means the event $a$ is
-observed without replacement; $b[a]$ means the event $b$ is observed
-but replaced with $a$ in an adjacent input.
+\indent
+\emph{Privacy analysis}.
+In the processing phase, the bit $f_a$ for an event type $a$ can only
+have two different distributions $U$ and $B$. If the event type is
+never observed, $\Pr_U[f_a = 0] = \Pr_U[f_a = 1] = \frac{1}{2}$; if
+the event type has been observed, we have $\Pr_B[f_a = 0] =
+\frac{3}{8}$ and $\Pr_B[f_a = 1] = \frac{5}{8}$. Note that
+$\frac{1}{2} \cdot \frac{3}{8} \leq \frac{1}{2} \leq 2 \cdot
+\frac{3}{8}$ and $\frac{1}{2} \cdot \frac{5}{8} \leq \frac{1}{2} \leq
+2 \cdot \frac{5}{8}$. For every $x \in \{ 0, 1 \}$, we have
+$\frac{1}{2} \cdot \Pr_B[f_a = x] \leq \Pr_U[f_a = x] \leq 2 \cdot
+\Pr_B[f_a = x]$. On neighboring observations, the distribution of
+$f_a$ can only be changed from $U$ to $B$, values of $f_a$ is $(2,
+0)$-differentially private and similarly for $f_b$. Finally, the
+$1/2$-geometric mechanism gives $(1/2, 0)$-differential privacy. By
+composition theorem, the steaming data analysis mechanism in
+Algorithm~\ref{algorithm:frequency-estimator} is $(1,
+0)$-differentially private.
+
+To model Algorithm~\ref{algorithm:frequency-estimator}, let
+us start with the formalization of observations. Note that the
+lengths of observations are finite but unbounded. An infinite Markov
+chain would be necessary if observations were formalized as states. 
+Recall that neighbors are observations of the same length. If we
+furthermore consider the formalization of neighbors, it is not clear
+how to construct a finite Markov chain for
+Algorithm~\ref{algorithm:frequency-estimator}.
+
+In order to obtain a finite model, we formalize neighbors as action
+sequences in Markov decision processes. Consider the action set $\{
+a[a], a[b], b[a], b[b] \}$. Informally, the action $x[y]$ means the
+event type $x$ is replaced by the event type $y$ in a neighbor. An
+action sequence thus represents an observation and one of its
+neighbors and vice versa. More formally, the action sequence
+$x_1[y_1]x_2[y_2]\cdots x_n[y_n]$ represents the observation
+$x_1x_2\cdots x_n$ and its neighbor $y_1y_2\cdots y_n$. For instance,
+the action sequence $a[a]a[b]a[a]b[a]b[b]$ represents the observation
+$aaabb$ and a neighbor $abaab$.
+
+Similar to Algorithm~\ref{algorithm:frequency-estimator}, our Markov
+decision process has internal states $00$, $01$, $10$, $11$ for the
+two bits $f_a$ and $f_b$ of observed event types
+(Figure~\ref{figure:adjacent-inputs}). The internal state $01$
+represents $(f_a, f_b) = (0, 1)$, for instance. From the
+initialization state $I$, the Markov decision process moves to one of
+the four internal states randomly with the uniform
+distribution. Consider the internal state $01$. If the actions $a[a]$
+or $a[b]$ are chosen, they denote that the event type $a$ is currently
+observed. The Markov decision process moves to the internal state $01$
+and $11$ with probability $.375 = \frac{3}{8}$ and $.625 =
+\frac{5}{8}$ respectively. Other transitions are defined similarly.
+
+On the other hand, the initialization state $\underline{I}$
+corresponds to the computation on neighbors. The Markov decision
+process moves to the interal states $\underline{00}$,
+$\underline{01}$, $\underline{10}$, $\underline{11}$ randomly with
+the uniform distribution. Consider, for example, the internal state
+$\underline{10}$. If the action $a[a]$ or $b[a]$ are chosen, they
+denote that the current event type is replaced with the eveny type
+$a$. The Markov decision process moves to the internal state
+$\underline{00}$ and $\underline{10}$ with probability $.375 =
+\frac{3}{8}$ and $.625 = \frac{5}{8}$ respectively. At an internal
+state, the Markov decision process can move to the output states $0$,
+$1$, or $2$ by the truncated $\frac{1}{2}$-geometric mechanism over
+$\underline{2}$ (Figure~\ref{figure:randomized-frequency}, \`{a} la
+Figure~\ref{figure:geometric-mechanism}). For any action sequence,
+observe that paths from the state $I$ and those from $\underline{I}$
+formalize the computation of Algorithm~\ref{algorithm:frequency-estimator} 
+on an observation and its neighbor represented by the sequence. We
+hence say the states $I$ and $\underline{I}$ are neighbors in the
+Markov decision process (Figure~\ref{figure:frequency-estimator}).
 
@@
       (01) edge [loop below] node { $a[\uscore]$, .375 } (01)
-      (01) edge [bend left=10] node { $a[\uscore]$, .675 } (11)
+      (01) edge [bend left=10] node { $a[\uscore]$, .625 } (11)
       (01) edge [loop left] node [above=1,rotate=90] { $b[\uscore]$, .625 } (01)
@@
       (01') edge [loop below] node { $\uscore[a]$, .375 } (01')
-      (01') edge [bend left=10] node { $\uscore[a]$, .675 } (11')
+      (01') edge [bend left=10] node { $\uscore[a]$, .625 } (11')
       (01') edge [loop left] node [above=1,rotate=90] { $\uscore[b]$, .625 } (01')
@@
 
-\todo{verify this is $1$-DP}
+\noindent
+\emph{Privacy Specification}.
+For classical differential privacy, we are interested in the output
+distribution. Algorithm~\ref{algorithm:frequency-estimator} has three
+different outputs. By privacy analysis, the state $I$ satisfies the
+following $\dpCTLstar$ formula:
+\[
+\dpriv{1}{0} \F 0 \wedge \dpriv{1}{0} \F 1 \wedge \dpriv{1}{0} \F 2
+\]
+For pan privacy, we moreover wish to measure privacy loss in internal
+states. For observations of length $k$, the probabilities of reaching
+each internal states need to be sufficiently close. The state $I$
+therefore must satisfy
+\[
+\dpriv{1}{0} \F 0 \wedge \dpriv{1}{0} \F 1 \wedge \dpriv{1}{0} \F 2
+\wedge
+\bigwedge_{k \in \bbfN}
+\dpriv{2}{0} \X^k 00 \wedge 
+\dpriv{2}{0} \X^k 01 \wedge 
+\dpriv{2}{0} \X^k 10 \wedge 
+\dpriv{2}{0} \X^k 11.
+\]

### 2018-01-28 mdp.tex
examples
remove product mdp from figure

@@
 
-\begin{figure}
+\begin{wrapfigure}{r}{.5\columnwidth}
   \centering
-  \begin{subfigure}{.40\columnwidth}
-    \resizebox{\columnwidth}{!}{
+    \resizebox{.4\columnwidth}{!}{
     \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
@@
     \caption{Markov Decision Process}
-    \label{figure:simple-mdp-mdp}
-  \end{subfigure}
-  \begin{subfigure}{.58\columnwidth}
-    \resizebox{\columnwidth}{!}{
-    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
-      distance=2cm,node/.style={circle,draw,inner sep=1pt}]
-      \node[node] (pq) at (-2.1, 2.1) { $+-$ };
-      \node[node] (qp) at ( 2.1,-2.1) { $-+$ };
-      \node[node] (RR) at (-2.1,-2.1) { $YY$ };
-      \node[node] (BB) at ( 2.1, 2.1) { $NN$ };
-      \node[node] (RB) at (- .7,- .7) { $YN$ };
-      \node[node] (BR) at (  .7,  .7) { $NY$ };
-
-      \path
-      (pq) edge [bend right=5] node [left=1,below=12,rotate=270] { $L, .1875$ } (RR)
-      (pq) edge [bend left=5]  node [above=5,right=1] { $L, .1875$ } (BB)
-
-      (pq) edge [bend left=5]  node [right=4,rotate=270] { $H, .16$ } (RR)
-      (pq) edge [bend right=5] node [below=5,right=1] { $H, .16$ } (BB)
-
-      (pq) edge [bend left=5]  node [left,below=10,rotate=295]  { $L, .5625$ } (RB)
-      (pq) edge [bend right=5] node [left,below,rotate=330]  { $L, .0625$ } (BR)
-
-      (pq) edge [bend right=5] node [right=2,above,rotate=300] { $H, .64$ } (RB)
-      (pq) edge [bend left=5]  node [right,above,rotate=340] { $H, .04$ } (BR)
-
-      (qp) edge [bend right=5] node [above=5,left=1] { $L, .1875$ } (RR)
-      (qp) edge [bend left=5]  node [above=30,left=5,rotate=90] { $L, .1875$ } (BB)
-
-      (qp) edge [bend left=5]  node [below=5,left=1] { $H, .16$ } (RR)
-      (qp) edge [bend right=5] node [above,right=5,rotate=90] { $H, .16$ } (BB)
-
-      (qp) edge [bend left=5]  node [left,below,rotate=330]  { $L, .0625$ } (RB)
-      (qp) edge [bend right=5] node [left=3,below=3,rotate=295]  { $L, .5625$ } (BR)
-
-      (qp) edge [bend right=5] node [left=5,above,rotate=330] { $H, .04$ } (RB)
-      (qp) edge [bend left=5]  node [right=1,above=1,rotate=295] { $H, .64$ } (BR)
-
-      (RR) edge [loop left] node [above] { $-, 1$ } (RR)
-      (BB) edge [loop right] node [above] { $-, 1$ } (BB)
-      (RB) edge [loop below] node [left] { $-, 1$ } (RB)
-      (BR) edge [loop above] node [right] { $-, 1$ } (BR)
-      ;
-      \end{tikzpicture}
-\hide{
-    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
-      distance=2cm,node/.style={circle,draw,inner sep=1pt}]
-      \node[node] (pq) at (   0,   0) { $+-$ };
-      \node[node] (RR) at (-  2,   0) { $YY$ };
-      \node[node] (BB) at (   2,   0) { $NN$ };
-      \node[node] (RB) at (   0, 1.5) { $YN$ };
-      \node[node] (BR) at (   0,-1.5) { $NY$ };
-
-      \path
-      (pq) edge [bend right=15] node [above] { $L, .1875$ } (RR)
-      (pq) edge [bend left=15]  node [above] { $L, .1875$ } (BB)
-
-      (pq) edge [bend left=15]  node [below] { $H, .16$ } (RR)
-      (pq) edge [bend right=15] node [below] { $H, .16$ } (BB)
-
-      (pq) edge [bend left=15]  node [left]  { $L, .5625$ } (RB)
-      (pq) edge [bend right=15] node [left]  { $L, .0625$ } (BR)
-
-      (pq) edge [bend right=15] node [right] { $H, .64$ } (RB)
-      (pq) edge [bend left=15]  node [right] { $H, .04$ } (BR)
-
-      (RR) edge [loop below] node [below] { $-, 1$ } (RR)
-      (BB) edge [loop above] node [above] { $-, 1$ } (BB)
-      (RB) edge [loop left] node [left] { $-, 1$ } (RB)
-      (BR) edge [loop right] node [right] { $-, 1$ } (BR)
-      ;
-      \end{tikzpicture}
-}
-    }
-    \caption{Product Markov Decision Process}
-    \label{figure:simple-mdp-product}
-  \end{subfigure}
-  \caption{Markov Decision Process and its Self-Product}
     \label{figure:simple-mdp}
-\end{figure}
+\end{wrapfigure}
 
@@
 distributions associated with actions.
+
+\lz{assume that we need reactive MDPs with absorbing states}
 

### 2018-01-28 applications.tex
add \lCase in algorithmic
fix privacy guarantee for geometric mechanism in application

@@
 by replacing event types any number of times, that is, if they are of
-the same length. \bw{Later, we will consider a different definition of
-neighborhood by restricting the number of replacements.} 
+the same length. %\bw{Later, we will consider a different definition of
+%neighborhood by restricting the number of replacements.} 
 We would like to inquire the number of event types in
@@
     \Comment{Processing}
-        \Case{$a$}
-          \State{$f_a \leftarrow 0, 1$ with probability
+        \lCase{$a$}{$f_a \leftarrow 0, 1$ with probability
             $\frac{3}{8}, \frac{5}{8}$ respectively}
-        \EndCase
-        \Case{$b$}
-          \State{$f_b \leftarrow 0, 1$ with probability
+        \lCase{$b$}{$f_b \leftarrow 0, 1$ with probability
             $\frac{3}{8}, \frac{5}{8}$ respectively}
-        \EndCase
     \EndMatch
@@
     \Comment{Output}
-      \Case{0}
-        \State{$\mathit{out} \leftarrow 0, 1, 2$ with probability
+      \lCase{0}{$\mathit{out} \leftarrow 0, 1, 2$ with probability
           $\frac{2}{3}, \frac{1}{6}, \frac{1}{6}$ respectively}
-      \EndCase
-      \Case{1}
-        \State{$\mathit{out} \leftarrow 0, 1, 2$ with probability
+      \lCase{1}{$\mathit{out} \leftarrow 0, 1, 2$ with probability
           $\frac{1}{3}, \frac{1}{3}, \frac{1}{3}$ respectively}
-      \EndCase
-      \Case{2}
-        \State{$\mathit{out} \leftarrow 0, 1, 2$ with probability
+      \lCase{2}{$\mathit{out} \leftarrow 0, 1, 2$ with probability
           $\frac{1}{6}, \frac{1}{6}, \frac{2}{3}$ respectively}
-      \EndCase
     \EndMatch
@@
 0)$-differentially private and similarly for $f_b$. Finally, the
-$1/2$-geometric mechanism gives $(1/2, 0)$-differential privacy. By
+$1/2$-geometric mechanism gives $(2, 0)$-differential privacy. By
 composition theorem, the steaming data analysis mechanism in
-Algorithm~\ref{algorithm:frequency-estimator} is $(1,
+Algorithm~\ref{algorithm:frequency-estimator} is $(4,
 0)$-differentially private.
@@
 \[
-\dpriv{1}{0} \F 0 \wedge \dpriv{1}{0} \F 1 \wedge \dpriv{1}{0} \F 2
+\dpriv{4}{0} \F 0 \wedge \dpriv{4}{0} \F 1 \wedge \dpriv{4}{0} \F 2
 \]
@@
 \[
-\dpriv{1}{0} \F 0 \wedge \dpriv{1}{0} \F 1 \wedge \dpriv{1}{0} \F 2
+\dpriv{4}{0} \F 0 \wedge \dpriv{4}{0} \F 1 \wedge \dpriv{4}{0} \F 2
 \wedge
 \bigwedge_{k \in \bbfN}
-\dpriv{2}{0} \X^k 00 \wedge 
-\dpriv{2}{0} \X^k 01 \wedge 
-\dpriv{2}{0} \X^k 10 \wedge 
-\dpriv{2}{0} \X^k 11.
+\dpriv{4}{0} \X^k 00 \wedge 
+\dpriv{4}{0} \X^k 01 \wedge 
+\dpriv{4}{0} \X^k 10 \wedge 
+\dpriv{4}{0} \X^k 11.
 \]

### 2018-01-28 mdp.tex
mdps

--- 2018-01-27 mdp.tex
@@
   \end{subfigure}
-  \caption{Markov Decision Process and its Self-Product}
+  \caption{Markov Decision Process and its Self-Product\lz{in the current formulation, self loops in this example are not needed}}
   \label{figure:simple-mdp}
@@
 
-Mechanisms are not necessarily closed randomized algorithms; they may
+\hide{Mechanisms are not necessarily closed randomized algorithms; they may
 perform different compution on users' requests.
+We use Markov decision processes to model such
+interactive mechanisms. Specifically, external inputs are modeled by
+actions. Behaviors associated with different inputs are modeled by
+distributions associated with actions.
+}
+
+As we describe earlier, the user has the possibility of asking different mechanisms.
 We use Markov decision processes to model such
@@
   \textmd{for all } t \textmd{ with } s \neighbor{S} t \textmd{ and
-   position-independent scheduler } \\
+   scheduler } \\
   & & \scheduler{S},
@@
 
-\lz{how about using "query sequence" instead of position-independent scheduler, and then argue that it makes sense for the resulting MDP model from differential privacy?
-This corresponds to a special scheduler that is applied on both neighboring states.
-}
 Recall that $M_{\scheduler{S}}$ is but a Markov chain. The semantics
@@
 defined as in Markov chains.
-The semantics of $\dpCTL$ on reactive Markov decision processes
+The semantics of $\dpCTLstar$ on Markov decision processes
 is again standard except the differentially private operator
@@
 \delta)$-close to all its neighbors in having paths satisfying $\phi$
-for arbitrary position-independent schedulers. That is, no
+for arbitrary schedulers. That is, no
 position-independent scheduler can force any of neighbors to have
@@
 \noindent
+\emph{Justification of the scheduler class.}
+\lz{related to language equivalence...}
+
+\noindent
 \emph{Discussion.}
+\lz{this part not needed?}
 Another definition of the differentially private operator

### 2018-01-28 preliminaries.tex
mdps

@@
 
-Let $\Act$ denote the set of \emph{actions}.
+Let $\Act$ denote a finite set of \emph{actions}.
 A \emph{(finite) Markov decision process} $M = (S, \Act, \wp, L)$ consists
@@
 \times S \rightarrow [0, 1]$ with $\sum_{t \in S} \wp(s, \alpha, t)
-\in \{ 0, 1 \}$ for every $s \in S$ and $\alpha \in \Act$ and $S$,
-$L$ as for Markov chains. Define
-$\Act(s) = \{ \alpha : \sum_{t \in S} \wp (s, \alpha, t) = 1
-\}$. That is, $\Act(s)$ is the set of \emph{enabled} actions in state
-$s$. A Markov devision process is \emph{reactive} if $\Act(s) = \Act$
+= 1$ for every $s \in S$ and $\alpha \in \Act$ and $S$,
+$L$ as for Markov chains.
+Note the model we consider is \emph{reactive}\footnote{In the literature, such model is referred to as \emph{probabilistic automata}\cite{}.
+In the model checking context, the notion MDPs is more popular.} as $\Act(s) = \Act$
 for every $s \in S$.
@@
 Let $M = (S, \Act, \wp, L)$ be a Markov decision process. A
-\emph{scheduler} for $M$ is a function $\scheduler{S} : S^+
-\rightarrow \Act$ with $\scheduler{S}(s_0s_1\cdots s_n) \in
-\Act(s_n)$. A scheduler is \emph{position-independent} if
-$\scheduler{S}(\sigma) = \scheduler{S}(\sigma')$ when $|\sigma| =
-|\sigma'|$. 
+\emph{scheduler} for $M$ is simply an infinite sequence  $\scheduler{S}\in \Act^\omega$.
 A path $\pi =

### 2018-01-28 mdp.tex

--- 2018-01-28 mdp.tex
@@
 
-Mechanisms are not necessarily closed randomized algorithms; they may
+\hide{Mechanisms are not necessarily closed randomized algorithms; they may
 perform different compution on users' requests.
@@
 distributions associated with actions.
+}
 
-\lz{assume that we need reactive MDPs with absorbing states}
+As we describe earlier, the user has the possibility of asking different mechanisms.
+We use Markov decision processes to model such
+interactive mechanisms. Specifically, external inputs are modeled by
+actions. Behaviors associated with different inputs are modeled by
+distributions associated with actions.
 

### 2018-01-28 applications.tex
use standard differential privacy definition e^{\epsilon}

@@
 0)$-differentially private and similarly for $f_b$. Finally, the
-$1/2$-geometric mechanism gives $(2, 0)$-differential privacy. By
-composition theorem, the steaming data analysis mechanism in
-Algorithm~\ref{algorithm:frequency-estimator} is $(4,
-0)$-differentially private.
+$1/2$-geometric mechanism gives $(\ln 2, 0)$-differential privacy. 
+Recall $2\ln 2 = \ln 4$,
+the steaming data analysis mechanism in
+Algorithm~\ref{algorithm:frequency-estimator} is $(\ln 4,
+0)$-differentially private by
+composition theorem.
 
@@
 \[
-\dpriv{4}{0} \F 0 \wedge \dpriv{4}{0} \F 1 \wedge \dpriv{4}{0} \F 2
+\dpriv{\ln 4}{0} \F 0 \wedge \dpriv{\ln 4}{0} \F 1 \wedge 
+\dpriv{\ln 4}{0} \F 2
 \]
@@
 \[
-\dpriv{4}{0} \F 0 \wedge \dpriv{4}{0} \F 1 \wedge \dpriv{4}{0} \F 2
+\dpriv{\ln 4}{0} \F 0 \wedge \dpriv{2\ln 4}{0} \F 1 \wedge 
+\dpriv{\ln 4}{0} \F 2
 \wedge
 \bigwedge_{k \in \bbfN}
-\dpriv{4}{0} \X^k 00 \wedge 
-\dpriv{4}{0} \X^k 01 \wedge 
-\dpriv{4}{0} \X^k 10 \wedge 
-\dpriv{4}{0} \X^k 11.
+\dpriv{\ln 4}{0} \X^k 00 \wedge 
+\dpriv{\ln 4}{0} \X^k 01 \wedge 
+\dpriv{\ln 4}{0} \X^k 10 \wedge 
+\dpriv{\ln 4}{0} \X^k 11.
 \]

### 2018-01-28 examples.tex
mdps
use standard differential privacy definition e^{\epsilon}

@@
 respectively; the states $Y$ and $N$ denote answers to the survey
-question. We have $AP=\{Y,N\}$, $L(Y)=\{Y\}$, $L(N)=\{N\}$, and $L(+)=L(-)=\emptyset$. States $+$ and $-$ are neighbors.\lz{draw it on the picture?}
+question.
+We have $AP=\{Y,N\}$, $L(Y)=\{Y\}$, $L(N)=\{N\}$, and $L(+)=L(-)=\emptyset$.
+States $+$ and $-$ are neighbors.\lz{In general, we omit a special state $\dagger$ with $L(\dagger)=\emptyset$, and all omitted transitions in the picture lead to this state.}
 
@@
 neighboring $x, x' \in \mathcal{X}$. The survey mechanism is
-$(3, 0)$-differentially private.
+$(\ln 3, 0)$-differentially private.
 
@@
 output distribution on any inputs $x, x'$ with $f (x) = f
-(x')$. Moreover, it is also $(\alpha^{\Delta (f)}, 0)$-differentially
+(x')$. Moreover, it is also $(- {\Delta (f)} \ln \alpha, 0)$-differentially
 private for any query $f$ with sensitivity $\Delta (f)$. The range of
@@
 $\underline{n}$. The truncated $\alpha$-geometric mechanism is
-again oblivious; it is also $(\alpha^{\Delta (f)}, 0)$-differentially
+again oblivious; it is also $(- {\Delta (f)} \ln \alpha, 0)$-differentially
 private for any query $f$ with sensitivity $\Delta (f)$.

### 2018-01-28 mdp.tex
use standard differential privacy definition e^{\epsilon}

@@
 \mathcal{X}$. Similarly, $\Pr[M_H (x) = N] \leq 4 \Pr[M_H (x') =
-N]$. The high-accuracy mechanism is $(4,0)$-differentially private.
+N]$. The high-accuracy mechanism is $(\ln 4,0)$-differentially private.
 The privacy guarantees vary from accuracy requests.
@@
   \textmd{for all } t \textmd{ with } s \neighbor{S} t \textmd{ and
-   scheduler } \\
-  & & \scheduler{S},
-   \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq \epsilon \cdot
+   scheduler } \scheduler{S}, 
+   \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq 
+  e^{\epsilon} \cdot
+   \\
+  & & 
    \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} + \delta
    \textmd{ and }
-   \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq \epsilon \cdot
-  \\
-  & &
+   \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq e^{\epsilon} \cdot
    \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} + \delta
@@
   \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq
-  \epsilon \cdot\\
+  e^{\epsilon} \cdot\\
   && \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} + \delta \textmd{ and }
   \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq
-  \epsilon \cdot \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}
+  e^{\epsilon} \cdot \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}
   + \delta
@@
 $\myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq
-\epsilon \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} + \delta$
+e^{\epsilon} \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} + \delta$
 for every scheduler $\scheduler{S}$.
@@
 \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\gX B} -
-\epsilon \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\gX B}
+e^{\epsilon} \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\gX B}
 \end{eqnarray}

### 2018-01-28 specifications.tex
use standard differential privacy definition e^{\epsilon}

@@
  Define the $\dpCTL$ formula
-$\psi = \dpriv{.5}{0} (\F out_0) \wedge \dpriv{.5}{0} (\F out_1) \wedge
-\cdots \wedge \dpriv{.5}{0} (\F out_5)$\lz{I think we better use $\X$ operator here ;)}. If the state $in_k$ satisfies
+$\psi = \dpriv{.5}{0} (\X out_0) \wedge \dpriv{.5}{0} (\X out_1) \wedge
+\cdots \wedge \dpriv{.5}{0} (\X out_5)$. If the state $in_k$ satisfies
 $\psi$ for $k = 0, \ldots, 5$, then the $.5$-geometric mechanism is
-$(.5, 0)$-differentially private.
+$(\ln 2, 0)$-differentially private.
 
@@
 According to the compositional theorem \cite[Theorem 3.16]{DR:14:AFDP}, for $(\epsilon_1,\delta_1)$-differentially private algorithm $M_1$ and
-$(\epsilon_2,\delta_2)$-differentially private algorithm $M_2$, their combination $M$ defined by $M(x)=(M_1(x), M_2(x))$ is $(\epsilon_1\epsilon_2,\delta_1+\delta_2)$-differentially private.
+$(\epsilon_2,\delta_2)$-differentially private algorithm $M_2$, their combination $M$ defined by $M(x)=(M_1(x), M_2(x))$ is $(\epsilon_1 + \epsilon_2,\delta_1+\delta_2)$-differentially private.
 The degradation is rooted in the repeated computation.
@@
 
-We consider the formula $\dpriv{9}{0} (\X (Y\wedge \X Y))$.
+We consider the formula $\dpriv{\ln 9}{0} (\X (Y\wedge \X Y))$.
 A path satisfies $\X (Y\wedge \X Y)$ if the second state satisfies $Y$ and the third state satisfies $Y$ as well.
 We verify that this formula is satisfied  for state $+$ and $-$.
-Moreover, the bound $\epsilon=9$ is tight, since the probability of satisfying $\X (Y\wedge \X Y)$ from states $+$ and $-$ are $\frac{9}{16}$ and $\frac{1}{16}$, respectively.
-Finally, the formula $\wedge_{a_1,a_2}\dpriv{9}{0} (\X (a_1\wedge \X a_2))$ corresponds to the desired compositional property for the model, where $a_1,a_2$ range over all atomic propositions $\{Y,N\}$.
+Moreover, the bound $\epsilon=\ln 9$ is tight, since the probability of satisfying $\X (Y\wedge \X Y)$ from states $+$ and $-$ are $\frac{9}{16}$ and $\frac{1}{16}$, respectively.
+Finally, the formula $\wedge_{a_1,a_2}\dpriv{\ln 9}{0} (\X (a_1\wedge \X a_2))$ corresponds to the desired compositional property for the model, where $a_1,a_2$ range over all atomic propositions $\{Y,N\}$.
 
 We consider two other formulae for comparision: 
-(a) Formula $\dpriv{3}{0} (\X \X Y)$.
+(a) Formula $\dpriv{\ln 3}{0} (\X \X Y)$.
 In this case we donot have privacy lost, even though we are querying twice: the reason is that the output of the first query is not utilized at all.
 It is easy to verify that it is indeed satisfied by $+$ and $-$.
-(b) Formula $\dpriv{3}{0} (\X (Y\wedge \dpriv{3}{0}(\X Y)))$.
-This is a nested $\dpCTL$ formula, where the inner state formula $\dpriv{3}{0}(\X Y)$ states the one step differential privacy.
+(b) Formula $\dpriv{\ln 3}{0} (\X (Y\wedge \dpriv{\ln 3}{0}(\X Y)))$.
+This is a nested $\dpCTL$ formula, where the inner state formula $\dpriv{\ln 3}{0}(\X Y)$ states the one step differential privacy.
 Observe the inner formula is satisfied by all neighbours.
@@
 One of the advantage of applying model checking is that we may get tighter bounds for compositional properties.
-Consider our survey example, and the property $\dpriv{1}{.5} (\X Y)$.
+Consider our survey example, and the property $\dpriv{0}{.5} (\X Y)$.
 Obviously, it holds in states $+$ and $-$.
-A careful check ensures that one cannot decrease the value of $\epsilon_1=1$ or $\delta_1=.5$ without increasing the other one.
+A careful check ensures that one cannot decrease the value of $\epsilon_1=0$ or $\delta_1=.5$ without increasing the other one.
 
 Now we consider the formula $\dpriv{\epsilon_2}{\delta_2} (\X (Y\wedge \X Y))$ under the model in Figure~\ref{figure:2-dp-mc}.
-Applying the compositional theorem, one can choose $\epsilon_2=\epsilon_1^2=1$ and $\delta_2=2\delta_1=1$.
-However, we can check easily that one gets better privacy parameter $(1,0.5)$ or $(0,\frac{9}{16})$.
+Applying the compositional theorem, one can choose $\epsilon_2=2\epsilon_1=0$ and $\delta_2=2\delta_1=1$.
+However, we can check easily that one gets better privacy parameter $(0,0.5)$ or $(-\infty,\frac{9}{16})$.
+\bw{use $-\infty$ instead of $\ln 0$.
+is it okay?}
 

### 2018-01-28 specifications.tex
change (3, 0)-dp to (\ln 3, 0)-dp

@@
 \paragraph{Standard Differential Privacy.}
-The formula $\dpriv{3}{0} (\X Y)$ holds in state $s$ if for  $t$ with $sNt$, the probability of satisfying $\X Y$ from $s$ and $t$ satisfy:
+The formula $\dpriv{\ln 3}{0} (\X Y)$ holds in state $s$ if for  $t$ with $sNt$, the probability of satisfying $\X Y$ from $s$ and $t$ satisfy:
 $\myPr{s}{K}{N}{\X Y}\le 3\myPr{t}{K}{N}{\X Y}$.
-The formula $\dpriv{3}{0} (\X Y)\wedge \dpriv{3}{0} (\X N)$ thus specifies the differential privacy property for states $+$ and $-$ in Example~\ref{exa:survey}.
+The formula $\dpriv{\ln 3}{0} (\X Y)\wedge \dpriv{\ln 3}{0} (\X N)$ thus specifies the differential privacy property for states $+$ and $-$ in Example~\ref{exa:survey}.
 
  Define the $\dpCTL$ formula
-$\psi = \dpriv{.5}{0} (\X out_0) \wedge \dpriv{.5}{0} (\X out_1) \wedge
-\cdots \wedge \dpriv{.5}{0} (\X out_5)$. If the state $in_k$ satisfies
+$\psi = \dpriv{\ln 2}{0} (\X out_0) \wedge \dpriv{\ln 2}{0} (\X out_1) \wedge
+\cdots \wedge \dpriv{\ln 2}{0} (\X out_5)$. If the state $in_k$ satisfies
 $\psi$ for $k = 0, \ldots, 5$, then the $.5$-geometric mechanism is

### 2018-01-28 applications.tex
model model checking story

@@
 We would like to inquire the number of event types in
-observations. For instance, the query result result is $1$ 
+observations. For instance, the query result is $1$
 for observations $a \cdots a$ and $b \cdots b$ of arbitrary positive

### 2018-01-28 mdp.tex
model model checking story

@@
 
-As we describe earlier, the user has the possibility of asking different mechanisms.
+As we describe earlier, the user has the possibility of querying multiple mechanisms.
 We use Markov decision processes to model such
@@
 decision process. Define the satisfaction
-relation $M, \neighbor{S}, s \models \Phi$ as follows.
+relation $M, \neighbor{S}, s \models \Phi$ for $\PJ{J}$ and $\dpriv{\epsilon}{\delta} \phi$ (others are standard) as follows.
 \begin{eqnarray*}
-  M, \neighbor{S}, s \models \top\\
-  M, \neighbor{S}, s \models p
-  & \textmd{ if } &
-  p \in L(s)\\
-  M, \neighbor{S}, s \models \neg \Phi
-  & \textmd{ if } &
-  M, \neighbor{S}, s \not\models \Phi\\
-  M, \neighbor{S}, s \models \Phi_0 \wedge \Phi_1
-  & \textmd{ if } &
-  M, \neighbor{S}, s \models \Phi_0 \textmd{ and }
-  M, \neighbor{S}, s \models \Phi_1\\
+%  M, \neighbor{S}, s \models \top\\
+%  M, \neighbor{S}, s \models p  & \textmd{ if } &  p \in L(s)\\
+%  M, \neighbor{S}, s \models \neg \Phi  & \textmd{ if } &  M, \neighbor{S}, s \not\models \Phi\\
+%  M, \neighbor{S}, s \models \Phi_0 \wedge \Phi_1  & \textmd{ if } &  M, \neighbor{S}, s \models \Phi_0 \textmd{ and }
+%  M, \neighbor{S}, s \models \Phi_1\\
   M, \neighbor{S}, s \models \PJ{J} \phi
@@
 \emph{Justification of the scheduler class.}
-\lz{related to language equivalence...}
-
-\noindent
-\emph{Discussion.}
-\lz{this part not needed?}
-Another definition of the differentially private operator
-might be:
+For MDPs a richer class of scheduler classes have been considered in the literature, see \cite{BK:08:PMC}.
+It may depend on the full history of state and action sequences, and the decision can be randomised.
+If we use such scheduler, another definition of the differentially private operator
+might be:\lz{put $bad$ on $\models$?}
 \begin{eqnarray*}
@@
 take different actions from different states. It might need different
-action sequences to distinguish neighbors. Two neighbors could still be
-differentiated by a action sequence and lose their privacy. We hence
+action sequences to distinguish \lz{or mimic?} neighbors. \lz{add However,?}Two neighbors could still be
+differentiated by an action sequence and lose their privacy.
+
+We consider only sequences $\scheduler{S}\in Act^\omega$, which correspond to the query sequences.
+In the differential privacy context, neighbor states exhibit same action behaviours.
+And we will see later that
+our semantics agrees with the notion of differential privacy in the literature for such systems.
+Wpe hence
 prefer the original definition.
@@
 
-In general, the computation of the maximal probability for MDPs is non-trivial. \lz{we argue that it is difficult and leave as future work}
-
-Finally, we the property
-$\wedge_{i\in N} \dpriv{\epsilon}{0}(\X^k \top) \wedge \dpriv{\epsilon}{0}(F \top)$ that is used to express the privacy for our threshold Markov chains.
-It is only a valid formula if we would allow infinite conjunctions.
-However, it is interesting to note the relation to language equivalence for probabilistic automata\lz{to be discussed}
+In general, the problem is non-trivial.
+We consider the formula $\phi:=F B$ with $B\subseteq S$ and assume that states in $B$ are absorbing (with only self-loops).
+For the case $\epsilon=0$, the condition reduces to
+$\myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{F B} -
+\myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{F B}\le \delta$.
+If $\delta=0$ it is the classical language equivalence problem for probabilistic automata~\cite{Rabin63}, which can be solved in polynomial time \cite{Tzeng92}.
+However, if $\delta>0$, the problem
+becomes a metric version of the language equivalence problem, see \cite{FengZ13} for the definition.
+To the best of our knowledge, it has not be investigated in the literature.
+Thus we leave the general model checking algorithm as future works.
+\lz{see this}
 

### 2018-01-28 preliminaries.tex
use standard differential privacy definition e^{\epsilon}
model model checking story

@@
 \[
-\Pr[M_f (x) = \tilde{r}] \leq \epsilon \Pr[M_f (x') =
-\tilde{r}] + \delta.\footnote{The standard definition requires
-$\Pr[M_f (x) = \tilde{r}] \leq e^\epsilon \Pr[M_f (x') =
-\tilde{r}] + \delta$~\cite{DMNS:06:CNSPD,D:06:DP}.
-Since $\epsilon = e^{\ln \epsilon}$, our $(\epsilon, \delta)$-closedness is
-equivalent to the standard $(\ln \epsilon,
-\delta)$-closedness~\cite{GRS:09:UUPM,GRS:12:UUPM}.}
+\Pr[M_f (x) = \tilde{r}] \leq e^{\epsilon} \Pr[M_f (x') =
+\tilde{r}] + \delta.
+%\footnote{The standard definition requires
+%$\Pr[M_f (x) = \tilde{r}] \leq e^\epsilon \Pr[M_f (x') =
+%\tilde{r}] + \delta$~\cite{DMNS:06:CNSPD,D:06:DP}.
+%Since $\epsilon = e^{\ln \epsilon}$, our $(\epsilon, \delta)$-closedness is
+%equivalent to the standard $(\ln \epsilon,
+%\delta)$-closedness~\cite{GRS:09:UUPM,GRS:12:UUPM}.}
 \]
@@
 
-Let $\Act$ denote a finite set of \emph{actions}.
-A \emph{(finite) Markov decision process} $M = (S, \Act, \wp, L)$ consists
-of a \emph{transition probability function} $\wp : S \times \Act
+A \emph{(finite) Markov decision process} 
+(MDP)~\footnote{The MDP we consider is \emph{reactive} in the sense that 
+all actions are enabled in every state. If equipped with a set of final states $F\subseteq S$, such model is
+referred to as \emph{probabilistic automaton} in the literature~\cite{Rabin63}, and language equivalence/inclusion
+problems have been investigated. In the model checking context, the notion MDPs is mostly used.}
+ $M = (S, \Act, \wp, L)$ consists
+of 
+a finite set of \emph{actions} $\Act$,
+a \emph{transition probability function} $\wp : S \times \Act
 \times S \rightarrow [0, 1]$ with $\sum_{t \in S} \wp(s, \alpha, t)
@@
 $L$ as for Markov chains.
-Note the model we consider is \emph{reactive}\footnote{In the literature, such model is referred to as \emph{probabilistic automata}\cite{}.
-In the model checking context, the notion MDPs is more popular.} as $\Act(s) = \Act$
-for every $s \in S$.
 A \emph{path} $\pi$ in $M$ is an infinite sequence $\pi_0 \alpha_1
@@
 
-Let $M = (S, \Act, \wp, L)$ be a Markov decision process. A
+Let $M = (S, \Act, \wp, L)$ be an MDP. A
 \emph{scheduler} for $M$ is simply an infinite sequence  $\scheduler{S}\in \Act^\omega$.
@@
 \pi_1 \cdots \pi_i)$ for all $i \geq 0$.
-Note that a Markov decision process with a
+Note that an MDP with a
 scheduler $\scheduler{S}$ induces a Markov chain $M_{\scheduler{S}} =

### 2018-01-28 specifications.tex
model model checking story

@@
 
- Define the $\dpCTL$ formula
+ Define the formula
 $\psi = \dpriv{\ln 2}{0} (\X out_0) \wedge \dpriv{\ln 2}{0} (\X out_1) \wedge
@@
 \bw{use $-\infty$ instead of $\ln 0$.
-is it okay?}
+is it okay?}\lz{lets just remove it ;)}
 
@@
 Here $\dpriv{\epsilon}{0}(F \top)$ is a reachability privacy: it states that related states should have similar probabilities of eventually reaching $\top$.
+
+Finally, we the property
+$\wedge_{i\in N} \dpriv{\epsilon}{0}(\X^k \top) \wedge \dpriv{\epsilon}{0}(F \top)$ that is used to express the privacy for our threshold Markov chains.
+It is only a valid formula if we would allow infinite conjunctions.
+However, it is interesting to note the relation to language equivalence for probabilistic automata\lz{to be discussed}
 

### 2018-01-29 mdp.tex
some changes

@@
 
-\paragraph{Model Checking Algorithm.}
-
+\paragraph{Discussion about the Model Checking Algorithm.}
 Our next task is to compute $\{ M, \neighbor{S}, s \models
@@
 If the path formula only contain next operators, it is easy to deal with.
-For $\phi:=\X B$ with $B\subseteq S$ only one step behaviour needs to be considered, thus only the first action in the query sequence is needed to compute the maximum
-\begin{eqnarray}
-  \label{eqn:max-nextB}
-\max_{\scheduler{S}}
-\myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\gX B} -
-e^{\epsilon} \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\gX B}
-\end{eqnarray}
-with $s, t \in S$. It can be easily generalized to nested next operators:  one needs only to enumerate
+For $\phi:=\X B$ with $B\subseteq S$ only one step behaviour needs to be considered, thus only the first action in the query sequence is needed.
+It can also be easily generalized to nested next operators:  one needs only to enumerate
 all actions query sequences of length obtained by the number of nexted $\X$ operators.
@@
 In general, the problem is non-trivial.
-We consider the formula $\phi:=F B$ with $B\subseteq S$ and assume that states in $B$ are absorbing (with only self-loops).
+For instance, we consider the formula $\phi:=F B$ with $B\subseteq S$ and assume that states in $B$ are absorbing (with only self-loops).
 For the case $\epsilon=0$, the condition reduces to
@@
 becomes a metric version of the language equivalence problem, see \cite{FengZ13} for the definition.
-To the best of our knowledge, it has not be investigated in the literature.
+To the best of our knowledge, the decision problem has not be investigated in the literature.
+If in addition $\epsilon>0$, it seems to be a problem that has not been considered in the MDP model checking community before.
 Thus we leave the general model checking algorithm as future works.

### 2018-01-29 semantics.tex
use standard differential privacy definition e^{\epsilon}
some changes

@@
   \textmd{for every } t \textmd{ with }  s \neighbor{S} t,
-      p(s) \leq \epsilon p(t) + \delta \textmd{ and }
-      p(t) \leq \\
-  & & \epsilon p(s)  + \delta  \textmd{ where } p(x) = \Pr [\{
-      \pi : K, \neighbor{S}, \pi \models \phi \textmd{ with}\\
-  & &  \pi_0 = x \}]
+      p(s) \leq e^{\epsilon} p(t) + \delta \textmd{ and }
+      p(t) \leq e^{\epsilon} p(s)  +  \\
+  & &  \delta \textmd{ where } p(x) = \Pr [\{
+      \pi : K, \neighbor{S}, \pi \models \phi \textmd{ with}
+      \pi_0 = x \}]
 \end{eqnarray*}
@@
 Moreover, the relation $K, \neighbor{S}, \pi \models \phi$ is defined as
-for the standard LTL formulas. We only give the semantics for $\X$ and $\until$ as follows.
+for the standard LTL formulas~\cite{which-to-cite?}. We only recall the semantics for $\X$ and $\until$ as follows.
 \begin{eqnarray*}
@@
   \textmd{for every } t \textmd{ with }  s \neighbor{S} t,
-      \myPr{s}{K}{N}{\phi} \leq \epsilon \myPr{t}{K}{N}{\phi} + \delta
+      \myPr{s}{K}{N}{\phi} \leq e^{\epsilon} \myPr{t}{K}{N}{\phi} + \delta
       \textmd{ and }\\
-  & &  \myPr{t}{K}{N}{\phi} \leq \epsilon \myPr{s}{K}{N}{\phi} + \delta
+  & &  \myPr{t}{K}{N}{\phi} \leq e^{\epsilon} \myPr{s}{K}{N}{\phi} + \delta
 \end{eqnarray*}
@@
 In Lemma~\ref{lemma:PJ-subroutines}, we abuse the notation slightly to
-admit path formulae of the form $\gX B$ (next $B$) and
+admit path formulae of the form $\gX B$ (next $B$) and\lz{what is $\guntil$?}
 $B \guntil C$ ($B$ until $C$) with $B, C \subseteq S$ as
@@
       \For{$t$ with $s \neighbor{K} t$}
-        \If{$p(s) \not\leq \epsilon p(t) + \delta$ or
-            $p(t) \not\leq \epsilon p(s) + \delta$}
+        \If{$p(s) \not\leq e^{\epsilon} p(t) + \delta$ or
+            $p(t) \not\leq e^{\epsilon} p(s) + \delta$}
         {remove $s$ from $R$}
@@
 Then, the probability reduces to a
-  reachability probability in the product Markov chain obtained from $K$ and $R$. There are more efficient algorithms without the product construction, see \cite{} for details.
+  reachability probability in the product Markov chain obtained from $K$ and $R$. There are more efficient algorithms without the product construction, see \cite{CY95,CSS03,BaierKKKMW16} for details.

### 2018-01-29 specifications.tex
some changes

@@
 \paragraph{Standard Differential Privacy.}
-The formula $\dpriv{\ln 3}{0} (\X Y)$ holds in state $s$ if for  $t$ with $sNt$, the probability of satisfying $\X Y$ from $s$ and $t$ satisfy:
+The formula $\dpriv{\ln 3}{0} (\X Y)$ \lz{$\X Y$ looks a bit weired?} holds in state $s$ if for  $t$ with $sNt$, the probability of satisfying $\X Y$ from $s$ and $t$ satisfy:
 $\myPr{s}{K}{N}{\X Y}\le 3\myPr{t}{K}{N}{\X Y}$.

### 2018-01-29 mdp.tex
small changes

@@
 \paragraph{Discussion about the Model Checking Algorithm.}
-Our next task is to compute $\{ M, \neighbor{S}, s \models
-\dpriv{\epsilon}{\delta} \phi \}$ for a Markov decision process $M =
-(S, \Act, \wp, L)$. Recall the semantics of $\dpriv{\epsilon}{\delta}
+For an MDP $M =
+(S, \Act, \wp, L)$, we want to compute $\{ M, \neighbor{S}, s \models
+\dpriv{\epsilon}{\delta} \phi \}$ . Recall the semantics of $\dpriv{\epsilon}{\delta}
 \phi$. Given $s, t$ with $s \neighbor{M} t$ and a path formula
@@
 However, if $\delta>0$, the problem
-becomes a metric version of the language equivalence problem, see \cite{FengZ13} for the definition.
-To the best of our knowledge, the decision problem has not be investigated in the literature.
+becomes an approximate version of the language equivalence problem~\cite{Tzeng92}.
+To the best of our knowledge, the decision problem is still open, and has only be investigated for the special case when all states are connected~\cite{Tzeng92}.
 If in addition $\epsilon>0$, it seems to be a problem that has not been considered in the MDP model checking community before.

### 2018-01-29 semantics.tex
small changes

@@
 
-The model checking algorithm for $\dpCTLstar$ can be treated as in the classical setting \cite{BK:08:PMC}, as all we need is to  compute the  probability
+The model checking algorithm for $\dpCTLstar$ can be treated as in the classical setting~\cite{BK:08:PMC}:  all we need is to  compute the  probability
   $\myPr{s}{K}{\neighbor{S}}{\phi}$ with general path formula $\phi$.

### 2018-01-29 introduction.tex

% privacy and differential privacy

In the era of data analysis, personal information is constantly
collected and analyzed by various parties. Privacy has become an
important issue for every individual. In order to address such
concerns, the research community has proposed several privacy
preserving mechanisms over the years (see~\cite{FWC:10:PPDP} for a
slightly outdated survey). Among these mechanisms, differential
privacy has attracted much attention from theoretical computer science
to industry~\cite{DR:14:AFDP,JLE:14:DPML,A:16:EPYU}.

Differential privacy formalizes the tradeoff between privacy and
utility in data analysis. Intuitively, a randomized data analysis
mechanism is differentially private if it behaves similarly on similar
input databases~\cite{DMNS:06:CNSPD,D:06:DP}. Consider, for example,
the Laplace mechanism where a random noise is added to analysis
results with the Laplace distribution~\cite{DR:14:AFDP}. Random noises
hide the differences of analysis results from similar databases.
Clearly, more
randomness gives more privacy but less utility in released noisy
results. Under the framework of differential privacy, data analysts
can balance the tradeoff rigorously in their data analysis
mechanisms~\cite{DR:14:AFDP,JLE:14:DPML}.

% model checking differentially private implementation

Designing differentially private mechanisms can be tedious for
sophisticated data analyses. Privacy leak has also been observed from
data analysis programs with differentially private
designs~\cite{M:12:SLSBDP,TKBW:17:PLAI}. This calls for formal
analysis of differential privacy on both designs and implementations.
In tis paper, we propose the logic $\dpCTLstar$ for specifying
differential privacy and investigate their model checking
algorithms. Data analysts can automatically verify their designs and
implementations with our techniques. Most interestingly, our
techniques can be adopted rather easily by existing probabilistic
model checkers. Hopefully, more interaction between model checking and
privacy analysis will follow.

% Markov chains and MDP as formal models, actions made by adversary

In differential privacy, data analysis mechanisms are
but randomized algorithms. We follow the standard practice in
probabilistic model checking to formalize such mechanisms by
by Markov chains or Markov decision processes. We give
formalizations of several differentially private mechanisms and
explain subtle privacy conditions in this paper.
When a data analysis mechanism does not
interact with its environment, it is formalized as a Markov
chain. Otherwise, its interactions are formalized by actions in
Markov decision processes. Our formalization effectively assumes that
actions are controlled by adversaries. It is subsequently necessary to
consider attacks from all possible adversaries in order to establish
differential privacy.
%\todo{What is a state?}

% new differential privacy state modal operator

Two ingredients are introduced to specify differentially private
behaviors. A reflexive and commutative user-defined binary relation
over states is required to formalize similarity. Two states are
similar if they are related by the binary relation. We moreover add
the path quantifier $\dpriv{\epsilon}{\delta}$ for similar
behaviors. Informally, a state satisfies $\dpriv{\epsilon}{\delta}
\phi$ if it has probabilistically similar path behavior $\phi$ as
its similar states. Consider, for instance, a binary data analysis
mechanism computing the likelihood ($\mathit{high}$ or $\mathit{low}$)
of an epidemic in a population. Then a state satisfies
$\dpriv{\epsilon}{\delta} (\F \mathit{high}) \wedge
\dpriv{\epsilon}{\delta} (\F \mathit{low})$ if the mechanism is
$(\epsilon, \delta)$-differentially private at the state.
%\todo{not clear here}\lz{here it should be $\X$?}

% differential privacy of infinite behaviors

An important byproduct of this work is to analyze privacy over
infinite behaviors. Consider the formula $\dpriv{\epsilon}{\delta}
(\G \F \mathit{high})$. A state satisfies the formula if it
is probabilistically similar to always almost surely
answer $\mathit{high}$ as similar states. Compared to the unbounded
but finite behavior specified by $\F \mathit{high}$,
$\G \F \mathit{high}$ specifies an infinite behavior.
Not only can differential privacy specify randomized algorithms, it
also naturally generalizes to reactive systems.
Privacy analysis applies to systems with infinite behaviors as well.

% construction and complexity

In order to verify $\dpCTLstar$ properties, we extend the standard
probabilistic model checking algorithms. For Markov chains, the states
satisfying a subformula $\dpriv{\epsilon}{\delta} \phi$ are computed
by a simple variant of the algorithm for Markov chains.
The time complexity of our model checking algorithm is
still polynomial. The logic $\dpCTLstar$ essentially obtains its
expressiveness for free. \bw{double check}
For Markov decision processes, we show that checking whether a state
satisfies $\dpriv{\epsilon}{\delta} \phi$ is undecidable.

% analyze mechanism

### 2018-01-29 main.tex
structures and some examples
more changes ;)
examples
remove product mdp from figure
use Times font to get more space
some changes
small changes

@@
 \captionsetup{compatibility=false}
+\usepackage{wrapfig}
 \usepackage{mathtools}
 \usepackage{tikz}
+\usetikzlibrary{arrows,calc}
 \usepackage{url}
-\usetikzlibrary{arrows,calc}
+\usepackage{times}
 
@@
 \begin{abstract}
-  We introduce the branching time temporal logic \dpCTL for specifying
-  differential privacy. Its model checking algorithm is also developed. 
+  We introduce the branching time temporal logic $\dpCTLstar$ for
+  specifying differential privacy. Several differentially private
+  mechanisms are formalized as Markov chains or Markov decision
+  processes. Using our formal models, subtle privacy conditions
+  are specified by $\dpCTLstar$. In order to verify privacy properties
+  automatically, the model checking problems are investigated.  We
+  give a model checking algorithm for Markov chains. Model checking
+  $\dpCTLstar$ properties on Markov decision processes however is
+  shown to be undecidable.
 \end{abstract}
@@
 
-\section{Examples}
+\section{Markov Chains  for Differentially Private Mechanisms}
 \label{section:examples}
@@
 
-\section{The Logic $\dpCTL$}
+\section{The Logic $\dpCTLstar$}
 \label{section:dpCTL}
@@
 
-\section{Model Checking Algorithms}
-\label{section:model-checking-algorithms}
-\input{algorithms}
+\section{Specifying Properties in $\dpCTLstar$}
+\label{section:specifying-properties}
+\input{specifications}
 
-\section{Applications}
+\section{Markov Decision Processes}
+\label{section:mdp}
+\input{mdp}
+
+%\section{Model Checking Algorithms}
+%\label{section:model-checking-algorithms}
+%\input{algorithms}
+
+\subsection{MDPs for Streaming Processes}
 \label{section:applications}

### 2018-01-29 mdp.tex
proof almost finished

@@
 diagnosed with the disease X. This gives an interactive mechanism
-corresponding to the Markov decision process shown in
-Figure~\ref{figure:simple-mdp-mdp}.
+corresponding to the MDP shown in
+Figure~\ref{figure:simple-mdp}.\lz{remove the loops}
 
@@
 \subsection{Semantics}
-The semantics over Markov chains generalizes to Markov decision
-processes as well. Note that two states with different enabled actions
-are trivially distinguishable; no privacy can be preserved. Without
-loss of generality, we consider reactive Markov decision processes.
-Let $M = (S, \Act, \wp, L)$ be a reactive Markov
-decision process. Define the satisfaction
+The logic $\dpCTLstar$ can be interpreted over MDPs. Semantics over Markov chains generalizes to MDPs as well. Note that two states with different enabled actions
+are trivially distinguishable; no privacy can be preserved.
+Let $M = (S, \Act, \wp, L)$ be an MDP. Define the satisfaction
 relation $M, \neighbor{S}, s \models \Phi$ for $\PJ{J}$ and $\dpriv{\epsilon}{\delta} \phi$ (others are standard) as follows.
@@
 defined as in Markov chains.
-The semantics of $\dpCTLstar$ on Markov decision processes
+The semantics of $\dpCTLstar$ on MDPs
 is again standard except the differentially private operator
@@
 
-\paragraph{Discussion about the Model Checking Algorithm.}
+\subsection{Model Checking Algorithm}
 For an MDP $M =
@@
 all actions query sequences of length obtained by the number of nexted $\X$ operators.
+In general, the problem is undecidable.
 
-In general, the problem is non-trivial.
-For instance, we consider the formula $\phi:=F B$ with $B\subseteq S$ and assume that states in $B$ are absorbing (with only self-loops).
-For the case $\epsilon=0$, the condition reduces to
+\begin{theorem}
+The $\dpCTLstar$ model checking problem for MDPs  is undecidable.
+\end{theorem}
+\begin{proof}
+The proof follows by a reduction from the \emph{threshold} problem for probabilistic automata.
+A \emph{probabilistic automaton}~\cite{Rabin63} is a tuple $\mathcal{A}= (S,\Sigma,M,s_0,F)$ where
+\begin{itemize}
+  \item $S$ is a finite set of states,
+  \item $\Sigma$ is the finite set of input alphabet,
+  \item $M: S\times\Sigma\times S \to [0,1]$ such that $\sum_{t\in S} M(s,\alpha,t)=1$ for all $s\in S$ and $\alpha\in \Sigma$,
+  \item $s_0\in S$ is the initial state,
+  \item $F\subseteq S$ is a set of accepting states.
+\end{itemize}
+
+Each input alphabet $\alpha$ induces a stochastic matrix $M(\alpha)$ in the obvious way.
+Let $\lambda$ denote the empty string.
+For $\eta\in\Sigma^*$ we define $M(\eta)$ inductively by: $M(\lambda)$ is the identity matrix, $M(x\eta')=M(x)M(\eta')$.
+Thus, $M(\eta)(s,s')$ denotes the probability of going from $s$ to $s'$ after reading $\eta$.
+Let $v_F$ denote the characteristic row vector for the set $F$, and $v_{s_0}$ denote the characteristic row vector for the set $\{s_0\}$.
+Then, the accepting probably of $\eta$ by $\mathcal{A}$ is defined as $v_{s_0}\cdot M(\eta)\cdot (v_F)^c$ where $(v_F)^c$ denotes the transpose of $v_F$.
+The following \emph{emptyness problem} is know to be undecidable~\cite{Paz71}
+%~\cite{Worrell08}:
+
+\paragraph{Emptiness problem:} Given a probabilistic automaton $\mathcal{A}= (S,\Sigma,M,s_0,F)$, whether there exists $\eta\in\Sigma^*$ such that $v_{s_0}\cdot M(\eta)\cdot (v_F)^c > 0$?
+
+Now we establish the proof by reducing the threshold problem to our $\dpCTLstar$ model checking problem.
+Given the probabilistic automaton  $\mathcal{A}= (S,\Sigma,M,s_0,F)$, assume we have a \emph{primed} copy $\mathcal{A}¡¯= (S¡¯,\Sigma,M',s_0¡¯,\emptyset)$.
+
+Now we construct our MDP  $M =
+(S\cupdot S', \Sigma, \wp, L)$ where $\wp(s,a,t)$ equals to $M(s,a,t)$ if $s,t\in S$ and to $M'(s,a,t)$ if $s,t\in S'$.
+We define the neighbor relation $N_S:=\{(s_0,s_0')\}$ by relating states $s_0,s_0'$.
+The labelling function $L$ is defined by $L(s)=at_F$ if $s\in F$ and $L(s)=\empty$ otherwise.
+
+\lz{shall we use $\diamond$ for finally?}
+Now we consider the formula  $\Phi = \dpriv{1}{0} (\diamond at_F)$.
+For the reduction we prove $s_0\models \dpriv{1}{0} (\diamond at_F)$ iff for all $\eta\in\Sigma^*$ it holds $v_{s_0}\cdot M(\eta)\cdot (v_F)^c \leq 0$.
+
+\lz{remember explain meaning of $\Sigma^\omega$}
+First we assume $s_0\models \dpriv{1}{0} (\diamond at_F)$. By $\dpCTLstar$ semantics we have that for all $\scheduler{S}\in\Sigma^\omega$,
+$\myPr{s_0}{M_{\scheduler{S}}}{\neighbor{S}}{\diamond at_F} \leq
+  e \cdot
+  \myPr{s_0'}{M_{\scheduler{S}}}{\neighbor{S}}{\diamond at_F}$. Observe $\myPr{s_0'}{M_{\scheduler{S}}}{\neighbor{S}}{\diamond at_F}=0$ by construction, thus we have
+$\myPr{s_0}{M_{\scheduler{S}}}{\neighbor{S}}{\diamond at_F} \leq
+ 0$. This implies  $v_{s_0}\cdot M(\eta)\cdot (v_F)^c \leq 0$ for all $\eta\in\Sigma^*$.
+
+ For the other direction, 
+\end{proof}
+
+For instance, we consider the formula $\phi:=F B$ with $B\subseteq S$ and assume that states in $B$ are absorbing (with only self-loops). For the case $\epsilon=0$, the condition reduces to
 $\myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{F B} -

### 2018-01-30 mdp.tex
proof completed

@@
 \begin{proof}
-The proof follows by a reduction from the \emph{threshold} problem for probabilistic automata.
+The proof follows by a reduction from the \emph{emptiness} problem for probabilistic automata.
 A \emph{probabilistic automaton}~\cite{Rabin63} is a tuple $\mathcal{A}= (S,\Sigma,M,s_0,F)$ where
@@
 Then, the accepting probably of $\eta$ by $\mathcal{A}$ is defined as $v_{s_0}\cdot M(\eta)\cdot (v_F)^c$ where $(v_F)^c$ denotes the transpose of $v_F$.
-The following \emph{emptyness problem} is know to be undecidable~\cite{Paz71}
+The following \emph{emptiness problem} is know to be undecidable~\cite{Paz71}
 %~\cite{Worrell08}:
@@
 
- For the other direction, 
+ For the other direction, assume that all $\eta\in\Sigma^*$ it holds $v_{s_0}\cdot M(\eta)\cdot (v_F)^c \leq 0$.
+For the sake of contradiction, assume that $s_0\not\models \dpriv{1}{0} (\diamond at_F)$.
+Since the relation $N_S$ contains only one pair, it implies that there exists a scheduler $\scheduler{S}\in\Sigma^\omega$ such that $\myPr{s_0}{M_{\scheduler{S}}}{\neighbor{S}}{\diamond at_F} >0$.
+It is then easy to construct a finite sequence $\eta\in\Sigma^*$ with  $v_{s_0}\cdot M(\eta)\cdot (v_F)^c > 0$, a contradiction.
 \end{proof}
 
-For instance, we consider the formula $\phi:=F B$ with $B\subseteq S$ and assume that states in $B$ are absorbing (with only self-loops). For the case $\epsilon=0$, the condition reduces to
+We consider some decidable special cases.
+Consider the formula $\phi:=F B$ with $B\subseteq S$ and assume that states in $B$ are absorbing (with only self-loops).
+For the case $\epsilon=0$, the condition reduces to
 $\myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{F B} -
@@
 To the best of our knowledge, the decision problem is still open, and has only be investigated for the special case when all states are connected~\cite{Tzeng92}.
-If in addition $\epsilon>0$, it seems to be a problem that has not been considered in the MDP model checking community before.
-Thus we leave the general model checking algorithm as future works.
-\lz{see this}
-
+p
 \hide{

### 2018-01-30 specifications.tex
proof completed

@@
 \paragraph{Reachability Privacy.}
+The formula $\dpriv{\epsilon}{\delta}(F \top)$ is a reachability privacy: it states that related states should have similar probabilities of eventually reaching $\top$.
+Note neighbor states may reach the goal state $\top$ with
+
 For our threshold \lz{to do} model, we have the differential privacy property $\wedge_{i=1}^{c} \dpriv{\epsilon}{0}(\X^k \top) \wedge \dpriv{\epsilon}{0}(F \top)$.
-Here $\dpriv{\epsilon}{0}(F \top)$ is a reachability privacy: it states that related states should have similar probabilities of eventually reaching $\top$.
 

### 2018-01-30 specifications.tex
add bluetooth DoS description

@@
 
-We consider the Bluetooth example from the PRISM benchmarks \cite{prism-bluetooth-page}. The model is a
-Markov chain, and it ...
-
-For instance,
-$\dpriv{\epsilon}{\delta} (\G   \F \mathit{discovered})$
-specifies states having $(\epsilon, \delta)$-close
-infinitely often $\mathit{discovered}$ behaviors to its neighbors.
+Consider the Bluetooth example in~\cite{prism-bluetooth-page} and the
+formula $\dpriv{\epsilon}{\delta}{\F \G \mathit{timeout}}$. Let us say
+two states are neighbors if the number of slave devices differs by at
+most one. Then the formula holds if the probability of having
+continuous time out (say, in the denial of service attack) is
+$(\epsilon, \delta)$-close for all neighbors. That is, adding a slave
+device cannot increase the chance of service failure significantly.

### 2018-01-30 introduction.tex
contributions

@@
 % analyze mechanism
+
+\paragraph{Contributions.} Our main contributions are three folds. (i)
+We introduce the logic $\dpCTLstar$ for reasoning about properties in 
+differential privacy. We use the logic to express classical
+differential privacy properties, and moreover, complex properties such as those 
+involving infinite behaviors can also be expressed. (ii) We model several differential privacy mechanisms, including the streaming process, in Markov chains or MDPs.
+(iii) We show that the model checking problem 
+problem for Markov chains is standard. For MDPs we show that it is undecidable.

### 2018-01-30 introduction.tex
explain complexity

@@
 by a simple variant of the algorithm for Markov chains.
-The time complexity of our model checking algorithm is
-still polynomial. The logic $\dpCTLstar$ essentially obtains its
-expressiveness for free. \bw{double check}
+The time complexity of our model checking algorithm is the same as
+those of $\CTLstar$ for Markov chains. The logic $\dpCTLstar$ obtains its
+expressiveness for free. 
 For Markov decision processes, we show that checking whether a state

### 2018-01-30 applications.tex
fix typo
remove notes

@@
 $1$, or $2$ by the truncated $\frac{1}{2}$-geometric mechanism over
-$\underline{2}$ (Figure~\ref{figure:randomized-frequency}, \`{a} la
+$\{ 0, 1, 2 \}$ (Figure~\ref{figure:randomized-frequency}, \`{a} la
 Figure~\ref{figure:geometric-mechanism}). For any action sequence,
@@
 \[
-\dpriv{\ln 4}{0} \F 0 \wedge \dpriv{\ln 4}{0} \F 1 \wedge
-\dpriv{\ln 4}{0} \F 2
+\Psi = \dpriv{\ln 4}{0} \F \mathit{at}_0 \wedge \dpriv{\ln 4}{0} \F
+\mathit{at}_1 \wedge \dpriv{\ln 4}{0} \F \mathit{at}_2
 \]
@@
 \[
-\dpriv{\ln 4}{0} \F 0 \wedge \dpriv{2\ln 4}{0} \F 1 \wedge
-\dpriv{\ln 4}{0} \F 2
+\Psi
 \wedge
 \bigwedge_{k \in \bbfN}
-\dpriv{\ln 4}{0} \X^k 00 \wedge
-\dpriv{\ln 4}{0} \X^k 01 \wedge
-\dpriv{\ln 4}{0} \X^k 10 \wedge
-\dpriv{\ln 4}{0} \X^k 11.
+\dpriv{\ln 4}{0} \X^k \mathit{at}_{00} \wedge
+\dpriv{\ln 4}{0} \X^k \mathit{at}_{01} \wedge
+\dpriv{\ln 4}{0} \X^k \mathit{at}_{10} \wedge
+\dpriv{\ln 4}{0} \X^k \mathit{at}_{11}.
 \]

### 2018-01-30 examples.tex
remove notes

@@
 respectively; the states $Y$ and $N$ denote answers to the survey
-question.
-We have $AP=\{Y,N\}$, $L(Y)=\{Y\}$, $L(N)=\{N\}$, and $L(+)=L(-)=\emptyset$.
-States $+$ and $-$ are neighbors.\lz{In general, we omit a special state $\dagger$ with $L(\dagger)=\emptyset$, and all omitted transitions in the picture lead to this state.}
-
+question. We have $AP=\{Y,N\}$, $L(Y)=\{Y\}$, $L(N)=\{N\}$, and $L(+)=L(-)=\emptyset$. States $+$ and $-$ are neighbors. 
 The random noise boosts the probability of answering
@@
 diagnoses. Inferences on individual diagnosis can be plausibly denied.
+{In general, we omit a special state $\dagger$ with $L(\dagger)=\emptyset$, and all omitted transitions in the picture lead to this state.}
 \end{example}

### 2018-01-30 mdp.tex
contributions
remove notes

@@
 
-      (F) edge [loop left] node [below] { $-, 1$ } (F)
-      (T) edge [loop right] node [below] { $-, 1$ } (T)
+%      (F) edge [loop left] node [below] { $-, 1$ } (F)
+%      (T) edge [loop right] node [below] { $-, 1$ } (T)
       ;
@@
 corresponding to the MDP shown in
-Figure~\ref{figure:simple-mdp}.\lz{remove the loops}
+Figure~\ref{figure:simple-mdp}.
 
@@
 If we use such scheduler, another definition of the differentially private operator
-might be:\lz{put $bad$ on $\models$?}
+might be:
 \begin{eqnarray*}
@@
 take different actions from different states. It might need different
-action sequences to distinguish \lz{or mimic?} neighbors. \lz{add However,?}Two neighbors could still be
+action sequences to distinguish neighbors. Two neighbors could still be
 differentiated by an action sequence and lose their privacy.
@@
 Now we establish the proof by reducing the threshold problem to our $\dpCTLstar$ model checking problem.
-Given the probabilistic automaton  $\mathcal{A}= (S,\Sigma,M,s_0,F)$, assume we have a \emph{primed} copy $\mathcal{A}¡¯= (S¡¯,\Sigma,M',s_0¡¯,\emptyset)$.
+Given the probabilistic automaton  $\mathcal{A}= (S,\Sigma,M,s_0,F)$, assume we have a \emph{primed} copy $\mathcal{A}'= (S',\Sigma,M',s_0',\emptyset)$.
 
@@
 To the best of our knowledge, the decision problem is still open, and has only be investigated for the special case when all states are connected~\cite{Tzeng92}.
-p
+
 \hide{

### 2018-01-30 preliminaries.tex
proof almost finished
remove notes

@@
 Let $\bbfZ$ and $\bbfZ^{\geq 0}$ be the sets of integers and
-non-negative integers respectively. For $n \geq 0$, $\underline{n} =
-\{ 0, 1, \ldots, n \} \subseteq \bbfZ^{\geq 0}$.
-
+non-negative integers respectively. 
+% For $n \geq 0$, $\underline{n} = \{ 0, 1, \ldots, n \} \subseteq
+% \bbfZ^{\geq 0}$.
+%\lz{not so often used: can we do withuot this notatin?}
 We briefly review the definitions of differential privacy, Markov
 chains, and Markov decision processes. For differential privacy, we
-follow the non-standard \todo{why?} definition in~\cite{GRS:09:UUPM,GRS:12:UUPM}. Our
+follow the standard definition in~\cite{DR:14:AFDP,GRS:09:UUPM,GRS:12:UUPM}. Our
 definitions of Markov chains and Markov decision processes are adopted
@@
 similar behabiors on neighboring databases. It will have similar
-output distributions when a row is added, deleted, or replaced by
-another in a given
+output distributions when a row is replaced by another in a given
 database. Consider, for instance, a database contains the row
@@
 mechanisms will behave similarly. Privacy of the individual is thus
-preserved by differentially private mechanisms. \todo{not clear}
+preserved by differentially private mechanisms.
 

### 2018-01-30 semantics.tex
remove notes

@@
 In Lemma~\ref{lemma:PJ-subroutines}, we abuse the notation slightly to
-admit path formulae of the form $\gX B$ (next $B$) and\lz{what is $\guntil$?}
+admit path formulae of the form $\gX B$ (next $B$) and
 $B \guntil C$ ($B$ until $C$) with $B, C \subseteq S$ as
@@
 propositions $B$ and $C$ for each $s \in B$ and $s \in C$
-respectively. We will also use graphical symbols $\gF C$ ($C$ in
+respectively. 
+\hide{
+We will also use graphical symbols $\gF C$ ($C$ in
 the future) and $\gG C$ (globally $C$) in such path formulae
 $\varphi$ to minimize confusion.
+}
 

### 2018-01-30 specifications.tex
remove notes

@@
 \paragraph{Standard Differential Privacy.}
-The formula $\dpriv{\ln 3}{0} (\X Y)$ \lz{$\X Y$ looks a bit weired?} holds in state $s$ if for  $t$ with $sNt$, the probability of satisfying $\X Y$ from $s$ and $t$ satisfy:
+The formula $\dpriv{\ln 3}{0} (\X \mathit{out}_Y)$ \lz{$\X Y$ looks a bit weired?} holds in state $s$ if for  $t$ with $sNt$, the probability of satisfying $\X Y$ from $s$ and $t$ satisfy:
 $\myPr{s}{K}{N}{\X Y}\le 3\myPr{t}{K}{N}{\X Y}$.
@@
 
-Now we consider the formula $\dpriv{\epsilon_2}{\delta_2} (\X (Y\wedge \X Y))$ under the model in Figure~\ref{figure:2-dp-mc}.
-Applying the compositional theorem, one can choose $\epsilon_2=2\epsilon_1=0$ and $\delta_2=2\delta_1=1$.
-However, we can check easily that one gets better privacy parameter $(0,0.5)$ or $(-\infty,\frac{9}{16})$.
-\bw{use $-\infty$ instead of $\ln 0$.
-is it okay?}\lz{lets just remove it ;)}
+Now we consider the formula $\dpriv{\epsilon_2}{\delta_2} (\X (Y\wedge
+\X Y))$ under the model in Figure~\ref{figure:2-dp-mc}. Applying the
+compositional theorem, one can choose $\epsilon_2=2\epsilon_1=0$ and
+$\delta_2=2\delta_1=1$. However, we can check easily that one gets
+better privacy parameter $(0,0.5)$.
+% or $(-\infty,\frac{9}{16})$. \bw{use $-\infty$ instead of $\ln 0$. is it okay?}\lz{lets just remove it ;)}
 
@@
 formula $\dpriv{\epsilon}{\delta}{\F \G \mathit{timeout}}$. Let us say
-two states are neighbors if the number of slave devices differs by at
-most one. Then the formula holds if the probability of having
+two states are neighbors if a slave device in a state is replaced by
+another device in the other state.
+Then the formula holds if the probability of having
 continuous time out (say, in the denial of service attack) is

### 2018-01-30 examples.tex
fix graphs

@@
 respectively; the states $Y$ and $N$ denote answers to the survey
-question. We have $AP=\{Y,N\}$, $L(Y)=\{Y\}$, $L(N)=\{N\}$, and $L(+)=L(-)=\emptyset$. States $+$ and $-$ are neighbors. 
+question. For each state $s$, we associate an atomic predicate
+$\mathit{at}_s$ (not shown in the figure).
+States $+$ and $-$ are neighbors. 
 The random noise boosts the probability of answering
@@
 diagnoses. Inferences on individual diagnosis can be plausibly denied.
-{In general, we omit a special state $\dagger$ with $L(\dagger)=\emptyset$, and all omitted transitions in the picture lead to this state.}
+In general, missing transitions (such as those from $Y$ and $N$) lead
+to a special state $\dagger$ with a self-loop. We omit the transitions
+and the state $\dagger$ for clarity.
 \end{example}
@@
 
-The \emph{truncated $\alpha$-geometric mechanism over $\underline{n}$}
+The \emph{truncated $\alpha$-geometric mechanism over $\{ 0, 1,
+  \ldots, n \}$}
 outputs $f (x) + Z$ where $Z$ is a random variable with the following
@@
 Note the range of the truncated $\alpha$-geometric mechanism is
-$\underline{n}$. The truncated $\alpha$-geometric mechanism is
+$\{ 0, 1, \ldots, n \}$. The truncated $\alpha$-geometric mechanism is
 again oblivious; it is also $(- {\Delta (f)} \ln \alpha, 0)$-differentially
 private for any query $f$ with sensitivity $\Delta (f)$.
-The truncated $1/2$-geometric mechanism over $\underline{5}$ is
+The truncated $1/2$-geometric mechanism over $\{ 0, 1, \ldots, 5 \}$ is
 given in Figure~\ref{figure:geometric-mechanism-table}.
@@
 denote the input value $k$ and output value $l$ respectively. Define
-the state set $S = \{ in_k, out_k : k \in \underline{n} \}$.
+the state set $S = \{ in_k, out_k : k \in \{ 0, 1, \ldots, n \} \}$.
 The probability transition $\wp (in_k, out_l)$ is the probability of the
@@
 the Markov chain of the truncated
-$1/2$-geometric mechanism over $\underline{5}$. Further, assume
-$L(out_i) = \{ out_i \}$ and $\Delta (f) = 1$. Then $in_k$ and $in_l$ are
+$1/2$-geometric mechanism over $\{ 0, 1, \ldots, 5 \}$. Assume
+$\Delta (f) = 1$. Then $in_k$ and $in_l$ are
 neighbors iff $| k - l | \leq 1$.
@@
 We consider the database set $D = \{0,1,2,3,4,5\}$, a threshold $t \in \{0,1,2,3,4,5\}$, queries $\{f| f $ with sensitivity $\Delta (f) = 1 \}$.
-In order to protect privacy, a truncated $1/2-$geometric mechanisms over $\underline{5}$ is applied to each query and the threshold, and what we observe is the first time that a query result is greater than or equal to the threshold after interference.
+In order to protect privacy, a truncated $1/2-$geometric mechanisms over $\{ 0, 1, \ldots, 5 \}$ is applied to each query and the threshold, and what we observe is the first time that a query result is greater than or equal to the threshold after interference.
 We are interested in the differential privacy property of the data in $D$, assuming that they are all neighbours.

### 2018-01-30 semantics.tex
remove \top from semantics

@@
 \begin{eqnarray*}
-  K, \neighbor{S}, s \models \top\\
+%  K, \neighbor{S}, s \models \top\\
   K, \neighbor{S}, s \models p
@@
   & &  \delta \textmd{ where } p(x) = \Pr [\{
-      \pi : K, \neighbor{S}, \pi \models \phi \textmd{ with}
+      \pi : K, \neighbor{S}, \pi \models \phi \textmd{ with }
       \pi_0 = x \}]

### 2018-01-30 examples.tex
change model states and labels

@@
       distance=2cm,node/.style={circle,draw}]
-      \node[node] (i0) at (0,  1) { $in_0$ };
+      \node[node] (i0) at (0,  1) { $s_0$ };
+      \node at (0, 1.5) { $\mathit{in}_0$ };
       \node at (0, 0) { $\vdots$ };
       \node at (.8, -.8) { $\vdots$ };
-      \node[node] (i5) at (0, -1) { $in_5$ };
+      \node[node] (i5) at (0, -1) { $s_5$ };
+      \node at (0, -.5) { $\mathit{in}_5$ };
 
-      \node[node] (o0) at (4,  2.5) { $out_0$ };
-      \node[node] (o1) at (4,  1.5) { $out_1$ };
-      \node[node] (o2) at (4,  0.5) { $out_2$ };
-      \node[node] (o3) at (4, -0.5) { $out_3$ };
-      \node[node] (o4) at (4, -1.5) { $out_4$ };
-      \node[node] (o5) at (4, -2.5) { $out_5$ };
+      \node[node] (o0) at (4,  2.5) { $t_0$ };
+      \node at (4, 3) { $\mathit{out}_0$ };
+      \node[node] (o1) at (4,  1.5) { $t_1$ };
+      \node at (4, 2) { $\mathit{out}_1$ };
+      \node[node] (o2) at (4,  0.5) { $t_2$ };
+      \node at (4, 1) { $\mathit{out}_2$ };
+      \node[node] (o3) at (4, -0.5) { $t_3$ };
+      \node at (4, 0) { $\mathit{out}_3$ };
+      \node[node] (o4) at (4, -1.5) { $t_4$ };
+      \node at (4, -1) { $\mathit{out}_4$ };
+      \node[node] (o5) at (4, -2.5) { $t_5$ };
+      \node at (4, -2) { $\mathit{out}_5$ };
 
@@
 \begin{example}
-Let the state $in_k$ and $out_l$
+Let the state $s_k$ and $t_l$
 denote the input value $k$ and output value $l$ respectively. Define
-the state set $S = \{ in_k, out_k : k \in \{ 0, 1, \ldots, n \} \}$.
-The probability transition $\wp (in_k, out_l)$ is the probability of the
+the state set $S = \{ s_k, t_k : k \in \{ 0, 1, \ldots, n \} \}$.
+The probability transition $\wp (s_k, t_l)$ is the probability of the
 output value $l$ on the input value $k$ as defined in the
-mechanism. Figure~\ref{figure:geometric-mechanism-markov-chain} gives
+mechanism. Moreover, we have $\mathit{in}_k \in L (s_k)$ and
+$\mathit{out}_k \in L (t_k)$ for $k \in \{ 0, 1, \ldots, n \}$.
+If $\Delta (f) = 1$, $s_k$ and $s_l$ are
+neighbors iff $| k - l | \leq 1$.
+Figure~\ref{figure:geometric-mechanism-markov-chain} gives
 the Markov chain of the truncated
-$1/2$-geometric mechanism over $\{ 0, 1, \ldots, 5 \}$. Assume
-$\Delta (f) = 1$. Then $in_k$ and $in_l$ are
-neighbors iff $| k - l | \leq 1$.
+$1/2$-geometric mechanism over $\{ 0, 1, \ldots, 5 \}$.
 \end{example}

### 2018-01-30 mdp.tex
proof adpated

@@
   \item $s_0\in S$ is the initial state,
-  \item $F\subseteq S$ is a set of accepting states.
+  \item $B\subseteq S$ is a set of accepting states.
 \end{itemize}
@@
 Thus, $M(\eta)(s,s')$ denotes the probability of going from $s$ to $s'$ after reading $\eta$.
-Let $v_F$ denote the characteristic row vector for the set $F$, and $v_{s_0}$ denote the characteristic row vector for the set $\{s_0\}$.
-Then, the accepting probably of $\eta$ by $\mathcal{A}$ is defined as $v_{s_0}\cdot M(\eta)\cdot (v_F)^c$ where $(v_F)^c$ denotes the transpose of $v_F$.
-The following \emph{emptiness problem} is know to be undecidable~\cite{Paz71}
+Let $v_B$ denote the characteristic row vector for the set $F$, and $v_{s_0}$ denote the characteristic row vector for the set $\{s_0\}$.
+Then, the accepting probably of $\eta$ by $\mathcal{A}$ is defined as $v_{s_0}\cdot M(\eta)\cdot (v_B)^c$ where $(v_B)^c$ denotes the transpose of $v_B$.
+The following \emph{emptiness problem} is know to be undecidable~\cite{Paz71}:
 %~\cite{Worrell08}:
 
-\paragraph{Emptiness problem:} Given a probabilistic automaton $\mathcal{A}= (S,\Sigma,M,s_0,F)$, whether there exists $\eta\in\Sigma^*$ such that $v_{s_0}\cdot M(\eta)\cdot (v_F)^c > 0$?
+\paragraph{Emptiness problem:} Given a probabilistic automaton $\mathcal{A}= (S,\Sigma,M,s_0,B)$, whether there exists $\eta\in\Sigma^*$ such that $v_{s_0}\cdot M(\eta)\cdot (v_B)^c > 0$?
 
 Now we establish the proof by reducing the threshold problem to our $\dpCTLstar$ model checking problem.
-Given the probabilistic automaton  $\mathcal{A}= (S,\Sigma,M,s_0,F)$, assume we have a \emph{primed} copy $\mathcal{A}'= (S',\Sigma,M',s_0',\emptyset)$.
+Given the probabilistic automaton  $\mathcal{A}= (S,\Sigma,M,s_0,B)$, assume we have a \emph{primed} copy $\mathcal{A}'= (S',\Sigma,M',s_0',\emptyset)$.
 
@@
 (S\cupdot S', \Sigma, \wp, L)$ where $\wp(s,a,t)$ equals to $M(s,a,t)$ if $s,t\in S$ and to $M'(s,a,t)$ if $s,t\in S'$.
-We define the neighbor relation $N_S:=\{(s_0,s_0')\}$ by relating states $s_0,s_0'$.
-The labelling function $L$ is defined by $L(s)=at_F$ if $s\in F$ and $L(s)=\empty$ otherwise.
+We define the neighbor relation $N_S:=\{(s_0,s_0'),(s_0',s_0)\}$ by relating states $s_0,s_0'$.
+The labelling function $L$ is defined by $L(s)=at_B$ if $s\in F$ and $L(s)=\empty$ otherwise.
 
-\lz{shall we use $\diamond$ for finally?}
-Now we consider the formula  $\Phi = \dpriv{1}{0} (\diamond at_F)$.
-For the reduction we prove $s_0\models \dpriv{1}{0} (\diamond at_F)$ iff for all $\eta\in\Sigma^*$ it holds $v_{s_0}\cdot M(\eta)\cdot (v_F)^c \leq 0$.
+Now we consider the formula  $\Phi = \dpriv{1}{0} (F at_B)$.
+For the reduction we prove $s_0\models \dpriv{1}{0} (F at_B)$ iff for all $\eta\in\Sigma^*$ it holds $v_{s_0}\cdot M(\eta)\cdot (v_B)^c \leq 0$.
 
-\lz{remember explain meaning of $\Sigma^\omega$}
-First we assume $s_0\models \dpriv{1}{0} (\diamond at_F)$. By $\dpCTLstar$ semantics we have that for all $\scheduler{S}\in\Sigma^\omega$,
-$\myPr{s_0}{M_{\scheduler{S}}}{\neighbor{S}}{\diamond at_F} \leq
+First we assume $s_0\models \dpriv{1}{0} (F at_B)$. By $\dpCTLstar$ semantics we have that for all $\scheduler{S}\in\Sigma^\omega$,
+$\myPr{s_0}{M_{\scheduler{S}}}{\neighbor{S}}{F at_B} \leq
   e \cdot
-  \myPr{s_0'}{M_{\scheduler{S}}}{\neighbor{S}}{\diamond at_F}$. Observe $\myPr{s_0'}{M_{\scheduler{S}}}{\neighbor{S}}{\diamond at_F}=0$ by construction, thus we have
-$\myPr{s_0}{M_{\scheduler{S}}}{\neighbor{S}}{\diamond at_F} \leq
- 0$. This implies  $v_{s_0}\cdot M(\eta)\cdot (v_F)^c \leq 0$ for all $\eta\in\Sigma^*$.
+  \myPr{s_0'}{M_{\scheduler{S}}}{\neighbor{S}}{F at_B}$. Observe $\myPr{s_0'}{M_{\scheduler{S}}}{\neighbor{S}}{F at_B}=0$ by construction, thus we have
+$\myPr{s_0}{M_{\scheduler{S}}}{\neighbor{S}}{F at_B} \leq
+ 0$. This implies  $v_{s_0}\cdot M(\eta)\cdot (v_B)^c \leq 0$ for all $\eta\in\Sigma^*$.
 
- For the other direction, assume that all $\eta\in\Sigma^*$ it holds $v_{s_0}\cdot M(\eta)\cdot (v_F)^c \leq 0$.
-For the sake of contradiction, assume that $s_0\not\models \dpriv{1}{0} (\diamond at_F)$.
-Since the relation $N_S$ contains only one pair, it implies that there exists a scheduler $\scheduler{S}\in\Sigma^\omega$ such that $\myPr{s_0}{M_{\scheduler{S}}}{\neighbor{S}}{\diamond at_F} >0$.
-It is then easy to construct a finite sequence $\eta\in\Sigma^*$ with  $v_{s_0}\cdot M(\eta)\cdot (v_F)^c > 0$, a contradiction.
+ For the other direction, assume that all $\eta\in\Sigma^*$ it holds $v_{s_0}\cdot M(\eta)\cdot (v_B)^c \leq 0$.
+For the sake of contradiction, assume that $s_0\not\models \dpriv{1}{0} (F at_B)$.
+Since the relation $N_S$ contains only one pair, it implies that there exists a scheduler $\scheduler{S}\in\Sigma^\omega$ such that $\myPr{s_0}{M_{\scheduler{S}}}{\neighbor{S}}{F at_B} >0$.
+It is then easy to construct a finite sequence $\eta\in\Sigma^*$ with  $v_{s_0}\cdot M(\eta)\cdot (v_B)^c > 0$, a contradiction.
 \end{proof}

### 2018-01-30 specifications.tex
proof adpated

@@
 The formula $\dpriv{\epsilon}{\delta}(F \top)$ is a reachability privacy: it states that related states should have similar probabilities of eventually reaching $\top$.
-Note neighbor states may reach the goal state $\top$ with
+Reachability privacy is often used in combination with other properties.
 
-For our threshold \lz{to do} model, we have the differential privacy property $\wedge_{i=1}^{c} \dpriv{\epsilon}{0}(\X^k \top) \wedge \dpriv{\epsilon}{0}(F \top)$.
-
-Finally, we the property
-$\wedge_{i\in N} \dpriv{\epsilon}{0}(\X^k \top) \wedge \dpriv{\epsilon}{0}(F \top)$ that is used to express the privacy for our threshold Markov chains.
+As an example, for our threshold \lz{to do} model, we consider the property $\wedge_{i=1}^{c} \dpriv{\epsilon}{0}(\X^k \top) \wedge \PJ{\le 10^{-6}}(\X^c F T) \wedge \dpriv{\epsilon}{0}(\X^c F \top)$.
+that is used to express the privacy for our threshold Markov chains.
 It is only a valid formula if we would allow infinite conjunctions.
 However, it is interesting to note the relation to language equivalence for probabilistic automata\lz{to be discussed}
+
+Later we shall illustrate how reachability privacy can be used to specify differential privacy property for MDPs.
 

### 2018-01-30 specifications.tex
fix graphs

@@
 \paragraph{Standard Differential Privacy.}
-The formula $\dpriv{\ln 3}{0} (\X \mathit{out}_Y)$ \lz{$\X Y$ looks a bit weired?} holds in state $s$ if for  $t$ with $sNt$, the probability of satisfying $\X Y$ from $s$ and $t$ satisfy:
-$\myPr{s}{K}{N}{\X Y}\le 3\myPr{t}{K}{N}{\X Y}$.
-The formula $\dpriv{\ln 3}{0} (\X Y)\wedge \dpriv{\ln 3}{0} (\X N)$ thus specifies the differential privacy property for states $+$ and $-$ in Example~\ref{exa:survey}.
+The formula $\dpriv{\ln 3}{0} (\X \mathit{out}_Y)$ holds in state $s$ if for  $t$ with $sNt$, the probability of satisfying $\X \mathit{out}_Y$ from $s$ and $t$ satisfy:
+$\myPr{s}{K}{N}{\X \mathit{out}_Y}\le 3\myPr{t}{K}{N}{\X \mathit{out}_Y}$.
+The formula $\dpriv{\ln 3}{0} (\X \mathit{out}_Y)\wedge \dpriv{\ln 3}{0} (\X \mathit{out}_N)$ thus specifies the differential privacy property for states $+$ and $-$ in Example~\ref{exa:survey}.
 
@@
 $\psi = \dpriv{\ln 2}{0} (\X out_0) \wedge \dpriv{\ln 2}{0} (\X out_1) \wedge
-\cdots \wedge \dpriv{\ln 2}{0} (\X out_5)$. If the state $in_k$ satisfies
+\cdots \wedge \dpriv{\ln 2}{0} (\X out_5)$. If the state $s_k$ satisfies
 $\psi$ for $k = 0, \ldots, 5$, then the $.5$-geometric mechanism is
@@
 To illustrate this property, we consider the extended survey example below where the curator allows two consequent queries.
-The corresponding model is depicted in Figure~\ref{figure:2-dp-mc}, where states $Y_1,Y_2$ are labelled by $Y$, and $N_1,N_2$ are labelled by $N$.
+The model is depicted in Figure~\ref{figure:2-dp-mc}.
+%where states $Y_1,Y_2$ are labelled by $Y$, and $N_1,N_2$ are labelled by $N$.
 
@@
     \node[node] (p) at ( -1.5,  0) { $+$ };
-    \node[node] (Y1) at ( 0, .8) { $Y_1$ };
-    \node[node] (N1) at ( 0, -.8) { $N_1$ };
-    \node[node] (Y2) at (1.5,  .8) { $Y_2$ };
-    \node[node] (N2) at ( 1.5, -.8) { $N_2$ };
+    \node[node] (Y1) at ( 0, .8) { $s_1$ };
+    \node at (0, 1.3) { $\mathit{out}_Y$ };
+    \node[node] (N1) at ( 0, -.8) { $t_1$ };
+    \node at (0, -.3) { $\mathit{out}_N$ };
+    \node[node] (Y2) at (1.5,  .8) { $s_2$ };
+    \node at (1.5, 1.3) { $\mathit{out}_Y$ };
+    \node[node] (N2) at ( 1.5, -.8) { $t_2$ };
+    \node at (1.5, -.3) { $\mathit{out}_N$ };
 
@@
     \node[node] (n) at ( -1.5,  0) { $-$ };
-    \node[node] (Y1) at ( 0, .8) { $Y_1$ };
-    \node[node] (N1) at ( 0, -.8) { $N_1$ };
-    \node[node] (Y2) at (1.5,  .8) { $Y_2$ };
-    \node[node] (N2) at ( 1.5, -.8) { $N_2$ };
+    \node[node] (Y1) at ( 0, .8) { $s_1$ };
+    \node at (0, 1.3) { $\mathit{out}_Y$ };
+    \node[node] (N1) at ( 0, -.8) { $t_1$ };
+    \node at (0, -.3) { $\mathit{out}_N$ };
+    \node[node] (Y2) at (1.5,  .8) { $s_2$ };
+    \node at (1.5, 1.3) { $\mathit{out}_Y$ };
+    \node[node] (N2) at ( 1.5, -.8) { $t_2$ };
+    \node at (1.5, -.3) { $\mathit{out}_N$ };
 
@@
 
-We consider the formula $\dpriv{\ln 9}{0} (\X (Y\wedge \X Y))$.
-A path satisfies $\X (Y\wedge \X Y)$ if the second state satisfies $Y$ and the third state satisfies $Y$ as well.
-We verify that this formula is satisfied  for state $+$ and $-$.
-Moreover, the bound $\epsilon=\ln 9$ is tight, since the probability of satisfying $\X (Y\wedge \X Y)$ from states $+$ and $-$ are $\frac{9}{16}$ and $\frac{1}{16}$, respectively.
-Finally, the formula $\wedge_{a_1,a_2}\dpriv{\ln 9}{0} (\X (a_1\wedge \X a_2))$ corresponds to the desired compositional property for the model, where $a_1,a_2$ range over all atomic propositions $\{Y,N\}$.
+We consider the formula $\dpriv{\ln 9}{0} (\X (\mathit{out}_Y\wedge \X \mathit{out}_Y))$.
+A path satisfies $\X (\mathit{out}_Y\wedge \X \mathit{out}_Y)$ if the second state satisfies $\mathit{out}_Y$ and the third state satisfies $\mathit{out}_Y$ as well.
+We verify that this formula is satisfied for states $+$ and $-$.
+Moreover, the bound $\epsilon=\ln 9$ is tight, since the probability of satisfying $\X (\mathit{out}_Y\wedge \X \mathit{out}_Y)$ from states $+$ and $-$ are $\frac{9}{16}$ and $\frac{1}{16}$, respectively.
+Finally, the formula $\wedge_{a_1,a_2}\dpriv{\ln 9}{0} (\X (a_1\wedge \X a_2))$ corresponds to the desired compositional property for the model, where $a_1,a_2$ range over all atomic propositions $\{\mathit{out}_Y, \mathit{out}_N\}$.
 
 We consider two other formulae for comparision:
-(a) Formula $\dpriv{\ln 3}{0} (\X \X Y)$.
+(a) Formula $\dpriv{\ln 3}{0} (\X \X \mathit{out}_Y)$.
 In this case we donot have privacy lost, even though we are querying twice: the reason is that the output of the first query is not utilized at all.
 It is easy to verify that it is indeed satisfied by $+$ and $-$.
-(b) Formula $\dpriv{\ln 3}{0} (\X (Y\wedge \dpriv{\ln 3}{0}(\X Y)))$.
-This is a nested $\dpCTL$ formula, where the inner state formula $\dpriv{\ln 3}{0}(\X Y)$ states the one step differential privacy.
+(b) Formula $\dpriv{\ln 3}{0} (\X (\mathit{out}_Y\wedge \dpriv{\ln 3}{0}(\X \mathit{out}_Y)))$.
+This is a nested $\dpCTL$ formula, where the inner state formula $\dpriv{\ln 3}{0}(\X \mathit{out}_Y)$ states the one step differential privacy.
 Observe the inner formula is satisfied by all neighbours.
@@
 One of the advantage of applying model checking is that we may get tighter bounds for compositional properties.
-Consider our survey example, and the property $\dpriv{0}{.5} (\X Y)$.
+Consider our survey example, and the property $\dpriv{0}{.5} (\X \mathit{out}_Y)$.
 Obviously, it holds in states $+$ and $-$.
@@
 
-Now we consider the formula $\dpriv{\epsilon_2}{\delta_2} (\X (Y\wedge
-\X Y))$ under the model in Figure~\ref{figure:2-dp-mc}. Applying the
+Now we consider the formula $\dpriv{\epsilon_2}{\delta_2} (\X 
+(\mathit{out}_Y\wedge \X \mathit{out}_Y))$ under the model in 
+Figure~\ref{figure:2-dp-mc}. Applying the
 compositional theorem, one can choose $\epsilon_2=2\epsilon_1=0$ and

### 2018-01-30 examples.tex
use out_Y and out_N

@@
       \node[node] (n) at ( 0, -.5) { $-$ };
-      \node[node] (Y) at (-1.5,  0) { $Y$ };
-      \node[node] (N) at ( 1.5,  0) { $N$ };
+      \node[node] (Y) at (-1.5,  0) { $s$ };
+      \node at (-1.5, .5) { $\mathit{out}_Y$ };
+      \node[node] (N) at ( 1.5,  0) { $t$ };
+      \node at ( 1.5, .5) { $\mathit{out}_N$ };
 
@@
 $+$ and $-$ denote the diagnoses are positive or negative
-respectively; the states $Y$ and $N$ denote answers to the survey
-question. For each state $s$, we associate an atomic predicate
-$\mathit{at}_s$ (not shown in the figure).
+respectively; the states $s$ and $t$ denote answers to the survey
+question and hence let $\mathit{out}_Y \in L (s)$ and $\mathit{out}_N
+\in L (t)$.
+For each state $x$, we moreover associate an atomic predicate
+$\mathit{at}_x$ (not shown in the figure). 
 States $+$ and $-$ are neighbors. 

### 2018-01-30 main.tex
remove author list
added conclusion

@@
 
+\newif\ifpublic
+%\publictrue
+
 \title{Model Checking Differentially Private Properties}
+\ifpublic
 \author{
@@
 }
+\else
+\author{\vspace*{-1.2cm}}
+\institute{\vspace*{-1.2cm}\ }
+\fi
 
@@
 \label{section:conclusions}
+We have introduced $\dpCTLstar$ to reason about properties in differential privacy, and have investigated the model checking problem.
+For Markov chains, the model checking problem has polynomial complexity.
+The general MDP model checking problem is undecidable.
+We have discussed that some special cases are decidable.
+An interesting work is to identify more
+decidable subclass. As an example, consider the formula $\bigwedge_{k \in \bbfN}
+\dpriv{\epsilon}{\delta} \X^k \mathit{at}_{00}$ which was used in specifying the pan privacy.
+For the case $\epsilon=\delta=0$, it reduces to a language equivalence problem for probabilistic automata.
+It is interesting to characterise other cases as well.
+ Another interesting line of further works is to consider continuous perturbation (such as Laplace distribution used in \cite{DR:14:AFDP}) -- we would need Markov models with continuous state space for the modelling.
 

### 2018-01-30 applications.tex
proof adpated
added conclusion
add labels

@@
 variant of the mechanisms by Markov decision processes. We will also
-specify pan privacy in our model using $\dpCTL$.
+specify pan privacy in our model using $\dpCTLstar$.
 
@@
 Similar to Algorithm~\ref{algorithm:frequency-estimator}, our Markov
-decision process has internal states $00$, $01$, $10$, $11$ for the
-two bits $f_a$ and $f_b$ of observed event types
-(Figure~\ref{figure:adjacent-inputs}). The internal state $01$
-represents $(f_a, f_b) = (0, 1)$, for instance. From the
-initialization state $I$, the Markov decision process moves to one of
+decision process has internal states $s_{00}$, $s_{01}$, $s_{10}$, 
+$s_{11}$ for the two bits $f_a$ and $f_b$ of observed event types
+(Figure~\ref{figure:adjacent-inputs}). The internal state $s_{01}$
+represents $(f_a, f_b) = (0, 1)$, for instance. Moreover, let $xy \in
+L (s_{xy})$ for $x, y \in \{ 0, 1 \}$. From the
+initialization state $s$, the Markov decision process moves to one of
 the four internal states randomly with the uniform
-distribution. Consider the internal state $01$. If the actions $a[a]$
+distribution. Consider the internal state $s_{01}$. If the actions $a[a]$
 or $a[b]$ are chosen, they denote that the event type $a$ is currently
-observed. The Markov decision process moves to the internal state $01$
-and $11$ with probability $.375 = \frac{3}{8}$ and $.625 =
+observed. The Markov decision process moves to the internal state $s_{01}$
+and $s_{11}$ with probability $.375 = \frac{3}{8}$ and $.625 =
 \frac{5}{8}$ respectively. Other transitions are defined similarly.
 
-On the other hand, the initialization state $\underline{I}$
+On the other hand, the initialization state $t$
 corresponds to the computation on neighbors. The Markov decision
-process moves to the interal states $\underline{00}$,
-$\underline{01}$, $\underline{10}$, $\underline{11}$ randomly with
-the uniform distribution. Consider, for example, the internal state
-$\underline{10}$. If the action $a[a]$ or $b[a]$ are chosen, they
+process moves to the interal states $t_{00}$,
+$t_{01}$, $t_{10}$, $t_{11}$ randomly with
+the uniform distribution. Similarly, $xy \in L (t_{xy})$ for $x, y \in
+\{ 0, 1 \}$.
+Consider, for example, the internal state
+$t_{10}$. If the action $a[a]$ or $b[a]$ are chosen, they
 denote that the current event type is replaced with the eveny type
 $a$. The Markov decision process moves to the internal state
-$\underline{00}$ and $\underline{10}$ with probability $.375 =
+$t_{00}$ and $t_{10}$ with probability $.375 =
 \frac{3}{8}$ and $.625 = \frac{5}{8}$ respectively. At an internal
@@
 Figure~\ref{figure:geometric-mechanism}). For any action sequence,
-observe that paths from the state $I$ and those from $\underline{I}$
+observe that paths from the state $s$ and those from $t$
 formalize the computation of Algorithm~\ref{algorithm:frequency-estimator}
 on an observation and its neighbor represented by the sequence. We
-hence say the states $I$ and $\underline{I}$ are neighbors in the
+hence say the states $s$ and $t$ are neighbors in the
 Markov decision process (Figure~\ref{figure:frequency-estimator}).
@@
       distance=2cm,node/.style={circle,draw}]
-      \node[node] (I) at  ( 0,  0) { $I$ };
-      \node[node] (00) at (-1.5,  1.5) { $00$ };
-      \node[node] (01) at (-1.5, -1.5) { $01$ };
-      \node[node] (10) at ( 1.5,  1.5) { $10$ };
-      \node[node] (11) at ( 1.5, -1.5) { $11$ };
+      \node[node] (I) at  ( 0,  0) { $s$ };
+      \node[node] (00) at (-1.5,  1.5) { $s_{00}$ };
+      \node at (-1.9, 1.9) { $00$ };
+      \node[node] (01) at (-1.5, -1.5) { $s_{01}$ };
+      \node at (-1.9, -1.9) { $01$ };
+      \node[node] (10) at ( 1.5,  1.5) { $s_{10}$ };
+      \node at ( 1.9,  1.9) { $10$ };
+      \node[node] (11) at ( 1.5, -1.5) { $s_{11}$ };
+      \node at ( 1.9, -1.9) { $11$ };
 
@@
 
-      \node[node] (I') at  ( 5.5,    0) { $\underline{I}$ };
-      \node[node] (00') at (   4,  1.5) { $\underline{00}$ };
-      \node[node] (01') at (   4, -1.5) { $\underline{01}$ };
-      \node[node] (10') at (   7,  1.5) { $\underline{10}$ };
-      \node[node] (11') at (   7, -1.5) { $\underline{11}$ };
+      \node[node] (I') at  ( 5.5,    0) { $t$ };
+      \node[node] (00') at (   4,  1.5) { $t_{00}$ };
+      \node at ( 3.6,  1.9) { $00$ };
+      \node[node] (01') at (   4, -1.5) { $t_{01}$ };
+      \node at ( 3.6, -1.9) { $01$ };
+      \node[node] (10') at (   7,  1.5) { $t_{10}$ };
+      \node at ( 7.4,  1.9) { $10$ };
+      \node[node] (11') at (   7, -1.5) { $t_{11}$ };
+      \node at ( 7.4, -1.9) { $11$ };
 
@@
 
-      \node[node] (00) at (-4.5, 1.5) { $00$ };
-      \node[node] (01) at (-1.5, 2) { $01$ };
-      \node[node] (10) at ( 1.5, 2) { $10$ };
-      \node[node] (11) at ( 4.5, 1.5) { $11$ };
+      \node[node] (00) at (-4.5, 1.5) { $s_{00}$ };
+      \node at (-4.9,  1.9) { $00$ };
+      \node[node] (01) at (-1.5, 2) { $s_{01}$ };
+      \node at (-1.9,  2.4) { $01$ };
+      \node[node] (10) at ( 1.5, 2) { $s_{10}$ };
+      \node at ( 1.9,  2.4) { $10$ };
+      \node[node] (11) at ( 4.5, 1.5) { $s_{11}$ };
+      \node at ( 4.9,  1.9) { $11$ };
 
-      \node[node] (00') at (-4.5, -1.5) { $\underline{00}$ };
-      \node[node] (01') at (-1.5, -2) { $\underline{01}$ };
-      \node[node] (10') at ( 1.5, -2) { $\underline{10}$ };
-      \node[node] (11') at ( 4.5, -1.5) { $\underline{11}$ };
+      \node[node] (00') at (-4.5, -1.5) { $t_{00}$ };
+      \node at (-4.9, -1.9) { $00$ };
+      \node[node] (01') at (-1.5, -2) { $t_{01}$ };
+      \node at (-1.9, -2.4) { $01$ };
+      \node[node] (10') at ( 1.5, -2) { $t_{10}$ };
+      \node at ( 1.9, -2.4) { $10$ };
+      \node[node] (11') at ( 4.5, -1.5) { $t_{11}$ };
+      \node at ( 4.9, -1.9) { $11$ };
 
@@
 \[
-\Psi = \dpriv{\ln 4}{0} \F \mathit{at}_0 \wedge \dpriv{\ln 4}{0} \F
-\mathit{at}_1 \wedge \dpriv{\ln 4}{0} \F \mathit{at}_2
+\Psi = \dpriv{\ln 4}{0} (\F \mathit{at}_0) \wedge \dpriv{\ln 4}{0} (\F
+\mathit{at}_1) \wedge \dpriv{\ln 4}{0} (\F \mathit{at}_2)p
 \]
@@
 \bigwedge_{k \in \bbfN}
-\dpriv{\ln 4}{0} \X^k \mathit{at}_{00} \wedge
-\dpriv{\ln 4}{0} \X^k \mathit{at}_{01} \wedge
-\dpriv{\ln 4}{0} \X^k \mathit{at}_{10} \wedge
-\dpriv{\ln 4}{0} \X^k \mathit{at}_{11}.
+\dpriv{\ln 4}{0} \X^k {00} \wedge
+\dpriv{\ln 4}{0} \X^k {01} \wedge
+\dpriv{\ln 4}{0} \X^k {10} \wedge
+\dpriv{\ln 4}{0} \X^k {11}.
 \]

### 2018-01-30 mdp.tex
add labels
proof appendix

@@
   \centering
-    \resizebox{.4\columnwidth}{!}{
+    \resizebox{.45\columnwidth}{!}{
     \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
@@
       \node[node] (q) at ( 0, -1.5) { $-$ };
-      \node[node] (F) at (-1.5,  0) { $Y$ };
-      \node[node] (T) at ( 1.5,  0) { $N$ };
+      \node[node] (F) at (-1.5,  0) { $s$ };
+      \node at (-2, .5) { $\mathit{out}_Y$ };
+      \node[node] (T) at ( 1.5,  0) { $t$ };
+      \node at ( 2, .5) { $\mathit{out}_N$ };
 
@@
 \end{theorem}
-\begin{proof}
-The proof follows by a reduction from the \emph{emptiness} problem for probabilistic automata.
-A \emph{probabilistic automaton}~\cite{Rabin63} is a tuple $\mathcal{A}= (S,\Sigma,M,s_0,F)$ where
-\begin{itemize}
-  \item $S$ is a finite set of states,
-  \item $\Sigma$ is the finite set of input alphabet,
-  \item $M: S\times\Sigma\times S \to [0,1]$ such that $\sum_{t\in S} M(s,\alpha,t)=1$ for all $s\in S$ and $\alpha\in \Sigma$,
-  \item $s_0\in S$ is the initial state,
-  \item $B\subseteq S$ is a set of accepting states.
-\end{itemize}
-
-Each input alphabet $\alpha$ induces a stochastic matrix $M(\alpha)$ in the obvious way.
-Let $\lambda$ denote the empty string.
-For $\eta\in\Sigma^*$ we define $M(\eta)$ inductively by: $M(\lambda)$ is the identity matrix, $M(x\eta')=M(x)M(\eta')$.
-Thus, $M(\eta)(s,s')$ denotes the probability of going from $s$ to $s'$ after reading $\eta$.
-Let $v_B$ denote the characteristic row vector for the set $F$, and $v_{s_0}$ denote the characteristic row vector for the set $\{s_0\}$.
-Then, the accepting probably of $\eta$ by $\mathcal{A}$ is defined as $v_{s_0}\cdot M(\eta)\cdot (v_B)^c$ where $(v_B)^c$ denotes the transpose of $v_B$.
-The following \emph{emptiness problem} is know to be undecidable~\cite{Paz71}:
-%~\cite{Worrell08}:
-
-\paragraph{Emptiness problem:} Given a probabilistic automaton $\mathcal{A}= (S,\Sigma,M,s_0,B)$, whether there exists $\eta\in\Sigma^*$ such that $v_{s_0}\cdot M(\eta)\cdot (v_B)^c > 0$?
-
-Now we establish the proof by reducing the threshold problem to our $\dpCTLstar$ model checking problem.
-Given the probabilistic automaton  $\mathcal{A}= (S,\Sigma,M,s_0,B)$, assume we have a \emph{primed} copy $\mathcal{A}'= (S',\Sigma,M',s_0',\emptyset)$.
-
-Now we construct our MDP  $M =
-(S\cupdot S', \Sigma, \wp, L)$ where $\wp(s,a,t)$ equals to $M(s,a,t)$ if $s,t\in S$ and to $M'(s,a,t)$ if $s,t\in S'$.
-We define the neighbor relation $N_S:=\{(s_0,s_0'),(s_0',s_0)\}$ by relating states $s_0,s_0'$.
-The labelling function $L$ is defined by $L(s)=at_B$ if $s\in F$ and $L(s)=\empty$ otherwise.
-
-Now we consider the formula  $\Phi = \dpriv{1}{0} (F at_B)$.
-For the reduction we prove $s_0\models \dpriv{1}{0} (F at_B)$ iff for all $\eta\in\Sigma^*$ it holds $v_{s_0}\cdot M(\eta)\cdot (v_B)^c \leq 0$.
-
-First we assume $s_0\models \dpriv{1}{0} (F at_B)$. By $\dpCTLstar$ semantics we have that for all $\scheduler{S}\in\Sigma^\omega$,
-$\myPr{s_0}{M_{\scheduler{S}}}{\neighbor{S}}{F at_B} \leq
-  e \cdot
-  \myPr{s_0'}{M_{\scheduler{S}}}{\neighbor{S}}{F at_B}$. Observe $\myPr{s_0'}{M_{\scheduler{S}}}{\neighbor{S}}{F at_B}=0$ by construction, thus we have
-$\myPr{s_0}{M_{\scheduler{S}}}{\neighbor{S}}{F at_B} \leq
- 0$. This implies  $v_{s_0}\cdot M(\eta)\cdot (v_B)^c \leq 0$ for all $\eta\in\Sigma^*$.
-
- For the other direction, assume that all $\eta\in\Sigma^*$ it holds $v_{s_0}\cdot M(\eta)\cdot (v_B)^c \leq 0$.
-For the sake of contradiction, assume that $s_0\not\models \dpriv{1}{0} (F at_B)$.
-Since the relation $N_S$ contains only one pair, it implies that there exists a scheduler $\scheduler{S}\in\Sigma^\omega$ such that $\myPr{s_0}{M_{\scheduler{S}}}{\neighbor{S}}{F at_B} >0$.
-It is then easy to construct a finite sequence $\eta\in\Sigma^*$ with  $v_{s_0}\cdot M(\eta)\cdot (v_B)^c > 0$, a contradiction.
-\end{proof}
 

### 2018-01-30 proof.tex
proof appendix

\begin{proof}
The proof follows by a reduction from the \emph{emptiness} problem for probabilistic automata.
A \emph{probabilistic automaton}~\cite{Rabin63} is a tuple $\mathcal{A}= (S,\Sigma,M,s_0,F)$ where
\begin{itemize}
  \item $S$ is a finite set of states,
  \item $\Sigma$ is the finite set of input alphabet,
  \item $M: S\times\Sigma\times S \to [0,1]$ such that $\sum_{t\in S} M(s,\alpha,t)=1$ for all $s\in S$ and $\alpha\in \Sigma$,
  \item $s_0\in S$ is the initial state,
  \item $B\subseteq S$ is a set of accepting states.
\end{itemize}

Each input alphabet $\alpha$ induces a stochastic matrix $M(\alpha)$ in the obvious way. Let $\lambda$ denote the empty string. For $\eta\in\Sigma^*$ we define $M(\eta)$ inductively by: $M(\lambda)$ is the identity matrix, $M(x\eta')=M(x)M(\eta')$. Thus, $M(\eta)(s,s')$ denotes the probability of going from $s$ to $s'$ after reading $\eta$. Let $v_B$ denote the characteristic row vector for the set $F$, and $v_{s_0}$ denote the characteristic row vector for the set $\{s_0\}$. Then, the accepting probably of $\eta$ by $\mathcal{A}$ is defined as $v_{s_0}\cdot M(\eta)\cdot (v_B)^c$ where $(v_B)^c$ denotes the transpose of $v_B$. The following \emph{emptiness problem} is know to be undecidable~\cite{Paz71}:
%~\cite{Worrell08}:

\paragraph{Emptiness problem:} Given a probabilistic automaton $\mathcal{A}= (S,\Sigma,M,s_0,B)$, whether there exists $\eta\in\Sigma^*$ such that $v_{s_0}\cdot M(\eta)\cdot (v_B)^c > 0$?

Now we establish the proof by reducing the threshold problem to our $\dpCTLstar$ model checking problem. Given the probabilistic automaton  $\mathcal{A}= (S,\Sigma,M,s_0,B)$, assume we have a \emph{primed} copy $\mathcal{A}'= (S',\Sigma,M',s_0',\emptyset)$.

Now we construct our MDP  $M =
(S\cupdot S', \Sigma, \wp, L)$ where $\wp(s,a,t)$ equals to $M(s,a,t)$ if $s,t\in S$ and to $M'(s,a,t)$ if $s,t\in S'$. We define the neighbor relation $N_S:=\{(s_0,s_0'),(s_0',s_0)\}$ by relating states $s_0,s_0'$. The labelling function $L$ is defined by $L(s)=at_B$ if $s\in F$ and $L(s)=\empty$ otherwise.

Now we consider the formula  $\Phi = \dpriv{1}{0} (F at_B)$. For the reduction we prove $s_0\models \dpriv{1}{0} (F at_B)$ iff for all $\eta\in\Sigma^*$ it holds $v_{s_0}\cdot M(\eta)\cdot (v_B)^c \leq 0$.

First we assume $s_0\models \dpriv{1}{0} (F at_B)$. By $\dpCTLstar$ semantics we have that for all $\scheduler{S}\in\Sigma^\omega$,
$\myPr{s_0}{M_{\scheduler{S}}}{\neighbor{S}}{F at_B} \leq
  e \cdot
  \myPr{s_0'}{M_{\scheduler{S}}}{\neighbor{S}}{F at_B}$. Observe $\myPr{s_0'}{M_{\scheduler{S}}}{\neighbor{S}}{F at_B}=0$ by construction, thus we have
$\myPr{s_0}{M_{\scheduler{S}}}{\neighbor{S}}{F at_B} \leq
 0$. This implies  $v_{s_0}\cdot M(\eta)\cdot (v_B)^c \leq 0$ for all $\eta\in\Sigma^*$.

 For the other direction, assume that all $\eta\in\Sigma^*$ it holds $v_{s_0}\cdot M(\eta)\cdot (v_B)^c \leq 0$. For the sake of contradiction, assume that $s_0\not\models \dpriv{1}{0} (F at_B)$. Since the relation $N_S$ contains only one pair, it implies that there exists a scheduler $\scheduler{S}\in\Sigma^\omega$ such that $\myPr{s_0}{M_{\scheduler{S}}}{\neighbor{S}}{F at_B} >0$. It is then easy to construct a finite sequence $\eta\in\Sigma^*$ with  $v_{s_0}\cdot M(\eta)\cdot (v_B)^c > 0$, a contradiction.
\end{proof}

### 2018-01-30 examples.tex
modify the algorithm

@@
 
+\begin{algorithm}
+	\begin{algorithmic}[1]
+		\Function{OnlineModel}{$d$, $t$}
+		\Match{the input threshold $t$}
+		\Comment{Assignments of $t'$ by $\frac{1}{4}$-geometric mechanism}
+		\lCase{$0$}{$t' \leftarrow 0, 1,2,3,4,5$ with probability
+			$\frac{4}{5}$,$\frac{3}{20}$,
+			$\frac{3}{80}$,$\frac{3}{320}$,
+			$\frac{3}{1280}$,$\frac{1}{1280}$ respectively}
+		\lCase{$1$}{$t' \leftarrow 0, 1,2,3,4,5$ with probability
+			$\frac{1}{5}$,$\frac{3}{5}$,
+			$\frac{3}{20}$,$\frac{3}{80}$,
+			$\frac{3}{320}$,$\frac{1}{320}$ respectively}
+		\lCase{$2$}{$t' \leftarrow 0, 1,2,3,4,5$ with probability
+			$\frac{1}{20}$,$\frac{3}{20}$,
+			$\frac{3}{5}$,$\frac{3}{20}$,
+			$\frac{3}{80}$,$\frac{1}{80}$ respectively}
+		\lCase{$3$}{$t' \leftarrow 0, 1,2,3,4,5$ with probability
+			$\frac{1}{80}$,$\frac{3}{80}$,
+			$\frac{3}{20}$,$\frac{3}{5}$,
+			$\frac{3}{20}$,$\frac{1}{20}$ respectively}
+		\lCase{$4$}{$t' \leftarrow 0, 1,2,3,4,5$ with probability
+			$\frac{1}{320}$,$\frac{3}{320}$,
+			$\frac{3}{80}$,$\frac{3}{20}$,
+			$\frac{3}{5}$,$\frac{1}{5}$ respectively}
+		\lCase{$5$}{$t' \leftarrow 0, 1,2,3,4,5$ with probability
+			$\frac{1}{1280}$,$\frac{3}{1280}$,
+			$\frac{3}{320}$,$\frac{3}{80}$,
+			$\frac{3}{20}$,$\frac{4}{5}$ respectively}
+		\EndMatch
+		\For{each query $f_i$}
+		\Match{the input databese $d$}
+		\Comment{Assignments of $f_{i}'$ by $\frac{1}{2}$-geometric mechanism}
+		\lCase{$0$}{$f_{i}' \leftarrow 0, 1,2,3,4,5$ with probability
+			$\frac{2}{3}$,$\frac{1}{6}$,
+			$\frac{1}{12}$,$\frac{1}{24}$,
+			$\frac{1}{48}$,$\frac{1}{48}$ respectively}
+		\lCase{$1$}{$f_{i}' \leftarrow 0, 1,2,3,4,5$ with probability
+			$\frac{1}{3}$,$\frac{1}{3}$,
+			$\frac{1}{6}$,$\frac{1}{12}$,
+			$\frac{1}{24}$,$\frac{1}{24}$ respectively}
+		\lCase{$2$}{$f_{i}' \leftarrow 0, 1,2,3,4,5$ with probability
+			$\frac{1}{6}$,$\frac{1}{6}$,
+			$\frac{1}{3}$,$\frac{1}{6}$,
+			$\frac{1}{12}$,$\frac{1}{12}$ respectively}
+		\lCase{$3$}{$f_{i}' \leftarrow 0, 1,2,3,4,5$ with probability
+			$\frac{1}{12}$,$\frac{1}{12}$,
+			$\frac{1}{6}$,$\frac{1}{3}$,
+			$\frac{1}{6}$,$\frac{1}{6}$ respectively}
+		\lCase{$4$}{$f_{i}' \leftarrow 0, 1,2,3,4,5$ with probability
+			$\frac{1}{24}$,$\frac{1}{24}$,
+			$\frac{1}{12}$,$\frac{1}{6}$,
+			$\frac{1}{3}$,$\frac{1}{3}$ respectively}
+		\lCase{$5$}{$f_{i}' \leftarrow 0, 1,2,3,4,5$ with probability
+			$\frac{1}{48}$,$\frac{1}{48}$,
+			$\frac{1}{24}$,$\frac{1}{12}$,
+			$\frac{1}{6}$,$\frac{2}{3}$ respectively}
+		\EndMatch
+		\If{$f_{i}'>t'$}
+		\State{$a_i=\top$}
+		\State{\Return}
+		\Else
+		\State{$a_i=\bot$}
+		\EndIf
+		\EndFor
+		\EndFunction
+
+	\end{algorithmic}
+	\caption{Input is a private database $d \in \{0,1,2,3,4,5\}$, an adaptively chosen stream of sensitivity 1 queries $f_1$, .
+.
+., a threshold $t \in \{0,1,2,3,4,5\}$.Output is a stream of answers $a_1$,...}
+	\label{algorithm:online-model}
+\end{algorithm}
+
 \subsection{Markov Chains for Differentially Private Mechanisms}
@@
 \paragraph{Online Model.}
-Below we describe an online model adapted from \cite{}: it is used in the scenario where a large amount of queries occur, but we only care for the queries which
+Below we describe an online model adapted from~\cite{DR:14:AFDP}: it is used in the scenario where a large amount of queries occur, but we only care for the queries which
 are above a given threshold.
 Other queries below the threshold are considered "irrelevent" and might be discarded.
-In \cite{}, since the model is set for continuous database, the Laplace mechanism is applied to both queries and the threshold.
+In ~\cite{DR:14:AFDP}, since the model is set for continuous database, the Laplace mechanism is applied to both queries and the threshold.
 In our model, the database is discrete and instead the geometric mechanism is applied.
@@
 We consider the database set $D = \{0,1,2,3,4,5\}$, a threshold $t \in \{0,1,2,3,4,5\}$, queries $\{f| f $ with sensitivity $\Delta (f) = 1 \}$.
-In order to protect privacy, a truncated $1/2-$geometric mechanisms over $\{ 0, 1, \ldots, 5 \}$ is applied to each query and the threshold, and what we observe is the first time that a query result is greater than or equal to the threshold after interference.
-We are interested in the differential privacy property of the data in $D$, assuming that they are all neighbours.
-The modeling can be done in guarded command language and some fragments are shown below.\todo{the algorithm seems redundant}
-
-\lz{use $\frac{2}{3}$, not $2/3$, and use $t'$, not $\hat{t}$}
-\begin{algorithm}
-    \begin{algorithmic}[1]
-
-        \Function{OnlineModel}{$d$, $t$}
-
-        \State{$\mathbf{If}$  $t=0$ $\rightarrow   2/3 : (\hat{t} = 0) + 1/6 : (\hat{t} = 1) + 1/12 : (\hat{t} = 2)$ }
-        \State{$+ 1/24 : (\hat{t} = 3) + 1/48 : (\hat{t} = 4) + 1/48: (\hat{t} = 5);$}
-        \State{$| $ $ t=1\rightarrow   1/3 : (\hat{t} = 0) + 1/3 : (\hat{t} = 1) + 1/6 : (\hat{t} = 2)$}
-        \State{$+ 1/12 : (\hat{t} = 3) + 1/24 : (\hat{t} = 4) + 1/24: (\hat{t} = 5);$}
-        \State{$...$  $...$}
-        \Comment{Assignments of $\hat{t}$ by the geometric mechanism}
-        \State{$\mathbf{Fi}$}
-
-        \For{each query $f_i$}
-        \State{$\mathbf{If}$  $d=0$ $\rightarrow   2/3 : (\hat{f} = 0) + 1/6 : (\hat{f} = 1) + 1/12 : (\hat{f} = 2)$ }
-        \State{$+ 1/24 : (\hat{f} = 3) + 1/48 : (\hat{f} = 4) + 1/48: (\hat{f} = 5);$}
-        \State{$| $ $ d=1\rightarrow   1/3 : (\hat{f} = 0) + 1/3 : (\hat{f} = 1) + 1/6 : (\hat{f} = 2)$}
-        \State{$+ 1/12 : (\hat{f} = 3) + 1/24 : (\hat{f} = 4) + 1/24: (\hat{f} = 5);$}
-        \State{$...$  $...$}
-        \Comment{Assignments of $\hat{f}$ by the geometric mechanism}
-        \State{$\mathbf{Fi}$}
-        \State{$\mathbf{If}$ $\hat{f}\ge \hat{t}$ $a_i=\top$ \Return }
-        \State{$|$   $\hat{f} < \hat{t}$ $a_i=\bot$}
-        \State{$\mathbf{Fi}$}
-        \EndFor
-        \EndFunction
-
-    \end{algorithmic}
-    \caption{Input is a private database $d \in \{0,1,2,3,4,5\}$, an adaptively chosen stream of sensitivity 1 queries $f_1$, .
-.
-., a threshold $t \in \{0,1,2,3,4,5\}$.Output is a stream of answers $a_1$,...}
-    \label{algorithm:online-model}
-\end{algorithm}
+In order to protect privacy, truncated $\alpha$-geometric mechanisms over $\underline{5}$ is applied to each query and the threshold, and what we observe is the first time that a query result is greater than or equal to the threshold after interference.
+We are interested in the differential privacy property of the data in $D$, assuming that a pair of data $d,d'$are neighbours, if $|d-d'|\le 1$.
+The algorithm is shown in Algorithm~\ref{algorithm:online-model}.

### 2018-01-30 specifications.tex
spec

@@
 
-As an example, for our threshold \lz{to do} model, we consider the property $\wedge_{i=1}^{c} \dpriv{\epsilon}{0}(\X^k \top) \wedge \PJ{\le 10^{-6}}(\X^c F T) \wedge \dpriv{\epsilon}{0}(\X^c F \top)$.
+As an example, for our threshold \lz{to do} model, we consider the property $\wedge_{i=1}^{7} \dpriv{3}{0}(\X^i \top) \wedge \PJ{\le .1}(\X^7 F T) \wedge \dpriv{3}{0}(\X^7 F \top)$.
 that is used to express the privacy for our threshold Markov chains.
@@
 
-Consider the Bluetooth example in~\cite{prism-bluetooth-page} and the
+Consider the Bluetooth example in~\cite{DKNP04} and the
 formula $\dpriv{\epsilon}{\delta}{\F \G \mathit{timeout}}$. Let us say

### 2018-01-30 applications.tex
fix typo
proof appendix
work through and 16 pages ;)

@@
-Streaming processing gives a motivating example for Markov
-decision processes. In stream processing, input data are retrieved and
+Streaming processing gives a motivating example for MDPs. In stream 
+processing, input data are retrieved and
 analyzed continuously. Such continual observation is commonly found
@@
 neighboring observations. In the following, we model a discrete
-variant of the mechanisms by Markov decision processes. We will also
+variant of the mechanisms by MDPs. We will also
 specify pan privacy in our model using $\dpCTLstar$.
@@
 In order to obtain a finite model, we formalize neighbors as action
-sequences in Markov decision processes. Consider the action set $\{
+sequences in MDPs. Consider the action set $\{
 a[a], a[b], b[a], b[b] \}$. Informally, the action $x[y]$ means the
@@
 
-Similar to Algorithm~\ref{algorithm:frequency-estimator}, our Markov
-decision process has internal states $s_{00}$, $s_{01}$, $s_{10}$, 
+Similar to Algorithm~\ref{algorithm:frequency-estimator}, our MDP has internal states $s_{00}$, $s_{01}$, $s_{10}$,
 $s_{11}$ for the two bits $f_a$ and $f_b$ of observed event types
@@
 L (s_{xy})$ for $x, y \in \{ 0, 1 \}$. From the
-initialization state $s$, the Markov decision process moves to one of
+initialization state $s$, the MDP moves to one of
 the four internal states randomly with the uniform
@@
 or $a[b]$ are chosen, they denote that the event type $a$ is currently
-observed. The Markov decision process moves to the internal state $s_{01}$
+observed. The MDP moves to the internal state $s_{01}$
 and $s_{11}$ with probability $.375 = \frac{3}{8}$ and $.625 =
@@
 On the other hand, the initialization state $t$
-corresponds to the computation on neighbors. The Markov decision
-process moves to the interal states $t_{00}$,
+corresponds to the computation on neighbors. The MDP moves to the internal states $t_{00}$,
 $t_{01}$, $t_{10}$, $t_{11}$ randomly with
@@
 denote that the current event type is replaced with the eveny type
-$a$. The Markov decision process moves to the internal state
+$a$. The MDP moves to the internal state
 $t_{00}$ and $t_{10}$ with probability $.375 =
 \frac{3}{8}$ and $.625 = \frac{5}{8}$ respectively. At an internal
-state, the Markov decision process can move to the output states $0$,
+state, the MDP can move to the output states $0$,
 $1$, or $2$ by the truncated $\frac{1}{2}$-geometric mechanism over
@@
 hence say the states $s$ and $t$ are neighbors in the
-Markov decision process (Figure~\ref{figure:frequency-estimator}).
+MDP (Figure~\ref{figure:frequency-estimator}).
 
-\begin{figure}
+\begin{figure}[tbh]
   \centering
@@
       ;
+
+      \hspace{.05\columnwidth}
 
@@
     \end{subfigure}
-  \caption{Frequency Estimator}
+  \caption{Frequency Estimator (some states are drawn twice).}
   \label{figure:frequency-estimator}
@@
 following $\dpCTLstar$ formula:
-\[
+$
 \Psi = \dpriv{\ln 4}{0} (\F \mathit{at}_0) \wedge \dpriv{\ln 4}{0} (\F
-\mathit{at}_1) \wedge \dpriv{\ln 4}{0} (\F \mathit{at}_2)p
-\]
+\mathit{at}_1) \wedge \dpriv{\ln 4}{0} (\F \mathit{at}_2)
+$.
 For pan privacy, we moreover wish to measure privacy loss in internal

### 2018-01-30 introduction.tex
remove notes
work through and 16 pages ;)

@@
 (\G \F \mathit{high})$. A state satisfies the formula if it
-is probabilistically similar to always almost surely
+is probabilistically similar to always almost surely\lz{infinitely often}
 answer $\mathit{high}$ as similar states. Compared to the unbounded
@@
 The time complexity of our model checking algorithm is the same as
-those of $\CTLstar$ for Markov chains. The logic $\dpCTLstar$ obtains its
+those of $\PCTLstar$ for Markov chains. The logic $\dpCTLstar$ obtains its
 expressiveness for free.
@@
 differential privacy properties, and moreover, complex properties such as those
-involving infinite behaviors can also be expressed. (ii) We model several differential privacy mechanisms, including the streaming process, in Markov chains or MDPs.
+involving infinite behaviors can also be expressed. (ii) We model several differential privacy mechanisms, including the streaming process, in Markov chains or Markov decision processes.
 (iii) We show that the model checking problem

### 2018-01-30 mdp.tex
work through and 16 pages ;)

@@
-\subsection{Markov Decision Processes}
+%\subsection{Markov Decision Processes}
 
@@
 As we describe earlier, the user has the possibility of querying multiple mechanisms.
-We use Markov decision processes to model such
+We use MDPs to model such
 interactive mechanisms. Specifically, external inputs are modeled by
@@
 \subsection{Semantics}
-The logic $\dpCTLstar$ can be interpreted over MDPs. Semantics over Markov chains generalizes to MDPs as well. Note that two states with different enabled actions
-are trivially distinguishable; no privacy can be preserved.
+The logic $\dpCTLstar$ can be interpreted over MDPs. Semantics over Markov chains generalizes to MDPs as well. %Note that two states with different enabled actions
+%are trivially distinguishable; no privacy can be preserved.
 Let $M = (S, \Act, \wp, L)$ be an MDP. Define the satisfaction
@@
 
-\begin{theorem}
+\begin{theorem}\label{theorem:mdp-model-checking}
 The $\dpCTLstar$ model checking problem for MDPs  is undecidable.

### 2018-01-30 syntax.tex
some changes
remove notes
work through and 16 pages ;)

@@
-Consider
+The syntax of $\dpCTL$ state and path formulae are given by:
 \begin{eqnarray*}
-  \Phi & ::= & \top \ |\ p \ |\ \neg \Phi \ |\ \Phi \wedge \Phi \ |\
+  \Phi & ::= & p \ |\ \neg \Phi \ |\ \Phi \wedge \Phi \ |\
                \PJ{J} \phi \ |\ \dpriv{\epsilon}{\delta} \phi\\
@@
 \end{eqnarray*}
-A \emph{state} formula $\Phi$ is either $\top$, an atomic proposition
+A \emph{state} formula $\Phi$ is either  an atomic proposition
 $p$, the negation of a sate formula, the conjunction of two state
@@
 ($\until$) operator enclosed by two path formulae.
+We often use syntactic sugar $F$ and $G$ defined by: $F \phi:= true \until \phi$ and $G\phi :=\neg F (\neg\phi)$.
 
 As in the classical setting, we consider the sublogic $\dpCTL$ by allowing only path formulae of the form $\X\Phi$ and $\Phi\until\Phi$.
+Moreover, one obtains $\PCTL$ and $\PCTLstar$ from $\dpCTL$ and $\dpCTLstar$
+by removing the differentially private operator  $\dpriv{\epsilon}{\delta}$.

### 2018-01-30 specifications.tex
add specification of the property

@@
 
-As an example, for our threshold \lz{to do} model, we consider the property $\wedge_{i=1}^{7} \dpriv{3}{0}(\X^i \top) \wedge \PJ{\le .1}(\X^7 F T) \wedge \dpriv{3}{0}(\X^7 F \top)$.
-that is used to express the privacy for our threshold Markov chains.
-It is only a valid formula if we would allow infinite conjunctions.
-However, it is interesting to note the relation to language equivalence for probabilistic automata\lz{to be discussed}
+As an example, for our threshold model(Algorithm ~\ref{algorithm:online-model}) , we consider the property $\wedge_{i=1}^{7} \dpriv{3}{0}(\X^i \top) \wedge \PJ{\le 0.025}(\X^7 F \top) \wedge \dpriv{3}{0.025}(\X^7 F \top)$, to express the privacy.
+The initial state of the corresponding Markov chain, where the input threshold $t$ is $3$, and the database $d$ is $4$, satisfies the above property by our computation.
+It means the probability of seeing $\top$  on any particular position, within 7 steps, is $(3, 0)$-differentially private for the database $d$ and its neighbours(3 and 5).
+It also states that the probability of seeing $\top$ after no less than 7 steps, is smaller than $0.025$ and is  $(3, 0.025)$-differentially private.
+It is only a valid formula if we would allow infinite conjunctions. However, it is interesting to note the relation to language equivalence for probabilistic automata\lz{to be discussed}
 

### 2018-01-30 specifications.tex
small changes

--- 2018-01-30 specifications.tex
@@
 
-As an example, for our threshold \lz{to do} model, we consider the property $\wedge_{i=1}^{7} \dpriv{3}{0}(\X^i \top) \wedge \PJ{\le .1}(\X^7 F T) \wedge \dpriv{3}{0}(\X^7 F \top)$.
-that is used to express the privacy for our threshold Markov chains.
-It is only a valid formula if we would allow infinite conjunctions.
-However, it is interesting to note the relation to language equivalence for probabilistic automata\lz{to be discussed}
+As an example, for our threshold model(Algorithm ~\ref{algorithm:online-model}) , we consider the property $\wedge_{i=1}^{6} \dpriv{3}{0}(\X^i \top) \wedge \PJ{\le 0.025}(\X^6 F \top) \wedge \dpriv{3}{0.025}(\X^6 F \top)$, to express the privacy.
+The initial state of the corresponding Markov chain, where the input threshold $t$ is $3$, and the database $d$ is $4$, satisfies the above property by our computation.
+It means the probability of seeing $\top$  on any particular position, within $6$ steps, is $(3, 0)$-differentially private for the database $d$ and its neighbours($3$ and $5$).
+It also states that the probability of seeing $\top$ after no less than $6$ steps, is smaller than $0.025$ and is  $(3, 0.025)$-differentially private.
+It is only a valid formula if we would allow infinite conjunctions. However, it is interesting to note the relation to language equivalence for probabilistic automata\lz{to be discussed}
 

### 2018-01-30 examples.tex
work through and 16 pages ;)
adaptions

@@
 \in L (t)$.
-For each state $x$, we moreover associate an atomic predicate
-$\mathit{at}_x$ (not shown in the figure). 
 States $+$ and $-$ are neighbors.
@@
 
-\begin{algorithm}
+\begin{algorithm}[tbh]
 	\begin{algorithmic}[1]
@@
 	\end{algorithmic}
-	\caption{Input is a private database $d \in \{0,1,2,3,4,5\}$, an adaptively chosen stream of sensitivity 1 queries $f_1$, .
-.
-., a threshold $t \in \{0,1,2,3,4,5\}$.Output is a stream of answers $a_1$,...}
+	\caption{Input is a private database $d \in \{0,1,\ldots,5\}$, an adaptively chosen stream of sensitivity 1 queries $f_1,f_2$,\ldots, a threshold $t \in \{0,1,\ldots,5\}$.Output is a stream of answers $a_1,a_2$,\ldots}
 	\label{algorithm:online-model}
@@
 are above a given threshold.
-Other queries below the threshold are considered "irrelevent" and might be discarded.
-In ~\cite{DR:14:AFDP}, since the model is set for continuous database, the Laplace mechanism is applied to both queries and the threshold.
+Other queries below the threshold are considered \emph{irrelevant} and might be discarded.
+In~\cite{DR:14:AFDP}, the model is set for continuous database, and the Laplace mechanism is applied to both queries and the threshold.
 In our model, the database is discrete and instead the geometric mechanism is applied.
 
-We consider the database set $D = \{0,1,2,3,4,5\}$, a threshold $t \in \{0,1,2,3,4,5\}$, queries $\{f| f $ with sensitivity $\Delta (f) = 1 \}$.
-In order to protect privacy, truncated $\alpha$-geometric mechanisms over $\underline{5}$ is applied to each query and the threshold, and what we observe is the first time that a query result is greater than or equal to the threshold after interference.
+We consider the database set $D = \{0,1\ldots,5\}$, a threshold $t \in \{0,1\ldots,5\}$, and queries $\{f| f $ with sensitivity $\Delta (f) = 1 \}$.
+In order to protect privacy, truncated $\alpha$-geometric mechanisms over $\{0,1\ldots,5\}$ is applied to each query and the threshold, and what we observe is the first time that a query result is greater than or equal to the threshold after interference.
 We are interested in the differential privacy property of the data in $D$, assuming that a pair of data $d,d'$are neighbours, if $|d-d'|\le 1$.

### 2018-01-30 semantics.tex
adaptions

@@
-Given a finite discrete-time Markov chain $K = (S, \wp, L)$, a
-\emph{neighborhood relation on $S$} $\neighbor{S} \subseteq S \times S$
+Given a finite Markov chain $K = (S, \wp, L)$, a
+\emph{neighborhood relation} $\neighbor{S} \subseteq S \times S$
 is a reflexive and symmetric relation on $S$. We will write $s
@@
 through iterative approximations. We summarize it in the following
-theorem~\cite{BK:08:PMC}:
+theorem (details see~\cite{BK:08:PMC}):
 
@@
 } are computable within time polynomial in
-  $|S|$.
+  $|S|$.\lz{and $|\wp|$?}
   \label{lemma:PJ-subroutines}
@@
   $\myPr{s}{K}{\neighbor{S}}{\phi}$ with general path formula $\phi$.
-For this purpose one first constructs a deterministic $\omega$-automaton $R$, such as deterministic Rabin automaton.
+For this purpose one first constructs a deterministic $\omega$-automaton $R$ corresponding to $\phi$, such as deterministic Rabin automaton.
 Then, the probability reduces to a

### 2018-01-30 specifications.tex
adaptions

--- 2018-01-30 specifications.tex
@@
 
-As an example, for our threshold \lz{to do} model, we consider the property $\wedge_{i=1}^{7} \dpriv{3}{0}(\X^i \top) \wedge \PJ{\le .1}(\X^7 F T) \wedge \dpriv{3}{0}(\X^7 F \top)$.
-that is used to express the privacy for our threshold Markov chains.
-It is only a valid formula if we would allow infinite conjunctions.
-However, it is interesting to note the relation to language equivalence for probabilistic automata\lz{to be discussed}
+As an example, for our threshold model (Algorithm~\ref{algorithm:online-model}), we consider the property $\Phi=\wedge_{i=1}^{6} \dpriv{3}{0}(\X^i \top) \wedge \PJ{\le 0.025}(\X^6 F \top) \wedge \dpriv{3}{0.025}(\X^6 F \top)$, to express the privacy.
+The state $s_{34}$ (with the input threshold $t=3$, and the database $d=4$) satisfies $\Phi$ by our computation.
+It means the probability of reaching $\top$  within $i$ steps with $i$ ranges over $\{1,\ldots,6\}$, is $(3, 0)$-differentially private for the database $d$ and its neighbours ($d=3$ and $d=5$).
+It also states that the probability of reaching $\top$ after no less than $6$ steps, is smaller than $0.025$ and is  $(3, 0.025)$-differentially private.
 

### 2018-01-30 introduction.tex
intro organisation done

@@
 (\G \F \mathit{high})$. A state satisfies the formula if it
-is probabilistically similar to always almost surely\lz{infinitely often}
+is probabilistically similar to  infinitely often
 answer $\mathit{high}$ as similar states. Compared to the unbounded
@@
 
-\paragraph{Contributions.} Our main contributions are three folds. (i)
+\paragraph{Contributions.} Our main contributions are threefold. (I)
 We introduce the logic $\dpCTLstar$ for reasoning about properties in
@@
 differential privacy properties, and moreover, complex properties such as those
-involving infinite behaviors can also be expressed. (ii) We model several differential privacy mechanisms, including the streaming process, in Markov chains or Markov decision processes.
-(iii) We show that the model checking problem
+involving infinite behaviors can also be expressed. (II) We model several differential privacy mechanisms, including the streaming process, in Markov chains or Markov decision processes.
+(III) We show that the model checking problem
 problem for Markov chains is standard. For MDPs we show that it is undecidable.

### 2018-01-30 main.tex
proof appendix
work through and 16 pages ;)
adaptions
intro organisation done

@@
 \else
-\author{\vspace*{-1.2cm}}
-\institute{\vspace*{-1.2cm}\ }
+\author{Author}
+\institute{Institute}
 \fi
@@
 
-This paper is organized as follows. After the introduction,
-preliminaries are given in Section~\ref{section:preliminaries}. The
-logic $\dpCTL$ is presented in
-Section~\ref{section:dpCTL}. The model cheking algorithms are developed in
-Section~\ref{section:model-checking-algorithms}. They are followed by
-applications (Section~\ref{section:applications}. Finally,
+\paragraph{Organisation of the paper.}
+Preliminaries are given in Section~\ref{section:preliminaries}. 
+In Section~\ref{section:examples} we discuss how differential privacy mechanisms can be modeled in Markov chains.
+The logic $\dpCTLstar$ and model checking algorithms for Markov chains are presented in
+Section~\ref{section:dpCTL}. 
+Section~\ref{section:specifying-properties} describes differential privacy properties using   $\dpCTLstar$.
+They are followed by the extension to Markov decision processes and an
+application in streaming process (Section~\ref{section:mdp}). Finally,
 Section~\ref{section:conclusions} concludes our presentation.
@@
 
+\newpage
+\appendix
+
+\section{Proof of Theorem~\ref{theorem:mdp-model-checking}}
+\input{proof}
+
 \end{document}

### 2018-01-30 proof.tex
remove at from the pre, added it where it is needed

@@
 The proof follows by a reduction from the \emph{emptiness} problem for probabilistic automata.
-A \emph{probabilistic automaton}~\cite{Rabin63} is a tuple $\mathcal{A}= (S,\Sigma,M,s_0,F)$ where
+A \emph{probabilistic automaton}~\cite{Rabin63} is a tuple $\mathcal{A}= (S,\Sigma,M,s_0,B)$ where
 \begin{itemize}
@@
 Thus, $M(\eta)(s,s')$ denotes the probability of going from $s$ to $s'$ after reading $\eta$.
-Let $v_B$ denote the characteristic row vector for the set $F$, and $v_{s_0}$ denote the characteristic row vector for the set $\{s_0\}$.
+Let $v_B$ denote the characteristic row vector for the set $B$, and $v_{s_0}$ denote the characteristic row vector for the set $\{s_0\}$.
 Then, the accepting probably of $\eta$ by $\mathcal{A}$ is defined as $v_{s_0}\cdot M(\eta)\cdot (v_B)^c$ where $(v_B)^c$ denotes the transpose of $v_B$.
@@
 
-Now we construct our MDP  $M =
+Let $AP:=\{at_B\}$. Now we construct our MDP  $M =
 (S\cupdot S', \Sigma, \wp, L)$ where $\wp(s,a,t)$ equals to $M(s,a,t)$ if $s,t\in S$ and to $M'(s,a,t)$ if $s,t\in S'$.
 We define the neighbor relation $N_S:=\{(s_0,s_0'),(s_0',s_0)\}$ by relating states $s_0,s_0'$.
-The labelling function $L$ is defined by $L(s)=at_B$ if $s\in F$ and $L(s)=\empty$ otherwise.
+The labelling function $L$ is defined by $L(s)=\{at_B\}$ if $s\in B$ and $L(s)=\emptyset$ otherwise.
 
@@
  For the other direction, assume that all $\eta\in\Sigma^*$ it holds $v_{s_0}\cdot M(\eta)\cdot (v_B)^c \leq 0$.
-For the sake of contradiction, assume that $s_0\not\models \dpriv{1}{0} (F at_B)$.
+We prove by contradiction.
+Assume that $s_0\not\models \dpriv{1}{0} (F at_B)$.
 Since the relation $N_S$ contains only one pair, it implies that there exists a scheduler $\scheduler{S}\in\Sigma^\omega$ such that $\myPr{s_0}{M_{\scheduler{S}}}{\neighbor{S}}{F at_B} >0$.

### 2018-01-31 examples.tex
...

@@
-As discussed in previous section, a simple way to design differentially private mechanisms is to add
+A simple way to design differentially private mechanisms is to add
 random noises.
-Thus, in this section, we discuss how to model differentially private mechanisms using Markov chains. We explain the main idea first with some examples.
+Now we describe our main ideas for  modelling differentially private mechanisms.
+A state $s$ stores at least the a database $x\in \mathcal{X}^n$, which is the data hold by a trusted \emph{curator}. A transition corresponds to the data analysis mechanisms.
+States are typically labelled with the corresponding outputs
+the curator generates. We explain the main idea with some examples.
 
-\paragraph{Survey Mechanisms.}
+\subsection{Survey Mechanisms}
+\label{subsec:survey}
+
+Consider the survey question: have you been diagnosed
+with the disease $X$? In order to protect surveyees' privacy, each
+answers the question as follows. The surveyee first flips a
+coin. If it is tail, she answers the question truthfully. Otherwise,
+she randomly answers \textit{Yes} or \textit{No}
+uniformly~\cite{DR:14:AFDP}.
 
@@
 
-Consider the survey question: have you been diagnosed
-with the disease $X$? In order to protect surveyees' privacy, each
-answers the question as follows. The surveyee first flips a
-coin. If it is tail, she answers the question truthfully. Otherwise,
-she randomly answers \textit{Yes} or \textit{No}
-uniformly~\cite{DR:14:AFDP}.
-
 \begin{example}\label{exa:survey}
   Figure~\ref{figure:2-dp} shows the survey
-mechanism and its corresponding Markov chain. In the figure, the states
+mechanism and its corresponding Markov chain. 
+The data universe
+$\mathcal{X}$ is $\{ +, - \}$. The mechanism $M$ is a randomized
+algorithm with inputs in $\mathcal{X}$ and outputs in $\{ Y, N
+\}$. In the figure, the states
 $+$ and $-$ denote the diagnoses are positive or negative
@@
 
-Let us recall briefly the analysis of the survey mechanism. The data universe
-$\mathcal{X}$ is $\{ +, - \}$. The mechanism $M$ is a randomized
-algorithm with inputs in $\mathcal{X}$ and outputs in $\{ Y, N
-\}$. For any $x \in \mathcal{X}$, we have $1/4 \leq \Pr[M(x) = Y] \leq
-3/4$. Hence for any neighboring $x, x' \in \mathcal{X}$, $\Pr[M(x) =
-Y] \leq 3/4 = 3 \cdot 1/4 \leq 3 \Pr[M (x') =
-Y]$. Similarly, $\Pr[M (x) = N] \leq 3 \Pr[M (x') = N]$ for any
-neighboring $x, x' \in \mathcal{X}$. The survey mechanism is
-$(\ln 3, 0)$-differentially private.
-
-\paragraph{Truncated $\alpha$-Geometric Mechanism.}
+\subsection{Truncated $\alpha$-Geometric Mechanism}\label{subsec:geometric}
 More sophisticated differentially private mechanisms are
@@
 
-\begin{figure}
+\begin{figure}[tbh]
   \centering
@@
   \end{subfigure}
-
   \caption{A Markov Chain for $1/2$-Geometric Mechanism}
-  \label{figure:geometric-mechanism}
+  \label{figure:geometric-mechanism}\vspace*{-.5cm}
 \end{figure}
@@
 
-\subsection{Markov Chains for Differentially Private Mechanisms}
-Now we describe our main ideas for  modelling differentially private mechanisms.
-A state $s$ stores at least the a database $x\in \mathcal{X}^n$, which is the data hold by a trusted \emph{curator}. A transition corresponds to the data analysis mechanisms.
-States are typically labelled with the corresponding outputs
-the curator generates.
-
+\subsection{Above Threshold Mechanisms}
 In the differentially privacy context, we differentiate between \emph{offline} and \emph{online} models.
@@
 We are interested in the differential privacy property of the data in $D$, assuming that a pair of data $d,d'$are neighbours, if $|d-d'|\le 1$.
-The algorithm is shown in Algorithm~\ref{algorithm:online-model}.
+The algorithm
+is shown in Algorithm~\ref{algorithm:online-model}.

### 2018-01-31 preliminaries.tex
work through and 16 pages ;)
...

@@
 We briefly review the definitions of differential privacy, Markov
-chains, and Markov decision processes. For differential privacy, we
+chains, and Markov decision processes~\cite{Put05}. For differential privacy, we
 follow the standard definition in~\cite{DR:14:AFDP,GRS:09:UUPM,GRS:12:UUPM}. Our
@@
 \subsection{Differential Privacy}
-
 We denote the data universe by $\mathcal{X}$; $x \in \mathcal{X}^n$ is

### 2018-01-31 semantics.tex
...

@@
 algorithms for $\PCTL$ by computing the states satisfying
-sub state-formulae inductively~\cite{BK:08:PMC}. It hence suffices to
+sub state-formulae inductively~\cite{HanssonJ94,BK:08:PMC}. It hence suffices to
 consider the inductive step where the states satisfying the subformula

### 2018-01-31 specifications.tex
...

@@
 \paragraph{Standard Differential Privacy.}
+Consider the survey mechanism (Subsection~\ref{subsec:survey}).
 The formula $\dpriv{\ln 3}{0} (\X \mathit{out}_Y)$ holds in state $s$ if for  $t$ with $sNt$, the probability of satisfying $\X \mathit{out}_Y$ from $s$ and $t$ satisfy:
@@
 The formula $\dpriv{\ln 3}{0} (\X \mathit{out}_Y)\wedge \dpriv{\ln 3}{0} (\X \mathit{out}_N)$ thus specifies the differential privacy property for states $+$ and $-$ in Example~\ref{exa:survey}.
+The survey mechanism is
+$(\ln 3, 0)$-differentially private.
+
+\hide{Let us recall briefly the analysis of the survey mechanism. For any $x \in \mathcal{X}$, we have $1/4 \leq \Pr[M(x) = Y] \leq
+3/4$. Hence for any neighboring $x, x' \in \mathcal{X}$, $\Pr[M(x) =
+Y] \leq 3/4 = 3 \cdot 1/4 \leq 3 \Pr[M (x') =
+Y]$. Similarly, $\Pr[M (x) = N] \leq 3 \Pr[M (x') = N]$ for any
+neighboring $x, x' \in \mathcal{X}$. 
+}
 
@@
 \cdots \wedge \dpriv{\ln 2}{0} (\X out_5)$. If the state $s_k$ satisfies
-$\psi$ for $k = 0, \ldots, 5$, then the $.5$-geometric mechanism is
+$\psi$ for $k = 0, \ldots, 5$, then the $.5$-geometric mechanism (Subsection~\ref{subsec:geometric}) is
 $(\ln 2, 0)$-differentially private.

### 2018-01-31 syntax.tex
...

@@
-The syntax of $\dpCTL$ state and path formulae are given by:
+The syntax of $\dpCTL$ state and path formulae is given by:
 \begin{eqnarray*}
@@
 As in the classical setting, we consider the sublogic $\dpCTL$ by allowing only path formulae of the form $\X\Phi$ and $\Phi\until\Phi$.
-Moreover, one obtains $\PCTL$ and $\PCTLstar$ from $\dpCTL$ and $\dpCTLstar$
+Moreover, one obtains $\PCTL$~\cite{pHanssonJ94} and $\PCTLstar$~\cite{BiancoA95} from $\dpCTL$ and $\dpCTLstar$
 by removing the differentially private operator  $\dpriv{\epsilon}{\delta}$.

### 2018-01-31 introduction.tex
...
merge comments

@@
 input databases~\cite{DMNS:06:CNSPD,D:06:DP}. Consider, for example,
-the Laplace mechanism where a random noise is added to analysis
-results with the Laplace distribution~\cite{DR:14:AFDP}. Random noises
+the Laplace mechanism where analysis results are perturbed by random
+noises with the Laplace distribution~\cite{DR:14:AFDP}. Random noises
 hide the differences of analysis results from similar databases.
 Clearly, more
-randomness gives more privacy but less utility in released noisy
+randomness gives more privacy but less utility in released perturbed
 results. Under the framework of differential privacy, data analysts
@@
 Designing differentially private mechanisms can be tedious for
-sophisticated data analyses. Privacy leak has also been observed from
+sophisticated data analyses. Privacy leak has also been observed in
 data analysis programs with differentially private
@@
 differential privacy and investigate their model checking
-algorithms. Data analysts can automatically verify their designs and
+problems. Data analysts can automatically verify their designs and
 implementations with our techniques. Most interestingly, our
 techniques can be adopted rather easily by existing probabilistic
-model checkers. Hopefully, more interaction between model checking and
-privacy analysis will follow.
+model checkers. Privacy checking with existing tools can be attained
+with minimal efforts and overheads. Hopefully, more interaction
+between model checking and privacy analysis will follow.
 
@@
 
+In order to illusrtrate applicability of our techniques, we give
+detailed formalizations of several data analysis mechanisms in this
+paper. 
 In differential privacy, data analysis mechanisms are
@@
 probabilistic model checking to formalize such mechanisms by
-by Markov chains or Markov decision processes. We give
+by Markov chains or Markov decision processes~\cite{Put05}. We give
 formalizations of several differentially private mechanisms and
-explain subtle privacy conditions in this paper.
+explain subtle privacy conditions.
 When a data analysis mechanism does not
@@
 behaviors. A reflexive and commutative user-defined binary relation
-over states is required to formalize similarity. Two states are
-similar if they are related by the binary relation. We moreover add
+over states is required to formalize similarity. We moreover add
 the path quantifier $\dpriv{\epsilon}{\delta}$ for similar
@@
 Not only can differential privacy specify randomized algorithms, it
-also naturally generalizes to reactive systems.
-Privacy analysis applies to systems with infinite behaviors as well.
+also generalizes to reactive systems naturally.
+More data analysis mechanisms are admitable for sophisticated formal
+privacy analysis through our techniques.
 
@@
 
-In order to verify $\dpCTLstar$ properties, we extend the standard
-probabilistic model checking algorithms. For Markov chains, the states
+We moreover extend the standard probabilistic model checking
+algorithms to verify $\dpCTLstar$ properties automatically. For Markov
+chains, states 
 satisfying a subformula $\dpriv{\epsilon}{\delta} \phi$ are computed
-by a simple variant of the algorithm for Markov chains.
-The time complexity of our model checking algorithm is the same as
+by a simple variant of the model checking algorithm for Markov chains.
+The time complexity of our algorithm is the same as
 those of $\PCTLstar$ for Markov chains. The logic $\dpCTLstar$ obtains its
-expressiveness for free.
+essentially expressiveness for free.
 For Markov decision processes, we show that checking whether a state
@@
 differential privacy properties, and moreover, complex properties such as those
-involving infinite behaviors can also be expressed. (II) We model several differential privacy mechanisms, including the streaming process, in Markov chains or Markov decision processes.
-(III) We show that the model checking problem
-problem for Markov chains is standard. For MDPs we show that it is undecidable.
+involving infinite behaviors can also be expressed. (II) We model several differential privacy mechanisms, including stream processing, in Markov chains or Markov decision processes.
+(III) We show that the model checking 
+problem for Markov chains is standard. For Markov decision processes we show that it is undecidable.

### 2018-01-31 main.tex
remove at from the pre, added it where it is needed
merge comments

@@
   are specified by $\dpCTLstar$. In order to verify privacy properties
-  automatically, the model checking problems are investigated.  We
+  automatically, model checking problems are investigated.  We
   give a model checking algorithm for Markov chains. Model checking
@@
 decidable subclass. As an example, consider the formula $\bigwedge_{k \in \bbfN}
-\dpriv{\epsilon}{\delta} \X^k \mathit{at}_{00}$ which was used in specifying the pan privacy.
+\dpriv{\epsilon}{\delta} (\X^k  00)$ which was used in specifying the pan privacy.
 For the case $\epsilon=\delta=0$, it reduces to a language equivalence problem for probabilistic automata.

### 2018-01-31 preliminaries.tex
merge comments

@@
 will have a similar output distribution when the individual is not in
-the database. Regardless of the presence of the individual, such
-mechanisms will behave similarly. Privacy of the individual is thus
+the database. Such mechanisms behave similarly regardless of the
+presence of any individual. Individual privacy is thus 
 preserved by differentially private mechanisms.
@@
 \times S \rightarrow [0, 1]$ with $\sum_{t \in S} \wp(s, \alpha, t)
-= 1$ for every $s \in S$ and $\alpha \in \Act$ and $S$,
-$L$ as for Markov chains.
+= 1$ for every $s \in S$ and $\alpha \in \Act$. $S$ and 
+$L$ are as for Markov chains.
 A \emph{path} $\pi$ in $M$ is an infinite sequence $\pi_0 \alpha_1
@@
 Let $M = (S, \Act, \wp, L)$ be an MDP. A
-\emph{scheduler} for $M$ is simply an infinite sequence  $\scheduler{S}\in \Act^\omega$.
+\emph{scheduler} for $M$ is a function $\scheduler{S} : S^+
+\rightarrow \Act$. A \emph{query scheduler} for $M$ is a function
+$\scheduler{Q} : \bbfZ^{\geq 0} \rightarrow \Act$. 
 A path $\pi =
-\pi_0 \alpha_1 \pi_1 \cdots \pi_n \alpha_{n+1} \cdots$ is an
+\pi_0 \alpha_1 \pi_1 \cdots$ $\pi_n$ $\alpha_{n+1} \cdots$ is an
 \emph{$\scheduler{S}$-path} if $\alpha_{i+1} = \scheduler{S}(\pi_0

### 2018-01-31 applications.tex
remove at from the pre, added it where it is needed
merge comments

@@
-Streaming processing gives a motivating example for MDPs. In stream 
+Stream processing gives a motivating example for MDPs. In stream 
 processing, input data are retrieved and
@@
 formulated and designed. Classical differential privacy however is
-insufficient in streaming processing. Note that observations are of
-indefinite lengths in such applications. All differentially private
+insufficient in streaming processing. Note that data are collected
+continuously and outcomes are requested at an indefinite future
+in such applications. All differentially private
 mechanism necessarily stay in their internal states indefinitely
@@
 neighboring observations. In the following, we model a discrete
-variant of the mechanisms by MDPs. We will also
+variant of a stream processing mechanisms by MDPs. We will also
 specify pan privacy in our model using $\dpCTLstar$.
@@
 event types. To compute the number of observed event types, the
-mechanism adopts the truncated $1/2$-geometric mechanism on the sum of
+mechanism adopts the truncated $\frac{1}{2}$-geometric mechanism on the sum of
 two bits for event types. If the sum is $0$, the mechanism outputs $0$
@@
 \Pr_B[f_a = x]$. On neighboring observations, the distribution of
-$f_a$ can only be changed from $U$ to $B$, values of $f_a$ is $(2,
+$f_a$ can only be changed from $U$ to $B$, values of $f_a$ is $(\ln 2,
 0)$-differentially private and similarly for $f_b$. Finally, the
-$1/2$-geometric mechanism gives $(\ln 2, 0)$-differential privacy.
+$\frac{1}{2}$-geometric mechanism gives $(\ln 2, 0)$-differential privacy.
 Recall $2\ln 2 = \ln 4$,
@@
 distribution. Algorithm~\ref{algorithm:frequency-estimator} has three
-different outputs. By privacy analysis, the state $I$ satisfies the
+different outputs. By privacy analysis, the state $s$ satisfies the
 following $\dpCTLstar$ formula:
@@
 \mathit{at}_1) \wedge \dpriv{\ln 4}{0} (\F \mathit{at}_2)
-$.
+$
+(Here 
+$\mathit{at}_i$, i=1,2,3, is an atomic proposition satisfied only at state $i$, which is not shown in the figure).
 For pan privacy, we moreover wish to measure privacy loss in internal
@@
 \bigwedge_{k \in \bbfN}
-\dpriv{\ln 4}{0} \X^k {00} \wedge
-\dpriv{\ln 4}{0} \X^k {01} \wedge
-\dpriv{\ln 4}{0} \X^k {10} \wedge
-\dpriv{\ln 4}{0} \X^k {11}.
+\dpriv{\ln 4}{0} (\X^k {00}) \wedge
+\dpriv{\ln 4}{0} (\X^k {01}) \wedge
+\dpriv{\ln 4}{0} (\X^k {10}) \wedge
+\dpriv{\ln 4}{0} (\X^k {11}).
 \]

### 2018-01-31 examples.tex
merge comments

@@
 uniformly~\cite{DR:14:AFDP}.
+
+Let us analyze the mechanism briefly. The data universe $\mathcal{X}$
+is $\{ +, - \}$. The mechanism $M$ is a randomized algorithm with
+inputs in $\mathcal{X}$ and outputs in $\{ Y, N \}$. For any $x \in
+\mathcal{X}$, we have $\frac{1}{4} \leq \Pr[M (x) = Y] \leq
+\frac{3}{4}$. Hence $\Pr[M (x) = Y] \leq \frac{3}{4} = 3 \cdot
+\frac{1}{4} \leq e^{\ln 3} \Pr[M (x') = Y]$ for any neighboring $x, x'
+\in \mathcal{X}$. Similarly, $\Pr[M (x) = N] \leq e^{\ln 3} \Pr[M (x')
+= N]$. The survey mechanism is $(\ln 3, 0)$-differentially private.
 
@@
   \end{subfigure}
-  \caption{Survey Mechanism with 2-Differential Privacy}
+  \caption{Survey Mechanism with $\ln 3$-Differential Privacy}
   \label{figure:2-dp}
@@
 diagnoses. Inferences on individual diagnosis can be plausibly denied.
-In general, missing transitions (such as those from $Y$ and $N$) lead
+Missing transitions (such as those from $s$ and $t$) lead
 to a special state $\dagger$ with a self-loop. We omit the transitions
@@
 output distribution on any inputs $x, x'$ with $f (x) = f
-(x')$. Moreover, it is also $(- {\Delta (f)} \ln \alpha, 0)$-differentially
+(x')$. Moreover, it is $(- {\Delta (f)} \ln \alpha, 0)$-differentially
 private for any query $f$ with sensitivity $\Delta (f)$. The range of
@@
 private for any query $f$ with sensitivity $\Delta (f)$.
-The truncated $1/2$-geometric mechanism over $\{ 0, 1, \ldots, 5 \}$ is
+The truncated $\frac{1}{2}$-geometric mechanism over $\{ 0, 1, \ldots, 5 \}$ is
 given in Figure~\ref{figure:geometric-mechanism-table}.
@@
   \end{subfigure}
-  \caption{A Markov Chain for $1/2$-Geometric Mechanism}
+  \caption{A Markov Chain for $\frac{1}{2}$-Geometric Mechanism}
   \label{figure:geometric-mechanism}\vspace*{-.5cm}
@@
 Similar to the survey mechanism, it is
-straightfoward to model the truncated $1/2$-geometric
+straightfoward to model the truncated $\frac{1}{2}$-geometric
 mechanisms as Markov chains.
@@
 the Markov chain of the truncated
-$1/2$-geometric mechanism over $\{ 0, 1, \ldots, 5 \}$.
+$\frac{1}{2}$-geometric mechanism over $\{ 0, 1, \ldots, 5 \}$.
 \end{example}

### 2018-01-31 semantics.tex
merge comments

@@
-Given a finite Markov chain $K = (S, \wp, L)$, a
+Given a Markov chain $K = (S, \wp, L)$, a
 \emph{neighborhood relation} $\neighbor{S} \subseteq S \times S$
@@
   \textmd{for every } t \textmd{ with }  s \neighbor{S} t,
-      \myPr{s}{K}{N}{\phi} \leq e^{\epsilon} \myPr{t}{K}{N}{\phi} + \delta
+      \myPr{s}{K}{\neighbor{S}}{\phi} \leq e^{\epsilon} 
+      \myPr{t}{K}{\neighbor{S}}{\phi} + \delta
       \textmd{ and }\\
-  & &  \myPr{t}{K}{N}{\phi} \leq e^{\epsilon} \myPr{s}{K}{N}{\phi} + \delta
+  & &  \myPr{t}{K}{\neighbor{S}}{\phi} \leq e^{\epsilon} 
+       \myPr{s}{K}{\neighbor{S}}{\phi} + \delta
 \end{eqnarray*}
@@
 \begin{lemma}
-  Let $K = (S, \wp, L)$ be a Markov chain, $s
+  Let $K = (S, \wp, L)$ be a Markov chain, $\neighbor{S}$ a
+  neighborhood relation on $S$, $s
   \in S$, and $B, C \subseteq S$. The probabilities
@@
 \begin{proposition}
-  Let $K = (S, \wp, L)$ be a Markov chain.
-  $\{ s : K, \neighbor{K}, s \models \dpriv{\epsilon}{\delta} \phi \}$ is
+  Let $K = (S, \wp, L)$ be a Markov chain and $\neighbor{S}$ a
+  neighborhood relation on $S$.
+  $\{ s : K, \neighbor{S}, s \models \dpriv{\epsilon}{\delta} \phi \}$ is
   computable within time polynomial in $|S|$.
@@
 \begin{corollary}
-  Let $K = (S, \wp, L)$ be a Markov chain and $\Phi$ a $\dpCTL$
-  formula. $\{ s : K, \neighbor{K}, s \models \Phi \}$ is
+  Let $K = (S, \wp, L)$ be a Markov chain, $\neighbor{S}$ a
+  neighborhood relation on $S$, and $\Phi$ a $\dpCTL$
+  formula. $\{ s : K, \neighbor{S}, s \models \Phi \}$ is
   computable within time polynomial in $|S|$ and $|\Phi|$.

### 2018-01-31 specifications.tex
merge comments

@@
 \cdots \wedge \dpriv{\ln 2}{0} (\X out_5)$. If the state $s_k$ satisfies
-$\psi$ for $k = 0, \ldots, 5$, then the $.5$-geometric mechanism (Subsection~\ref{subsec:geometric}) is
+$\psi$ for $k = 0, \ldots, 5$, then the $\frac{1}{2}$-geometric mechanism (Subsection~\ref{subsec:geometric}) is
 $(\ln 2, 0)$-differentially private.
@@
 The degradation is rooted in the repeated computation.
-To illustrate this property, we consider the extended survey example below where the curator allows two consequent queries.
+To illustrate this property, we consider the extended survey example below where the mechanism allows two consequent queries.
 The model is depicted in Figure~\ref{figure:2-dp-mc}.
@@
 (a) Formula $\dpriv{\ln 3}{0} (\X \X \mathit{out}_Y)$.
-In this case we donot have privacy lost, even though we are querying twice: the reason is that the output of the first query is not utilized at all.
+In this case we do not have privacy loss, even though we are querying twice: the reason is that the output of the first query is not utilized at all.
 It is easy to verify that it is indeed satisfied by $+$ and $-$.
@@
 Observe the inner formula is satisfied by all neighbours.
-Thus, the outer formula also has no privacy lost.
+Thus, the outer formula also has no privacy loss.
 
-\paragraph{Tighter Privacy Bounds for Compositional Properties.}
-One of the advantage of applying model checking is that we may get tighter bounds for compositional properties.
+\paragraph{Tighter Privacy Bounds for Composition.}
+One of the advantage of applying model checking is that we may get tighter bounds for composition.
 Consider our survey example, and the property $\dpriv{0}{.5} (\X \mathit{out}_Y)$.
 Obviously, it holds in states $+$ and $-$.
-A careful check ensures that one cannot decrease the value of $\epsilon_1=0$ or $\delta_1=.5$ without increasing the other one.
+A careful check ensures that one cannot decrease the value of $\epsilon_1=0$ or $\delta_1=.5$ without increasing the other.
 
@@
 Figure~\ref{figure:2-dp-mc}. Applying the
-compositional theorem, one can choose $\epsilon_2=2\epsilon_1=0$ and
+compositional theorem, one has $\epsilon_2=2\epsilon_1=0$ and
 $\delta_2=2\delta_1=1$. However, we can check easily that one gets
@@
 \paragraph{Infinite Differential Privacy.}
-Due to the power of linear temporal logic, our logic $\dpCTLstar$ generalizes differential privacy to infinite
+Thanks to the expressiveness of linear temporal logic, our logic $\dpCTLstar$ generalizes differential privacy to infinite
 behaviors naturally. Classical differential privacy discusses properties about
-randomized terminating mechanisms. Sophisticated mechanisms for
-streaming data have also been discussed in social networks and
-internet of things. Such mechanisms inherently have infinite
-computation. Their privacy guarantees are approximated by those of the
-prefixes of computation.
+randomized terminating mechanisms. 
+% Sophisticated mechanisms for streaming data have also been discussed 
+% in social networks and internet of things. Such mechanisms inherently 
+% have infinite computation. Their privacy guarantees are approximated 
+% by those of the prefixes of computation.
 With $\dpCTLstar$, the subformula $\phi$ in $\dpriv{\epsilon}{\delta} \phi$

### 2018-01-31 applications.tex
adapted schedulers ect

@@
 privacy even in their internal states~\cite{DNPR:10:DPCO,DNPRY:10:PPSA}.
-The idea is to show reachability probabilities are similar on
+The idea is to show \emph{reachability probabilities} are similar on
 neighboring observations. In the following, we model a discrete
@@
     \end{subfigure}
-  \caption{Frequency Estimator (some states are drawn twice).}
+  \caption{MDP for the Frequency Estimator (some states are drawn twice).}
   \label{figure:frequency-estimator}

### 2018-01-31 mdp.tex
adapted schedulers ect

@@
 %\subsection{Markov Decision Processes}
-
-\begin{wrapfigure}{r}{.5\columnwidth}
+\begin{wrapfigure}[14]{r}{.45\columnwidth}
   \centering
@@
 
-As we describe earlier, the user has the possibility of querying multiple mechanisms.
-We use MDPs to model such
+We use MDPs to model 
 interactive mechanisms. Specifically, external inputs are modeled by
@@
 In the figure, the states $+$,
-$-$, $Y$, and $N$ are interpreted as before. The actions $L$ and $H$
+$-$, $s$, and $t$ are interpreted as before. The actions $L$ and $H$
 denote low- and high-accuracy requests respectively. Let $M_H$ denote
 the Markov chain derived by high-accuracy requests.
-Observe that $1/5 \leq \Pr[M_H(x) = Y] \leq 4/5$ for any $x
-\in \mathcal{X}$. Hence $\Pr[M_H (x) = Y] \leq 4/5 = 4 \cdot 1/5 \leq
-4 \Pr[M_H (x') = Y]$ for any neighboring $x, x' \in
-\mathcal{X}$. Similarly, $\Pr[M_H (x) = N] \leq 4 \Pr[M_H (x') =
-N]$. The high-accuracy mechanism is $(\ln 4,0)$-differentially private.
-The privacy guarantees vary from accuracy requests.
+Note that the high-accuracy mechanism is $(\ln 4,0)$-differentially private.
+Unlike non-interactive mechanisms, the privacy guarantees vary from accuracy requests.
 
@@
   \textmd{for all } t \textmd{ with } s \neighbor{S} t \textmd{ and
-   scheduler } \scheduler{S},
-   \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq
+   query scheduler } \scheduler{Q},
+   \myPr{s}{M_{\scheduler{Q}}}{\neighbor{S}}{\phi} \leq
   e^{\epsilon} \cdot
@@
   & &
-   \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} + \delta
+   \myPr{t}{M_{\scheduler{Q}}}{\neighbor{S}}{\phi} + \delta
    \textmd{ and }
-   \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq e^{\epsilon} \cdot
-   \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} + \delta
+   \myPr{t}{M_{\scheduler{Q}}}{\neighbor{S}}{\phi} \leq e^{\epsilon} \cdot
+   \myPr{s}{M_{\scheduler{Q}}}{\neighbor{S}}{\phi} + \delta
 \end{eqnarray*}
@@
 for arbitrary schedulers. That is, no
-position-independent scheduler can force any of neighbors to have
+query scheduler can force any of neighbors to have
 probabilistically distinguishable behaviors.
@@
 \emph{Justification of the scheduler class.}
-For MDPs a richer class of scheduler classes have been considered in the literature, see \cite{BK:08:PMC}.
-It may depend on the full history of state and action sequences, and the decision can be randomised.
+For defining semantics for the $\mathcal{D}$-operator, we used query scheduler.
 If we use such scheduler, another definition of the differentially private operator
@@
 
-We consider only sequences $\scheduler{S}\in Act^\omega$, which correspond to the query sequences.
+We consider only query scheduler $\scheduler{Q}\in Act^\omega$, which correspond to the query sequences.
 In the differential privacy context, neighbor states exhibit same action behaviours.
@@
 our semantics agrees with the notion of differential privacy in the literature for such systems.
-Wpe hence
+We hence
 prefer the original definition.
 
-\subsection{Model Checking Algorithm}
+\subsection{The Model Checking Problem}
 For an MDP $M =
@@
 \dpriv{\epsilon}{\delta} \phi \}$ . Recall the semantics of $\dpriv{\epsilon}{\delta}
-\phi$. Given $s, t$ with $s \neighbor{M} t$ and a path formula
+\phi$. Given $s, t$ with $s \neighbor{S} t$ and a path formula
 $\phi$, we need to decide whether
-$\myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq
-e^{\epsilon} \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} + \delta$
-for every scheduler $\scheduler{S}$.
-If the path formula only contain next operators, it is easy to deal with.
-For $\phi:=\X B$ with $B\subseteq S$ only one step behaviour needs to be considered, thus only the first action in the query sequence is needed.
+$\myPr{s}{M_{\scheduler{Q}}}{\neighbor{S}}{\phi} \leq
+e^{\epsilon} \myPr{t}{M_{\scheduler{Q}}}{\neighbor{S}}{\phi} + \delta$
+for every query scheduler $\scheduler{Q}$.
+If the path formula only contains next operators, it is easy to deal with.
+When $\phi$ is $\gX B$ with $B\subseteq S$, only one step behaviour needs to be considered.
+Thus only the first action in the query sequence is needed.
 It can also be easily generalized to nested next operators:  one needs only to enumerate
 all actions query sequences of length obtained by the number of nexted $\X$ operators.
-In general, the problem is undecidable.
+The problem however is undecidable in general.
 

### 2018-01-31 proof.tex
adapted schedulers ect

@@
 
-First we assume $s_0\models \dpriv{1}{0} (F at_B)$. By $\dpCTLstar$ semantics we have that for all $\scheduler{S}\in\Sigma^\omega$,
-$\myPr{s_0}{M_{\scheduler{S}}}{\neighbor{S}}{F at_B} \leq
+First we assume $s_0\models \dpriv{1}{0} (F at_B)$. By $\dpCTLstar$ semantics we have that for all query scheduler $\scheduler{Q}\in\Sigma^\omega$,
+$\myPr{s_0}{M_{\scheduler{Q}}}{\neighbor{S}}{F at_B} \leq
   e \cdot
-  \myPr{s_0'}{M_{\scheduler{S}}}{\neighbor{S}}{F at_B}$. Observe $\myPr{s_0'}{M_{\scheduler{S}}}{\neighbor{S}}{F at_B}=0$ by construction, thus we have
-$\myPr{s_0}{M_{\scheduler{S}}}{\neighbor{S}}{F at_B} \leq
+  \myPr{s_0'}{M_{\scheduler{Q}}}{\neighbor{S}}{F at_B}$. Since the set of accepting state in the primed copy is empty, we have $\myPr{s_0'}{M_{\scheduler{Q}}}{\neighbor{S}}{F at_B}=0$, thus we have
+$\myPr{s_0}{M_{\scheduler{Q}}}{\neighbor{S}}{F at_B} \leq
  0$. This implies  $v_{s_0}\cdot M(\eta)\cdot (v_B)^c \leq 0$ for all $\eta\in\Sigma^*$.
@@
 Assume that $s_0\not\models \dpriv{1}{0} (F at_B)$.
-Since the relation $N_S$ contains only one pair, it implies that there exists a scheduler $\scheduler{S}\in\Sigma^\omega$ such that $\myPr{s_0}{M_{\scheduler{S}}}{\neighbor{S}}{F at_B} >0$.
+Since the relation $N_S$ contains only one pair, it implies that there exists a query scheduler $\scheduler{Q}\in\Sigma^\omega$ such that $\myPr{s_0}{M_{\scheduler{Q}}}{\neighbor{S}}{F at_B} >0$.
 It is then easy to construct a finite sequence $\eta\in\Sigma^*$ with  $v_{s_0}\cdot M(\eta)\cdot (v_B)^c > 0$, a contradiction.

### 2018-01-31 semantics.tex
adapted schedulers ect

@@
 Moreover, the relation $K, \neighbor{S}, \pi \models \phi$ is defined as
-for the standard LTL formulas~\cite{which-to-cite?}. We only recall the semantics for $\X$ and $\until$ as follows.
+for the standard LTL formulas~\cite{which-to-cite?}. We only recall the semantics for $\X$ and $\until$ as follows:
 \begin{eqnarray*}
@@
 } are computable within time polynomial in
-  $|S|$.\lz{and $|\wp|$?}
+  $|S|$.
   \label{lemma:PJ-subroutines}
@@
 The algorithm requires at most $O
-(|S|^2)$ additional steps. We hence have the following results:
+(|S|^2)$ additional steps. 
+We hence have the following results:
 
-\begin{algorithm}
+\begin{algorithm}[tbh]
   \begin{algorithmic}[1]

### 2018-01-31 specifications.tex
adapted schedulers ect

@@
 Obviously, it holds in states $+$ and $-$.
-A careful check ensures that one cannot decrease the value of $\epsilon_1=0$ or $\delta_1=.5$ without increasing the other.
+A careful check ensures that one cannot decrease  $\delta_1=.5$ without increasing $\epsilon$.
 
@@
 better privacy parameter $(0,0.5)$.
-% or $(-\infty,\frac{9}{16})$. \bw{use $-\infty$ instead of $\ln 0$. is it okay?}\lz{lets just remove it ;)}
 
@@
 
-Later we shall illustrate how reachability privacy can be used to specify differential privacy property for MDPs.
+Later we use reachability privacy to specify differential privacy property for MDPs.
 

### 2018-01-31 syntax.tex
adapted schedulers ect

@@
 As in the classical setting, we consider the sublogic $\dpCTL$ by allowing only path formulae of the form $\X\Phi$ and $\Phi\until\Phi$.
-Moreover, one obtains $\PCTL$~\cite{pHanssonJ94} and $\PCTLstar$~\cite{BiancoA95} from $\dpCTL$ and $\dpCTLstar$
+Moreover, one obtains $\PCTL$~\cite{HanssonJ94} and $\PCTLstar$~\cite{BiancoA95} from $\dpCTL$ and $\dpCTLstar$
 by removing the differentially private operator  $\dpriv{\epsilon}{\delta}$.

### 2018-01-31 examples.tex
adapted schedulers ect
small modifications of examples.tex

@@
 Now we describe our main ideas for  modelling differentially private mechanisms.
-A state $s$ stores at least the a database $x\in \mathcal{X}^n$, which is the data hold by a trusted \emph{curator}. A transition corresponds to the data analysis mechanisms.
+A state $s$ stores at least a database $x\in \mathcal{X}^n$, which is the data held by a trusted \emph{curator}. A transition corresponds to a data analysis mechanism.
 States are typically labelled with the corresponding outputs
@@
 mechanism and its corresponding Markov chain.
-The data universe
-$\mathcal{X}$ is $\{ +, - \}$. The mechanism $M$ is a randomized
-algorithm with inputs in $\mathcal{X}$ and outputs in $\{ Y, N
-\}$. In the figure, the states
+In the figure, the states
 $+$ and $-$ denote the diagnoses are positive or negative
@@
 	\end{algorithmic}
-	\caption{Input is a private database $d \in \{0,1,\ldots,5\}$, an adaptively chosen stream of sensitivity 1 queries $f_1,f_2$,\ldots, a threshold $t \in \{0,1,\ldots,5\}$.Output is a stream of answers $a_1,a_2$,\ldots}
+	\caption{Input is a private database $d \in \{0,1,\ldots,5\}$, an adaptively chosen stream of sensitivity 1 queries $f_1,f_2$,\ldots, a threshold $t \in \{0,1,\ldots,5\}$.
+Output is a stream of answers $a_1,a_2$,\ldots}
 	\label{algorithm:online-model}
@@
 In this setting it is obvious that queries do not change database themselves.
-To maintain differentially privacy, the curator however can decide to change data analysis mechanisms accordingly.
+To maintain differential privacy, the curator however can decide to change data analysis mechanisms accordingly.
 For instance, after a certain output is disclosed or a certain number of queries, it may decide to disable further queries.
 
-\paragraph{Online Model.}
 Below we describe an online model adapted from~\cite{DR:14:AFDP}: it is used in the scenario where a large amount of queries occur, but we only care for the queries which
 are above a given threshold.
-Other queries below the threshold are considered \emph{irrelevant} and might be discarded.
-In~\cite{DR:14:AFDP}, the model is set for continuous database, and the Laplace mechanism is applied to both queries and the threshold.
-In our model, the database is discrete and instead the geometric mechanism is applied.
+Other queries below the threshold are considered \emph{irrelevant} and won't cause privacy disclosure.
+In~\cite{DR:14:AFDP}, the model is set for continuous database, and the Laplace mechanisms are applied to both queries and the threshold.
+In our model, the database is discrete and instead the geometric mechanisms are applied.
 
 We consider the database set $D = \{0,1\ldots,5\}$, a threshold $t \in \{0,1\ldots,5\}$, and queries $\{f| f $ with sensitivity $\Delta (f) = 1 \}$.
-In order to protect privacy, truncated $\alpha$-geometric mechanisms over $\{0,1\ldots,5\}$ is applied to each query and the threshold, and what we observe is the first time that a query result is greater than or equal to the threshold after interference.
-We are interested in the differential privacy property of the data in $D$, assuming that a pair of data $d,d'$are neighbours, if $|d-d'|\le 1$.
-The algorithm
-is shown in Algorithm~\ref{algorithm:online-model}.
+In order to protect privacy, truncated $\alpha$-geometric mechanisms over $\{0,1\ldots,5\}$ are applied to each query and the threshold, and what we observe is the first time that a query result is greater than or equal to the threshold after interference.
+We are interested in the differentially private property of the data in $D$, assuming that a pair of data $d,d'$are neighbours, if $|d-d'|\le 1$.
+The algorithm is shown in Algorithm~\ref{algorithm:online-model} and a property will be specefied in Section~\ref{section:specifying-properties}.

### 2018-01-31 introduction.tex
add related work

@@
 
+\paragraph{Related Work.}
+An early on formal verification of differential privacy
+is~\cite{TKD:11:FVDPIS}. The work formalizes differential privacy in
+the framework of information leakage. The connection between
+differential privacy and information leakage is investigated
+in~\cite{AACDP:15:OILDPM,GMP:16:PDPFPS}. Type systems for differential
+privacy have been developed
+in~\cite{WHRP:17:FADP,GHHNP:13:LDPDP,RP:10:DMTGS}. A light-weight
+technique for checking differential privacy can be found
+in~\cite{ZK:17:LTADPP}. Lots of formal Coq proofs in differential
+privacy are reported
+in~\cite{BFGGHS:16:APCDP,BGGHS:16:PDPPC,BKOB:12:PRRDP,BDGKZ:13:VCDPASM,BGAHKS:14:PDPHL,BGAHRS:15:HOARRT,BFGAGHS:16:DPBP}. This
+work emphasizes on model checking differential privacy. We develop a
+framework to formalize and analyze differential privacy in Markov
+chains and Markov decision processes.
+
 \paragraph{Contributions.} Our main contributions are threefold. (I)

### 2018-01-31 proof.tex
notations

@@
 Assume that $s_0\not\models \dpriv{1}{0} (F at_B)$.
-Since the relation $N_S$ contains only one pair, it implies that there exists a query scheduler $\scheduler{Q}\in\Sigma^\omega$ such that $\myPr{s_0}{M_{\scheduler{Q}}}{\neighbor{S}}{F at_B} >0$.
+Since the relation $N_S=\{(s_0,s_0'),(s_0',s_0)\}$, there exists $(s_0,s_0')$,
+  and a query scheduler $\scheduler{Q}\in\Sigma^\omega$ such that 
+ \[
+ \myPr{s_0}{M_{\scheduler{Q}}}{\neighbor{S}}{F at_B} \not\leq
+  e \cdot
+  \myPr{s_0'}{M_{\scheduler{Q}}}{\neighbor{S}}{F at_B}
+ \]
+ which implies $\myPr{s_0}{M_{\scheduler{Q}}}{\neighbor{S}}{F at_B} >0$.
 It is then easy to construct a finite sequence $\eta\in\Sigma^*$ with  $v_{s_0}\cdot M(\eta)\cdot (v_B)^c > 0$, a contradiction.

### 2018-01-31 specifications.tex
notations

@@
 The state $s_{34}$ (with the input threshold $t=3$, and the database $d=4$) satisfies $\Phi$ by our computation.
-It means the probability of reaching $\top$  within $i$ steps with $i$ ranges over $\{1,\ldots,6\}$, is $(3, 0)$-differentially private for the database $d$ and its neighbours ($d=3$ and $d=5$).
+It means the probability of reaching $\top$ at $i$ steps with $i$ ranges over $\{1,\ldots,6\}$, is $(3, 0)$-differentially private for the database $d$ and its neighbours ($d=3$ and $d=5$).
 It also states that the probability of reaching $\top$ after no less than $6$ steps, is smaller than $0.025$ and is  $(3, 0.025)$-differentially private.
@@
 continuous time out (say, in the denial of service attack) is
-$(\epsilon, \delta)$-close for all neighbors. That is, adding a slave
+$(\epsilon, \delta)$-close for all neighbors. That is, replacing a slave
 device cannot increase the chance of service failure significantly.

### 2018-01-31 introduction.tex
fix short sentences

@@
 implementations with our techniques. Most interestingly, our
-techniques can be adopted rather easily by existing probabilistic
-model checkers. Privacy checking with existing tools can be attained
-with minimal efforts and overheads. Hopefully, more interaction
-between model checking and privacy analysis will follow.
+techniques can be adopted easily by existing probabilistic
+model checkers. Privacy checking with existing tools is attainable
+with minimal efforts. More interaction
+between model checking and privacy analysis hopefully will follow.
 
@@
 Markov decision processes. Our formalization effectively assumes that
-actions are controlled by adversaries. It is subsequently necessary to
-consider attacks from all possible adversaries in order to establish
+actions are controlled by adversaries. It is thus necessary to
+consider all privacy attacks from adversaries in order to establish
 differential privacy.
@@
 those of $\PCTLstar$ for Markov chains. The logic $\dpCTLstar$ obtains its
-essentially expressiveness for free.
-For Markov decision processes, we show that checking whether a state
+expressiveness essentially for free.
+For Markov decision processes, checking whether a state
 satisfies $\dpriv{\epsilon}{\delta} \phi$ is undecidable.
@@
 differential privacy properties, and moreover, complex properties such as those
-involving infinite behaviors can also be expressed. (II) We model several differential privacy mechanisms, including stream processing, in Markov chains or Markov decision processes.
+involving infinite behaviors can also be expressed. (II) We model several differentially private mechanisms, including stream processing, in Markov chains or Markov decision processes.
 (III) We show that the model checking 

### 2018-01-31 main.tex
merge comments
adapted schedulers ect
fix short sentences

@@
 Preliminaries are given in Section~\ref{section:preliminaries}.
-In Section~\ref{section:examples} we discuss how differential privacy mechanisms can be modeled in Markov chains.
+In Section~\ref{section:examples} we discuss how differentially private mechanisms can be modeled in Markov chains.
 The logic $\dpCTLstar$ and model checking algorithms for Markov chains are presented in
@@
 
-\subsection{MDPs for Streaming Processes}
+\subsection{MDPs for Stream Processing}
 \label{section:applications}
@@
 We have introduced $\dpCTLstar$ to reason about properties in differential privacy, and have investigated the model checking problem.
-For Markov chains, the model checking problem has polynomial complexity.
+For Markov chains, the model checking problem has the same complexity as for $\PCTLstar$.
 The general MDP model checking problem is undecidable.

### 2018-01-31 introduction.tex
use noindent

@@
 
-\paragraph{Related Work.}
+\noindent
+\emph{Related Work.}
 An early on formal verification of differential privacy
@@
 
-\paragraph{Contributions.} Our main contributions are threefold. (I)
+\noindent
+\emph{Contributions.} Our main contributions are threefold. (I)
 We introduce the logic $\dpCTLstar$ for reasoning about properties in

### 2018-01-31 applications.tex
Correct some spelling errors and others

@@
 neighboring observations. In the following, we model a discrete
-variant of a stream processing mechanisms by MDPs. We will also
+variant of a stream processing mechanism by MDPs. We will also
 specify pan privacy in our model using $\dpCTLstar$.
@@
 $t_{10}$. If the action $a[a]$ or $b[a]$ are chosen, they
-denote that the current event type is replaced with the eveny type
+denote that the current event type is replaced with the event type
 $a$. The MDP moves to the internal state

### 2018-01-31 examples.tex
notations
Correct some spelling errors and others

@@
 Similar to the survey mechanism, it is
-straightfoward to model the truncated $\frac{1}{2}$-geometric
+straightforward to model the truncated $\frac{1}{2}$-geometric
 mechanisms as Markov chains.
@@
 	\begin{algorithmic}[1]
-		\Function{OnlineModel}{$d$, $t$}
+		\Procedure{AboveThreshold}{$d$, $t$}
 		\Match{the input threshold $t$}
@@
 		\EndFor
-		\EndFunction
+		\EndProcedure
 
@@
 
-\subsection{Above Threshold Mechanisms}
+\subsection{Above Threshold Mechanism}
 In the differentially privacy context, we differentiate between \emph{offline} and \emph{online} models.
@@
 In order to protect privacy, truncated $\alpha$-geometric mechanisms over $\{0,1\ldots,5\}$ are applied to each query and the threshold, and what we observe is the first time that a query result is greater than or equal to the threshold after interference.
-We are interested in the differentially private property of the data in $D$, assuming that a pair of data $d,d'$are neighbours, if $|d-d'|\le 1$.
+We are interested in the differentially private property of the data in $D$, assuming that a pair of data $d,d'$are neighbors, if $|d-d'|\le 1$.
 The algorithm is shown in Algorithm~\ref{algorithm:online-model} and a property will be specefied in Section~\ref{section:specifying-properties}.

### 2018-01-31 introduction.tex
bib
Correct some spelling errors and others

@@
 analysis of differential privacy on both designs and implementations.
-In tis paper, we propose the logic $\dpCTLstar$ for specifying
+In this paper, we propose the logic $\dpCTLstar$ for specifying
 differential privacy and investigate their model checking
@@
 with minimal efforts. More interaction
-between model checking and privacy analysis hopefully will follow.
+between model checking~\cite{ManneP92,BK:08:PMC} and privacy analysis hopefully will follow.
 
@@
 but randomized algorithms. We follow the standard practice in
-probabilistic model checking to formalize such mechanisms by
+probabilistic model checking to formalize such mechanisms
 by Markov chains or Markov decision processes~\cite{Put05}. We give
@@
 \emph{Related Work.}
-An early on formal verification of differential privacy
+An early attempt on formal verification of differential privacy
 is~\cite{TKD:11:FVDPIS}. The work formalizes differential privacy in

### 2018-01-31 specifications.tex
Correct some spelling errors and others

@@
 $\myPr{s}{K}{N}{\X \mathit{out}_Y}\le 3\myPr{t}{K}{N}{\X \mathit{out}_Y}$.
-The formula $\dpriv{\ln 3}{0} (\X \mathit{out}_Y)\wedge \dpriv{\ln 3}{0} (\X \mathit{out}_N)$ thus specifies the differential privacy property for states $+$ and $-$ in Example~\ref{exa:survey}.
+The formula $\dpriv{\ln 3}{0} (\X \mathit{out}_Y)\wedge \dpriv{\ln 3}{0} (\X \mathit{out}_N)$ thus specifies the differentially private property for states $+$ and $-$ in Example~\ref{exa:survey}.
 The survey mechanism is
@@
 $\delta_2=2\delta_1=1$. However, we can check easily that one gets
-better privacy parameter $(0,0.5)$.
+better privacy parameter $(0,.5)$.
 
@@
 
-As an example, for our threshold model (Algorithm~\ref{algorithm:online-model}), we consider the property $\Phi=\wedge_{i=1}^{6} \dpriv{3}{0}(\X^i \top) \wedge \PJ{\le 0.025}(\X^6 F \top) \wedge \dpriv{3}{0.025}(\X^6 F \top)$, to express the privacy.
+As an example, for our threshold model (Algorithm~\ref{algorithm:online-model}), we consider the property $\Phi=\wedge_{i=1}^{6} \dpriv{3}{0}(\X^i \top) \wedge \PJ{\le .025}(\X^6 F \top) \wedge \dpriv{3}{.025}(\X^6 F \top)$, to express the privacy.
 The state $s_{34}$ (with the input threshold $t=3$, and the database $d=4$) satisfies $\Phi$ by our computation.
 It means the probability of reaching $\top$ at $i$ steps with $i$ ranges over $\{1,\ldots,6\}$, is $(3, 0)$-differentially private for the database $d$ and its neighbours ($d=3$ and $d=5$).
-It also states that the probability of reaching $\top$ after no less than $6$ steps, is smaller than $0.025$ and is  $(3, 0.025)$-differentially private.
+It also states that the probability of reaching $\top$ after no less than $6$ steps, is smaller than $.025$ and is  $(3, .025)$-differentially private.
 
@@
 Consider the Bluetooth example in~\cite{DKNP04} and the
-formula $\dpriv{\epsilon}{\delta}{\F \G \mathit{timeout}}$. Let us say
+formula $\dpriv{\epsilon}{\delta}{(\F \G \mathit{timeout})}$. Let us say
 two states are neighbors if a slave device in a state is replaced by

### 2018-01-31 syntax.tex
notations
Correct some spelling errors and others

@@
-The syntax of $\dpCTL$ state and path formulae is given by:
+The syntax of $\dpCTLstar$ state and path formulae is given by:
 \begin{eqnarray*}
@@
 A \emph{state} formula $\Phi$ is either  an atomic proposition
-$p$, the negation of a sate formula, the conjunction of two state
+$p$, the negation of a state formula, the conjunction of two state
 formulae, a \emph{probabilistic} operator $\PJ{J}$ with $J$

### 2018-02-01 dpCTL.tex
minor fixes

@@
 from $s$ is close to the probabilities of
-having $\phi$-paths from its neighboring states.
+having $\phi$-paths from its neighbors.
 %We give semantics for both Markov chains and Markov decision processes.

### 2018-02-01 examples.tex
minor fixes

@@
 
-\subsection{Survey Mechanisms}
+\subsection{Survey Mechanism}
 \label{subsec:survey}
@@
 she randomly answers \textit{Yes} or \textit{No}
-uniformly~\cite{DR:14:AFDP}.
+uniformly (Figure~\ref{figure:2-dp-table})~\cite{DR:14:AFDP}.
 
@@
 = N]$. The survey mechanism is $(\ln 3, 0)$-differentially private.
+The random noise boosts the probability of answering
+\textit{Yes} or \textit{No} to at least $1/4$ regardless of
+diagnoses. Inferences on individual diagnosis can be plausibly denied.
 
@@
 
-\begin{example}\label{exa:survey}
-  Figure~\ref{figure:2-dp} shows the survey
-mechanism and its corresponding Markov chain.
+Figure~\ref{figure:2-dp-mdp} shows the corresponding Markov chain.
 In the figure, the states
@@
 States $+$ and $-$ are neighbors.
-The random noise boosts the probability of answering
-\textit{Yes} or \textit{No} to at least $1/4$ regardless of
-diagnoses. Inferences on individual diagnosis can be plausibly denied.
 Missing transitions (such as those from $s$ and $t$) lead
@@
 and the state $\dagger$ for clarity.
-\end{example}
 
@@
 output distribution on any inputs $x, x'$ with $f (x) = f
-(x')$. Moreover, it is $(- {\Delta (f)} \ln \alpha, 0)$-differentially
+(x')$. It is $(- {\Delta (f)} \ln \alpha, 0)$-differentially
 private for any query $f$ with sensitivity $\Delta (f)$. The range of
@@
     \]
-    \caption{$1/2$-Geometric Mechanism}
+    \caption{$\frac{1}{2}$-Geometric Mechanism}
     \label{figure:geometric-mechanism-table}
@@
 straightforward to model the truncated $\frac{1}{2}$-geometric
-mechanisms as Markov chains.
-
-\begin{example}
-Let the state $s_k$ and $t_l$
+mechanisms as Markov chains. Let the state $s_k$ and $t_l$
 denote the input value $k$ and output value $l$ respectively. Define
@@
 $\frac{1}{2}$-geometric mechanism over $\{ 0, 1, \ldots, 5 \}$.
-\end{example}
 

### 2018-02-01 introduction.tex
minor fixes

@@
 technique for checking differential privacy can be found
-in~\cite{ZK:17:LTADPP}. Lots of formal Coq proofs in differential
+in~\cite{ZK:17:LTADPP}. Lots of formal Coq proofs about differential
 privacy are reported
@@
 (III) We show that the model checking
-problem for Markov chains is standard. For Markov decision processes we show that it is undecidable.
+problem for Markov chains is standard. For Markov decision processes, we show that it is undecidable.

### 2018-02-01 semantics.tex
bib
Correct some spelling errors and others
minor fixes

@@
 Moreover, the relation $K, \neighbor{S}, \pi \models \phi$ is defined as
-for the standard LTL formulas~\cite{which-to-cite?}. We only recall the semantics for $\X$ and $\until$ as follows:
+for the standard LTL formulas~\cite{ManneP92}. We only recall the semantics for $\X$ and $\until$ as follows:
 \begin{eqnarray*}
@@
 recall that  $\myPr{s}{K}{N}{\phi}$ is the probability of having
-paths satisfying $\phi$ from $s$. Thus, a state $s$ satisfies
+paths satisfying $\phi$ from $s$. A state $s$ satisfies
 $\dpriv{\epsilon}{\delta} \phi$ if having paths satisfying $\phi$
 from $s$ is $(\epsilon, \delta)$-close to having paths satisfying
-$\phi$ from every neighbor of $s$. In other words, it is
+$\phi$ from every neighbor of $s$. Informally, it is
 probabilistically similar to observe paths satisfying $\phi$ from $s$
-and from all neighbors of $s$.
+and from all its neighbors.
 
 \subsection{Model Checking Algorithms}
-We first describe the model checking algorithm for $\dpCTL$, which follows the classical
+We describe the model checking algorithm for $\dpCTL$. The algorithm follows the classical
 algorithms for $\PCTL$ by computing the states satisfying
@@
 consider the inductive step where the states satisfying the subformula
-$\dpriv{\epsilon}{\delta} (\phi)$ is computed.
+$\dpriv{\epsilon}{\delta} (\phi)$ are to be computed.
 
@@
 computing $\myPr{s}{K}{\neighbor{S}}{\phi}$ for state $s \in S$.
-These probabilities can be computed by solving linear equations or
+These probabilities can be obtained by solving linear equations or
 through iterative approximations. We summarize it in the following
@@
 In order to determine the set $\{ s : K, \neighbor{S}, s \models
-\dpriv{\epsilon}{\delta} \varphi \}$, our algorithm first computes
+\dpriv{\epsilon}{\delta} \varphi \}$, our algorithm computes
 the probabilities $p (s) = \myPr{s}{K}{\neighbor{S}}{\varphi}$ for every
@@
 $p (t)$ are not $(\epsilon, \delta)$-close, the state $s$
-is removed from the result. Hence
+is removed from the result. 
 Algorithm~\ref{algorithm:sat-dpriv-dtmc} returns all states which are
@@
   \begin{algorithmic}[1]
-    \Function{SAT}{$K$, $\neighbor{S}$, $\phi$}
+    \Procedure{SAT}{$K$, $\neighbor{S}$, $\phi$}
     \Match{$\phi$}
@@
     \For{$s \in S$}
-      \For{$t$ with $s \neighbor{K} t$}
+      \For{$t$ with $s \neighbor{S} t$}
         \If{$p(s) \not\leq e^{\epsilon} p(t) + \delta$ or
@@
     \State{\Return $R$}
-    \EndFunction
+    \EndProcedure
   \end{algorithmic}
-  \caption{SAT($\phi$, $\neighbor{K}$)}
+  \caption{SAT($K$, $\neighbor{S}$, $\phi$)}
   \label{algorithm:sat-dpriv-dtmc}

### 2018-02-01 specifications.tex
minor fixes

@@
 \paragraph{Standard Differential Privacy.}
-Consider the survey mechanism (Subsection~\ref{subsec:survey}).
-The formula $\dpriv{\ln 3}{0} (\X \mathit{out}_Y)$ holds in state $s$ if for  $t$ with $sNt$, the probability of satisfying $\X \mathit{out}_Y$ from $s$ and $t$ satisfy:
-$\myPr{s}{K}{N}{\X \mathit{out}_Y}\le 3\myPr{t}{K}{N}{\X \mathit{out}_Y}$.
-The formula $\dpriv{\ln 3}{0} (\X \mathit{out}_Y)\wedge \dpriv{\ln 3}{0} (\X \mathit{out}_N)$ thus specifies the differentially private property for states $+$ and $-$ in Example~\ref{exa:survey}.
-The survey mechanism is
-$(\ln 3, 0)$-differentially private.
+Consider the survey mechanism (Section~\ref{subsec:survey}).
+The formula $\dpriv{\ln 3}{0} (\X \mathit{out}_Y)$ holds in state $u$ if for  $v$ with $u N v$, the probability of satisfying $\X \mathit{out}_Y$ from $u$ and $v$ satisfy:
+$\myPr{u}{K}{N}{\X \mathit{out}_Y}\le 3\myPr{v}{K}{N}{\X \mathit{out}_Y}$.
+The formula $\dpriv{\ln 3}{0} (\X \mathit{out}_Y)\wedge \dpriv{\ln 3}{0} (\X \mathit{out}_N)$ thus specifies differential privacy for states $+$ and $-$.
+The survey mechanism is $(\ln 3, 0)$-differentially private.
 
@@
 
- Define the formula
+For the $\frac{1}{2}$-geometric mechanism (Section~\ref{subsec:geometric}),
+define the formula
 $\psi = \dpriv{\ln 2}{0} (\X out_0) \wedge \dpriv{\ln 2}{0} (\X out_1) \wedge
 \cdots \wedge \dpriv{\ln 2}{0} (\X out_5)$. If the state $s_k$ satisfies
-$\psi$ for $k = 0, \ldots, 5$, then the $\frac{1}{2}$-geometric mechanism (Subsection~\ref{subsec:geometric}) is
+$\psi$ for $k = 0, \ldots, 5$, then the $\frac{1}{2}$-geometric mechanism is
 $(\ln 2, 0)$-differentially private.
 
-\paragraph{Compositional Properties.}
-Compositional aspect is one of the building block for differentially private algorithms.
+\paragraph{Compositionality.}
+Compositionality is one of the building block for differential privacy.
 According to the compositional theorem \cite[Theorem 3.16]{DR:14:AFDP}, for $(\epsilon_1,\delta_1)$-differentially private algorithm $M_1$ and
 $(\epsilon_2,\delta_2)$-differentially private algorithm $M_2$, their combination $M$ defined by $M(x)=(M_1(x), M_2(x))$ is $(\epsilon_1 + \epsilon_2,\delta_1+\delta_2)$-differentially private.
-The degradation is rooted in the repeated computation.
-To illustrate this property, we consider the extended survey example below where the mechanism allows two consequent queries.
+The degradation is rooted in the repeated releases of information.
+To illustrate this property, we consider the extended survey mechanism which allows two consequent queries.
 The model is depicted in Figure~\ref{figure:2-dp-mc}.
@@
 
-We consider the formula $\dpriv{\ln 9}{0} (\X (\mathit{out}_Y\wedge \X \mathit{out}_Y))$.
+Consider the formula $\dpriv{\ln 9}{0} (\X (\mathit{out}_Y\wedge \X \mathit{out}_Y))$.
 A path satisfies $\X (\mathit{out}_Y\wedge \X \mathit{out}_Y)$ if the second state satisfies $\mathit{out}_Y$ and the third state satisfies $\mathit{out}_Y$ as well.
@@
 Moreover, the bound $\epsilon=\ln 9$ is tight, since the probability of satisfying $\X (\mathit{out}_Y\wedge \X \mathit{out}_Y)$ from states $+$ and $-$ are $\frac{9}{16}$ and $\frac{1}{16}$, respectively.
-Finally, the formula $\wedge_{a_1,a_2}\dpriv{\ln 9}{0} (\X (a_1\wedge \X a_2))$ corresponds to the desired compositional property for the model, where $a_1,a_2$ range over all atomic propositions $\{\mathit{out}_Y, \mathit{out}_N\}$.
+Finally, the formula $\wedge_{a_1,a_2}\dpriv{\ln 9}{0} (\X (a_1\wedge \X a_2))$ specifies differential privacy for the model, where $a_1,a_2$ range over all atomic propositions $\{\mathit{out}_Y, \mathit{out}_N\}$.
 
-We consider two other formulae for comparision:
-(a) Formula $\dpriv{\ln 3}{0} (\X \X \mathit{out}_Y)$.
+In comparison, let us consider two slightly different formulae:
+(a) $\dpriv{\ln 3}{0} (\X \X \mathit{out}_Y)$.
 In this case we do not have privacy loss, even though we are querying twice: the reason is that the output of the first query is not utilized at all.
 It is easy to verify that it is indeed satisfied by $+$ and $-$.
-(b) Formula $\dpriv{\ln 3}{0} (\X (\mathit{out}_Y\wedge \dpriv{\ln 3}{0}(\X \mathit{out}_Y)))$.
+(b) $\dpriv{\ln 3}{0} (\X (\mathit{out}_Y\wedge \dpriv{\ln 3}{0}(\X \mathit{out}_Y)))$.
 This is a nested $\dpCTL$ formula, where the inner state formula $\dpriv{\ln 3}{0}(\X \mathit{out}_Y)$ states the one step differential privacy.
-Observe the inner formula is satisfied by all neighbours.
+Observe the inner formula is satisfied by all neighbors.
 Thus, the outer formula also has no privacy loss.
@@
 \paragraph{Tighter Privacy Bounds for Composition.}
-One of the advantage of applying model checking is that we may get tighter bounds for composition.
-Consider our survey example, and the property $\dpriv{0}{.5} (\X \mathit{out}_Y)$.
+An advantage of applying model checking is that we may get tighter bounds for composition.
+Consider the survey mechanism, and the property $\dpriv{0}{.5} (\X \mathit{out}_Y)$.
 Obviously, it holds in states $+$ and $-$.
 A careful check ensures that one cannot decrease  $\delta_1=.5$ without increasing $\epsilon$.
-
-Now we consider the formula $\dpriv{\epsilon_2}{\delta_2} (\X
-(\mathit{out}_Y\wedge \X \mathit{out}_Y))$ under the model in
+Now consider the formula $\dpriv{\epsilon_2}{\delta_2} (\X
+(\mathit{out}_Y\wedge \X \mathit{out}_Y))$ on the model in
 Figure~\ref{figure:2-dp-mc}. Applying the
@@
 \paragraph{Reachability Privacy.}
-The formula $\dpriv{\epsilon}{\delta}(F \top)$ is a reachability privacy: it states that related states should have similar probabilities of eventually reaching $\top$.
+The formula $\dpriv{\epsilon}{\delta}(F \top)$ is a reachability privacy.
+It specifies that neighbors should have similar probabilities of eventually reaching $\top$.
 Reachability privacy is often used in combination with other properties.
-
-As an example, for our threshold model (Algorithm~\ref{algorithm:online-model}), we consider the property $\Phi=\wedge_{i=1}^{6} \dpriv{3}{0}(\X^i \top) \wedge \PJ{\le .025}(\X^6 F \top) \wedge \dpriv{3}{.025}(\X^6 F \top)$, to express the privacy.
+In the above threshold mechanism in Section~\ref{subsec:threshold},
+consider the property $\Phi=\wedge_{i=1}^{6} \dpriv{3}{0}(\X^i \top) \wedge \PJ{\le .025}(\X^6 F \top) \wedge \dpriv{3}{.025}(\X^6 F \top)$.
 The state $s_{34}$ (with the input threshold $t=3$, and the database $d=4$) satisfies $\Phi$ by our computation.
@@
 
-Later we use reachability privacy to specify differential privacy property for MDPs.
+%Later we use reachability privacy to specify differential privacy property for MDPs.
 

### 2018-02-01 mdp.tex
Correct some spelling errors and others
minor fixes

@@
 answers are requested, the surveyee answers \textit{Yes} with
-probability $4/5$ and \textit{No} with probability $1/5$ when she has
-positive diagnosis. She answers \textit{Yes} with probability $1/5$
-and \textit{No} with probability $4/5$ when she is not
+probability $\frac{4}{5}$ and \textit{No} with probability $\frac{1}{5}$ when she has
+positive diagnosis. She answers \textit{Yes} with probability $\frac{1}{5}$
+and \textit{No} with probability $\frac{4}{5}$ when she is not
 diagnosed with the disease X. This gives an interactive mechanism
@@
 \delta)$-close to all its neighbors in having paths satisfying $\phi$
-for arbitrary schedulers. That is, no
+for query schedulers. That is, no
 query scheduler can force any of neighbors to have
@@
 For defining semantics for the $\mathcal{D}$-operator, we used query scheduler.
-If we use such scheduler, another definition of the differentially private operator
+If we use arbitrary scheduler, another definition of the differentially private operator
 might be:
@@
 
+The proof is put in Appendix.
 We consider some decidable special cases.

### 2018-02-01 applications.tex
finalize

@@
 
-In pan privacy, data analysis mechanisms are required to respect
+In pan privacy, mechanisms are required to respect
 privacy even in their internal states~\cite{DNPR:10:DPCO,DNPRY:10:PPSA}.
@@
 neighboring observations. In the following, we model a discrete
-variant of a stream processing mechanism by MDPs. We will also
-specify pan privacy in our model using $\dpCTLstar$.
+variant of a stream processing mechanism by an MDP and
+specify pan privacy using $\dpCTLstar$.
 
@@
 A sequence of event types is said to be an observation. Two
-obsevations are {neighbors} if one can be obtained from the other
+observations are {neighbors} if one can be obtained from the other
 by replacing event types any number of times, that is, if they are of
@@
 Recall $2\ln 2 = \ln 4$,
-the steaming data analysis mechanism in
+the steaming mechanism in
 Algorithm~\ref{algorithm:frequency-estimator} is $(\ln 4,
@@
 a[a], a[b], b[a], b[b] \}$. Informally, the action $x[y]$ means the
-event type $x$ is replaced by the event type $y$ in a neighbor. An
+event type $x$ is replaced by the event type $y$. An
 action sequence thus represents an observation and one of its
-neighbors and vice versa. More formally, the action sequence
+neighbors. Concretely, 
 $x_1[y_1]x_2[y_2]\cdots x_n[y_n]$ represents the observation
@@
 states. For observations of length $k$, the probabilities of reaching
-each internal states need to be sufficiently close. The state $I$
-therefore must satisfy
+each internal states need to be sufficiently close. Define
 \[
-\Psi
-\wedge
-\bigwedge_{k \in \bbfN}
-\dpriv{\ln 4}{0} (\X^k {00}) \wedge
-\dpriv{\ln 4}{0} (\X^k {01}) \wedge
-\dpriv{\ln 4}{0} (\X^k {10}) \wedge
-\dpriv{\ln 4}{0} (\X^k {11}).
+\Psi_i =
+\dpriv{\ln 4}{0} (\X^i {00}) \wedge
+\dpriv{\ln 4}{0} (\X^i {01}) \wedge
+\dpriv{\ln 4}{0} (\X^i {10}) \wedge
+\dpriv{\ln 4}{0} (\X^i {11}).
 \]
+The state $s$ therefore must satisfy $\Psi$ and $\Psi_i$ for every $i
+\in \bbfZ^{\geq 0}$.

### 2018-02-01 examples.tex
minor fixes
finalize

@@
+To model differentially private mechanisms by Markov chains, we
+formalize inputs (such as databases or query results) as states. For
+each input, its output distribution is modeled as a probabilistic
+transition. Finally, neighbors are specified by a neighborhood
+relation. States are typically labelled with atomic propositions to
+signify their intended interpretation (such as inputs or outputs). We
+demonstrate the key ideas in examples.
+
+\hide{
 A simple way to design differentially private mechanisms is to add
@@
 the curator generates. We explain the main idea with some examples.
+}
 
@@
 The random noise boosts the probability of answering
-\textit{Yes} or \textit{No} to at least $1/4$ regardless of
+\textit{Yes} or \textit{No} to at least $\frac{1}{4}$ regardless of
 diagnoses. Inferences on individual diagnosis can be plausibly denied.
@@
 	\begin{algorithmic}[1]
-		\Procedure{AboveThreshold}{$d$, $t$}
-		\Match{the input threshold $t$}
-		\Comment{Assignments of $t'$ by $\frac{1}{4}$-geometric mechanism}
+		\Procedure{AboveThreshold}{$d$, $\{ f_1, f_2, \ldots
+                  \}$, $t$}
+		\Match{$t$}
+		\Comment{obtain $t'$ by $\frac{1}{4}$-geometric mechanism}
 		\lCase{$0$}{$t' \leftarrow 0, 1,2,3,4,5$ with probability
@@
 		\For{each query $f_i$}
-		\Match{the input databese $d$}
-		\Comment{Assignments of $f_{i}'$ by $\frac{1}{2}$-geometric mechanism}
-		\lCase{$0$}{$f_{i}' \leftarrow 0, 1,2,3,4,5$ with probability
+                \State{$r_i \leftarrow f_i (d)$}
+		\Match{ $r_i$}
+		\Comment{obtain $r_{i}'$ by $\frac{1}{2}$-geometric mechanism}
+		\lCase{$0$}{$r_{i}' \leftarrow 0, 1,2,3,4,5$ with probability
 			$\frac{2}{3}$,$\frac{1}{6}$,
@@
 			$\frac{1}{48}$,$\frac{1}{48}$ respectively}
-		\lCase{$1$}{$f_{i}' \leftarrow 0, 1,2,3,4,5$ with probability
+		\lCase{$1$}{$r_{i}' \leftarrow 0, 1,2,3,4,5$ with probability
 			$\frac{1}{3}$,$\frac{1}{3}$,
@@
 			$\frac{1}{24}$,$\frac{1}{24}$ respectively}
-		\lCase{$2$}{$f_{i}' \leftarrow 0, 1,2,3,4,5$ with probability
+		\lCase{$2$}{$r_{i}' \leftarrow 0, 1,2,3,4,5$ with probability
 			$\frac{1}{6}$,$\frac{1}{6}$,
@@
 			$\frac{1}{12}$,$\frac{1}{12}$ respectively}
-		\lCase{$3$}{$f_{i}' \leftarrow 0, 1,2,3,4,5$ with probability
+		\lCase{$3$}{$r_{i}' \leftarrow 0, 1,2,3,4,5$ with probability
 			$\frac{1}{12}$,$\frac{1}{12}$,
@@
 			$\frac{1}{6}$,$\frac{1}{6}$ respectively}
-		\lCase{$4$}{$f_{i}' \leftarrow 0, 1,2,3,4,5$ with probability
+		\lCase{$4$}{$r_{i}' \leftarrow 0, 1,2,3,4,5$ with probability
 			$\frac{1}{24}$,$\frac{1}{24}$,
@@
 			$\frac{1}{3}$,$\frac{1}{3}$ respectively}
-		\lCase{$5$}{$f_{i}' \leftarrow 0, 1,2,3,4,5$ with probability
+		\lCase{$5$}{$r_{i}' \leftarrow 0, 1,2,3,4,5$ with probability
 			$\frac{1}{48}$,$\frac{1}{48}$,
@@
 		\EndMatch
-		\If{$f_{i}'>t'$}
+                \State{\textbf{if} $r'_i \geq t'$ \textbf{then} 
+                  \textbf{halt} with $a_i = \top$ \textbf{else}
+                  $a_i = \bot$}
+\hide{
+		\If{$r_{i}'>t'$}
 		\State{$a_i=\top$}
-		\State{\Return}
+		\State{\textbf{halt}}
 		\Else
@@
 		\EndIf
+}
 		\EndFor
@@
 	\end{algorithmic}
-	\caption{Input is a private database $d \in \{0,1,\ldots,5\}$, an adaptively chosen stream of sensitivity 1 queries $f_1,f_2$,\ldots, a threshold $t \in \{0,1,\ldots,5\}$.
-Output is a stream of answers $a_1,a_2$,\ldots}
+	\caption{Input: private database $d$, 
+          queries $f_i : \{ 0, 1, \ldots, 5 \} \rightarrow \{ 0, 1,
+          \ldots, 5 \}$ with sensitivity $1$, threshold $t \in \{ 0,
+          1, \ldots, 5 \}$; Output: $a_1, a_2, \ldots$}
 	\label{algorithm:online-model}
@@
 \subsection{Above Threshold Mechanism}
-In the differentially privacy context, we differentiate between \emph{offline} and \emph{online} models.
-In the offline (or \emph{non-interactive}) model the curator releases the outputs only once and plays no further role.
-Thus the models we constructed in previous section are offline models.
-On the contrary, in the online (or \emph{interactive}) model the curator can permit further queries.
-In this setting it is obvious that queries do not change database themselves.
-To maintain differential privacy, the curator however can decide to change data analysis mechanisms accordingly.
-For instance, after a certain output is disclosed or a certain number of queries, it may decide to disable further queries.
+\label{subsec:threshold}
+In differential privacy, we differentiate between {offline} and
+{online} mechanisms. 
+An offline mechanism releases outputs only once and plays no further
+role. The mechanisms we constructed previously are offline mechanisms. 
+On the other hand, online mechanisms permit further queries. 
+Obviously, queries do not change databases themselves. To maintain
+privacy, online mechanisms may decide to release
+information differently. For instance, they may disable further
+queries after a certain output is disclosed.
 
-Below we describe an online model adapted from~\cite{DR:14:AFDP}: it is used in the scenario where a large amount of queries occur, but we only care for the queries which
-are above a given threshold.
-Other queries below the threshold are considered \emph{irrelevant} and won't cause privacy disclosure.
-In~\cite{DR:14:AFDP}, the model is set for continuous database, and the Laplace mechanisms are applied to both queries and the threshold.
-In our model, the database is discrete and instead the geometric mechanisms are applied.
+Below we describe an online mechanism adapted from~\cite{DR:14:AFDP}:
+it is used in the scenario where a large amount of queries occur, but
+we only care for the queries which 
+are above a given threshold. Queries below the threshold are
+considered {irrelevant} and only disclose minimal information.
+In~\cite{DR:14:AFDP}, the mechanism is design for continuous queries.
+The Laplace mechanism is applied. Here, our mechanism is designed for
+bounded and discrete queries. The geometric mechanism is applied instead.
 
-We consider the database set $D = \{0,1\ldots,5\}$, a threshold $t \in \{0,1\ldots,5\}$, and queries $\{f| f $ with sensitivity $\Delta (f) = 1 \}$.
-In order to protect privacy, truncated $\alpha$-geometric mechanisms over $\{0,1\ldots,5\}$ are applied to each query and the threshold, and what we observe is the first time that a query result is greater than or equal to the threshold after interference.
-We are interested in the differentially private property of the data in $D$, assuming that a pair of data $d,d'$are neighbors, if $|d-d'|\le 1$.
-The algorithm is shown in Algorithm~\ref{algorithm:online-model} and a property will be specefied in Section~\ref{section:specifying-properties}.
+Consider a threshold $t \in
+\{ 0, 1 \ldots, 5\}$, and queries $\{ f_i : f_i $ with sensitivity $\Delta
+(f_i) = 1 \}$. In order to protect privacy, our mechanism applies the
+truncated $\frac{1}{4}$-geometric mechanism to obtain a noisy
+threshold $t'$. For each query, the truncated $\frac{1}{2}$-geometric
+mechanism is applied to its result. If the noisy result is not less
+than the noisy threshold, our mechanism outputs $\top$ and
+halts. Otherwise, it outputs $\bot$ and continues to the next query
+(Algorithm~\ref{algorithm:online-model}). The above threshold
+mechanism outputs a sequence of the form $\bot \cdots \bot \top$. On
+similar databases, we would like to show that the discrete above
+threshold mechanism outputs the same sequence with similar
+probabilities.

### 2018-02-01 introduction.tex
finalize

@@
 
-In order to illusrtrate applicability of our techniques, we give
+In order to illustrate applicability of our techniques, we give
 detailed formalizations of several data analysis mechanisms in this
@@
 \noindent
-\emph{Related Work.}
+\paragraph{Related Work.}
 An early attempt on formal verification of differential privacy
@@
 \noindent
-\emph{Contributions.} Our main contributions are threefold. (I)
-We introduce the logic $\dpCTLstar$ for reasoning about properties in
-differential privacy. We use the logic to express classical
-differential privacy properties, and moreover, complex properties such as those
-involving infinite behaviors can also be expressed. (II) We model several differentially private mechanisms, including stream processing, in Markov chains or Markov decision processes.
-(III) We show that the model checking
+\paragraph{Contributions.} Our main contributions are threefold. 
+\begin{enumerate}
+\item We introduce the logic $\dpCTLstar$ for reasoning about properties in
+differential privacy. The logic is able to to express classical
+differentially private properties. Complex properties such as those
+involving infinite behaviors can also be expressed;
+\item We model several differentially private mechanisms, including stream processing, in Markov chains or Markov decision processes; and
+\item We show that the model checking
 problem for Markov chains is standard. For Markov decision processes, we show that it is undecidable.
+\end{enumerate}

### 2018-02-01 main.tex
remove end if to save space
minor fixes
finalize

@@
 \usepackage{times}
+
+\algtext*{EndWhile}% Remove "end while" text
+\algtext*{EndIf}% Remove "end if" text
+\algtext*{EndFor}% Remove "end if" text
 
@@
 
-\paragraph{Organisation of the paper.}
+\noindent
+\paragraph{Organization of the paper.}
 Preliminaries are given in Section~\ref{section:preliminaries}.
@@
 They are followed by the extension to Markov decision processes and an
-application in streaming process (Section~\ref{section:mdp}). Finally,
+application in stream processing (Section~\ref{section:mdp}). Finally,
 Section~\ref{section:conclusions} concludes our presentation.
@@
 
-\section{Markov Decision Processes}
+\section{Markov Decision Processes for Differentially Private Mechanisms}
 \label{section:mdp}
@@
 An interesting work is to identify more
-decidable subclass. As an example, consider the formula $\bigwedge_{k \in \bbfN}
+decidable subclass. As an example, consider the formula $\bigwedge_{k \in \bbfZ^{\geq 0}}
 \dpriv{\epsilon}{\delta} (\X^k  00)$ which was used in specifying the pan privacy.
 For the case $\epsilon=\delta=0$, it reduces to a language equivalence problem for probabilistic automata.
-It is interesting to characterise other cases as well.
+It is interesting to characterize other cases as well.
  Another interesting line of further works is to consider continuous perturbation (such as Laplace distribution used in \cite{DR:14:AFDP}) -- we would need Markov models with continuous state space for the modelling.

### 2018-02-01 mdp.tex
finalize

@@
 \begin{wrapfigure}[14]{r}{.45\columnwidth}
+%\begin{figure}
   \centering
@@
 \end{wrapfigure}
+%\end{figure}
 
 \hide{Mechanisms are not necessarily closed randomized algorithms; they may
-perform different compution on users' requests.
+perform different computation on users' requests.
 We use Markov decision processes to model such
@@
 Consider again the survey mechanism. Suppose we would like to design
-an interactive mechanism which adjusts random noises on surveyers'
-requests. When the surveyer requests low-accuracy answers, the
+an interactive mechanism which adjusts random noises on surveyors'
+requests. When the surveyor requests low-accuracy answers, the
 surveyee uses the survey mechanism as before. When high-accuracy
@@
 \noindent
-\emph{Justification of the scheduler class.}
-For defining semantics for the $\mathcal{D}$-operator, we used query scheduler.
-If we use arbitrary scheduler, another definition of the differentially private operator
-might be:
+\paragraph{Justification of query schedulers.}
+We use query schedulers in the semantics for the differentially
+private operator. A definition with history-dependent schedulers
+might be
 \begin{eqnarray*}
@@
 A state satisfies $\dpriv{\epsilon}{\delta}^{\mathit{bad}} \phi$ if
-no scheduler could differentiate the probabilities of paths satisfying
-$\phi$ from neighbors. Such a definition would allow strategies to
-take different actions from different states. It might need different
-action sequences to distinguish neighbors. Two neighbors could still be
-differentiated by an action sequence and lose their privacy.
-
-We consider only query scheduler $\scheduler{Q}\in Act^\omega$, which correspond to the query sequences.
-In the differential privacy context, neighbor states exhibit same action behaviours.
-And we will see later that
-our semantics agrees with the notion of differential privacy in the literature for such systems.
-We hence
+no history-dependent scheduler could differentiate the probabilities of paths satisfying
+$\phi$ from neighbors. Such a definition would allow schedulers to
+take different actions from different states. Two neighbors might be
+differentiated by different action sequences erroneously.
+A query scheduler $\scheduler{Q} : \bbfZ^{\geq 0} \rightarrow \omega$
+corresponds to a query sequence. In differential privacy,
+neighbors exhibit similar action behaviors.
+Our semantics agrees with the informal interpretation of differential
+privacy for such systems. We hence
 prefer the original definition.
@@
 If the path formula only contains next operators, it is easy to deal with.
-When $\phi$ is $\gX B$ with $B\subseteq S$, only one step behaviour needs to be considered.
+When $\phi$ is $\gX B$ with $B\subseteq S$, only one step behavior needs to be considered.
 Thus only the first action in the query sequence is needed.
 It can also be easily generalized to nested next operators:  one needs only to enumerate
-all actions query sequences of length obtained by the number of nexted $\X$ operators.
+all actions query sequences of length obtained by the number of nested $\X$ operators.
 The problem however is undecidable in general.
@@
 The proof is put in Appendix.
-We consider some decidable special cases.
+We discuss some decidable special cases.
 Consider the formula $\phi:=F B$ with $B\subseteq S$ and assume that states in $B$ are absorbing (with only self-loops).
 For the case $\epsilon=0$, the condition reduces to
-$\myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{F B} -
-\myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{F B}\le \delta$.
+$\myPr{s}{M_{\scheduler{Q}}}{\neighbor{S}}{F B} -
+\myPr{t}{M_{\scheduler{Q}}}{\neighbor{S}}{F B}\le \delta$.
 If $\delta=0$ it is the classical language equivalence problem for probabilistic automata~\cite{Rabin63}, which can be solved in polynomial time \cite{Tzeng92}.

### 2018-02-01 preliminaries.tex
notations
Correct some spelling errors and others
finalize

@@
 % \bbfZ^{\geq 0}$.
-%\lz{not so often used: can we do withuot this notatin?}
+%\lz{not so often used: can we do without this notation?}
 We briefly review the definitions of differential privacy, Markov
@@
 $\Pr[M_f(x) = \tilde{r}] = \Pr[M_f(x') = \tilde{r}]$ for every
-$r \in \tilde{R}$ when $f (x) = f (x')$. In words, outputs of an
+$\tilde{r} \in \tilde{R}$ when $f (x) = f (x')$. In words, outputs of an
 oblivious mechanism depend on the query result $f (x)$ but not on the
@@
 the smaller they are, the behaviors are more similar.
-Informally, a differentially private mechamism has probabilistically
-similar behabiors on neighboring databases. It will have similar
+Informally, a differentially private mechanism has probabilistically
+similar behaviors on neighboring databases. It will have similar
 output distributions when a row is replaced by another in a given
@@
 
-Let $M = (S, \Act, \wp, L)$ be an MDP. A
+Let $M = (S, \Act, \wp, L)$ be an MDP. A (history-dependent)
 \emph{scheduler} for $M$ is a function $\scheduler{S} : S^+
 \rightarrow \Act$. A \emph{query scheduler} for $M$ is a function
-$\scheduler{Q} : \bbfZ^{\geq 0} \rightarrow \Act$. 
+$\scheduler{Q} : \bbfZ^{\geq 0} \rightarrow \Act$. Intuitively, its decision depends only on the length of the history.
 A path $\pi =

### 2018-02-01 semantics.tex
finalize

@@
 Moreover, the relation $K, \neighbor{S}, \pi \models \phi$ is defined as
-for the standard LTL formulas~\cite{ManneP92}. We only recall the semantics for $\X$ and $\until$ as follows:
+for the standard linear temporal logic formulas~\cite{ManneP92}. We only recall the semantics for $\X$ and $\until$ as follows:
 \begin{eqnarray*}
@@
 probabilistically similar to observe paths satisfying $\phi$ from $s$
-and from all its neighbors.
+and from its neighbors.
 

### 2018-02-01 specifications.tex
finalize

@@
 
-\paragraph{Standard Differential Privacy.}
+\paragraph{Classical Differential Privacy.}
 Consider the survey mechanism (Section~\ref{subsec:survey}).
@@
 In comparison, let us consider two slightly different formulae:
-(a) $\dpriv{\ln 3}{0} (\X \X \mathit{out}_Y)$.
+\begin{itemize}
+\item $\dpriv{\ln 3}{0} (\X \X \mathit{out}_Y)$.
 In this case we do not have privacy loss, even though we are querying twice: the reason is that the output of the first query is not utilized at all.
 It is easy to verify that it is indeed satisfied by $+$ and $-$.
-(b) $\dpriv{\ln 3}{0} (\X (\mathit{out}_Y\wedge \dpriv{\ln 3}{0}(\X \mathit{out}_Y)))$.
-This is a nested $\dpCTL$ formula, where the inner state formula $\dpriv{\ln 3}{0}(\X \mathit{out}_Y)$ states the one step differential privacy.
+\item $\dpriv{\ln 3}{0} (\X (\mathit{out}_Y\wedge \dpriv{\ln 3}{0}(\X \mathit{out}_Y)))$.
+This is a nested $\dpCTL$ formula, where the inner state formula $\dpriv{\ln 3}{0}(\X \mathit{out}_Y)$ specifies the one step differential privacy.
 Observe the inner formula is satisfied by all neighbors.
 Thus, the outer formula also has no privacy loss.
+\end{itemize}
 
@@
 $\delta_2=2\delta_1=1$. However, we can check easily that one gets
-better privacy parameter $(0,.5)$.
+better privacy parameter $(0,.5)$ using the model checking algorithm.
+In general, compositional theorems for differential privacy only give
+asymptotic upper bounds. Privacy parameters $\epsilon$ and $\delta$
+must be calculated carefully and often overly pessimistic. Our
+algorithm allows data analysts to choose such parameters more
+tightly.
 
@@
 Reachability privacy is often used in combination with other properties.
-In the above threshold mechanism in Section~\ref{subsec:threshold},
-consider the property $\Phi=\wedge_{i=1}^{6} \dpriv{3}{0}(\X^i \top) \wedge \PJ{\le .025}(\X^6 F \top) \wedge \dpriv{3}{.025}(\X^6 F \top)$.
-The state $s_{34}$ (with the input threshold $t=3$, and the database $d=4$) satisfies $\Phi$ by our computation.
-It means the probability of reaching $\top$ at $i$ steps with $i$ ranges over $\{1,\ldots,6\}$, is $(3, 0)$-differentially private for the database $d$ and its neighbours ($d=3$ and $d=5$).
-It also states that the probability of reaching $\top$ after no less than $6$ steps, is smaller than $.025$ and is  $(3, .025)$-differentially private.
+In the above threshold mechanism (Section~\ref{subsec:threshold}),
+recall that outputs are of the form $\bot \cdots \bot \top$. In order
+to establish differential privacy, one wishes to show that $\Phi_i$
+holds for every $i \in \bbfZ^{\geq 0}$ where $\Phi_i =
+\dpriv{3}{0}(\X^i \top)$. That is, the mechanism halts at the $i$-th
+step with similar probabilities for every $i$. 
+It is not clear how to verify them automatically.
+
+Now consider $\Phi=\wedge_{i=1}^{6} \dpriv{3}{0}(\X^i \top) 
+\wedge \PJ{\le .025}(\X^6 F \top) \wedge \dpriv{3}{.025}(\X^6 F
+\top)$. The state with the threshold $t=3$, and the query result $r=4$ 
+satisfies $\Phi$ by our computation. It means the
+probability of reaching $\top$ at $i$ steps with $i$ ranges over
+$\{1,\ldots,6\}$, is $(3, 0)$-differentially private for the database
+$d$ and its neighbors ($r=3$ and $r=5$). It also states that the
+probability of reaching $\top$ after no less than $6$ steps, is
+smaller than $.025$ and $(3, .025)$-differentially private as well. 
+Rather than checking the probabilities of all possible outputs, $\Phi$
+focuses on finitely many short outputs. It says the mechanism enjoys $(3,
+0)$-differentially privacy for the short outputs, and long outputs are
+unlikely. Moreover, all neighbors are unlikely to have long outputs. 
+Using this trick, one can bound the probability of long outputs
+arbitrarily (say, less than $10^{-6}$). Our technique can establish
+practical privacy guarantees automatically.
 
@@
 device cannot increase the chance of service failure significantly.
+In many applications, data analyses are carried out constantly. 
+Privacy analysis over infinite behaviors is crucial. 
+We will revisit this issue again in
+Section~\ref{section:applications}.

### 2018-02-01 syntax.tex
minor fixes
finalize

@@
  by a path formula. A
-\emph{path} formula $\phi$ is simply a LTL formula, with temporal operator \emph{next}  ($\X$)
+\emph{path} formula $\phi$ is simply a linear temporal logic formula, with temporal operator \emph{next}  ($\X$)
 and  \emph{until} operator
 ($\until$) operator enclosed by two path formulae.
-We often use syntactic sugar $F$ and $G$ defined by: $F \phi:= true \until \phi$ and $G\phi :=\neg F (\neg\phi)$.
+We moreover define $F \phi \equiv true \until \phi$ and $G\phi \equiv \neg F (\neg\phi)$.
 

### 2018-04-19 introduction.tex
revise

@@
 sophisticated data analyses. Privacy leak has also been observed in
-data analysis programs with differentially private
-designs~\cite{M:12:SLSBDP,TKBW:17:PLAI}. This calls for formal
+data analysis programs implementing differentialy
+privacy~\cite{M:12:SLSBDP,TKBW:17:PLAI}. This calls for formal 
 analysis of differential privacy on both designs and implementations.
@@
 (\G \F \mathit{high})$. A state satisfies the formula if it
-is probabilistically similar to  infinitely often
-answer $\mathit{high}$ as similar states. Compared to the unbounded
+is as likely to answer $\mathit{high}$ infinitely often
+as similar states. Compared to the unbounded
 but finite behavior specified by $\F \mathit{high}$,

### 2018-04-19 preliminaries.tex
revise

@@
 input database $x$. The order of rows in a database, for instance, is
-irrelevant to oblivious mechanisms. A database $x$ is
-\emph{$(\epsilon, \delta)$-close} to $x'$ in a mechanism
-$M_f$ if for every $\tilde{r} \in \tilde{R}$,
+irrelevant to oblivious mechanisms for counting queries. Let $x, x'$
+be databses and $\tilde{r} \in \tilde{R}$. We say the probability of
+the mechanism $M_f$ outputing $\tilde{r}$ on $x$ is
+\emph{$(\epsilon, \delta)$-close} to those on $x'$ if
 \[
@@
   private} % $(\epsilon, \delta \approx 0)$
-if for every $x, x' \in \mathcal{X}^n$ with $d(x, x') \leq 1$,
-$x$ is $(\epsilon, \delta)$-close to $x'$ in $M_f$.
+if for every $x, x' \in \mathcal{X}^n$ with $d(x, x') \leq 1$
+and $\tilde{r} \in \tilde{R}$, the probability of
+$M_f$ outputing $\tilde{r}$ on $x$ is $(\epsilon, \delta)$-close to
+those on $x'$.
 
-The parameters $\epsilon$ and $\delta$ quantify probabilistically
-similar behaviors;
+The non-negative parameters $\epsilon$ and $\delta$ quantify
+probabilistically similar behaviors;
 the smaller they are, the behaviors are more similar.

### 2018-04-19 examples.tex
revise

@@
 To model differentially private mechanisms by Markov chains, we
-formalize inputs (such as databases or query results) as states. For
+formalize inputs (such as databases or query results) as states. 
+Randomized computation is modeled by probabilistic transitions. Atomic
+propositions are used to designate intended interpretation on states
+(such as inputs or outputs). We demonstrates the ideas in examples.
+
+\hide{
+For
 each input, its output distribution is modeled as a probabilistic
@@
 
-\hide{
 A simple way to design differentially private mechanisms is to add
@@
 \frac{3}{4}$. Hence $\Pr[M (x) = Y] \leq \frac{3}{4} = 3 \cdot
-\frac{1}{4} \leq e^{\ln 3} \Pr[M (x') = Y]$ for any neighboring $x, x'
+\frac{1}{4} \leq e^{\ln 3} \Pr[M (x') = Y]$ for any neighbors $x, x'
 \in \mathcal{X}$. Similarly, $\Pr[M (x) = N] \leq e^{\ln 3} \Pr[M (x')
@@
 In the figure, the states
-$+$ and $-$ denote the diagnoses are positive or negative
+$+$ and $-$ denote positive or negative diagnoses 
 respectively; the states $s$ and $t$ denote answers to the survey
-question and hence let $\mathit{out}_Y \in L (s)$ and $\mathit{out}_N
+question and hence $\mathit{out}_Y \in L (s)$ and $\mathit{out}_N
 \in L (t)$.
@@
 Missing transitions (such as those from $s$ and $t$) lead
-to a special state $\dagger$ with a self-loop. We omit the transitions
+to a special state $\dagger$ with a self-loop. We omit such transitions
 and the state $\dagger$ for clarity.
@@
 available. Consider a query
-$f : \mathcal{X}^n \rightarrow \bbfZ^{\geq 0}$. Let $\alpha \in (0, 1)$.
+$f : \mathcal{X}^n \rightarrow \{ 0, 1, \ldots, m \}$. Let $\alpha \in (0, 1)$.
 The \emph{$\alpha$-geometric mechanism}
@@
 however is $\bbfZ$. It may give nonsensical outputs such as
-negative integers for queries with positive ranges.
+negative integers for non-negative queries.
 
 The \emph{truncated $\alpha$-geometric mechanism over $\{ 0, 1,
-  \ldots, n \}$}
+  \ldots, m \}$}
 outputs $f (x) + Z$ where $Z$ is a random variable with the following
@@
     \frac{1 - \alpha}{1 + \alpha}\alpha^{|z|} &
-    \textmd{ if } -f (x) < z < n - f (x)\\
-    \frac{\alpha^{n - f (x)}}{1 + \alpha} & \textmd{ if } z = n-f (x)\\
-    0 & \textmd{ if } z > n - f (x)
+    \textmd{ if } -f (x) < z < m - f (x)\\
+    \frac{\alpha^{m - f (x)}}{1 + \alpha} & \textmd{ if } z = m-f (x)\\
+    0 & \textmd{ if } z > m - f (x)
   \end{array}
@@
 Note the range of the truncated $\alpha$-geometric mechanism is
-$\{ 0, 1, \ldots, n \}$. The truncated $\alpha$-geometric mechanism is
+$\{ 0, 1, \ldots, m \}$. The truncated $\alpha$-geometric mechanism is
 again oblivious; it is also $(- {\Delta (f)} \ln \alpha, 0)$-differentially
@@
 
-Similar to the survey mechanism, it is
-straightforward to model the truncated $\frac{1}{2}$-geometric
-mechanisms as Markov chains. Let the state $s_k$ and $t_l$
+Similar to the survey mechanism, it is straightforward to model the
+truncated $\frac{1}{2}$-geometric mechanism as a Markov chain. 
+Recall that the truncated $\frac{1}{2}$-geometric mechanism is
+oblivious. It suffices to consider $\{ f (x) : x \in \mathcal{X}^n \}$
+instead of $\mathcal{X}^n$ as inputs. 
+Let the state $s_k$ and $t_l$
 denote the input value $k$ and output value $l$ respectively. Define
-the state set $S = \{ s_k, t_k : k \in \{ 0, 1, \ldots, n \} \}$.
+the state set $S = \{ s_k, t_k : k \in \{ 0, 1, \ldots, m \} \}$.
 The probability transition $\wp (s_k, t_l)$ is the probability of the
@@
 $\mathit{out}_k \in L (t_k)$ for $k \in \{ 0, 1, \ldots, n \}$.
-If $\Delta (f) = 1$, $s_k$ and $s_l$ are
-neighbors iff $| k - l | \leq 1$.
+If $\Delta (f) = 1$, $| f (x) - f (x') | \leq 1$ for every neighbors
+$x, x' \in \mathcal{X}^n$. Subsequently, $s_k$ and $s_l$ are
+neighbors iff $| k - l | \leq 1$ in our model.
 Figure~\ref{figure:geometric-mechanism-markov-chain} gives
-the Markov chain of the truncated
+the Markov chain for the truncated
 $\frac{1}{2}$-geometric mechanism over $\{ 0, 1, \ldots, 5 \}$.
@@
 \label{subsec:threshold}
-In differential privacy, we differentiate between {offline} and
-{online} mechanisms. 
-An offline mechanism releases outputs only once and plays no further
-role. The mechanisms we constructed previously are offline mechanisms. 
-On the other hand, online mechanisms permit further queries. 
+In differential privacy, an offline mechanism releases outputs only
+once and plays no further role; an online (interactive) mechanism
+allows analysts to ask queries adaptively based on previous
+responses. The mechanisms we constructed previously are offline
+mechanisms. On the other hand, online mechanisms permit further queries. 
 Obviously, queries do not change databases themselves. To maintain
@@
 
-Below we describe an online mechanism adapted from~\cite{DR:14:AFDP}:
-it is used in the scenario where a large amount of queries occur, but
-we only care for the queries which 
-are above a given threshold. Queries below the threshold are
-considered {irrelevant} and only disclose minimal information.
-In~\cite{DR:14:AFDP}, the mechanism is design for continuous queries.
-The Laplace mechanism is applied. Here, our mechanism is designed for
-bounded and discrete queries. The geometric mechanism is applied instead.
+Below we describe an online mechanism adapted from~\cite{DR:14:AFDP}.
+Given a threshold and a series of adaptive queries, we only care for
+the queries whose results are above the threshold; queries below the
+threshold only disclose minimal information and hence are considered
+irrelevant. Since query results above the threshold convey useful
+information, let us assume the mechanism will halt on the first such
+query result for simplicity. The question is to design a
+privacy-preserving mechanism to have similar behaviors on similar
+databases. In~\cite{DR:14:AFDP}, a mechanism is design for continuous
+queries by applying the Laplace mechanism. We will develop a mechanism
+for bounded and discrete queries using the geometric mechanism.
 
@@
 (Algorithm~\ref{algorithm:online-model}). The above threshold
-mechanism outputs a sequence of the form $\bot \cdots \bot \top$. On
+mechanism outputs a sequence of the form $\bot^* \top$. On
 similar databases, we would like to show that the discrete above

### 2018-04-19 semantics.tex
revise

@@
 paths satisfying $\phi$ from $s$. A state $s$ satisfies
-$\dpriv{\epsilon}{\delta} \phi$ if having paths satisfying $\phi$
-from $s$ is $(\epsilon, \delta)$-close to having paths satisfying
-$\phi$ from every neighbor of $s$. Informally, it is
+$\dpriv{\epsilon}{\delta} \phi$ if the probability of
+having paths satisfying $\phi$ from $s$ is $(\epsilon, \delta)$-close
+those from every neighbor of $s$. Informally, it is
 probabilistically similar to observe paths satisfying $\phi$ from $s$

### 2018-04-19 specifications.tex
revise

@@
 \paragraph{Classical Differential Privacy.}
-Consider the survey mechanism (Section~\ref{subsec:survey}).
-The formula $\dpriv{\ln 3}{0} (\X \mathit{out}_Y)$ holds in state $u$ if for  $v$ with $u N v$, the probability of satisfying $\X \mathit{out}_Y$ from $u$ and $v$ satisfy:
-$\myPr{u}{K}{N}{\X \mathit{out}_Y}\le 3\myPr{v}{K}{N}{\X \mathit{out}_Y}$.
-The formula $\dpriv{\ln 3}{0} (\X \mathit{out}_Y)\wedge \dpriv{\ln 3}{0} (\X \mathit{out}_N)$ thus specifies differential privacy for states $+$ and $-$.
+Consider the survey mechanism (Section~\ref{subsec:survey}). For  $v$
+with $u N v$, we have $\myPr{u}{K}{N}{\X \mathit{out}_Y}\le
+3\myPr{v}{K}{N}{\X \mathit{out}_Y}$ for the probability of satisfying
+$\X \mathit{out}_Y$ from $u$ and $v$.
+The formula $\dpriv{\ln 3}{0} (\X \mathit{out}_Y)$ holds in state $u$
+and similarly for $\dpriv{\ln 3}{0} (\X \mathit{out}_N)$. 
+Recall that differential privacy requires similar output distributions
+on neighbors. The formula $\dpriv{\ln 3}{0} (\X
+\mathit{out}_Y)\wedge \dpriv{\ln 3}{0} (\X \mathit{out}_N)$ thus
+specifies differential privacy for states $+$ and $-$. 
 The survey mechanism is $(\ln 3, 0)$-differentially private.
@@
 Compositionality is one of the building block for differential privacy.
-According to the compositional theorem \cite[Theorem 3.16]{DR:14:AFDP}, for $(\epsilon_1,\delta_1)$-differentially private algorithm $M_1$ and
-$(\epsilon_2,\delta_2)$-differentially private algorithm $M_2$, their combination $M$ defined by $M(x)=(M_1(x), M_2(x))$ is $(\epsilon_1 + \epsilon_2,\delta_1+\delta_2)$-differentially private.
+According to the compositional theorem \cite[Theorem 3.16]{DR:14:AFDP}, for $(\epsilon_1,\delta_1)$-differentially private mechanism $M_1$ and
+$(\epsilon_2,\delta_2)$-differentially private mechanism $M_2$, their combination $M$ defined by $M(x)=(M_1(x), M_2(x))$ is $(\epsilon_1 + \epsilon_2,\delta_1+\delta_2)$-differentially private.
 The degradation is rooted in the repeated releases of information.
 To illustrate this property, we consider the extended survey mechanism which allows two consequent queries.
+In this mechanism, an input is either $+$ or $-$; but 
+outputs are $\mathit{out}_Y\mathit{out}_Y$,
+$\mathit{out}_Y\mathit{out}_N$, $\mathit{out}_N\mathit{out}_Y$, or
+$\mathit{out}_N\mathit{out}_N$. 
 The model is depicted in Figure~\ref{figure:2-dp-mc}.
@@
     \centering
-    \begin{subfigure}{.4\columnwidth}
-\resizebox{\columnwidth}{!}{
+\resizebox{.70\columnwidth}{!}{
     \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
     distance=1cm,node/.style={circle,draw}]
-    \node[node] (p) at ( -1.5,  0) { $+$ };
-    \node[node] (Y1) at ( 0, .8) { $s_1$ };
+    \node[node] (p) at  (  -3,  0) { $+$ };
+    \node[node] (pY1) at (-1.5, .8) { $s_1$ };
+    \node at (-1.5, 1.3) { $\mathit{out}_Y$ };
+    \node[node] (pN1) at (-1.5, -.8) { $t_1$ };
+    \node at (-1.5, -.3) { $\mathit{out}_N$ };
+
+    \node[node] (Y2) at ( 0,  .8) { $s_2$ };
     \node at (0, 1.3) { $\mathit{out}_Y$ };
-    \node[node] (N1) at ( 0, -.8) { $t_1$ };
+    \node[node] (N2) at ( 0, -.8) { $t_2$ };
     \node at (0, -.3) { $\mathit{out}_N$ };
-    \node[node] (Y2) at (1.5,  .8) { $s_2$ };
+
+    \node[node] (n) at  (   3,  0) { $-$ };
+    \node[node] (nY1) at ( 1.5, .8) { $\bar{s}_1$ };
     \node at (1.5, 1.3) { $\mathit{out}_Y$ };
-    \node[node] (N2) at ( 1.5, -.8) { $t_2$ };
+    \node[node] (nN1) at ( 1.5, -.8) { $\bar{t}_1$ };
     \node at (1.5, -.3) { $\mathit{out}_N$ };
@@
     \path
-    (p) edge node [above] { $\frac{3}{4}$ } (Y1)
-    (p) edge node [below] { $\frac{1}{4}$ } (N1)
-    (Y1) edge node [above] { $\frac{3}{4}$ } (Y2)
-    (Y1) edge node [above] { $\frac{1}{4}$ } (N2)
-    (N1) edge node [below] { $\frac{3}{4}$ } (Y2)
-    (N1) edge node [below] { $\frac{1}{4}$ } (N2)
-    ;
-    \end{tikzpicture}
-}
-%        \caption{Markov Chain of }
-%        \label{figure:2-dp-mc1}
-    \end{subfigure}
-    \hspace{.05\columnwidth}
-    \begin{subfigure}{.4\columnwidth}
-        \resizebox{\columnwidth}{!}{
-            \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
-    distance=1cm,node/.style={circle,draw}]
-    \node[node] (n) at ( -1.5,  0) { $-$ };
-    \node[node] (Y1) at ( 0, .8) { $s_1$ };
-    \node at (0, 1.3) { $\mathit{out}_Y$ };
-    \node[node] (N1) at ( 0, -.8) { $t_1$ };
-    \node at (0, -.3) { $\mathit{out}_N$ };
-    \node[node] (Y2) at (1.5,  .8) { $s_2$ };
-    \node at (1.5, 1.3) { $\mathit{out}_Y$ };
-    \node[node] (N2) at ( 1.5, -.8) { $t_2$ };
-    \node at (1.5, -.3) { $\mathit{out}_N$ };
-
-    \path
-    (n) edge node [above] { $\frac{1}{4}$ } (Y1)
-    (n) edge node [below] { $\frac{3}{4}$ } (N1)
-    (Y1) edge node [above] { $\frac{1}{4}$ } (Y2)
-    (Y1) edge node [above] { $\frac{3}{4}$ } (N2)
-    (N1) edge node [below] { $\frac{1}{4}$ } (Y2)
-    (N1) edge node [below] { $\frac{3}{4}$ } (N2)
+    (p) edge node [above] { $\frac{3}{4}$ } (pY1)
+    (p) edge node [below] { $\frac{1}{4}$ } (pN1)
+    (pY1) edge node [above] { $\frac{3}{4}$ } (Y2)
+    (pY1) edge node [above] { $\frac{1}{4}$ } (N2)
+    (pN1) edge node [below] { $\frac{3}{4}$ } (Y2)
+    (pN1) edge node [below] { $\frac{1}{4}$ } (N2)
+    (n) edge node [above] { $\frac{1}{4}$ } (nY1)
+    (n) edge node [below] { $\frac{3}{4}$ } (nN1)
+    (nY1) edge node [above] { $\frac{1}{4}$ } (Y2)
+    (nY1) edge node [above] { $\frac{3}{4}$ } (N2)
+    (nN1) edge node [below] { $\frac{1}{4}$ } (Y2)
+    (nN1) edge node [below] { $\frac{3}{4}$ } (N2)
     ;
@@
 %        \label{figure:2-dp-mc2}
-    \end{subfigure}
     \caption{Markov Chain of Double Surveys}
@@
 
-In comparison, let us consider two slightly different formulae:
+Let us consider two slightly different formulae for comparison:
 \begin{itemize}
-\item $\dpriv{\ln 3}{0} (\X \X \mathit{out}_Y)$.
-In this case we do not have privacy loss, even though we are querying twice: the reason is that the output of the first query is not utilized at all.
-It is easy to verify that it is indeed satisfied by $+$ and $-$.
+\item $\dpriv{\ln 3}{0} (\X \X \mathit{out}_Y)$. In this case we claim
+  there is no privacy loss, even though there are two queries. The reason is that the output of the first query is not observed at all. It is easy to verify that it is indeed satisfied by $+$ and $-$.
 \item $\dpriv{\ln 3}{0} (\X (\mathit{out}_Y\wedge \dpriv{\ln 3}{0}(\X \mathit{out}_Y)))$.
@@
 \paragraph{Tighter Privacy Bounds for Composition.}
-An advantage of applying model checking is that we may get tighter bounds for composition.
-Consider the survey mechanism, and the property $\dpriv{0}{.5} (\X \mathit{out}_Y)$.
-Obviously, it holds in states $+$ and $-$.
-A careful check ensures that one cannot decrease  $\delta_1=.5$ without increasing $\epsilon$.
+An advantage of applying model checking is that we may get tighter
+bounds for composition. Consider the survey mechanism, and the
+property $\dpriv{0}{.5} (\X \mathit{out}_Y)$. Obviously, it holds in
+states $+$ and $-$ since $\myPr{u}{K}{N}{\mathit{out}_Y} =
+\frac{3}{4}$, $\frac{1}{4}$ for $u = +, -$ respectively (Figure~\ref{figure:2-dp}). A careful check ensures that one cannot decrease  $\delta_1=.5$ without increasing $\epsilon$.
 Now consider the formula $\dpriv{\epsilon_2}{\delta_2} (\X
-(\mathit{out}_Y\wedge \X \mathit{out}_Y))$ on the model in
+(\mathit{out}_Y\wedge \X \mathit{out}_Y))$ in
 Figure~\ref{figure:2-dp-mc}. Applying the
@@
 $\delta_2=2\delta_1=1$. However, we can check easily that one gets
-better privacy parameter $(0,.5)$ using the model checking algorithm.
+better privacy parameter $(0,.5)$ using the model checking algorithm
+because $\myPr{u}{K}{N}{\mathit{out}_Y} = \frac{9}{16}$,
+$\frac{1}{16}$ for $u = +, -$ respectively.
 In general, compositional theorems for differential privacy only give
 asymptotic upper bounds. Privacy parameters $\epsilon$ and $\delta$
-must be calculated carefully and often overly pessimistic. Our
+must be calculated carefully and often overly pessimistically. Our
 algorithm allows data analysts to choose such parameters more
@@
 probability of reaching $\top$ at $i$ steps with $i$ ranges over
-$\{1,\ldots,6\}$, is $(3, 0)$-differentially private for the database
-$d$ and its neighbors ($r=3$ and $r=5$). It also states that the
+$\{1,\ldots,6\}$, is $(3, 0)$-differentially private for 
+neighbors ($r=3$ and $r=5$). It also states that the
 probability of reaching $\top$ after no less than $6$ steps, is

### 2018-04-19 mdp.tex
revise

@@
 $\dpriv{\epsilon}{\delta}$. For any path formula $\phi$,
-$\dpriv{\epsilon}{\delta} \phi$ specifies states which are $(\epsilon,
-\delta)$-close to all its neighbors in having paths satisfying $\phi$
-for query schedulers. That is, no
+$\dpriv{\epsilon}{\delta} \phi$ specifies states whose probability
+of having paths satisfying $\phi$ are $(\epsilon, \delta)$-close to
+those of all its neighbors for query schedulers. That is, no
 query scheduler can force any of neighbors to have
@@
 A state satisfies $\dpriv{\epsilon}{\delta}^{\mathit{bad}} \phi$ if
-no history-dependent scheduler could differentiate the probabilities of paths satisfying
-$\phi$ from neighbors. Such a definition would allow schedulers to
-take different actions from different states. Two neighbors might be
-differentiated by different action sequences erroneously.
-A query scheduler $\scheduler{Q} : \bbfZ^{\geq 0} \rightarrow \omega$
-corresponds to a query sequence. In differential privacy,
-neighbors exhibit similar action behaviors.
+no history-dependent scheduler could differentiate the probabilities
+of haveing paths satisfying
+$\phi$ from neighbors. Recall that a history-dependent scheduler takes
+actions by previous states. Such a definition would allow
+schedulers to take different actions from different states. 
+Two neighbors could be differentiated by different action sequences
+erroneously.
+A query scheduler $\scheduler{Q} : \bbfZ^{\geq 0} \rightarrow \Act$,
+on the other hand, corresponds to a query sequence. A state satisfies
+$\dpriv{\epsilon}{\delta} \phi$ if no query sequence can differentiate
+the probabilities of having paths satisfying $\phi$ from neighbors.
 Our semantics agrees with the informal interpretation of differential
@@
 \myPr{t}{M_{\scheduler{Q}}}{\neighbor{S}}{F B}\le \delta$.
-If $\delta=0$ it is the classical language equivalence problem for probabilistic automata~\cite{Rabin63}, which can be solved in polynomial time \cite{Tzeng92}.
+If $\delta=0$ it is the classical language equivalence problem for probabilistic automata~\cite{Rabin63}, which can be solved in polynomial time.
 However, if $\delta>0$, the problem
-becomes an approximate version of the language equivalence problem~\cite{Tzeng92}.
+becomes an approximate version of the language equivalence problem.
 To the best of our knowledge, the decision problem is still open, and has only be investigated for the special case when all states are connected~\cite{Tzeng92}.

### 2018-04-25 examples.tex
add mdp for above threshold

@@
 
+\begin{figure}
+  \centering
+    \resizebox{\columnwidth}{!}{
+    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
+      distance=2cm,node/.style={circle,draw}]
+      \node[node] (t0r1) at (-4.75, 0) { $t_0r_1$ };
+
+      \draw (-3.95, 2.1) rectangle (-.05, 4.9);
+      \node at (-3.25, 4.5) { $\vdots $};
+      \node[node] (t'0r1) at (-3.25, 3.5) { $t'_0r_1$ };
+      \node at (-3.25, 2.7) { $\vdots $};
+      \draw (-3.95, -1.9) rectangle (-.05, 1.9);
+      \node[node] (t'1r0) at (-3.25, 1.2) { $t'_1r_0$ };
+      \node[node] (t'1r1) at (-3.25, 0) { $t'_1r_1$ };
+      \node[node] (t'1r2) at (-3.25, -1.2) { $t'_1r_2$ };
+      \node at (-1.95, 0) { $\cdots$ };
+      \node at (-1.95, -1.2) { $\cdots$ };
+      \draw (-3.95, -2.1) rectangle (-.05, -4.9);
+      \node at (-3.25, -2.5) { $\vdots$ };
+      \node[node] (t'2r1) at (-3.25, -3.5) { $t'_2r_1$ };
+      \node at (-3.25, -4.3) { $\vdots$ };
+
+      \node[node] (t'1r'0) at (-.75, 1.2) { $t'_1r'_0$ };
+      \node[node] (t'1r'1) at (-.75, 0) { $t'_1r'_1$ };
+      \node[node] (t'1r'2) at (-.75, -1.2) { $t'_1r'_2$ };
+
+      \node[node] (tt0rr0) at (4.75, 1.2) { $\underline{t}_0\underline{r}_0$ };
+
+      \draw (3.95, 2.1) rectangle (.05, 4.9);
+      \node[node] (tt'0rr0) at (3.25, 4.2) { $\underline{t}'_0\underline{r}_0$ };
+      \node at (3.25, 3) { $\vdots$ };
+      \draw (3.95, -.9) rectangle (.05, 1.9);
+      \node[node] (tt'1rr0) at (3.25, 1.2) { $\underline{t}'_1\underline{r}_0$ };
+      \node at (3.25, 0) { $\vdots$ };
+
+      \draw (3.95, -1.1) rectangle (.05, -4.9);
+      \node[node] (tt'2rr0) at (3.25, -1.8) { $\underline{t}'_2\underline{r}_0$ };
+      \node[node] (tt'2rr1) at (3.25, -3.0) { $\underline{t}'_2\underline{r}_1$ };
+      \node[node] (tt'2rr2) at (3.25, -4.2) { $\underline{t}'_2\underline{r}_2$ };
+
+      \node at (1.95, -4.2) { $\cdots$ };
+
+      \node[node] (tt'2rr'0) at (.75, -1.8) { $\underline{t}'_2\underline{r}'_0$ };
+      \node[node] (tt'2rr'1) at (.75, -3.0) { $\underline{t}'_2\underline{r}'_1$ };
+      \node[node] (tt'2rr'2) at (.75, -4.2) { $\underline{t}'_2\underline{r}'_2$ };
+
+      \path
+      (t0r1) edge [bend left=20] node [above] { $\frac{1}{1}$ } (t'0r1)
+      (t0r1) edge node [above] { $\frac{1}{1}$ } (t'1r1)
+      (t0r1) edge [bend right=20] node [below] { $\frac{1}{1}$ } (t'2r1)
+
+%      (t'1r1) edge node [above] { $\frac{1}{1}$ } (t'1r'0)
+%      (t'1r1) edge node [above] { $\frac{1}{1}$ } (t'1r'1)
+%      (t'1r1) edge node [above] { $\frac{1}{1}$ } (t'1r'2)
+
+      (t'1r'0) edge [double] node [above] { $0|0, 0|1$ } (t'1r0)
+      (t'1r'0) edge [double, bend left=10] node [above] 
+              { $1|0, 1|1, 1|2$ } (t'1r1)
+      (t'1r'0) edge [double, bend left=20] node [below] 
+              { $2|1, 2|2$ } (t'1r2)
+
+      (tt0rr0) edge [bend right=20] node [above] { $\frac{1}{1}$ } (tt'0rr0)
+      (tt0rr0) edge node [above] { $\frac{1}{1}$ } (tt'1rr0)
+      (tt0rr0) edge [bend left=20] node [below] { $\frac{1}{1}$ } (tt'2rr0)
+
+      (tt'2rr'0) edge [double] node [above] { $0|0, 1|0$ } (tt'2rr0)
+      (tt'2rr'0) edge [double] node [above] { $0|1, 1|1, 2|1$ } (tt'2rr1)
+      (tt'2rr'0) edge [double] node [above] { $1|2, 2|2$ } (tt'2rr2)
+
+      (tt'2rr'1) edge [double] node [above] { $0|0, 1|0$ } (tt'2rr0)
+      (tt'2rr'1) edge [double] node [above] { $0|1, 1|1, 2|1$ } (tt'2rr1)
+      (tt'2rr'1) edge [double] node [above] { $1|2, 2|2$ } (tt'2rr2)
+
+      ;
+      \end{tikzpicture}
+    }
+  \caption{Markov Decision Process for Above Threshold}
+  \label{figure:mdp-above-threshold}
+\end{figure}
+
 \subsection{Above Threshold Mechanism}

### 2018-04-26 above-threshold.tex
remove streaming and model above threshold to mdp

--- 2018-04-19 examples.tex
@@
-To model differentially private mechanisms by Markov chains, we
-formalize inputs (such as databases or query results) as states. 
-Randomized computation is modeled by probabilistic transitions. Atomic
-propositions are used to designate intended interpretation on states
-(such as inputs or outputs). We demonstrates the ideas in examples.
-
-\hide{
-For
-each input, its output distribution is modeled as a probabilistic
-transition. Finally, neighbors are specified by a neighborhood
-relation. States are typically labelled with atomic propositions to
-signify their intended interpretation (such as inputs or outputs). We
-demonstrate the key ideas in examples.
-
-A simple way to design differentially private mechanisms is to add
-random noises.
-Now we describe our main ideas for  modelling differentially private mechanisms.
-A state $s$ stores at least a database $x\in \mathcal{X}^n$, which is the data held by a trusted \emph{curator}. A transition corresponds to a data analysis mechanism.
-States are typically labelled with the corresponding outputs
-the curator generates. We explain the main idea with some examples.
-}
-
-\subsection{Survey Mechanism}
-\label{subsec:survey}
-
-Consider the survey question: have you been diagnosed
-with the disease $X$? In order to protect surveyees' privacy, each
-answers the question as follows. The surveyee first flips a
-coin. If it is tail, she answers the question truthfully. Otherwise,
-she randomly answers \textit{Yes} or \textit{No}
-uniformly (Figure~\ref{figure:2-dp-table})~\cite{DR:14:AFDP}.
-
-Let us analyze the mechanism briefly. The data universe $\mathcal{X}$
-is $\{ +, - \}$. The mechanism $M$ is a randomized algorithm with
-inputs in $\mathcal{X}$ and outputs in $\{ Y, N \}$. For any $x \in
-\mathcal{X}$, we have $\frac{1}{4} \leq \Pr[M (x) = Y] \leq
-\frac{3}{4}$. Hence $\Pr[M (x) = Y] \leq \frac{3}{4} = 3 \cdot
-\frac{1}{4} \leq e^{\ln 3} \Pr[M (x') = Y]$ for any neighbors $x, x'
-\in \mathcal{X}$. Similarly, $\Pr[M (x) = N] \leq e^{\ln 3} \Pr[M (x')
-= N]$. The survey mechanism is $(\ln 3, 0)$-differentially private.
-The random noise boosts the probability of answering
-\textit{Yes} or \textit{No} to at least $\frac{1}{4}$ regardless of
-diagnoses. Inferences on individual diagnosis can be plausibly denied.
-
-\begin{figure}
-  \centering
-  \begin{subfigure}{.48\columnwidth}
-      \[
-      \begin{array}{|c|c|c|}
-        \hline
-        &
-        \multicolumn{2}{c|}{output}
-        \\
-        \hline
-        state & Y & N \\
-        \hline
-        + & \frac{3}{4} = \frac{1}{2} \cdot \frac{1}{2} + \frac{1}{2} \cdot 1
-          & \frac{1}{4} = \frac{1}{2} \cdot \frac{1}{2} + \frac{1}{2} \cdot 0
-        \\
-        \hline
-        - & \frac{1}{4} = \frac{1}{2} \cdot \frac{1}{2} + \frac{1}{2} \cdot 0
-          & \frac{3}{4} = \frac{1}{2} \cdot \frac{1}{2} + \frac{1}{2} \cdot 1
-        \\
-        \hline
-      \end{array}
-      \]
-    \caption{Survey Mechanism}
-    \label{figure:2-dp-table}
-  \end{subfigure}
-  \hspace{.05\columnwidth}
-  \begin{subfigure}{.40\columnwidth}
-    \resizebox{\columnwidth}{!}{
-    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
-      distance=2cm,node/.style={circle,draw}]
-      \node[node] (p) at ( 0,  .5) { $+$ };
-      \node[node] (n) at ( 0, -.5) { $-$ };
-      \node[node] (Y) at (-1.5,  0) { $s$ };
-      \node at (-1.5, .5) { $\mathit{out}_Y$ };
-      \node[node] (N) at ( 1.5,  0) { $t$ };
-      \node at ( 1.5, .5) { $\mathit{out}_N$ };
-
-      \path
-      (p) edge node [above] { $\frac{3}{4}$ } (Y)
-      (p) edge node [above] { $\frac{1}{4}$ } (N)
-      (n) edge node [below] { $\frac{1}{4}$ } (Y)
-      (n) edge node [below] { $\frac{3}{4}$ } (N)
-      ;
-      \end{tikzpicture}
-    }
-    \caption{Corresponding Markov Chain}
-    \label{figure:2-dp-mdp}
-  \end{subfigure}
-  \caption{Survey Mechanism with $\ln 3$-Differential Privacy}
-  \label{figure:2-dp}
-\end{figure}
-
-Figure~\ref{figure:2-dp-mdp} shows the corresponding Markov chain.
-In the figure, the states
-$+$ and $-$ denote positive or negative diagnoses 
-respectively; the states $s$ and $t$ denote answers to the survey
-question and hence $\mathit{out}_Y \in L (s)$ and $\mathit{out}_N
-\in L (t)$.
-States $+$ and $-$ are neighbors.
-Missing transitions (such as those from $s$ and $t$) lead
-to a special state $\dagger$ with a self-loop. We omit such transitions
-and the state $\dagger$ for clarity.
-
-\subsection{Truncated $\alpha$-Geometric Mechanism}\label{subsec:geometric}
-More sophisticated differentially private mechanisms are
-available. Consider a query
-$f : \mathcal{X}^n \rightarrow \{ 0, 1, \ldots, m \}$. Let $\alpha \in (0, 1)$.
-The \emph{$\alpha$-geometric mechanism}
-outputs $f(x) + Y$ where $Y$ is a random variable with the geometric
-distribution~\cite{GRS:09:UUPM,GRS:12:UUPM} :
-\[
-\Pr[Y = y] = \frac{1 - \alpha}{1 + \alpha}\alpha^{|y|}
-\textmd{ for } y \in \bbfZ.
-\]
-The $\alpha$-geometric mechanism is oblivious since it has the same
-output distribution on any inputs $x, x'$ with $f (x) = f
-(x')$. It is $(- {\Delta (f)} \ln \alpha, 0)$-differentially
-private for any query $f$ with sensitivity $\Delta (f)$. The range of
-the mechanism
-however is $\bbfZ$. It may give nonsensical outputs such as
-negative integers for non-negative queries.
-
-The \emph{truncated $\alpha$-geometric mechanism over $\{ 0, 1,
-  \ldots, m \}$}
-outputs $f (x) + Z$ where $Z$ is a random variable with the following
-distribution:
-\[
-\Pr[Z = z] =
-\left\{
-  \begin{array}{ll}
-    0 & \textmd{ if } z < - f (x) \\
-    \frac{\alpha^{f (x)}}{1 + \alpha} & \textmd{ if } z = -f (x)\\
-    \frac{1 - \alpha}{1 + \alpha}\alpha^{|z|} &
-    \textmd{ if } -f (x) < z < m - f (x)\\
-    \frac{\alpha^{m - f (x)}}{1 + \alpha} & \textmd{ if } z = m-f (x)\\
-    0 & \textmd{ if } z > m - f (x)
-  \end{array}
-\right.
-\]
-Note the range of the truncated $\alpha$-geometric mechanism is
-$\{ 0, 1, \ldots, m \}$. The truncated $\alpha$-geometric mechanism is
-again oblivious; it is also $(- {\Delta (f)} \ln \alpha, 0)$-differentially
-private for any query $f$ with sensitivity $\Delta (f)$.
-The truncated $\frac{1}{2}$-geometric mechanism over $\{ 0, 1, \ldots, 5 \}$ is
-given in Figure~\ref{figure:geometric-mechanism-table}.
-
-\begin{figure}[tbh]
-  \centering
-  \begin{subfigure}{.40\columnwidth}
-    \[
-    \begin{array}{|c|c|c|c|c|c|c|}
-      \hline
-      \textmd{input/output} & 0 & 1 & 2 & 3 & 4 & 5\\
-      \hline
-      0 & {2}/{3}  & {1}/{6} & {1}/{12} & {1}/{24}  & {1}/{48} & {1}/{48}\\
-      \hline
-      1 & {1}/{3}  & {1}/{3} & {1}/{6} & {1}/{12}  & {1}/{24} & {1}/{24}\\
-      \hline
-      2 & {1}/{6}  & {1}/{6} & {1}/{3} & {1}/{6}   & {1}/{12} & {1}/{12}\\
-      \hline
-      3 & {1}/{12} & {1}/{12} & {1}/{6} & {1}/{3}   & {1}/{6} & {1}/{6}\\
-      \hline
-      4 & {1}/{24} & {1}/{24} & {1}/{12} & {1}/{6}   & {1}/{3} & {1}/{3}\\
-      \hline
-      5 & {1}/{48} & {1}/{48} & {1}/{24} & {1}/{12} & {1}/{6} & {2}/{3}\\
-      \hline
-    \end{array}
-    \]
-    \caption{$\frac{1}{2}$-Geometric Mechanism}
-    \label{figure:geometric-mechanism-table}
-  \end{subfigure}
-  \hspace{.08\columnwidth}
-  \begin{subfigure}{.50\columnwidth}
-    \centering
-  \resizebox{.85\columnwidth}{!}{
-    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
-      distance=2cm,node/.style={circle,draw}]
-      \node[node] (i0) at (0,  1) { $s_0$ };
-      \node at (0, 1.5) { $\mathit{in}_0$ };
-      \node at (0, 0) { $\vdots$ };
-      \node at (.8, -.8) { $\vdots$ };
-      \node[node] (i5) at (0, -1) { $s_5$ };
-      \node at (0, -.5) { $\mathit{in}_5$ };
-
-      \node[node] (o0) at (4,  2.5) { $t_0$ };
-      \node at (4, 3) { $\mathit{out}_0$ };
-      \node[node] (o1) at (4,  1.5) { $t_1$ };
-      \node at (4, 2) { $\mathit{out}_1$ };
-      \node[node] (o2) at (4,  0.5) { $t_2$ };
-      \node at (4, 1) { $\mathit{out}_2$ };
-      \node[node] (o3) at (4, -0.5) { $t_3$ };
-      \node at (4, 0) { $\mathit{out}_3$ };
-      \node[node] (o4) at (4, -1.5) { $t_4$ };
-      \node at (4, -1) { $\mathit{out}_4$ };
-      \node[node] (o5) at (4, -2.5) { $t_5$ };
-      \node at (4, -2) { $\mathit{out}_5$ };
-
-      \path
-      (i0) edge node [right=30,above=10] { $2/3$ } (o0)
-      (i0) edge node [right=30,above=3] { $1/6$ } (o1)
-      (i0) edge node [right=30,above=-3] { $1/12$ } (o2)
-      (i0) edge node [right=30,below=-5] { $1/24$ } (o3)
-      (i0) edge node [right=30,below] { $1/48$ } (o4)
-      (i0) edge node [right=30,below=8] { $1/48$ } (o5)
-
-      \hide{
-      (i1) edge node [above] {  } (o0)
-      (i1) edge node [above] {  } (o2)
-      (i1) edge node [above] {  } (o3)
-      (i1) edge node [above] {  } (o4)
-      (i1) edge node [above] {  } (o5)
-
-      (i2) edge node [above] {  } (o0)
-      (i2) edge node [above] {  } (o2)
-      (i2) edge node [above] {  } (o3)
-      (i2) edge node [above] {  } (o4)
-      (i2) edge node [above] {  } (o5)
-
-      (i3) edge node [above] {  } (o0)
-      (i3) edge node [above] {  } (o2)
-      (i3) edge node [above] {  } (o3)
-      (i3) edge node [above] {  } (o4)
-      (i3) edge node [above] {  } (o5)
-
-      (i4) edge node [above] {  } (o0)
-      (i4) edge node [above] {  } (o2)
-      (i4) edge node [above] {  } (o3)
-      (i4) edge node [above] {  } (o4)
-      (i4) edge node [above] {  } (o5)
-      }
-
-      (i5) edge node [right=10,above=16] { $1/48$ } (o0)
-      \hide{
-      (i5) edge node [above] {  } (o2)
-      (i5) edge node [above] {  } (o3)
-      (i5) edge node [above] {  } (o4)
-      }
-      (i5) edge node [right=15,below=8] { $2/3$ } (o5)
-      ;
-      \end{tikzpicture}
-    }
-    \caption{Markov Chain}
-    \label{figure:geometric-mechanism-markov-chain}
-  \end{subfigure}
-  \caption{A Markov Chain for $\frac{1}{2}$-Geometric Mechanism}
-  \label{figure:geometric-mechanism}\vspace*{-.5cm}
-\end{figure}
-
-Similar to the survey mechanism, it is straightforward to model the
-truncated $\frac{1}{2}$-geometric mechanism as a Markov chain. 
-Recall that the truncated $\frac{1}{2}$-geometric mechanism is
-oblivious. It suffices to consider $\{ f (x) : x \in \mathcal{X}^n \}$
-instead of $\mathcal{X}^n$ as inputs. 
-Let the state $s_k$ and $t_l$
-denote the input value $k$ and output value $l$ respectively. Define
-the state set $S = \{ s_k, t_k : k \in \{ 0, 1, \ldots, m \} \}$.
-The probability transition $\wp (s_k, t_l)$ is the probability of the
-output value $l$ on the input value $k$ as defined in the
-mechanism. Moreover, we have $\mathit{in}_k \in L (s_k)$ and
-$\mathit{out}_k \in L (t_k)$ for $k \in \{ 0, 1, \ldots, n \}$.
-If $\Delta (f) = 1$, $| f (x) - f (x') | \leq 1$ for every neighbors
-$x, x' \in \mathcal{X}^n$. Subsequently, $s_k$ and $s_l$ are
-neighbors iff $| k - l | \leq 1$ in our model.
-Figure~\ref{figure:geometric-mechanism-markov-chain} gives
-the Markov chain for the truncated
-$\frac{1}{2}$-geometric mechanism over $\{ 0, 1, \ldots, 5 \}$.
+\subsection{Above Threshold Mechanism}
+\label{subsec:threshold}
 
@@
 
-\subsection{Above Threshold Mechanism}
-\label{subsec:threshold}
+\begin{figure}
+  \centering
+    \resizebox{.8\columnwidth}{!}{
+    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
+      distance=2cm,node/.style={circle,draw}]
+      \node[node] (t0r1) at (0, 4.75) { $t_0r_1$ };
+
+      \draw (2.9, 3.95) rectangle (4.9, .05);
+      \node at (4.65, 3.25) { $\cdots $};
+      \node[node] (t'0r1) at (3.9, 3.25) { $t'_0r_1$ };
+      \node at (3.2, 3.25) { $\cdots $};
+      \draw (-2.8, 3.95) rectangle (2.8, .05);
+      \node[node] (t'1r0) at (2.1, 3.25) { $t'_1r_0$ };
+      \node[node] (t'1r1) at (0, 3.25) { $t'_1r_1$ };
+      \node[node] (t'1r2) at (-2.1, 3.25) { $t'_1r_2$ };
+      \node at (0, 2.1) { $\vdots$ };
+      \node at (-2.1, 2.1) { $\vdots$ };
+      \draw (-2.9, 3.95) rectangle (-4.9, .05);
+      \node at (-3.15, 3.25) { $\cdots$ };
+      \node[node] (t'2r1) at (-3.9, 3.25) { $t'_2r_1$ };
+      \node at (-4.6, 3.25) { $\cdots$ };
+
+      \node[node] (t'1r'0) at (2.1, .75) { $t'_1r'_0$ };
+      \node[node] (t'1r'1) at (0, .75) { $t'_1r'_1$ };
+      \node[node] (t'1r'2) at (-2.1, .75) { $t'_1r'_2$ };
+
+      \node[node] (tt0rr0) at (2.1, -4.75) { $\underline{t}_0\underline{r}_0$ };
+
+      \draw (2.9, -3.95) rectangle (4.9, -.05);
+      \node[node] (tt'0rr0) at (4.2, -3.25) { $\underline{t}'_0\underline{r}_0$ };
+      \node at (3.35, -3.25) { $\cdots$ };
+      \draw (.8, -3.95) rectangle (2.8, -.05);
+      \node[node] (tt'1rr0) at (2.1, -3.25) { $\underline{t}'_1\underline{r}_0$ };
+      \node at (1.25, -3.25) { $\cdots$ };
+
+      \draw (.7, -3.95) rectangle (-4.9, -.05);
+      \node[node] (tt'2rr0) at (0, -3.25) { $\underline{t}'_2\underline{r}_0$ };
+      \node[node] (tt'2rr1) at (-2.1, -3.25) { $\underline{t}'_2\underline{r}_1$ };
+      \node[node] (tt'2rr2) at (-4.2, -3.25) { $\underline{t}'_2\underline{r}_2$ };
+
+      \node at (-4.2, -1.9) { $\vdots$ };
+
+      \node[node] (tt'2rr'0) at (0, -.75) { $\underline{t}'_2\underline{r}'_0$ };
+      \node[node] (tt'2rr'1) at (-2.1, -.75) { $\underline{t}'_2\underline{r}'_1$ };
+      \node[node] (tt'2rr'2) at (-4.2, -.75) { $\underline{t}'_2\underline{r}'_2$ };
+
+      \path
+      (t0r1) edge [bend left=20] node [above] { $\frac{1}{1}$ } (t'0r1)
+      (t0r1) edge node [above] { $\frac{1}{1}$ } (t'1r1)
+      (t0r1) edge [bend right=20] node [below] { $\frac{1}{1}$ } (t'2r1)
+
+%      (t'1r1) edge node [above] { $\frac{1}{1}$ } (t'1r'0)
+%      (t'1r1) edge node [above] { $\frac{1}{1}$ } (t'1r'1)
+%      (t'1r1) edge node [above] { $\frac{1}{1}$ } (t'1r'2)
+
+      (t'1r'0) edge [double] node [above] { $0|0, 0|1$ } (t'1r0)
+      (t'1r'0) edge [double, bend left=10] node [above] 
+              { $1|0, 1|1, 1|2$ } (t'1r1)
+      (t'1r'0) edge [double, bend left=20] node [below] 
+              { $2|1, 2|2$ } (t'1r2)
+
+      (tt0rr0) edge [bend right=20] node [above] { $\frac{1}{1}$ } (tt'0rr0)
+      (tt0rr0) edge node [above] { $\frac{1}{1}$ } (tt'1rr0)
+      (tt0rr0) edge [bend left=20] node [below] { $\frac{1}{1}$ } (tt'2rr0)
+
+      (tt'2rr'0) edge [double] node [above] { $0|0, 1|0$ } (tt'2rr0)
+      (tt'2rr'0) edge [double] node [above] { $0|1, 1|1, 2|1$ } (tt'2rr1)
+      (tt'2rr'0) edge [double] node [above] { $1|2, 2|2$ } (tt'2rr2)
+
+      (tt'2rr'1) edge [double] node [above] { $0|0, 1|0$ } (tt'2rr0)
+      (tt'2rr'1) edge [double] node [above] { $0|1, 1|1, 2|1$ } (tt'2rr1)
+      (tt'2rr'1) edge [double] node [above] { $1|2, 2|2$ } (tt'2rr2)
+
+      ;
+      \end{tikzpicture}
+    }
+  \caption{Markov Decision Process for Above Threshold}
+  \label{figure:mdp-above-threshold}
+\end{figure}
+
 In differential privacy, an offline mechanism releases outputs only

### 2018-04-26 examples.tex
adjust size
rotate figure
remove streaming and model above threshold to mdp

--- 2018-04-19 examples.tex
@@
 
-\begin{algorithm}[tbh]
-	\begin{algorithmic}[1]
-		\Procedure{AboveThreshold}{$d$, $\{ f_1, f_2, \ldots
-                  \}$, $t$}
-		\Match{$t$}
-		\Comment{obtain $t'$ by $\frac{1}{4}$-geometric mechanism}
-		\lCase{$0$}{$t' \leftarrow 0, 1,2,3,4,5$ with probability
-			$\frac{4}{5}$,$\frac{3}{20}$,
-			$\frac{3}{80}$,$\frac{3}{320}$,
-			$\frac{3}{1280}$,$\frac{1}{1280}$ respectively}
-		\lCase{$1$}{$t' \leftarrow 0, 1,2,3,4,5$ with probability
-			$\frac{1}{5}$,$\frac{3}{5}$,
-			$\frac{3}{20}$,$\frac{3}{80}$,
-			$\frac{3}{320}$,$\frac{1}{320}$ respectively}
-		\lCase{$2$}{$t' \leftarrow 0, 1,2,3,4,5$ with probability
-			$\frac{1}{20}$,$\frac{3}{20}$,
-			$\frac{3}{5}$,$\frac{3}{20}$,
-			$\frac{3}{80}$,$\frac{1}{80}$ respectively}
-		\lCase{$3$}{$t' \leftarrow 0, 1,2,3,4,5$ with probability
-			$\frac{1}{80}$,$\frac{3}{80}$,
-			$\frac{3}{20}$,$\frac{3}{5}$,
-			$\frac{3}{20}$,$\frac{1}{20}$ respectively}
-		\lCase{$4$}{$t' \leftarrow 0, 1,2,3,4,5$ with probability
-			$\frac{1}{320}$,$\frac{3}{320}$,
-			$\frac{3}{80}$,$\frac{3}{20}$,
-			$\frac{3}{5}$,$\frac{1}{5}$ respectively}
-		\lCase{$5$}{$t' \leftarrow 0, 1,2,3,4,5$ with probability
-			$\frac{1}{1280}$,$\frac{3}{1280}$,
-			$\frac{3}{320}$,$\frac{3}{80}$,
-			$\frac{3}{20}$,$\frac{4}{5}$ respectively}
-		\EndMatch
-		\For{each query $f_i$}
-                \State{$r_i \leftarrow f_i (d)$}
-		\Match{ $r_i$}
-		\Comment{obtain $r_{i}'$ by $\frac{1}{2}$-geometric mechanism}
-		\lCase{$0$}{$r_{i}' \leftarrow 0, 1,2,3,4,5$ with probability
-			$\frac{2}{3}$,$\frac{1}{6}$,
-			$\frac{1}{12}$,$\frac{1}{24}$,
-			$\frac{1}{48}$,$\frac{1}{48}$ respectively}
-		\lCase{$1$}{$r_{i}' \leftarrow 0, 1,2,3,4,5$ with probability
-			$\frac{1}{3}$,$\frac{1}{3}$,
-			$\frac{1}{6}$,$\frac{1}{12}$,
-			$\frac{1}{24}$,$\frac{1}{24}$ respectively}
-		\lCase{$2$}{$r_{i}' \leftarrow 0, 1,2,3,4,5$ with probability
-			$\frac{1}{6}$,$\frac{1}{6}$,
-			$\frac{1}{3}$,$\frac{1}{6}$,
-			$\frac{1}{12}$,$\frac{1}{12}$ respectively}
-		\lCase{$3$}{$r_{i}' \leftarrow 0, 1,2,3,4,5$ with probability
-			$\frac{1}{12}$,$\frac{1}{12}$,
-			$\frac{1}{6}$,$\frac{1}{3}$,
-			$\frac{1}{6}$,$\frac{1}{6}$ respectively}
-		\lCase{$4$}{$r_{i}' \leftarrow 0, 1,2,3,4,5$ with probability
-			$\frac{1}{24}$,$\frac{1}{24}$,
-			$\frac{1}{12}$,$\frac{1}{6}$,
-			$\frac{1}{3}$,$\frac{1}{3}$ respectively}
-		\lCase{$5$}{$r_{i}' \leftarrow 0, 1,2,3,4,5$ with probability
-			$\frac{1}{48}$,$\frac{1}{48}$,
-			$\frac{1}{24}$,$\frac{1}{12}$,
-			$\frac{1}{6}$,$\frac{2}{3}$ respectively}
-		\EndMatch
-                \State{\textbf{if} $r'_i \geq t'$ \textbf{then} 
-                  \textbf{halt} with $a_i = \top$ \textbf{else}
-                  $a_i = \bot$}
-\hide{
-		\If{$r_{i}'>t'$}
-		\State{$a_i=\top$}
-		\State{\textbf{halt}}
-		\Else
-		\State{$a_i=\bot$}
-		\EndIf
-}
-		\EndFor
-		\EndProcedure
+\subsection{Subsampling}
+\label{subsection:subsampling}
 
-	\end{algorithmic}
-	\caption{Input: private database $d$, 
-          queries $f_i : \{ 0, 1, \ldots, 5 \} \rightarrow \{ 0, 1,
-          \ldots, 5 \}$ with sensitivity $1$, threshold $t \in \{ 0,
-          1, \ldots, 5 \}$; Output: $a_1, a_2, \ldots$}
-	\label{algorithm:online-model}
-\end{algorithm}
-
-\subsection{Above Threshold Mechanism}
-\label{subsec:threshold}
-In differential privacy, an offline mechanism releases outputs only
-once and plays no further role; an online (interactive) mechanism
-allows analysts to ask queries adaptively based on previous
-responses. The mechanisms we constructed previously are offline
-mechanisms. On the other hand, online mechanisms permit further queries. 
-Obviously, queries do not change databases themselves. To maintain
-privacy, online mechanisms may decide to release
-information differently. For instance, they may disable further
-queries after a certain output is disclosed.
-
-Below we describe an online mechanism adapted from~\cite{DR:14:AFDP}.
-Given a threshold and a series of adaptive queries, we only care for
-the queries whose results are above the threshold; queries below the
-threshold only disclose minimal information and hence are considered
-irrelevant. Since query results above the threshold convey useful
-information, let us assume the mechanism will halt on the first such
-query result for simplicity. The question is to design a
-privacy-preserving mechanism to have similar behaviors on similar
-databases. In~\cite{DR:14:AFDP}, a mechanism is design for continuous
-queries by applying the Laplace mechanism. We will develop a mechanism
-for bounded and discrete queries using the geometric mechanism.
-
-Consider a threshold $t \in
-\{ 0, 1 \ldots, 5\}$, and queries $\{ f_i : f_i $ with sensitivity $\Delta
-(f_i) = 1 \}$. In order to protect privacy, our mechanism applies the
-truncated $\frac{1}{4}$-geometric mechanism to obtain a noisy
-threshold $t'$. For each query, the truncated $\frac{1}{2}$-geometric
-mechanism is applied to its result. If the noisy result is not less
-than the noisy threshold, our mechanism outputs $\top$ and
-halts. Otherwise, it outputs $\bot$ and continues to the next query
-(Algorithm~\ref{algorithm:online-model}). The above threshold
-mechanism outputs a sequence of the form $\bot^* \top$. On
-similar databases, we would like to show that the discrete above
-threshold mechanism outputs the same sequence with similar
-probabilities.
+\input{subsampling}

### 2018-04-26 subsampling.tex
add subsampling algorithm

\begin{algorithm}
  \begin{algorithmic}[1]
    \Function{Subsampling}{$d$, $f$}
    \Require{$d \in \{ R, B \}^3$, $f : \{ R, B \}^3 \rightarrow \{ R,
      B \}$}
    \State{Subsample $m$ data sets $\hat{d}_1, \hat{d}_2, \ldots,
      \hat{d}_m$ from $d$ where each row of $d$ is chosen with
      probability $q$}
    \State{$z \leftarrow \{ f (\hat{d}_1), f (\hat{d}_2), \ldots, f
      (\hat{d}_m) \}$}
    \State{$\mathit{count}_R, \mathit{count}_B \leftarrow 
            | \{ i : f (\hat{d}_i) = R \} |,
            | \{ i : f (\hat{d}_i) = B \} |$}
    \State{$r \leftarrow | \mathit{count}_R - \mathit{count}_B | /
      (4mq) - 1$}
    \State{$\hat{r} \leftarrow r + \mathit{Lap}(\frac{1}{\epsilon})$}
    \If{$\hat{r} > \ln (1/\delta)/\epsilon$}
      \State{\textbf{if} $\mathit{count}_R \geq \mathit{count}_B$
             \textbf{then} \Return { $R$ } \textbf{else} \Return {$B$} } 
    \Else
      \State{\Return {$\bot$}}
    \EndIf
  \EndFunction
  \end{algorithmic}
  \caption{Subsampling}
  \label{algorithm:subsampling}
\end{algorithm}

### 2018-05-01 subsampling.tex
add Markov chain figure

@@
       (4mq) - 1$}
-    \State{$\hat{r} \leftarrow r + \mathit{Lap}(\frac{1}{\epsilon})$}
-    \If{$\hat{r} > \ln (1/\delta)/\epsilon$}
+    \If{${r} + \mathit{Lap}(\frac{1}{\epsilon}) > \ln (1/\delta)/\epsilon$}
       \State{\textbf{if} $\mathit{count}_R \geq \mathit{count}_B$
@@
 \end{algorithm}
+
+\begin{figure}
+  \centering
+    \resizebox{.9\columnwidth}{!}{
+    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
+      distance=2cm,node/.style={circle,draw}]
+
+      \node[node] (rrr) at (-5,  2.25) { $rrr$ };
+%      \node at (-4.4, 2) { $\vdots$ };
+      \node[node] (rrb) at (-5,  0.75) { $rrb$ };
+%      \node at (-4.4, .75) { $\vdots$ };
+      \node[node] (rbb) at (-5, -0.75) { $rbb$ };
+%      \node at (-4.4, -.5) { $\vdots$ };
+      \node[node] (bbb) at (-5, -2.25) { $bbb$ };
+%      \node at (-4.4, -1.75) { $\vdots$ };
+
+      \node[node] (rmb0) at (-2.5,  2.25) { $m,0$ };
+      \node at (-2.5, 0) { $\vdots$ };
+      \node[node] (r0bm) at (-2.5, -2.25) { $0,m$ };
+
+      \node[node] (Rm) at (0,  1.5) { $r_m$ };
+      \node at (0, 0) { $\vdots$ };
+      \node[node] (RM) at (0, -1.5) { $r_M$ };
+
+      \node[node] (top) at (2.5,  0.75) { $\mathit{out}$ };
+      \node[node] (bot) at (2.5, -0.75) { $\bot$ };
+
+      \node[node] (outR) at (5,  1.25) { $R$ };
+      \node[node] (outB) at (5, .25) { $B$ };
+
+      \path
+      (rrr) edge node [above] { } (rmb0)
+      (rrr) edge node [above] { } (r0bm)
+      (rrb) edge node [above] { } (rmb0)
+      (rrb) edge node [above] { } (r0bm)
+      (rbb) edge node [above] { } (rmb0)
+      (rbb) edge node [above] { } (r0bm)
+      (bbb) edge node [above] { } (rmb0)
+      (bbb) edge node [above] { } (r0bm)
+
+      (rmb0) edge node [above] {} (RM)
+      (r0bm) edge node [above] {} (RM)
+
+      (Rm) edge node [above] {} (top)
+      (Rm) edge node [above] {} (bot)
+      (RM) edge node [above] {} (top)
+      (RM) edge node [above] {} (bot)
+
+      (top) edge (outR)
+      (top) edge (outB)
+      ;
+
+      \end{tikzpicture}
+    }
+  \caption{Markov Chain for Subsampling}
+  \label{figure:mc-subsampling}
+\end{figure}

### 2018-05-02 examples.tex
add description of subsampling

@@
 \]
+%Observe that $\Pr[Y=0] \rightarrow 1$ as $\alpha \rightarrow 0$. The
+%$\alpha$-geometric mechanism is less noisy for smaller $\alpha$.
 The $\alpha$-geometric mechanism is oblivious since it has the same
@@
 (x')$. It is $(- {\Delta (f)} \ln \alpha, 0)$-differentially
-private for any query $f$ with sensitivity $\Delta (f)$. The range of
-the mechanism
-however is $\bbfZ$. It may give nonsensical outputs such as
-negative integers for non-negative queries.
+private for any query $f$ with sensitivity $\Delta (f)$. Observe that
+the privay guarantee $(-{\Delta (f)} \ln \alpha, 0)$ depends on the
+sensitivity of the query $f$. To achieve $(\epsilon, 0)$-differential
+privacy using the $\alpha$-geometric mechanism, one first decides the
+sensitivity of the query and then computes the parameter $\alpha =
+e^{-{\epsilon}/{{\Delta (f)}}}$.
 
+The range of the mechanism is $\bbfZ$. It may give nonsensical
+outputs such as negative integers for non-negative queries.
 The \emph{truncated $\alpha$-geometric mechanism over $\{ 0, 1,
   \ldots, m \}$}
-outputs $f (x) + Z$ where $Z$ is a random variable with the following
+outputs $f (x) + Z$ where $Z$ is a random variable with the 
 distribution:
@@
 
-\subsection{Subsampling}
+\subsection{Subsampling Majority}
 \label{subsection:subsampling}

### 2018-05-02 subsampling.tex
add description of subsampling

@@
+The sensitivity of queries is required to apply the (truncated)
+$\alpha$-geometric mechanism. Recall that the sensitivity of a query
+is the maximal difference of query results on any two neighbors. There are
+two problems for such mechanisms depending on query sensitivity. First, 
+sensitivity of queries can be hard to compute or estimate. Second,
+the sensitivity over arbitrary neighbors can be too conservative for
+the actual dataset in use. One therefore would like to have
+mechanisms independent of query sensitivity.
+
+Subsampling is a technique to design such mechanisms. Concretely, let
+us consider $\mathcal{X} = \{ R, B \}$ (for red and blue teams) and a
+dataset $d \in \mathcal{X}^n$. Suppose we would like to ask which team
+is the majority in the dataset while respecting individual
+privacy. This can be achieved as follows. The mechanism first samples $m$
+sub-datasets $\hat{d}_1, \hat{d}_2, \ldots, \hat{d}_m$ from
+$d$. It then computes $m$ query results from the $m$
+sub-datasets. Finally, the mechanism reports a perturbed result in a
+privacy respecting way
+(Algorithm~\ref{algorithm:subsampling-majority}).
+
 \begin{algorithm}
   \begin{algorithmic}[1]
-    \Function{Subsampling}{$d$, $f$}
-    \Require{$d \in \{ R, B \}^3$, $f : \{ R, B \}^3 \rightarrow \{ R,
+    \Function{SubsamplingMajority}{$d$, $f$}
+    \Require{$d \in \{ R, B \}^n$, $f : \{ R, B \}^* \rightarrow \{ R,
       B \}$}
+    \State{$q, m \leftarrow 
+            \frac{\epsilon}{64\ln (1/\delta)},
+            \frac{\log (n/\delta)}{q^2}$}
     \State{Subsample $m$ data sets $\hat{d}_1, \hat{d}_2, \ldots,
@@
       probability $q$}
-    \State{$z \leftarrow \{ f (\hat{d}_1), f (\hat{d}_2), \ldots, f
-      (\hat{d}_m) \}$}
     \State{$\mathit{count}_R, \mathit{count}_B \leftarrow 
@@
   \end{algorithmic}
-  \caption{Subsampling}
-  \label{algorithm:subsampling}
+  \caption{Subsampling Majority}
+  \label{algorithm:subsampling-majority}
 \end{algorithm}
@@
 
-      \node[node] (top) at (2.5,  0.75) { $\mathit{out}$ };
+      \node[node] (top) at (2.5,  0.75) { $\top$ };
       \node[node] (bot) at (2.5, -0.75) { $\bot$ };
@@
     }
-  \caption{Markov Chain for Subsampling}
+  \caption{Markov Chain for Subsampling Majority}
   \label{figure:mc-subsampling}

### 2018-05-02 subsampling.tex
explain Markov chain for subsampling

@@
 is the majority in the dataset while respecting individual
-privacy. This can be achieved as follows. The mechanism first samples $m$
+privacy. This can be achieved as
+follows~(Algorithm~\ref{algorithm:subsampling-majority}). The
+mechanism first samples $m$ 
 sub-datasets $\hat{d}_1, \hat{d}_2, \ldots, \hat{d}_m$ from
-$d$. It then computes $m$ query results from the $m$
-sub-datasets. Finally, the mechanism reports a perturbed result in a
-privacy respecting way
-(Algorithm~\ref{algorithm:subsampling-majority}).
+$d$~(line~\ref{algorithm:subsampling-majority-subsample}). 
+It then computes the majority of each sub-dataset and obtains $m$
+sub-results. Let $\mathit{count}_R$ and $\mathit{count}_B$ be the
+number of sub-datasets with the majority $R$ and $B$
+respectively~(line~\ref{algorithm:subsampling-majority-count}). 
+Since there are $m$ sub-datasets, we have
+$\mathit{count}_R + \mathit{count}_B = m$. To ensure differential
+privacy, the mechanism makes sure the difference $| \mathit{count}_R -
+\mathit{count}_B |$ is significantly large after perturbation. In
+line~\ref{algorithm:subsampling-majority-noise},
+$\mathit{Lap}(p)$ denotes a continuous random
+variable with the Laplace distribution whose probability density
+function is $f(x) = \frac{1}{2p}e^{-|x|/p}$.
+If so,
+it reports the majority of the $m$
+sub-results~(line~\ref{algorithm:subsampling-majority-output}). Otherwise,
+no information is
+revealed~(line~\ref{algorithm:subsampling-majority-mute}).
 
@@
       probability $q$}
+    \label{algorithm:subsampling-majority-subsample}
     \State{$\mathit{count}_R, \mathit{count}_B \leftarrow 
@@
             | \{ i : f (\hat{d}_i) = B \} |$}
+    \label{algorithm:subsampling-majority-count}
     \State{$r \leftarrow | \mathit{count}_R - \mathit{count}_B | /
       (4mq) - 1$}
+    \label{algorithm:subsampling-majority-difference}
     \If{${r} + \mathit{Lap}(\frac{1}{\epsilon}) > \ln (1/\delta)/\epsilon$}
+    \label{algorithm:subsampling-majority-noise}
       \State{\textbf{if} $\mathit{count}_R \geq \mathit{count}_B$
              \textbf{then} \Return { $R$ } \textbf{else} \Return {$B$} } 
+      \label{algorithm:subsampling-majority-output}
     \Else
       \State{\Return {$\bot$}}
+      \label{algorithm:subsampling-majority-mute}
     \EndIf
@@
 \end{algorithm}
+
+Fix the dataset size $n$ and privacy parameters $\epsilon$, $\delta$,
+the subsampling majority mechanism can be modeled by a Markov chain.
+Figure~\ref{figure:subsampling-majority-markov-chain} gives a sketch
+of the Markov chain for $n = 3$. The leftmost four states represent
+all possible datasets. From each dataset, we compute the probability
+of having $(\mathit{count}_R, \mathit{count}_B) = (i, j)$ with $i + j
+= m$ from $m$ sub-datasets. Since $| \mathit{count}_R -
+\mathit{count}_B |$ can have only finitely many values, the number of
+possible values of $r$ 
+(line~\ref{algorithm:subsampling-majority-difference}) is in a finite
+set $\{ r_m, \ldots, r_M \}$. For each $r \in \{ r_m, \ldots, r_M \}$,
+the probability of having $r + \mathit{Lap}(\frac{1}{\epsilon}) > \ln
+(1/\delta)/\epsilon$ (line~\ref{algorithm:subsampling-majority-noise})
+is equal to the probability of $\mathit{Lap}(\frac{1}{\epsilon}) > \ln
+(1/\delta)/\epsilon - r$. This is equal to $\int^{\infty}_{\ln
+  (1/\delta)/\epsilon - r} \frac{\epsilon}{2}e^{-\epsilon|x|}$.
 
@@
   \caption{Markov Chain for Subsampling Majority}
-  \label{figure:mc-subsampling}
+  \label{figure:subsampling-majority-markov-chain}
 \end{figure}

### 2018-05-16 above-threshold.tex
shrink figure

@@
 
-\begin{algorithm}[tbh]
+In differential privacy, an offline mechanism releases outputs only
+once and plays no further role; an online (interactive) mechanism
+allows analysts to ask queries adaptively based on previous
+responses. The mechanisms we constructed previously are offline
+mechanisms. On the other hand, online mechanisms permit further queries. 
+To maintain
+privacy, online mechanisms may decide to release
+information differently. For instance, they may disable further
+queries after a certain output is disclosed.
+
+Below we describe an online mechanism adapted from~\cite{DR:14:AFDP}.
+Given a threshold and a series of adaptive queries, we only care for
+the queries whose results are above the threshold; queries below the
+threshold only disclose minimal information and hence are considered
+irrelevant. Since query results above the threshold convey useful
+information, let us assume the mechanism will halt on the first such
+query result for simplicity. The question is to design a
+privacy-preserving mechanism to have similar behaviors on similar
+databases. In~\cite{DR:14:AFDP}, a mechanism is design for continuous
+queries by applying the Laplace mechanism. We will develop a mechanism
+for bounded and discrete queries using the geometric mechanism.
+
+\begin{algorithm}
 	\begin{algorithmic}[1]
@@
 
-\begin{figure}
-  \centering
-    \resizebox{.8\columnwidth}{!}{
-    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
-      distance=2cm,node/.style={circle,draw}]
-      \node[node] (t0r1) at (0, 4.75) { $t_0r_1$ };
-
-      \draw (2.9, 3.95) rectangle (4.9, .05);
-      \node at (4.65, 3.25) { $\cdots $};
-      \node[node] (t'0r1) at (3.9, 3.25) { $t'_0r_1$ };
-      \node at (3.2, 3.25) { $\cdots $};
-      \draw (-2.8, 3.95) rectangle (2.8, .05);
-      \node[node] (t'1r0) at (2.1, 3.25) { $t'_1r_0$ };
-      \node[node] (t'1r1) at (0, 3.25) { $t'_1r_1$ };
-      \node[node] (t'1r2) at (-2.1, 3.25) { $t'_1r_2$ };
-      \node at (0, 2.1) { $\vdots$ };
-      \node at (-2.1, 2.1) { $\vdots$ };
-      \draw (-2.9, 3.95) rectangle (-4.9, .05);
-      \node at (-3.15, 3.25) { $\cdots$ };
-      \node[node] (t'2r1) at (-3.9, 3.25) { $t'_2r_1$ };
-      \node at (-4.6, 3.25) { $\cdots$ };
-
-      \node[node] (t'1r'0) at (2.1, .75) { $t'_1r'_0$ };
-      \node[node] (t'1r'1) at (0, .75) { $t'_1r'_1$ };
-      \node[node] (t'1r'2) at (-2.1, .75) { $t'_1r'_2$ };
-
-      \node[node] (tt0rr0) at (2.1, -4.75) { $\underline{t}_0\underline{r}_0$ };
-
-      \draw (2.9, -3.95) rectangle (4.9, -.05);
-      \node[node] (tt'0rr0) at (4.2, -3.25) { $\underline{t}'_0\underline{r}_0$ };
-      \node at (3.35, -3.25) { $\cdots$ };
-      \draw (.8, -3.95) rectangle (2.8, -.05);
-      \node[node] (tt'1rr0) at (2.1, -3.25) { $\underline{t}'_1\underline{r}_0$ };
-      \node at (1.25, -3.25) { $\cdots$ };
-
-      \draw (.7, -3.95) rectangle (-4.9, -.05);
-      \node[node] (tt'2rr0) at (0, -3.25) { $\underline{t}'_2\underline{r}_0$ };
-      \node[node] (tt'2rr1) at (-2.1, -3.25) { $\underline{t}'_2\underline{r}_1$ };
-      \node[node] (tt'2rr2) at (-4.2, -3.25) { $\underline{t}'_2\underline{r}_2$ };
-
-      \node at (-4.2, -1.9) { $\vdots$ };
-
-      \node[node] (tt'2rr'0) at (0, -.75) { $\underline{t}'_2\underline{r}'_0$ };
-      \node[node] (tt'2rr'1) at (-2.1, -.75) { $\underline{t}'_2\underline{r}'_1$ };
-      \node[node] (tt'2rr'2) at (-4.2, -.75) { $\underline{t}'_2\underline{r}'_2$ };
-
-      \path
-      (t0r1) edge [bend left=20] node [above] { $\frac{1}{1}$ } (t'0r1)
-      (t0r1) edge node [above] { $\frac{1}{1}$ } (t'1r1)
-      (t0r1) edge [bend right=20] node [below] { $\frac{1}{1}$ } (t'2r1)
-
-%      (t'1r1) edge node [above] { $\frac{1}{1}$ } (t'1r'0)
-%      (t'1r1) edge node [above] { $\frac{1}{1}$ } (t'1r'1)
-%      (t'1r1) edge node [above] { $\frac{1}{1}$ } (t'1r'2)
-
-      (t'1r'0) edge [double] node [above] { $0|0, 0|1$ } (t'1r0)
-      (t'1r'0) edge [double, bend left=10] node [above] 
-              { $1|0, 1|1, 1|2$ } (t'1r1)
-      (t'1r'0) edge [double, bend left=20] node [below] 
-              { $2|1, 2|2$ } (t'1r2)
-
-      (tt0rr0) edge [bend right=20] node [above] { $\frac{1}{1}$ } (tt'0rr0)
-      (tt0rr0) edge node [above] { $\frac{1}{1}$ } (tt'1rr0)
-      (tt0rr0) edge [bend left=20] node [below] { $\frac{1}{1}$ } (tt'2rr0)
-
-      (tt'2rr'0) edge [double] node [above] { $0|0, 1|0$ } (tt'2rr0)
-      (tt'2rr'0) edge [double] node [above] { $0|1, 1|1, 2|1$ } (tt'2rr1)
-      (tt'2rr'0) edge [double] node [above] { $1|2, 2|2$ } (tt'2rr2)
-
-      (tt'2rr'1) edge [double] node [above] { $0|0, 1|0$ } (tt'2rr0)
-      (tt'2rr'1) edge [double] node [above] { $0|1, 1|1, 2|1$ } (tt'2rr1)
-      (tt'2rr'1) edge [double] node [above] { $1|2, 2|2$ } (tt'2rr2)
-
-      ;
-      \end{tikzpicture}
-    }
-  \caption{Markov Decision Process for Above Threshold}
-  \label{figure:mdp-above-threshold}
-\end{figure}
-
-In differential privacy, an offline mechanism releases outputs only
-once and plays no further role; an online (interactive) mechanism
-allows analysts to ask queries adaptively based on previous
-responses. The mechanisms we constructed previously are offline
-mechanisms. On the other hand, online mechanisms permit further queries. 
-Obviously, queries do not change databases themselves. To maintain
-privacy, online mechanisms may decide to release
-information differently. For instance, they may disable further
-queries after a certain output is disclosed.
-
-Below we describe an online mechanism adapted from~\cite{DR:14:AFDP}.
-Given a threshold and a series of adaptive queries, we only care for
-the queries whose results are above the threshold; queries below the
-threshold only disclose minimal information and hence are considered
-irrelevant. Since query results above the threshold convey useful
-information, let us assume the mechanism will halt on the first such
-query result for simplicity. The question is to design a
-privacy-preserving mechanism to have similar behaviors on similar
-databases. In~\cite{DR:14:AFDP}, a mechanism is design for continuous
-queries by applying the Laplace mechanism. We will develop a mechanism
-for bounded and discrete queries using the geometric mechanism.
-
 Consider a threshold $t \in
@@
 probabilities.
+
+\begin{figure}
+  \centering
+    \resizebox{.8\columnwidth}{!}{
+    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
+      distance=2cm,node/.style={circle,draw}]
+      \node[node] (t0r1) at (0, 4.25) { $t_0r_1$ };
+
+      \draw (2.9, 3.45) rectangle (4.9, .05);
+      \node at (4.65, 2.75) { $\cdots $};
+      \node[node] (t'0r1) at (3.9, 2.75) { $t'_0r_1$ };
+      \node at (3.2, 2.75) { $\cdots $};
+      \draw (-2.8, 3.45) rectangle (2.8, .05);
+      \node[node] (t'1r0) at (2.1, 2.75) { $t'_1r_0$ };
+      \node[node] (t'1r1) at (0, 2.75) { $t'_1r_1$ };
+      \node[node] (t'1r2) at (-2.1, 2.75) { $t'_1r_2$ };
+      \node at (0, 1.85) { $\vdots$ };
+      \node at (-2.1, 1.85) { $\vdots$ };
+      \draw (-2.9, 3.45) rectangle (-4.9, .05);
+      \node at (-3.15, 2.75) { $\cdots$ };
+      \node[node] (t'2r1) at (-3.9, 2.75) { $t'_2r_1$ };
+      \node at (-4.6, 2.75) { $\cdots$ };
+
+      \node[node] (t'1r'0) at (2.1, .75) { $t'_1r'_0$ };
+      \node[node] (t'1r'1) at (0, .75) { $t'_1r'_1$ };
+      \node[node] (t'1r'2) at (-2.1, .75) { $t'_1r'_2$ };
+
+      \node[node] (tt0rr0) at (2.1, -4.25) { $\underline{t}_0\underline{r}_0$ };
+
+      \draw (2.9, -3.45) rectangle (4.9, -.05);
+      \node[node] (tt'0rr0) at (4.2, -2.75) { $\underline{t}'_0\underline{r}_0$ };
+      \node at (3.35, -2.75) { $\cdots$ };
+      \draw (.8, -3.45) rectangle (2.8, -.05);
+      \node[node] (tt'1rr0) at (2.1, -2.75) { $\underline{t}'_1\underline{r}_0$ };
+      \node at (1.25, -2.75) { $\cdots$ };
+
+      \draw (.7, -3.45) rectangle (-4.9, -.05);
+      \node[node] (tt'2rr0) at (0, -2.75) { $\underline{t}'_2\underline{r}_0$ };
+      \node[node] (tt'2rr1) at (-2.1, -2.75) { $\underline{t}'_2\underline{r}_1$ };
+      \node[node] (tt'2rr2) at (-4.2, -2.75) { $\underline{t}'_2\underline{r}_2$ };
+
+      \node at (0, -1.65) { $\vdots$ };
+
+      \node[node] (tt'2rr'0) at (0, -.75) { $\underline{t}'_2\underline{r}'_0$ };
+      \node[node] (tt'2rr'1) at (-2.1, -.75) { $\underline{t}'_2\underline{r}'_1$ };
+      \node at (-4.2, -1.65) { $\vdots$ };
+
+      \node[node] (tt'2rr'2) at (-4.2, -.75) { $\underline{t}'_2\underline{r}'_2$ };
+
+      \path
+      (t0r1) edge node [above] { $\frac{1}{1}$ } (t'0r1)
+      (t0r1) edge node [left] { $\frac{1}{1}$ } (t'1r1)
+      (t0r1) edge node [above] { $\frac{1}{1}$ } (t'2r1)
+
+%      (t'1r1) edge node [above] { $\frac{1}{1}$ } (t'1r'0)
+%      (t'1r1) edge node [above] { $\frac{1}{1}$ } (t'1r'1)
+%      (t'1r1) edge node [above] { $\frac{1}{1}$ } (t'1r'2)
+
+      (t'1r'0) edge [double] node [above]
+              { \hide{$0|0, 0|1$} } (t'1r0)
+      (t'1r'0) edge [double] node [above] 
+              { \hide{$1|0, 1|1, 1|2$} } (t'1r1)
+      (t'1r'0) edge [double] node [below] 
+              { \hide{$2|1, 2|2$} } (t'1r2)
+
+      (tt0rr0) edge node [below] { $\frac{1}{1}$ } (tt'0rr0)
+      (tt0rr0) edge node [left] { $\frac{1}{1}$ } (tt'1rr0)
+      (tt0rr0) edge node [below] { $\frac{1}{1}$ } (tt'2rr0)
+
+%      (tt'2rr'0) edge [double] node [above] { $0|0, 1|0$ } (tt'2rr0)
+%      (tt'2rr'0) edge [double] node [above] { $0|1, 1|1, 2|1$ } (tt'2rr1)
+%      (tt'2rr'0) edge [double] node [above] { $1|2, 2|2$ } (tt'2rr2)
+
+      (tt'2rr'1) edge [double] node [above]
+                { \hide{$0|0, 1|0$} } (tt'2rr0)
+      (tt'2rr'1) edge [double] node [above]
+                { \hide{$0|1, 1|1, 2|1$} } (tt'2rr1)
+      (tt'2rr'1) edge [double] node [above]
+                { \hide{$1|2, 2|2$} } (tt'2rr2)
+
+      ;
+      \end{tikzpicture}
+    }
+  \caption{Markov Decision Process for Above Threshold}
+  \label{figure:mdp-above-threshold}
+\end{figure}

### 2018-05-16 subsampling.tex
shrink figure

@@
 of the Markov chain for $n = 3$. The leftmost four states represent
-all possible datasets. From each dataset, we compute the probability
-of having $(\mathit{count}_R, \mathit{count}_B) = (i, j)$ with $i + j
-= m$ from $m$ sub-datasets. Since $| \mathit{count}_R -
-\mathit{count}_B |$ can have only finitely many values, the number of
-possible values of $r$ 
-(line~\ref{algorithm:subsampling-majority-difference}) is in a finite
-set $\{ r_m, \ldots, r_M \}$. For each $r \in \{ r_m, \ldots, r_M \}$,
+all possible datasets. Given a dataset, $m$ samples are taken with
+replacement. Outcomes of these samples are denoted by
+$(\mathit{count}_R, \mathit{count}_B)$. There are only $m + 1$
+outcomes from $(m, 0), (m - 1, 1), \ldots, (0, m)$. Each outcome is
+represetned by a state in
+Figure~\ref{figure:subsampling-majority-markov-chain}. From each
+dataset, the probability distribution of all outcomes gives the
+transition probability in the figure. Next, observe that  $|
+\mathit{count}_R - \mathit{count}_B |$ can have only finitely many
+values. The values of $r$ 
+(line~\ref{algorithm:subsampling-majority-difference}) hence belong to
+a finite set $\{ r_m, \ldots, r_M \}$ with the minimum $r_m$ and
+maximum $r_M$. For each $r \in \{ r_m, \ldots,
+r_M \}$, it is represented by a state in
+Figure~\ref{figure:subsampling-majority-markov-chain}. For instance,
+both outcomes $(m, 0)$ and $(0, m)$ transit to the state $r_M = 1/(4q)
+- 1$ with probability $1$. For each $r \in \{ r_m, \ldots, r_M \}$,
 the probability of having $r + \mathit{Lap}(\frac{1}{\epsilon}) > \ln
@@
 (1/\delta)/\epsilon - r$. This is equal to $\int^{\infty}_{\ln
-  (1/\delta)/\epsilon - r} \frac{\epsilon}{2}e^{-\epsilon|x|}$.
+  (1/\delta)/\epsilon - r} \frac{\epsilon}{2}e^{-\epsilon|x|} dx$.
+From each state $r \in \{ r_m, \ldots, r_M \}$, it thus goes to the state
+$\top$ with probability $\int^{\infty}_{\ln (1/\delta)/\epsilon - r}
+\frac{\epsilon}{2}e^{-\epsilon|x|} dx$ and to the state $\bot$ with
+probability $1 - \int^{\infty}_{\ln (1/\delta)/\epsilon - r}
+\frac{\epsilon}{2}e^{-\epsilon|x|} dx$. Finally, the Markov chain
+moves to the state $R$ if $\mathit{count}_R \geq \mathit{count}_B$;
+otherwise, it moves to $B$.
 

### 2018-05-31 mdp.tex
remove streaming and model above threshold to mdp
add todos
explain how to computer upper bound by classical schedulers

@@
 Unlike non-interactive mechanisms, the privacy guarantees vary from accuracy requests.
+
+\input{above-threshold}
 
@@
 }
+
+\todo{explain classical schedulers as upperbound of difference}
+
+Despite of the negative result in Theorem~\ref{theorem:mdp-model-checking},
+a sufficient condition for $M, \neighbor{S}, s \models
+\dpriv{\epsilon}{\delta} \phi$ can still be checked
+efficiently. To see this, observe that for any state $s \in S$ and
+query scheduler $\scheduler{Q}$, we have
+\[
+  \min_{\scheduler{S}} \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} 
+  \leq
+  \myPr{s}{M_{\scheduler{Q}}}{\neighbor{S}}{\phi} 
+  \leq
+  \max_{\scheduler{S}} \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} 
+\]
+where the minimum and maximum are taken over all (not necessarily
+history independent) schedulers $\scheduler{S}$. Hence,
+\begin{eqnarray*}
+  \myPr{s}{M_{\scheduler{Q}}}{\neighbor{S}}{\phi} -
+      e^\epsilon \myPr{t}{M_{\scheduler{Q}}}{\neighbor{S}}{\phi}
+  \leq
+    \max_{\scheduler{S}} \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}
+    - e^\epsilon \cdot
+    \min_{\scheduler{S}} \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}
+\end{eqnarray*}
+for any $s, t \in S$ and query scheduler $\scheduler{Q}$. We have the
+following proposition:
+\begin{proposition}\label{proposition:sufficient-condition-mdp-model-checking}
+  Let $M = (S, \Act, \wp, L)$ be an MDP, $\neighbor{S}$ a
+  neighborhood relation on $S$.
+  $M, \neighbor{S}, s \models \dpriv{\epsilon}{\delta} \phi$ if
+  \begin{itemize}
+  \item $\max\limits_{\scheduler{S}} \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}
+    - e^\epsilon \cdot
+    \min\limits_{\scheduler{S}} \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}
+    \leq \delta$; and
+  \item $\max\limits_{\scheduler{S}} \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}
+    - e^\epsilon \cdot
+    \min\limits_{\scheduler{S}} \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}
+    \leq \delta$.
+  \end{itemize}
+  for any $s, t \in S$ with $s \neighbor{S} t$.
+\end{proposition}
+
+For any $s \in S$, recall that $\min\limits_{\scheduler{S}}
+\myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}$ can be efficiently
+computed by iterative methods or linear programming~\cite{BK:08:PMC}. 
+Using
+Proposition~\ref{proposition:sufficient-condition-mdp-model-checking},
+$M, \neighbor{S}, s \models \dpriv{\epsilon}{\delta} \phi$ can be
+checked soundly.
+
+\todo{explain experiments}

### 2018-05-31 mdp.tex
minor fix

@@
 $M, \neighbor{S}, s \models \dpriv{\epsilon}{\delta} \phi$ can be
-checked soundly.
+checked soundly and efficiently.
 

### 2018-06-05 above-threshold.tex
add todos
change figure
explain how to computer upper bound by classical schedulers
reorganize sections

@@
-\subsection{Above Threshold Mechanism}
-\label{subsec:threshold}
-
-In differential privacy, an offline mechanism releases outputs only
-once and plays no further role; an online (interactive) mechanism
-allows analysts to ask queries adaptively based on previous
-responses. The mechanisms we constructed previously are offline
-mechanisms. On the other hand, online mechanisms permit further queries. 
-To maintain
-privacy, online mechanisms may decide to release
-information differently. For instance, they may disable further
-queries after a certain output is disclosed.
-
-Below we describe an online mechanism adapted from~\cite{DR:14:AFDP}.
-Given a threshold and a series of adaptive queries, we only care for
+Below we describe an online mechanism from~\cite{DR:14:AFDP}.
+Given a threshold and a series of adaptive queries, we care for
 the queries whose results are above the threshold; queries below the
@@
 information, let us assume the mechanism will halt on the first such
-query result for simplicity. The question is to design a
-privacy-preserving mechanism to have similar behaviors on similar
-databases. In~\cite{DR:14:AFDP}, a mechanism is design for continuous
+query result for simplicity. 
+%The question is to design a privacy-preserving mechanism to have 
+%similar behaviors on similar databases. 
+In~\cite{DR:14:AFDP}, a mechanism is design for continuous
 queries by applying the Laplace mechanism. We will develop a mechanism
@@
 (f_i) = 1 \}$. In order to protect privacy, our mechanism applies the
-truncated $\frac{1}{4}$-geometric mechanism to obtain a noisy
+truncated $\frac{1}{4}$-geometric mechanism to obtain a perturbed
 threshold $t'$. For each query, the truncated $\frac{1}{2}$-geometric
-mechanism is applied to its result. If the noisy result is not less
-than the noisy threshold, our mechanism outputs $\top$ and
+mechanism is applied to its result. If the perturbed result is not less
+than the perturbed threshold, our mechanism outputs $\top$ and
 halts. Otherwise, it outputs $\bot$ and continues to the next query
@@
 mechanism outputs a sequence of the form $\bot^* \top$. On
-similar databases, we would like to show that the discrete above
+similar databases, we would like to show that the above
 threshold mechanism outputs the same sequence with similar
@@
       \path
-      (t0r1) edge node [above] { $\frac{1}{1}$ } (t'0r1)
-      (t0r1) edge node [left] { $\frac{1}{1}$ } (t'1r1)
-      (t0r1) edge node [above] { $\frac{1}{1}$ } (t'2r1)
+      (t0r1) edge node [above] {  } (t'0r1)
+      (t0r1) edge node [left] {  } (t'1r1)
+      (t0r1) edge node [above] {  } (t'2r1)
 
@@
 
-      (tt0rr0) edge node [below] { $\frac{1}{1}$ } (tt'0rr0)
-      (tt0rr0) edge node [left] { $\frac{1}{1}$ } (tt'1rr0)
-      (tt0rr0) edge node [below] { $\frac{1}{1}$ } (tt'2rr0)
+      (tt0rr0) edge node [below] {  } (tt'0rr0)
+      (tt0rr0) edge node [left] {  } (tt'1rr0)
+      (tt0rr0) edge node [below] {  } (tt'2rr0)
 

### 2018-06-05 main.tex
remove streaming and model above threshold to mdp
reorganize sections

@@
 Preliminaries are given in Section~\ref{section:preliminaries}.
-In Section~\ref{section:examples} we discuss how differentially private mechanisms can be modeled in Markov chains.
-The logic $\dpCTLstar$ and model checking algorithms for Markov chains are presented in
-Section~\ref{section:dpCTL}.
+In Section~\ref{section:examples} we discuss how differentially private mechanisms can be modeled as Markov chains.
+The logic $\dpCTLstar$ and its syntax are presented in Section~\ref{section:dpCTL}. The semantics over Markov chains and its model checking algorithm are given in Section~\ref{section:semantics}.
 Section~\ref{section:specifying-properties} describes differential privacy properties using $\dpCTLstar$.
-They are followed by the extension to Markov decision processes and an
-application in stream processing (Section~\ref{section:mdp}). Finally,
+More examples of differentially private mechanisms as Markov decision processes are given in Section~\ref{section:mdp}.
+The semantics over Markov decision processes and undecidability of model checking is discussed in Section~\ref{section:semantics-mdp}.
+Finally,
 Section~\ref{section:conclusions} concludes our presentation.
@@
 
-\section{Markov Chains  for Differentially Private Mechanisms}
+\section{Differentially Private Mechanisms as Markov Chains}
 \label{section:examples}
 \input{examples}
+
+\subsection{Subsampling Majority}
+\label{subsection:subsampling}
+
+\input{subsampling}
 
@@
 
-\subsection{Semantics}
-\label{subsection:semantics}
+\section{$\dpCTLstar$ for Markov Chains}
+\label{section:semantics}
 \input{semantics}
@@
 
-\section{Markov Decision Processes for Differentially Private Mechanisms}
+\section{Differentially Private Mechanisms as Markov Decision Processes}
 \label{section:mdp}
 \input{mdp}
+
+\subsection{Above Threshold Mechanism}
+\label{subsec:threshold}
+\input{above-threshold}
+
+\section{$\dpCTLstar$ for Markov Decision Processes}
+\label{section:semantics-mdp}
+\input{semantics-mdp}
 
@@
 
-\subsection{MDPs for Stream Processing}
-\label{section:applications}
-\input{applications}
+%\subsection{MDPs for Stream Processing}
+%\label{section:applications}
+%\input{applications}
 
@@
 For Markov chains, the model checking problem has the same complexity as for $\PCTLstar$.
-The general MDP model checking problem is undecidable.
-We have discussed that some special cases are decidable.
-An interesting work is to identify more
-decidable subclass. As an example, consider the formula $\bigwedge_{k \in \bbfZ^{\geq 0}}
+The general MDP model checking problem however is undecidable.
+We have discussed some decidable special cases and an efficient and sufficient condition to check differentially private subformulas.
+An interesting furture work is to identify more
+decidable subclasses and sufficient conditions.
+As an example, consider the formula $\bigwedge_{k \in \bbfZ^{\geq 0}}
 \dpriv{\epsilon}{\delta} (\X^k  00)$ which was used in specifying the pan privacy.
@@
 It is interesting to characterize other cases as well.
- Another interesting line of further works is to consider continuous perturbation (such as Laplace distribution used in \cite{DR:14:AFDP}) -- we would need Markov models with continuous state space for the modelling.
+Another interesting line of further works is to consider continuous perturbation (such as Laplace distribution used in \cite{DR:14:AFDP}). We would need Markov models with continuous state space.
 

### 2018-06-05 mdp.tex
reorganize sections

--- 2018-04-19 mdp.tex
@@
-%\subsection{Markov Decision Processes}
+In differential privacy, an \emph{offline} mechanism releases outputs only
+once and plays no further role; an \emph{online} (or \emph{interactive}) mechanism
+allows analysts to ask queries adaptively based on previous
+responses. The mechanisms we constructed previously are offline
+mechanisms. On the other hand, online mechanisms permit further queries. 
+To maintain
+privacy, online mechanisms may decide to release
+information differently. For instance, they may disable further
+queries after a certain output is disclosed.
+
 \begin{wrapfigure}[14]{r}{.45\columnwidth}
@@
 Unlike non-interactive mechanisms, the privacy guarantees vary from accuracy requests.
-
-\subsection{Semantics}
-The logic $\dpCTLstar$ can be interpreted over MDPs. Semantics over Markov chains generalizes to MDPs as well. %Note that two states with different enabled actions
-%are trivially distinguishable; no privacy can be preserved.
-Let $M = (S, \Act, \wp, L)$ be an MDP. Define the satisfaction
-relation $M, \neighbor{S}, s \models \Phi$ for $\PJ{J}$ and $\dpriv{\epsilon}{\delta} \phi$ (others are standard) as follows.
-\begin{eqnarray*}
-%  M, \neighbor{S}, s \models \top\\
-%  M, \neighbor{S}, s \models p  & \textmd{ if } &  p \in L(s)\\
-%  M, \neighbor{S}, s \models \neg \Phi  & \textmd{ if } &  M, \neighbor{S}, s \not\models \Phi\\
-%  M, \neighbor{S}, s \models \Phi_0 \wedge \Phi_1  & \textmd{ if } &  M, \neighbor{S}, s \models \Phi_0 \textmd{ and }
-%  M, \neighbor{S}, s \models \Phi_1\\
-  M, \neighbor{S}, s \models \PJ{J} \phi
-  & \textmd{ if } &
-  \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \in J
-  \textmd{ for every scheduler } \scheduler{S}\\
-  M, \neighbor{S}, s \models \dpriv{\epsilon}{\delta} \phi
-  & \textmd{ if } &
-  \textmd{for all } t \textmd{ with } s \neighbor{S} t \textmd{ and
-   query scheduler } \scheduler{Q},
-   \myPr{s}{M_{\scheduler{Q}}}{\neighbor{S}}{\phi} \leq
-  e^{\epsilon} \cdot
-   \\
-  & &
-   \myPr{t}{M_{\scheduler{Q}}}{\neighbor{S}}{\phi} + \delta
-   \textmd{ and }
-   \myPr{t}{M_{\scheduler{Q}}}{\neighbor{S}}{\phi} \leq e^{\epsilon} \cdot
-   \myPr{s}{M_{\scheduler{Q}}}{\neighbor{S}}{\phi} + \delta
-\end{eqnarray*}
-
-Recall that $M_{\scheduler{S}}$ is but a Markov chain. The semantics
-of $M_{\scheduler{S}}, \neighbor{S}, \pi \models \phi$ and hence the
-probability $\myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}$ are
-defined as in Markov chains.
-The semantics of $\dpCTLstar$ on MDPs
-is again standard except the differentially private operator
-$\dpriv{\epsilon}{\delta}$. For any path formula $\phi$,
-$\dpriv{\epsilon}{\delta} \phi$ specifies states whose probability
-of having paths satisfying $\phi$ are $(\epsilon, \delta)$-close to
-those of all its neighbors for query schedulers. That is, no
-query scheduler can force any of neighbors to have
-probabilistically distinguishable behaviors.
-
-\noindent
-\paragraph{Justification of query schedulers.}
-We use query schedulers in the semantics for the differentially
-private operator. A definition with history-dependent schedulers
-might be
-\begin{eqnarray*}
-  M, \neighbor{S}, s \models \dpriv{\epsilon}{\delta}^{\mathit{bad}} \phi
-  & \textmd{ if } &
-  \textmd{for all } t \textmd{ with } s \neighbor{S} t \textmd{ and
-  scheduler } \scheduler{S},
-  \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq
-  e^{\epsilon} \cdot\\
-  && \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} + \delta \textmd{ and }
-  \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq
-  e^{\epsilon} \cdot \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}
-  + \delta
-\end{eqnarray*}
-A state satisfies $\dpriv{\epsilon}{\delta}^{\mathit{bad}} \phi$ if
-no history-dependent scheduler could differentiate the probabilities
-of haveing paths satisfying
-$\phi$ from neighbors. Recall that a history-dependent scheduler takes
-actions by previous states. Such a definition would allow
-schedulers to take different actions from different states. 
-Two neighbors could be differentiated by different action sequences
-erroneously.
-A query scheduler $\scheduler{Q} : \bbfZ^{\geq 0} \rightarrow \Act$,
-on the other hand, corresponds to a query sequence. A state satisfies
-$\dpriv{\epsilon}{\delta} \phi$ if no query sequence can differentiate
-the probabilities of having paths satisfying $\phi$ from neighbors.
-Our semantics agrees with the informal interpretation of differential
-privacy for such systems. We hence
-prefer the original definition.
-
-\subsection{The Model Checking Problem}
-For an MDP $M =
-(S, \Act, \wp, L)$, we want to compute $\{ M, \neighbor{S}, s \models
-\dpriv{\epsilon}{\delta} \phi \}$ . Recall the semantics of $\dpriv{\epsilon}{\delta}
-\phi$. Given $s, t$ with $s \neighbor{S} t$ and a path formula
-$\phi$, we need to decide whether
-$\myPr{s}{M_{\scheduler{Q}}}{\neighbor{S}}{\phi} \leq
-e^{\epsilon} \myPr{t}{M_{\scheduler{Q}}}{\neighbor{S}}{\phi} + \delta$
-for every query scheduler $\scheduler{Q}$.
-If the path formula only contains next operators, it is easy to deal with.
-When $\phi$ is $\gX B$ with $B\subseteq S$, only one step behavior needs to be considered.
-Thus only the first action in the query sequence is needed.
-It can also be easily generalized to nested next operators:  one needs only to enumerate
-all actions query sequences of length obtained by the number of nested $\X$ operators.
-The problem however is undecidable in general.
-
-\begin{theorem}\label{theorem:mdp-model-checking}
-The $\dpCTLstar$ model checking problem for MDPs  is undecidable.
-\end{theorem}
-
-The proof is put in Appendix.
-We discuss some decidable special cases.
-Consider the formula $\phi:=F B$ with $B\subseteq S$ and assume that states in $B$ are absorbing (with only self-loops).
-For the case $\epsilon=0$, the condition reduces to
-$\myPr{s}{M_{\scheduler{Q}}}{\neighbor{S}}{F B} -
-\myPr{t}{M_{\scheduler{Q}}}{\neighbor{S}}{F B}\le \delta$.
-If $\delta=0$ it is the classical language equivalence problem for probabilistic automata~\cite{Rabin63}, which can be solved in polynomial time.
-However, if $\delta>0$, the problem
-becomes an approximate version of the language equivalence problem.
-To the best of our knowledge, the decision problem is still open, and has only be investigated for the special case when all states are connected~\cite{Tzeng92}.
-
-\hide{
-Since the scheduler attaining the maximally probable
-behavior from a state may be different from the scheduler attaining
-the maximally probable behavior from its neighbors, two
-neighbors may still be distinguished by a scheduler. The
-weaker definition does not preserve differential privacy. We hence
-prefer the original definition.
-}

### 2018-06-05 preliminaries.tex
reorganize sections

@@
 identical except at most one row. A \emph{query} $f$ is a function
-from $\mathcal{X}^n$ to its range $R$. The \emph{sensitivity} of the
+from $\mathcal{X}^n$ to its range $R \subseteq \bbfZ$. The \emph{sensitivity} of the
 query $f$ (written $\Delta (f)$) is $\max_{d(x, x') \leq 1} | f (x) -
@@
 output distributions when a row is replaced by another in a given
-database. Consider, for instance, a database contains the row
-corresponding to an individual. A differentially private mechanism
-will have a similar output distribution when the individual is not in
-the database. Such mechanisms behave similarly regardless of the
-presence of any individual. Individual privacy is thus
-preserved by differentially private mechanisms.
+database. 
+%Consider, for instance, a database contains the row
+%corresponding to an individual. A differentially private mechanism
+%will have a similar output distribution when the individual is not in
+%the database. Such mechanisms behave similarly regardless of the
+%presence of any individual. 
+Individual privacy is thus preserved by differentially private mechanisms.
 

### 2018-06-05 semantics-mdp.tex
reorganize sections

The logic $\dpCTLstar$ can be interpreted over MDPs. Semantics over Markov chains generalizes to MDPs as well. %Note that two states with different enabled actions
%are trivially distinguishable; no privacy can be preserved.
Let $M = (S, \Act, \wp, L)$ be an MDP. Define the satisfaction
relation $M, \neighbor{S}, s \models \Phi$ for $\PJ{J}$ and $\dpriv{\epsilon}{\delta} \phi$ (others are standard) as follows.
\begin{eqnarray*}
%  M, \neighbor{S}, s \models \top\\
%  M, \neighbor{S}, s \models p  & \textmd{ if } &  p \in L(s)\\
%  M, \neighbor{S}, s \models \neg \Phi  & \textmd{ if } &  M, \neighbor{S}, s \not\models \Phi\\
%  M, \neighbor{S}, s \models \Phi_0 \wedge \Phi_1  & \textmd{ if } &  M, \neighbor{S}, s \models \Phi_0 \textmd{ and }
%  M, \neighbor{S}, s \models \Phi_1\\
  M, \neighbor{S}, s \models \PJ{J} \phi
  & \textmd{ if } &
  \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \in J
  \textmd{ for every scheduler } \scheduler{S}\\
  M, \neighbor{S}, s \models \dpriv{\epsilon}{\delta} \phi
  & \textmd{ if } &
  \textmd{for all } t \textmd{ with } s \neighbor{S} t \textmd{ and
   query scheduler } \scheduler{Q},
   \myPr{s}{M_{\scheduler{Q}}}{\neighbor{S}}{\phi} \leq
  e^{\epsilon} \cdot
   \\
  & &
   \myPr{t}{M_{\scheduler{Q}}}{\neighbor{S}}{\phi} + \delta
   \textmd{ and }
   \myPr{t}{M_{\scheduler{Q}}}{\neighbor{S}}{\phi} \leq e^{\epsilon} \cdot
   \myPr{s}{M_{\scheduler{Q}}}{\neighbor{S}}{\phi} + \delta
\end{eqnarray*}

Recall that $M_{\scheduler{S}}$ is but a Markov chain. The semantics
of $M_{\scheduler{S}}, \neighbor{S}, \pi \models \phi$ and hence the
probability $\myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}$ are
defined as in Markov chains.
The semantics of $\dpCTLstar$ on MDPs
is again standard except the differentially private operator
$\dpriv{\epsilon}{\delta}$. For any path formula $\phi$,
$\dpriv{\epsilon}{\delta} \phi$ specifies states whose probability
of having paths satisfying $\phi$ are $(\epsilon, \delta)$-close to
those of all its neighbors for query schedulers. That is, no
query scheduler can force any of neighbors to have
probabilistically distinguishable behaviors.

\noindent
\paragraph{Justification of query schedulers.}
We use query schedulers in the semantics for the differentially
private operator. A definition with history-dependent schedulers
might be
\begin{eqnarray*}
  M, \neighbor{S}, s \models \dpriv{\epsilon}{\delta}^{\mathit{bad}} \phi
  & \textmd{ if } &
  \textmd{for all } t \textmd{ with } s \neighbor{S} t \textmd{ and
  scheduler } \scheduler{S},
  \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq
  e^{\epsilon} \cdot\\
  && \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} + \delta \textmd{ and }
  \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} \leq
  e^{\epsilon} \cdot \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}
  + \delta
\end{eqnarray*}
A state satisfies $\dpriv{\epsilon}{\delta}^{\mathit{bad}} \phi$ if
no history-dependent scheduler could differentiate the probabilities
of haveing paths satisfying
$\phi$ from neighbors. Recall that a history-dependent scheduler takes
actions by previous states. Such a definition would allow
schedulers to take different actions from different states. 
Two neighbors could be differentiated by different action sequences
erroneously.
A query scheduler $\scheduler{Q} : \bbfZ^{\geq 0} \rightarrow \Act$,
on the other hand, corresponds to a query sequence. A state satisfies
$\dpriv{\epsilon}{\delta} \phi$ if no query sequence can differentiate
the probabilities of having paths satisfying $\phi$ from neighbors.
Our semantics agrees with the informal interpretation of differential
privacy for such systems. We hence
prefer the original definition.

\subsection{Model Checking}
For an MDP $M =
(S, \Act, \wp, L)$, we want to compute $\{ s : M, \neighbor{S}, s \models
\dpriv{\epsilon}{\delta} \phi \}$ . Recall the semantics of $\dpriv{\epsilon}{\delta}
\phi$. Given $s, t$ with $s \neighbor{S} t$ and a path formula
$\phi$, we need to decide whether
$\myPr{s}{M_{\scheduler{Q}}}{\neighbor{S}}{\phi} \leq
e^{\epsilon} \myPr{t}{M_{\scheduler{Q}}}{\neighbor{S}}{\phi} + \delta$
for every query scheduler $\scheduler{Q}$.
If the path formula only contains next operators, it is easy to deal with. When $\phi$ is $\gX B$ with $B\subseteq S$, only one step behavior needs to be considered. Thus only the first action in the query sequence is needed. It can also be easily generalized to nested next operators:  one needs only to enumerate
all actions query sequences of length obtained by the number of nested $\X$ operators.
The problem however is undecidable in general.

\begin{theorem}\label{theorem:mdp-model-checking}
The $\dpCTLstar$ model checking problem for MDPs  is undecidable.
\end{theorem}

The proof is in Appendix. We discuss some decidable special cases. Consider the formula $\phi:=F B$ with $B\subseteq S$ and assume that states in $B$ are absorbing (with only self-loops). For the case $\epsilon=0$, the condition reduces to
$\myPr{s}{M_{\scheduler{Q}}}{\neighbor{S}}{F B} -
\myPr{t}{M_{\scheduler{Q}}}{\neighbor{S}}{F B}\le \delta$. If $\delta=0$ it is the classical language equivalence problem for probabilistic automata~\cite{Rabin63}, which can be solved in polynomial time. However, if $\delta>0$, the problem
becomes an approximate version of the language equivalence problem. To the best of our knowledge, the decision problem is still open, and has only be investigated for the special case when all states are connected~\cite{Tzeng92}.

\hide{
Since the scheduler attaining the maximally probable
behavior from a state may be different from the scheduler attaining
the maximally probable behavior from its neighbors, two
neighbors may still be distinguished by a scheduler. The
weaker definition does not preserve differential privacy. We hence
prefer the original definition.
}

Despite of the negative result in Theorem~\ref{theorem:mdp-model-checking},
a sufficient condition for $M, \neighbor{S}, s \models
\dpriv{\epsilon}{\delta} \phi$ is available.
To see this, observe that for any state $s \in S$ and
query scheduler $\scheduler{Q}$, we have
\[
  \min_{\scheduler{S}} \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} 
  \leq
  \myPr{s}{M_{\scheduler{Q}}}{\neighbor{S}}{\phi} 
  \leq
  \max_{\scheduler{S}} \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi} 
\]
where the minimum and maximum are taken over all schedulers
$\scheduler{S}$. Hence, 
\begin{eqnarray*}
  \myPr{s}{M_{\scheduler{Q}}}{\neighbor{S}}{\phi} -
      e^\epsilon \myPr{t}{M_{\scheduler{Q}}}{\neighbor{S}}{\phi}
  \leq
    \max_{\scheduler{S}} \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}
    - e^\epsilon \cdot
    \min_{\scheduler{S}} \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}
\end{eqnarray*}
for any $s, t \in S$ and query scheduler $\scheduler{Q}$. We have the
following proposition:
\begin{proposition}\label{proposition:sufficient-condition-mdp-model-checking}
  Let $M = (S, \Act, \wp, L)$ be an MDP, $\neighbor{S}$ a
  neighborhood relation on $S$.
  $M, \neighbor{S}, s \models \dpriv{\epsilon}{\delta} \phi$ if
  for any $s, t \in S$ with $s \neighbor{S} t$, we have
  $\max\limits_{\scheduler{S}} \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}
    - e^\epsilon \cdot
    \min\limits_{\scheduler{S}} \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}
    \leq \delta$, and
  $\max\limits_{\scheduler{S}} \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}
    - e^\epsilon \cdot
    \min\limits_{\scheduler{S}} \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}
    \leq \delta$.
\end{proposition}

For any $s \in S$, recall that $\max\limits_{\scheduler{S}}
\myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}$ and
$\min\limits_{\scheduler{S}}
\myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}$ can be efficiently 
computed~\cite{BK:08:PMC}. 
Using
Proposition~\ref{proposition:sufficient-condition-mdp-model-checking},
$M, \neighbor{S}, s \models \dpriv{\epsilon}{\delta} \phi$ can be
checked soundly and efficiently.

\todo{explain experiments}

### 2018-06-05 specifications.tex
shrink figure
reorganize sections

@@
 \paragraph{Compositionality.}
-Compositionality is one of the building block for differential privacy.
-According to the compositional theorem \cite[Theorem 3.16]{DR:14:AFDP}, for $(\epsilon_1,\delta_1)$-differentially private mechanism $M_1$ and
-$(\epsilon_2,\delta_2)$-differentially private mechanism $M_2$, their combination $M$ defined by $M(x)=(M_1(x), M_2(x))$ is $(\epsilon_1 + \epsilon_2,\delta_1+\delta_2)$-differentially private.
+Compositionality is one of the building block for differential privacy. For any $(\epsilon_1,\delta_1)$-differentially private mechanism $M_1$ and
+$(\epsilon_2,\delta_2)$-differentially private mechanism $M_2$, their combination $M(x)=(M_1(x), M_2(x))$ is $(\epsilon_1 + \epsilon_2,\delta_1+\delta_2)$-differentially private by the compositional theorem \cite[Theorem 3.16]{DR:14:AFDP}.
 The degradation is rooted in the repeated releases of information.
-To illustrate this property, we consider the extended survey mechanism which allows two consequent queries.
+To illustrate this property, we consider the extended survey mechanism which allows two consecutive queries.
 In this mechanism, an input is either $+$ or $-$; but 
@@
 
+\hide{
 \paragraph{Reachability Privacy.}
@@
 practical privacy guarantees automatically. 
+}
 
@@
 
+\hide{
 \paragraph{Infinite Differential Privacy.}
@@
 Section~\ref{section:applications}.
+}

### 2018-06-05 subsampling.tex
reorganize sections

@@
 is the maximal difference of query results on any two neighbors. There are
-two problems for such mechanisms depending on query sensitivity. First, 
+two problems for mechanisms depending on query sensitivity. First, 
 sensitivity of queries can be hard to compute or estimate. Second,
@@
 
-Subsampling is a technique to design such mechanisms. Concretely, let
+Subsampling is a technique to design such mechanisms~\cite{DR:14:AFDP}. Concretely, let
 us consider $\mathcal{X} = \{ R, B \}$ (for red and blue teams) and a
@@
 function is $f(x) = \frac{1}{2p}e^{-|x|/p}$.
-If so,
-it reports the majority of the $m$
+If the perturbed difference is sufficiently large,
+the mechanism reports the majority of the $m$
 sub-results~(line~\ref{algorithm:subsampling-majority-output}). Otherwise,
@@
 
-      \node[node] (rrr) at (-5,  2.25) { $rrr$ };
+      \node[node] (rrr) at (-5,  1.5) { $rrr$ };
 %      \node at (-4.4, 2) { $\vdots$ };
-      \node[node] (rrb) at (-5,  0.75) { $rrb$ };
+      \node[node] (rrb) at (-5.5,  0.55) { $rrb$ };
 %      \node at (-4.4, .75) { $\vdots$ };
-      \node[node] (rbb) at (-5, -0.75) { $rbb$ };
+      \node[node] (rbb) at (-5.5, -0.55) { $rbb$ };
 %      \node at (-4.4, -.5) { $\vdots$ };
-      \node[node] (bbb) at (-5, -2.25) { $bbb$ };
+      \node[node] (bbb) at (-5, -1.5) { $bbb$ };
 %      \node at (-4.4, -1.75) { $\vdots$ };
 
-      \node[node] (rmb0) at (-2.5,  2.25) { $m,0$ };
+      \node[node] (rmb0) at (-2.5,  1.5) { $m,0$ };
       \node at (-2.5, 0) { $\vdots$ };
-      \node[node] (r0bm) at (-2.5, -2.25) { $0,m$ };
+      \node[node] (r0bm) at (-2.5, -1.5) { $0,m$ };
 

### 2018-06-05 syntax.tex
reorganize sections

@@
 $p$, the negation of a state formula, the conjunction of two state
-formulae, a \emph{probabilistic} operator $\PJ{J}$ with $J$
-an interval in $[0, 1]$ followed by a path formula, or a
+formulae, the \emph{probabilistic} operator $\PJ{J}$ with $J$
+an interval in $[0, 1]$ followed by a path formula, or the
 \emph{differentially private} operator $\dpriv{\epsilon}{\delta}$

### 2018-06-05 above-threshold.tex
explain above threshold mdp

@@
 \end{figure}
+
+It is not hard to model the above threshold mechanism as a Markov
+decision process (Figure~\ref{figure:mdp-above-threshold}). In the
+figure, we sketch the model where the threshold and query results are
+in $\{ 0, 1, 2 \}$. The model simulates two computation in parallel:
+one for the dataset, the other for its neighbor. The state $t_ir_j$
+represents the input threshold $i$ and the first query result
+$j$; the state $t'_ir'_j$ represents the perturbed threshold $i$ and
+the perturbed query result $j$. Other states are similar. Consider the
+state $t_0r_1$. After applying the truncated $\frac{1}{4}$-geometric
+mechanism, it goes to one of the states $t'_0r_1$, $t'_1r_1$
+accordingly. From the state $t'_1r_1$, for instance, it moves to one
+of $t'_1r'_0$, $t'_1r'_1$, $t'_1r'_2$ by applying the truncated
+$\frac{1}{2}$-geometric mechanism to the query result. If it arrives
+at $t'_1r'_1$ or $t'_1r'_2$, the perturbed query result is not less
+than the perturbed threshold. The model enters the state of self
+loop. Otherwise, it moves to one of $t'_1r_0$, $t'_1r_1$, or $t'_1r_2$
+by actions non-deterministically (double arrows). The computation of
+the neighbor is similar. We just use the underlined symbols to
+represent threshold and query results. For instance, the state
+$\underline{t}'_2\underline{r}'_1$ represents the perturbed threshold
+$2$ and the perturbed query result $1$ in the neighbor.
+
+Now, the non-deterministic choices in the two computation cannot be 
+independent. Recall that the sensitivity of each query $f_i$ is
+$1$. If the top computation moves to the state, say, $t'_1r_0$, it
+means the next query result on the dataset is $0$. Subsequently, the
+bottom computation can only move to $\underline{t}'_i\underline{r}_0$
+or $\underline{t}'_i\underline{r}_1$ depending on the perturbed
+threshold. This is where action symbols are useful. Define the action
+symbols $\{ m\underline{n} : |m - n| \leq 1 \}$. The action
+$m\underline{n}$ represents that the next query results for the dataset
+and its neighbor are $m$ and $n$ respectively. For instance, the
+non-deterministic choice from $t'_1r'_0$ to $t'_1r_0$ is associated
+with two actions $0\underline{0}$ and $0\underline{1}$ (but not
+$0\underline{2}$). Similarly, the choice from
+$\underline{t}'_2\underline{r}'_1$ to
+$\underline{t}'_2\underline{r}_0$ is associated with the actions
+$0\underline{0}$ and $1\underline{0}$ (but not $2\underline{0}$). On
+action $0\underline{0}$, both computation will move to $t'_ir_0$ and
+$\underline{t}'_j\underline{r}_0$ respectively by the perturbed
+thresholds. Action symbols make sure the two computation on neighbor
+datasets are modeled properly. It remains to define the neighborhood
+relation. Since the sensitivity is $1$, it suffices to consider 
+$\{ (t_ir_m, t_ir_m),
+    (\underline{t}_i\underline{r}_n, \underline{t}_i\underline{r}_n),
+    (t_ir_m, \underline{t}_i\underline{r}_n), 
+    (\underline{t}_i\underline{r}_n, t_ir_m) : | m - n | \leq 1 \}$.

### 2018-06-06 semantics-mdp.tex
Write something on experiments

@@
 
-\todo{explain experiments}
+We model Algorithm~\ref{algorithm:online-model} as in Section~\ref{subsec:threshold} 
+and apply implementation on model checking to check whether the
+mechanism is differentially private. An advantage of model checking is that we
+can check property of any exact value of parameters $\epsilon$ and $\delta$.
+On the one hand,
+smaller values for particular neighborhood database could be obtained.
+For instance, for the state $t_3r_5$ and its neighborhood state $\underline{t}_3\underline{r}_4$, we verify the state
+satisfies the property $\bigwedge_{k \in \bbfZ^{\geq 0}} \dpriv{0}{0.17} ((\X^k\bot) \top)$(Note that with k rising, the reachability probability decreases to approximately 0).
+On the other hand, we could verify that for all the neighborhood database, i.e, the neighborhood states, the property $\bigwedge_{k \in \bbfZ^{\geq 0}} \dpriv{1}{0.74} ((\X^k\bot) \top)$ is satisfied.
+This indicates that the mechanism in Algorithm~\ref{algorithm:online-model} is differential private with $\epsilon = 1$ and $\delta = 0.74$.
+The parameter $\delta$ here seems large, due to the fact that
+Propsition~\ref{proposition:sufficient-condition-mdp-model-checking} is a sufficient 
+condition.

### 2018-06-07 above-threshold.tex
Write something on experiments
polish and spell check

@@
 the queries whose results are above the threshold; queries below the
-threshold only disclose minimal information and hence are considered
+threshold only disclose minimal information and hence 
 irrelevant. Since query results above the threshold convey useful
@@
 %similar behaviors on similar databases. 
-In~\cite{DR:14:AFDP}, a mechanism is design for continuous
+In~\cite{DR:14:AFDP}, a mechanism is designed for continuous
 queries by applying the Laplace mechanism. We will develop a mechanism
-for bounded and discrete queries using the geometric mechanism.
+for discrete bounded queries using the truncated geometric mechanism.
 
@@
 	\caption{Input: private database $d$, 
-          queries $f_i : \{ 0, 1, \ldots, 5 \} \rightarrow \{ 0, 1,
+          queries $f_i : d \rightarrow \{ 0, 1,
           \ldots, 5 \}$ with sensitivity $1$, threshold $t \in \{ 0,
@@
 
-Consider a threshold $t \in
-\{ 0, 1 \ldots, 5\}$, and queries $\{ f_i : f_i $ with sensitivity $\Delta
+Assume that we have a threshold $t \in \{ 0, 1, \ldots, 5 \}$
+and queries $\{ f_i : f_i $ with sensitivity $\Delta
 (f_i) = 1 \}$. In order to protect privacy, our mechanism applies the
 truncated $\frac{1}{4}$-geometric mechanism to obtain a perturbed
-threshold $t'$. For each query, the truncated $\frac{1}{2}$-geometric
-mechanism is applied to its result. If the perturbed result is not less
-than the perturbed threshold, our mechanism outputs $\top$ and
-halts. Otherwise, it outputs $\bot$ and continues to the next query
+threshold $t'$. For each query $f_i$, the truncated $\frac{1}{2}$-geometric
+mechanism is applied to its result $r_i$. If the perturbed result
+$r'_i$ is not less
+than the perturbed threshold $t'$, the mechanism halts with the input
+$\top$. Otherwise, it outputs $\bot$ and continues to the next query
 (Algorithm~\ref{algorithm:online-model}). The above threshold
 mechanism outputs a sequence of the form $\bot^* \top$. On
-similar databases, we would like to show that the above
+similar datasets, we want to show that the above
 threshold mechanism outputs the same sequence with similar
@@
 state $t_0r_1$. After applying the truncated $\frac{1}{4}$-geometric
-mechanism, it goes to one of the states $t'_0r_1$, $t'_1r_1$
+mechanism, it goes to one of the states $t'_0r_1$, $t'_1r_1$, $t'_2r_1$
 accordingly. From the state $t'_1r_1$, for instance, it moves to one
@@
 at $t'_1r'_1$ or $t'_1r'_2$, the perturbed query result is not less
-than the perturbed threshold. The model enters the state of self
-loop. Otherwise, it moves to one of $t'_1r_0$, $t'_1r_1$, or $t'_1r_2$
-by actions non-deterministically (double arrows). The computation of
-the neighbor is similar. We just use the underlined symbols to
+than the perturbed threshold. The model halts with the output $\top$
+by entering the state of a self loop. 
+Otherwise, it moves to one of $t'_1r_0$, $t'_1r_1$, or $t'_1r_2$
+non-deterministically (double arrows). The computation of
+its neighbor is similar. We just use the underlined symbols to
 represent threshold and query results. For instance, the state
@@
 or $\underline{t}'_i\underline{r}_1$ depending on the perturbed
-threshold. This is where action symbols are useful. Define the action
-symbols $\{ m\underline{n} : |m - n| \leq 1 \}$. The action
-$m\underline{n}$ represents that the next query results for the dataset
+threshold. This is where actions are useful. Define the actions
+$\{ m\underline{n} : |m - n| \leq 1 \}$. The action
+$m\underline{n}$ represents that the next query result for the dataset
 and its neighbor are $m$ and $n$ respectively. For instance, the
@@
 $0\underline{0}$ and $1\underline{0}$ (but not $2\underline{0}$). On
-action $0\underline{0}$, both computation will move to $t'_ir_0$ and
-$\underline{t}'_j\underline{r}_0$ respectively by the perturbed
-thresholds. Action symbols make sure the two computation on neighbor
-datasets are modeled properly. It remains to define the neighborhood
-relation. Since the sensitivity is $1$, it suffices to consider 
+the action $0\underline{0}$, the top computation moves to $t'_ir_0$ and
+the bottom computation moves to $\underline{t}'_j\underline{r}_0$ by
+their respective perturbed thresholds.
+Actions make sure the two computation on neighbors
+is modeled properly. It remains to define the neighborhood
+relation. Recall the sensitivity is $1$. Consider the neighborhood
+relation 
 $\{ (t_ir_m, t_ir_m),
@@
     (\underline{t}_i\underline{r}_n, t_ir_m) : | m - n | \leq 1 \}$.
+That is, two states are neighbors if they represent two inputs of the
+same threshold and query results with difference at most one.

### 2018-06-07 examples.tex
reorganize sections
polish and spell check

@@
 To model differentially private mechanisms by Markov chains, we
-formalize inputs (such as databases or query results) as states. 
+formalize inputs (such as datasets or query results) as states. 
 Randomized computation is modeled by probabilistic transitions. Atomic
 propositions are used to designate intended interpretation on states
-(such as inputs or outputs). We demonstrates the ideas in examples.
+(such as inputs or outputs). We demonstrates these ideas in examples.
 
@@
 coin. If it is tail, she answers the question truthfully. Otherwise,
-she randomly answers \textit{Yes} or \textit{No}
+she randomly answers $1$ or $0$
 uniformly (Figure~\ref{figure:2-dp-table})~\cite{DR:14:AFDP}.
@@
 is $\{ +, - \}$. The mechanism $M$ is a randomized algorithm with
-inputs in $\mathcal{X}$ and outputs in $\{ Y, N \}$. For any $x \in
-\mathcal{X}$, we have $\frac{1}{4} \leq \Pr[M (x) = Y] \leq
-\frac{3}{4}$. Hence $\Pr[M (x) = Y] \leq \frac{3}{4} = 3 \cdot
-\frac{1}{4} \leq e^{\ln 3} \Pr[M (x') = Y]$ for any neighbors $x, x'
-\in \mathcal{X}$. Similarly, $\Pr[M (x) = N] \leq e^{\ln 3} \Pr[M (x')
-= N]$. The survey mechanism is $(\ln 3, 0)$-differentially private.
+inputs in $\mathcal{X}$ and outputs in $\{ 1, 0 \}$. For any $x \in
+\mathcal{X}$, we have $\frac{1}{4} \leq \Pr[M (x) = 1] \leq
+\frac{3}{4}$. Hence $\Pr[M (x) = 1] \leq \frac{3}{4} = 3 \cdot
+\frac{1}{4} \leq e^{\ln 3} \Pr[M (x') = 1]$ for any neighbors $x, x'
+\in \mathcal{X}$. Similarly, $\Pr[M (x) = 0] \leq e^{\ln 3} \Pr[M (x')
+= 0]$. The survey mechanism is $(\ln 3, 0)$-differentially private.
 The random noise boosts the probability of answering
-\textit{Yes} or \textit{No} to at least $\frac{1}{4}$ regardless of
+$1$ or $0$ to at least $\frac{1}{4}$ regardless of
 diagnoses. Inferences on individual diagnosis can be plausibly denied.
@@
       \[
-      \begin{array}{|c|c|c|}
-        \hline
+      \begin{array}{c|c|c|c|}
+        \multicolumn{2}{c}{} &
+        \multicolumn{2}{c}{output}
+        \\
+        \cline{2-4}
+        &  & 1 & 0 \\
+        \cline{2-4}
+        \multirow{2}*{\rotatebox{90}{state}} 
         &
-        \multicolumn{2}{c|}{output}
-        \\
-        \hline
-        state & Y & N \\
-        \hline
         + & \frac{3}{4} = \frac{1}{2} \cdot \frac{1}{2} + \frac{1}{2} \cdot 1
@@
         \\
-        \hline
+        \cline{2-4}
+        &
         - & \frac{1}{4} = \frac{1}{2} \cdot \frac{1}{2} + \frac{1}{2} \cdot 0
@@
         \\
-        \hline
+        \cline{2-4}
       \end{array}
@@
       \node[node] (Y) at (-1.5,  0) { $s$ };
-      \node at (-1.5, .5) { $\mathit{out}_Y$ };
+      \node at (-1.5, .5) { $\mathit{out}_1$ };
       \node[node] (N) at ( 1.5,  0) { $t$ };
-      \node at ( 1.5, .5) { $\mathit{out}_N$ };
+      \node at ( 1.5, .5) { $\mathit{out}_0$ };
 
@@
 private for any query $f$ with sensitivity $\Delta (f)$. Observe that
-the privay guarantee $(-{\Delta (f)} \ln \alpha, 0)$ depends on the
+the privacy guarantee $(-{\Delta (f)} \ln \alpha, 0)$ depends on the
 sensitivity of the query $f$. To achieve $(\epsilon, 0)$-differential
@@
     \[
-    \begin{array}{|c|c|c|c|c|c|c|}
-      \hline
-      \textmd{input/output} & 0 & 1 & 2 & 3 & 4 & 5\\
-      \hline
-      0 & {2}/{3}  & {1}/{6} & {1}/{12} & {1}/{24}  & {1}/{48} & {1}/{48}\\
-      \hline
-      1 & {1}/{3}  & {1}/{3} & {1}/{6} & {1}/{12}  & {1}/{24} & {1}/{24}\\
-      \hline
-      2 & {1}/{6}  & {1}/{6} & {1}/{3} & {1}/{6}   & {1}/{12} & {1}/{12}\\
-      \hline
-      3 & {1}/{12} & {1}/{12} & {1}/{6} & {1}/{3}   & {1}/{6} & {1}/{6}\\
-      \hline
-      4 & {1}/{24} & {1}/{24} & {1}/{12} & {1}/{6}   & {1}/{3} & {1}/{3}\\
-      \hline
-      5 & {1}/{48} & {1}/{48} & {1}/{24} & {1}/{12} & {1}/{6} & {2}/{3}\\
-      \hline
+    \begin{array}{c|c|c|c|c|c|c|c|}
+      \multicolumn{2}{c}{} &
+      \multicolumn{6}{c}{output} \\
+      \cline{2-8}
+      & & 0 & 1 & 2 & 3 & 4 & 5\\
+      \cline{2-8}
+      \multirow{6}*{\rotatebox{90}{input}}
+      & 0 & {2}/{3}  & {1}/{6} & {1}/{12} & {1}/{24}  & {1}/{48} & {1}/{48}\\
+      \cline{2-8}
+      & 1 & {1}/{3}  & {1}/{3} & {1}/{6} & {1}/{12}  & {1}/{24} & {1}/{24}\\
+      \cline{2-8}
+      & 2 & {1}/{6}  & {1}/{6} & {1}/{3} & {1}/{6}   & {1}/{12} & {1}/{12}\\
+      \cline{2-8}
+      & 3 & {1}/{12} & {1}/{12} & {1}/{6} & {1}/{3}   & {1}/{6} & {1}/{6}\\
+      \cline{2-8}
+      & 4 & {1}/{24} & {1}/{24} & {1}/{12} & {1}/{6}   & {1}/{3} & {1}/{3}\\
+      \cline{2-8}
+      & 5 & {1}/{48} & {1}/{48} & {1}/{24} & {1}/{12} & {1}/{6} & {2}/{3}\\
+      \cline{2-8}
     \end{array}
@@
 oblivious. It suffices to consider $\{ f (x) : x \in \mathcal{X}^n \}$
-instead of $\mathcal{X}^n$ as inputs. 
+as inputs. 
 Let the state $s_k$ and $t_l$
-denote the input value $k$ and output value $l$ respectively. Define
-the state set $S = \{ s_k, t_k : k \in \{ 0, 1, \ldots, m \} \}$.
+denote the input $k$ and output $l$ respectively. Define
+$S = \{ s_k, t_k : k \in \{ 0, 1, \ldots, m \} \}$.
 The probability transition $\wp (s_k, t_l)$ is the probability of the
-output value $l$ on the input value $k$ as defined in the
+output $l$ on the input $k$ as defined in the
 mechanism. Moreover, we have $\mathit{in}_k \in L (s_k)$ and
@@
 $\frac{1}{2}$-geometric mechanism over $\{ 0, 1, \ldots, 5 \}$.
-
-\subsection{Subsampling Majority}
-\label{subsection:subsampling}
-
-\input{subsampling}

### 2018-06-07 introduction.tex
reorganize sections
polish and spell check

@@
 mechanism is differentially private if it behaves similarly on similar
-input databases~\cite{DMNS:06:CNSPD,D:06:DP}. Consider, for example,
+input datasets~\cite{DMNS:06:CNSPD,D:06:DP}. Consider, for example,
 the Laplace mechanism where analysis results are perturbed by random
 noises with the Laplace distribution~\cite{DR:14:AFDP}. Random noises
-hide the differences of analysis results from similar databases.
-Clearly, more
-randomness gives more privacy but less utility in released perturbed
+hide the differences of analysis results from similar datasets.
+Clearly, more noises
+give more privacy but less utility in released perturbed
 results. Under the framework of differential privacy, data analysts
@@
 sophisticated data analyses. Privacy leak has also been observed in
-data analysis programs implementing differentialy
+data analysis programs implementing differential
 privacy~\cite{M:12:SLSBDP,TKBW:17:PLAI}. This calls for formal 
@@
 probabilistic model checking to formalize such mechanisms
-by Markov chains or Markov decision processes~\cite{Put05}. We give
-formalizations of several differentially private mechanisms and
-explain subtle privacy conditions.
+as Markov chains or Markov decision processes~\cite{Put05}. 
+%We give formalizations of several differentially private mechanisms
+%and explain subtle privacy conditions.
 When a data analysis mechanism does not
@@
 Two ingredients are introduced to specify differentially private
-behaviors. A reflexive and commutative user-defined binary relation
+behaviors. A reflexive and symmetric user-defined binary relation
 over states is required to formalize similarity. We moreover add
-the path quantifier $\dpriv{\epsilon}{\delta}$ for similar
+the path quantifier $\dpriv{\epsilon}{\delta}$ for specifying similar
 behaviors. Informally, a state satisfies $\dpriv{\epsilon}{\delta}
-\phi$ if it has probabilistically similar path behavior $\phi$ as
-its similar states. Consider, for instance, a binary data analysis
+\phi$ if its probability of having path $\phi$ is close to
+those of similar states. Consider, for instance, a binary data analysis
 mechanism computing the likelihood ($\mathit{high}$ or $\mathit{low}$)
-of an epidemic in a population. Then a state satisfies
+of an epidemic. Then a state satisfying
 $\dpriv{\epsilon}{\delta} (\F \mathit{high}) \wedge
-\dpriv{\epsilon}{\delta} (\F \mathit{low})$ if the mechanism is
-$(\epsilon, \delta)$-differentially private at the state.
+\dpriv{\epsilon}{\delta} (\F \mathit{low})$ denotes similar states
+have the same outcomes with similar probabilities.
 %\todo{not clear here}\lz{here it should be $\X$?}
 
+\hide{
 % differential privacy of infinite behaviors
@@
 also generalizes to reactive systems naturally.
-More data analysis mechanisms are admitable for sophisticated formal
+More data analysis mechanisms are admissible for sophisticated formal
 privacy analysis through our techniques.
+}
 
@@
 differential privacy. The logic is able to to express classical
-differentially private properties. Complex properties such as those
-involving infinite behaviors can also be expressed;
-\item We model several differentially private mechanisms, including stream processing, in Markov chains or Markov decision processes; and
+differentially private properties;
+%Complex properties such as those involving infinite behaviors can
+%also be expressed; 
+\item We model several differentially private mechanisms in Markov chains or Markov decision processes; and
 \item We show that the model checking

### 2018-06-07 main.tex
polish and spell check

@@
 \pagestyle{plain}
+\usepackage{multirow}
 \usepackage{algorithm,algpseudocode}
@@
 Preliminaries are given in Section~\ref{section:preliminaries}.
-In Section~\ref{section:examples} we discuss how differentially private mechanisms can be modeled as Markov chains.
+In Section~\ref{section:examples} we discuss how differentially private mechanisms are modeled as Markov chains.
 The logic $\dpCTLstar$ and its syntax are presented in Section~\ref{section:dpCTL}. The semantics over Markov chains and its model checking algorithm are given in Section~\ref{section:semantics}.
-Section~\ref{section:specifying-properties} describes differential privacy properties using $\dpCTLstar$.
+Section~\ref{section:specifying-properties} discusses differential privacy properties using $\dpCTLstar$.
 More examples of differentially private mechanisms as Markov decision processes are given in Section~\ref{section:mdp}.
-The semantics over Markov decision processes and undecidability of model checking is discussed in Section~\ref{section:semantics-mdp}.
+The semantics over Markov decision processes and undecidability of model checking is given in Section~\ref{section:semantics-mdp}.
 Finally,
@@
 \label{section:conclusions}
-We have introduced $\dpCTLstar$ to reason about properties in differential privacy, and have investigated the model checking problem.
+We have introduced $\dpCTLstar$ to reason about properties in differential privacy, and have investigated its model checking problems.
 For Markov chains, the model checking problem has the same complexity as for $\PCTLstar$.
 The general MDP model checking problem however is undecidable.
-We have discussed some decidable special cases and an efficient and sufficient condition to check differentially private subformulas.
-An interesting furture work is to identify more
+We have discussed some decidable special cases and an efficient yet sufficient condition to check differentially private subformulae.
+An interesting future work is to identify more
 decidable subclasses and sufficient conditions.
-As an example, consider the formula $\bigwedge_{k \in \bbfZ^{\geq 0}}
-\dpriv{\epsilon}{\delta} (\X^k  00)$ which was used in specifying the pan privacy.
+As an example, consider the extended $\dpCTLstar$ formula $\bigwedge_{k \in \bbfZ^{\geq 0}}
+\dpriv{\epsilon}{\delta} (\X^k  \top)$.
 For the case $\epsilon=\delta=0$, it reduces to a language equivalence problem for probabilistic automata.

### 2018-06-07 mdp.tex
polish and spell check

@@
 allows analysts to ask queries adaptively based on previous
-responses. The mechanisms we constructed previously are offline
-mechanisms. On the other hand, online mechanisms permit further queries. 
+responses. The mechanisms considered previously are offline
+mechanisms. 
+Since offline mechanisms only release one query result, they are
+relatively easy to analyze. For online mechanisms, one has to consider
+all possible adaptive queries. We therefore use MDPs to model these
+non-deterministic behaviors.
+Specifically, adaptive queries are modeled by
+actions. Randomized computation associated with different queries is
+modeled by distributions associated with actions.
+
+\hide{
+On the other hand, online mechanisms permit further queries. 
 To maintain
@@
 queries after a certain output is disclosed.
+}
+
+Consider again the survey mechanism. Suppose we would like to design
+an interactive mechanism which adjusts random noises on surveyors'
+requests. When the surveyor requests low-accuracy answers, the
+surveyee uses the survey mechanism as before. When high-accuracy
+answers are requested, the surveyee answers $1$ with
+probability $\frac{4}{5}$ and $0$ with probability $\frac{1}{5}$ when she has
+positive diagnosis. She answers $1$ with probability $\frac{1}{5}$
+and $0$ with probability $\frac{4}{5}$ when she is not
+diagnosed with the disease X. This gives an interactive mechanism
+corresponding to the MDP shown in
+Figure~\ref{figure:simple-mdp}.
 
@@
 
-We use MDPs to model 
-interactive mechanisms. Specifically, external inputs are modeled by
-actions. Behaviors associated with different inputs are modeled by
-distributions associated with actions.
-
-Consider again the survey mechanism. Suppose we would like to design
-an interactive mechanism which adjusts random noises on surveyors'
-requests. When the surveyor requests low-accuracy answers, the
-surveyee uses the survey mechanism as before. When high-accuracy
-answers are requested, the surveyee answers \textit{Yes} with
-probability $\frac{4}{5}$ and \textit{No} with probability $\frac{1}{5}$ when she has
-positive diagnosis. She answers \textit{Yes} with probability $\frac{1}{5}$
-and \textit{No} with probability $\frac{4}{5}$ when she is not
-diagnosed with the disease X. This gives an interactive mechanism
-corresponding to the MDP shown in
-Figure~\ref{figure:simple-mdp}.
-
 In the figure, the states $+$,
 $-$, $s$, and $t$ are interpreted as before. The actions $L$ and $H$
-denote low- and high-accuracy requests respectively. Let $M_H$ denote
-the Markov chain derived by high-accuracy requests.
-Note that the high-accuracy mechanism is $(\ln 4,0)$-differentially private.
-Unlike non-interactive mechanisms, the privacy guarantees vary from accuracy requests.
+denote low- and high-accuracy queries respectively. 
+Note that the high-accuracy survey mechanism is $(\ln
+4,0)$-differentially private. Unlike non-interactive mechanisms, the
+privacy guarantees vary from queries with different accuracies.

### 2018-06-07 preliminaries.tex
polish and spell check

@@
-Let $\bbfZ$ and $\bbfZ^{\geq 0}$ be the sets of integers and
-non-negative integers respectively.
+Let $\bbfZ$ and $\bbfZ^{\geq 0}$ be the sets of integers and non-negative integers respectively.
 % For $n \geq 0$, $\underline{n} = \{ 0, 1, \ldots, n \} \subseteq
@@
 We denote the data universe by $\mathcal{X}$; $x \in \mathcal{X}^n$ is
-a database with $n$ rows from the data universe. Two databases $x$ and
+a dataset with $n$ rows from the data universe. Two datasets $x$ and
 $x'$ are \emph{neighbors} (denoted by $d(x, x') \leq 1$) if they are
@@
 $\tilde{r} \in \tilde{R}$ when $f (x) = f (x')$. In words, outputs of an
-oblivious mechanism depend on the query result $f (x)$ but not on the
-input database $x$. The order of rows in a database, for instance, is
-irrelevant to oblivious mechanisms for counting queries. Let $x, x'$
-be databses and $\tilde{r} \in \tilde{R}$. We say the probability of
-the mechanism $M_f$ outputing $\tilde{r}$ on $x$ is
+oblivious mechanism depend on the query result $f (x)$.
+The order of rows in a dataset, for instance, is
+irrelevant to oblivious mechanisms. Let $x, x'$
+be datasets and $\tilde{r} \in \tilde{R}$. We say the probability of
+the mechanism $M_f$ outputting $\tilde{r}$ on $x$ is
 \emph{$(\epsilon, \delta)$-close} to those on $x'$ if
@@
 and $\tilde{r} \in \tilde{R}$, the probability of
-$M_f$ outputing $\tilde{r}$ on $x$ is $(\epsilon, \delta)$-close to
+$M_f$ outputting $\tilde{r}$ on $x$ is $(\epsilon, \delta)$-close to
 those on $x'$.
@@
 The non-negative parameters $\epsilon$ and $\delta$ quantify
-probabilistically similar behaviors;
+mechanism behaviors probabilistically;
 the smaller they are, the behaviors are more similar.
 Informally, a differentially private mechanism has probabilistically
-similar behaviors on neighboring databases. It will have similar
-output distributions when a row is replaced by another in a given
-database. 
+similar behaviors on neighboring datasets. It will have similar
+output distributions when any row is replaced by another in a given
+dataset. 
 %Consider, for instance, a database contains the row
@@
 (MDP)~\footnote{The MDP we consider is \emph{reactive} in the sense that
-all actions are enabled in every state. If equipped with a set of final states $F\subseteq S$, such model is
-referred to as \emph{probabilistic automaton} in the literature~\cite{Rabin63}, and language equivalence/inclusion
-problems have been investigated. In the model checking context, the notion MDPs is mostly used.}
+all actions are enabled in every state.
+% If equipped with a set of final states $F\subseteq S$, such model is
+% referred to as \emph{probabilistic automaton} in the
+% literature~\cite{Rabin63}, and language equivalence/inclusion
+% problems have been investigated. In the model checking context, the
+% notion MDPs is mostly used.
+}
 $M = (S, \Act, \wp, L)$ consists
@@
 \rightarrow \Act$. A \emph{query scheduler} for $M$ is a function
-$\scheduler{Q} : \bbfZ^{\geq 0} \rightarrow \Act$. Intuitively, its decision depends only on the length of the history.
+$\scheduler{Q} : S^+ \rightarrow \Act$ such that
+$\scheduler{Q}(\sigma) = \scheduler{Q}(\sigma')$ for any $\sigma,
+\sigma' \in S^+$ of the same length.
+Intuitively, its decision depends only on the length of the history.
 A path $\pi =

### 2018-06-07 semantics-mdp.tex
polish and spell check

--- 2018-06-05 semantics-mdp.tex
@@
-The logic $\dpCTLstar$ can be interpreted over MDPs. Semantics over Markov chains generalizes to MDPs as well. %Note that two states with different enabled actions
+The logic $\dpCTLstar$ can be interpreted over MDPs. 
+%Semantics over Markov chains generalizes to MDPs as well. Note that
+%two states with different enabled actions 
 %are trivially distinguishable; no privacy can be preserved.
-Let $M = (S, \Act, \wp, L)$ be an MDP. Define the satisfaction
-relation $M, \neighbor{S}, s \models \Phi$ for $\PJ{J}$ and $\dpriv{\epsilon}{\delta} \phi$ (others are standard) as follows.
+Let $M = (S, \Act, \wp, L)$ be an MDP and $\neighbor{S}$ a
+neighborhood relation on $S$. Define the satisfaction
+relation $M, \neighbor{S}, s \models \Phi$ for $\PJ{J} \phi$ and
+$\dpriv{\epsilon}{\delta} \phi$ as follows (others are straightforward).
 \begin{eqnarray*}
@@
 those of all its neighbors for query schedulers. That is, no
-query scheduler can force any of neighbors to have
-probabilistically distinguishable behaviors.
+query scheduler can force any of neighbors to distinguish the
+specified path behavior probabilistically.
 
@@
 A state satisfies $\dpriv{\epsilon}{\delta}^{\mathit{bad}} \phi$ if
-no history-dependent scheduler could differentiate the probabilities
-of haveing paths satisfying
+no history-dependent scheduler can differentiate the probabilities
+of having paths satisfying
 $\phi$ from neighbors. Recall that a history-dependent scheduler takes
@@
 schedulers to take different actions from different states. 
-Two neighbors could be differentiated by different action sequences
-erroneously.
-A query scheduler $\scheduler{Q} : \bbfZ^{\geq 0} \rightarrow \Act$,
+Two neighbors could hence be differentiated by different action sequences.
+The specification might be too strong for our purposes.
+A query scheduler $\scheduler{Q} : S^+ \rightarrow \Act$,
 on the other hand, corresponds to a query sequence. A state satisfies
@@
 the probabilities of having paths satisfying $\phi$ from neighbors.
+Recall query schedulers only depend on lengths of histories. Two
+neighbors cannot be distinguished by the same action sequence of any
+length if they satisfy a differentially private subformula.
 Our semantics agrees with the informal interpretation of differential
-privacy for such systems. We hence
-prefer the original definition.
+privacy for such systems. We therefore consider only query schedulers
+in our definition.
 
 \subsection{Model Checking}
-For an MDP $M =
-(S, \Act, \wp, L)$, we want to compute $\{ s : M, \neighbor{S}, s \models
-\dpriv{\epsilon}{\delta} \phi \}$ . Recall the semantics of $\dpriv{\epsilon}{\delta}
+Given an MDP $M =
+(S, \Act, \wp, L)$, a neighborhood relation $\neighbor{S}$, $s \in
+S$, and a path formula $\phi$, consider the problem of deciding $M,
+\neighbor{S}, s \models 
+\dpriv{\epsilon}{\delta} \phi$ . Recall the semantics of $\dpriv{\epsilon}{\delta}
 \phi$. Given $s, t$ with $s \neighbor{S} t$ and a path formula
@@
 for every query scheduler $\scheduler{Q}$.
-If the path formula only contains next operators, it is easy to deal with.
-When $\phi$ is $\gX B$ with $B\subseteq S$, only one step behavior needs to be considered.
-Thus only the first action in the query sequence is needed.
-It can also be easily generalized to nested next operators:  one needs only to enumerate
-all actions query sequences of length obtained by the number of nested $\X$ operators.
+When $\phi$ is $\gX B$ with $B\subseteq S$, only the first action in
+the query sequence needs to be considered. This can also be easily generalized to nested next operators:  one needs only to enumerate
+all actions query sequences of a fixed length.
+% obtained by the number of nested $\X$ operators.
 The problem however is undecidable in general.
@@
 We discuss some decidable special cases.
-Consider the formula $\phi:=F B$ with $B\subseteq S$ and assume that states in $B$ are absorbing (with only self-loops).
+Consider the formula $\phi:=F B$ with $B\subseteq S$ and assume that states in $B$ with only self-loops.
 For the case $\epsilon=0$, the condition reduces to
@@
 However, if $\delta>0$, the problem
-becomes an approximate version of the language equivalence problem.
-To the best of our knowledge, the decision problem is still open, and has only be investigated for the special case when all states are connected~\cite{Tzeng92}.
+becomes an approximate version of the language equivalence problem. To
+the best of our knowledge, its decidability is still open except for the special case where all states are connected~\cite{Tzeng92}.
 
@@
 \dpriv{\epsilon}{\delta} \phi$ is available.
-To see this, observe that for any state $s \in S$ and
+To see this, observe that for $s \in S$ and
 query scheduler $\scheduler{Q}$, we have
@@
   \myPr{s}{M_{\scheduler{Q}}}{\neighbor{S}}{\phi} -
-      e^\epsilon \myPr{t}{M_{\scheduler{Q}}}{\neighbor{S}}{\phi}
+      e^\epsilon \cdot \myPr{t}{M_{\scheduler{Q}}}{\neighbor{S}}{\phi}
   \leq
@@
   $M, \neighbor{S}, s \models \dpriv{\epsilon}{\delta} \phi$ if
-  for any $s, t \in S$ with $s \neighbor{S} t$, we have
   $\max\limits_{\scheduler{S}} \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}
@@
     \min\limits_{\scheduler{S}} \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}
-    \leq \delta$, and
+    \leq \delta$ and
   $\max\limits_{\scheduler{S}} \myPr{t}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}
@@
     \min\limits_{\scheduler{S}} \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}
-    \leq \delta$.
+    \leq \delta$ for any $s, t \in S$ with $s \neighbor{S} t$.
 \end{proposition}
 
-For any $s \in S$, recall that $\max\limits_{\scheduler{S}}
+For $s \in S$, recall that $\max\limits_{\scheduler{S}}
 \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}$ and
@@
 computed~\cite{BK:08:PMC}. 
-Using
+By
 Proposition~\ref{proposition:sufficient-condition-mdp-model-checking},
@@
 
-\todo{explain experiments}
+We model the above threshold algorithm (Algorithm~\ref{algorithm:online-model})
+and apply
+Proposition~\ref{proposition:sufficient-condition-mdp-model-checking} 
+check whether the mechanism is differentially private using the
+classical $\PCTL$ model checking algorithm for MDPs. 
+Since concrete values of the parameters $\epsilon$ and $\delta$ are
+computed, tighter bounds for specific neighbors can be obtained.
+For instance, for the state $t_3r_5$ and its neighbor
+$\underline{t}_3\underline{r}_4$, we verify the property $\bigwedge_{k
+  \in \bbfZ^{\geq 0}} \dpriv{0}{0.17} ((\X^k\bot) \top)$ is
+satisfied. Note the reachability probability goes to $0$ as $k$ goes
+to infinity. By repeating the computation, we verify that 
+the property $\bigwedge_{k \in \bbfZ^{\geq 0}} \dpriv{1}{0.74}
+((\X^k\bot) \top)$ is satisfied for all neighbors. 
+Subsequently, the above threshold mechanism in
+Algorithm~\ref{algorithm:online-model} is $(1, 0.74)$-differentially
+private. Compared to the parameters for the neighbors $t_3r_5$ and
+$\underline{t}_3\underline{r}_4$, the parameter $\delta$ appears to be
+significantly large. It means that there are two neighbors with
+drastically different output distributions from our mechanism.
+Moreover, recall that
+Proposition~\ref{proposition:sufficient-condition-mdp-model-checking}
+is a sufficient condition. It only gives an upper bound of privacy
+parameters. Automatically computing tighter bounds may be possible.

### 2018-06-07 semantics.tex
reorganize sections
polish and spell check

@@
 Moreover, the relation $K, \neighbor{S}, \pi \models \phi$ is defined as
-for the standard linear temporal logic formulas~\cite{ManneP92}. We only recall the semantics for $\X$ and $\until$ as follows:
+for the standard linear temporal logic formulae~\cite{ManneP92}. We only recall the semantics for $\X$ and $\until$:
 \begin{eqnarray*}
@@
 
-To simplify our notation, for any Markov chain $K$, neighborhood
-relation $N$ on $S$, a state $s \in S$, and a path formula $\phi$,
+For any Markov chain $K$, neighborhood
+relation $N$ on $S$, $s \in S$, and a path formula $\phi$,
 define
@@
 
-\subsection{Model Checking Algorithms}
+\subsection{Model Checking}
 We describe the model checking algorithm for $\dpCTL$. The algorithm follows the classical
@@
 states satisfying the subformula $\PJ{J} \phi$ are obtained by
-computing $\myPr{s}{K}{\neighbor{S}}{\phi}$ for state $s \in S$.
+computing $\myPr{s}{K}{\neighbor{S}}{\phi}$ for $s \in S$.
 These probabilities can be obtained by solving linear equations or
@@
 In order to determine the set $\{ s : K, \neighbor{S}, s \models
-\dpriv{\epsilon}{\delta} \varphi \}$, our algorithm computes
-the probabilities $p (s) = \myPr{s}{K}{\neighbor{S}}{\varphi}$ for every
+\dpriv{\epsilon}{\delta} \phi \}$, our algorithm computes
+the probabilities $p (s) = \myPr{s}{K}{\neighbor{S}}{\phi}$ for every
 $s \in S$ (Algorithm~\ref{algorithm:sat-dpriv-dtmc}). For each $s \in S$,
@@
 The model checking algorithm for $\dpCTLstar$ can be treated as in the classical setting~\cite{BK:08:PMC}:  all we need is to  compute the  probability
-  $\myPr{s}{K}{\neighbor{S}}{\phi}$ with general path formula $\phi$.
-For this purpose one first constructs a deterministic $\omega$-automaton $R$ corresponding to $\phi$, such as deterministic Rabin automaton.
-Then, the probability reduces to a
+  $\myPr{s}{K}{\neighbor{S}}{\phi}$ with general path formula $\phi$. For this purpose one first constructs a deterministic $\omega$-automaton $R$ for $\phi$. Then, the probability reduces to a
   reachability probability in the product Markov chain obtained from $K$ and $R$. There are more efficient algorithms without the product construction, see \cite{CY95,CSS03,BaierKKKMW16} for details.

### 2018-06-07 specifications.tex
polish and spell check

@@
 
-\paragraph{Classical Differential Privacy.}
+\paragraph{Differential Privacy.}
 Consider the survey mechanism (Section~\ref{subsec:survey}). For  $v$
-with $u N v$, we have $\myPr{u}{K}{N}{\X \mathit{out}_Y}\le
-3\myPr{v}{K}{N}{\X \mathit{out}_Y}$ for the probability of satisfying
-$\X \mathit{out}_Y$ from $u$ and $v$.
-The formula $\dpriv{\ln 3}{0} (\X \mathit{out}_Y)$ holds in state $u$
-and similarly for $\dpriv{\ln 3}{0} (\X \mathit{out}_N)$. 
+with $u N v$, we have $\myPr{u}{K}{N}{\X \mathit{out}_1}\le
+3\myPr{v}{K}{N}{\X \mathit{out}_1}$ for the probability of satisfying
+$\X \mathit{out}_1$ from $u$ and $v$.
+The formula $\dpriv{\ln 3}{0} (\X \mathit{out}_1)$ holds in state $u$
+and similarly for $\dpriv{\ln 3}{0} (\X \mathit{out}_0)$. 
 Recall that differential privacy requires similar output distributions
 on neighbors. The formula $\dpriv{\ln 3}{0} (\X
-\mathit{out}_Y)\wedge \dpriv{\ln 3}{0} (\X \mathit{out}_N)$ thus
+\mathit{out}_1)\wedge \dpriv{\ln 3}{0} (\X \mathit{out}_0)$ thus
 specifies differential privacy for states $+$ and $-$. 
@@
 \paragraph{Compositionality.}
-Compositionality is one of the building block for differential privacy. For any $(\epsilon_1,\delta_1)$-differentially private mechanism $M_1$ and
-$(\epsilon_2,\delta_2)$-differentially private mechanism $M_2$, their combination $M(x)=(M_1(x), M_2(x))$ is $(\epsilon_1 + \epsilon_2,\delta_1+\delta_2)$-differentially private by the compositional theorem \cite[Theorem 3.16]{DR:14:AFDP}.
+Compositionality is one of the building blocks for differential privacy. For any $(\epsilon_1,\delta_1)$-differentially private mechanism $M_1$ and
+$(\epsilon_2,\delta_2)$-differentially private mechanism $M_2$, their combination $(M_1(x), M_2(x))$ is $(\epsilon_1 + \epsilon_2,\delta_1+\delta_2)$-differentially private by the compositional theorem \cite[Theorem 3.16]{DR:14:AFDP}.
 The degradation is rooted in the repeated releases of information.
@@
 In this mechanism, an input is either $+$ or $-$; but 
-outputs are $\mathit{out}_Y\mathit{out}_Y$,
-$\mathit{out}_Y\mathit{out}_N$, $\mathit{out}_N\mathit{out}_Y$, or
-$\mathit{out}_N\mathit{out}_N$. 
+outputs are $\mathit{out}_1\mathit{out}_1$,
+$\mathit{out}_1\mathit{out}_0$, $\mathit{out}_0\mathit{out}_1$, or
+$\mathit{out}_0\mathit{out}_0$. 
 The model is depicted in Figure~\ref{figure:2-dp-mc}.
@@
     \node[node] (pY1) at (-1.5, .8) { $s_1$ };
-    \node at (-1.5, 1.3) { $\mathit{out}_Y$ };
+    \node at (-1.5, 1.3) { $\mathit{out}_1$ };
     \node[node] (pN1) at (-1.5, -.8) { $t_1$ };
-    \node at (-1.5, -.3) { $\mathit{out}_N$ };
+    \node at (-1.5, -.3) { $\mathit{out}_0$ };
 
     \node[node] (Y2) at ( 0,  .8) { $s_2$ };
-    \node at (0, 1.3) { $\mathit{out}_Y$ };
+    \node at (0, 1.3) { $\mathit{out}_1$ };
     \node[node] (N2) at ( 0, -.8) { $t_2$ };
-    \node at (0, -.3) { $\mathit{out}_N$ };
+    \node at (0, -.3) { $\mathit{out}_0$ };
 
@@
     \node[node] (nY1) at ( 1.5, .8) { $\bar{s}_1$ };
-    \node at (1.5, 1.3) { $\mathit{out}_Y$ };
+    \node at (1.5, 1.3) { $\mathit{out}_1$ };
     \node[node] (nN1) at ( 1.5, -.8) { $\bar{t}_1$ };
-    \node at (1.5, -.3) { $\mathit{out}_N$ };
+    \node at (1.5, -.3) { $\mathit{out}_0$ };
 
@@
 
-Consider the formula $\dpriv{\ln 9}{0} (\X (\mathit{out}_Y\wedge \X \mathit{out}_Y))$.
-A path satisfies $\X (\mathit{out}_Y\wedge \X \mathit{out}_Y)$ if the second state satisfies $\mathit{out}_Y$ and the third state satisfies $\mathit{out}_Y$ as well.
+Consider the formula $\dpriv{\ln 9}{0} (\X (\mathit{out}_1\wedge \X \mathit{out}_1))$.
+A path satisfies $\X (\mathit{out}_1\wedge \X \mathit{out}_1)$ if the second state satisfies $\mathit{out}_1$ and the third state satisfies $\mathit{out}_1$ as well.
 We verify that this formula is satisfied for states $+$ and $-$.
-Moreover, the bound $\epsilon=\ln 9$ is tight, since the probability of satisfying $\X (\mathit{out}_Y\wedge \X \mathit{out}_Y)$ from states $+$ and $-$ are $\frac{9}{16}$ and $\frac{1}{16}$, respectively.
-Finally, the formula $\wedge_{a_1,a_2}\dpriv{\ln 9}{0} (\X (a_1\wedge \X a_2))$ specifies differential privacy for the model, where $a_1,a_2$ range over all atomic propositions $\{\mathit{out}_Y, \mathit{out}_N\}$.
+Moreover, the bound $\epsilon=\ln 9$ is tight, since the probability of satisfying $\X (\mathit{out}_1\wedge \X \mathit{out}_1)$ from states $+$ and $-$ are $\frac{9}{16}$ and $\frac{1}{16}$, respectively.
+Finally, the formula $\wedge_{a_1,a_2}\dpriv{\ln 9}{0} (\X (a_1\wedge \X a_2))$ specifies differential privacy for the model, where $a_1,a_2$ range over all atomic propositions $\{\mathit{out}_1, \mathit{out}_0\}$.
 
@@
 \begin{itemize}
-\item $\dpriv{\ln 3}{0} (\X \X \mathit{out}_Y)$. In this case we claim
+\item $\dpriv{\ln 3}{0} (\X \X \mathit{out}_1)$. In this case we claim
   there is no privacy loss, even though there are two queries. The reason is that the output of the first query is not observed at all. It is easy to verify that it is indeed satisfied by $+$ and $-$.
-\item $\dpriv{\ln 3}{0} (\X (\mathit{out}_Y\wedge \dpriv{\ln 3}{0}(\X \mathit{out}_Y)))$.
-This is a nested $\dpCTL$ formula, where the inner state formula $\dpriv{\ln 3}{0}(\X \mathit{out}_Y)$ specifies the one step differential privacy.
+\item $\dpriv{\ln 3}{0} (\X (\mathit{out}_1\wedge \dpriv{\ln 3}{0}(\X \mathit{out}_1)))$.
+This is a nested $\dpCTL$ formula, where the inner state formula $\dpriv{\ln 3}{0}(\X \mathit{out}_1)$ specifies the one-step differential privacy.
 Observe the inner formula is satisfied by all neighbors.
@@
 bounds for composition. Consider the survey mechanism, and the
-property $\dpriv{0}{.5} (\X \mathit{out}_Y)$. Obviously, it holds in
-states $+$ and $-$ since $\myPr{u}{K}{N}{\mathit{out}_Y} =
+property $\dpriv{0}{.5} (\X \mathit{out}_1)$. Obviously, it holds in
+states $+$ and $-$ since $\myPr{u}{K}{N}{\mathit{out}_1} =
 \frac{3}{4}$, $\frac{1}{4}$ for $u = +, -$ respectively (Figure~\ref{figure:2-dp}). A careful check ensures that one cannot decrease  $\delta_1=.5$ without increasing $\epsilon$.
 Now consider the formula $\dpriv{\epsilon_2}{\delta_2} (\X
-(\mathit{out}_Y\wedge \X \mathit{out}_Y))$ in
+(\mathit{out}_1\wedge \X \mathit{out}_1))$ in
 Figure~\ref{figure:2-dp-mc}. Applying the
@@
 better privacy parameter $(0,.5)$ using the model checking algorithm
-because $\myPr{u}{K}{N}{\mathit{out}_Y} = \frac{9}{16}$,
+because $\myPr{u}{K}{N}{\mathit{out}_1} = \frac{9}{16}$,
 $\frac{1}{16}$ for $u = +, -$ respectively.
@@
 asymptotic upper bounds. Privacy parameters $\epsilon$ and $\delta$
-must be calculated carefully and often overly pessimistically. Our
-algorithm allows data analysts to choose such parameters more
-tightly.
+must be calculated carefully and often pessimistically. Our
+algorithm allows data analysts to choose better parameters.
 

### 2018-06-07 subsampling.tex
polish and spell check

@@
 The sensitivity of queries is required to apply the (truncated)
-$\alpha$-geometric mechanism. Recall that the sensitivity of a query
-is the maximal difference of query results on any two neighbors. There are
-two problems for mechanisms depending on query sensitivity. First, 
-sensitivity of queries can be hard to compute or estimate. Second,
+$\alpha$-geometric mechanism. Recall that the sensitivity 
+is the maximal difference of query results on any two neighbors. 
+Two practical problems may arise for
+mechanisms depending on query sensitivity. First, 
+sensitivity of queries can be hard to compute. Second,
 the sensitivity over arbitrary neighbors can be too conservative for
@@
 Subsampling is a technique to design such mechanisms~\cite{DR:14:AFDP}. Concretely, let
-us consider $\mathcal{X} = \{ R, B \}$ (for red and blue teams) and a
+us consider $\mathcal{X} = \{ R, B \}$ (for red and blue team members) and a
 dataset $d \in \mathcal{X}^n$. Suppose we would like to ask which team
@@
 $\mathit{Lap}(p)$ denotes a continuous random
-variable with the Laplace distribution whose probability density
-function is $f(x) = \frac{1}{2p}e^{-|x|/p}$.
+variable with the probability density
+function $f(x) = \frac{1}{2p}e^{-|x|/p}$ of Laplace distribution.
 If the perturbed difference is sufficiently large,
-the mechanism reports the majority of the $m$
-sub-results~(line~\ref{algorithm:subsampling-majority-output}). Otherwise,
-no information is
+the mechanism reports $1$ if the majority of the $m$
+sub-results is
+$R$ or $0$ if it is
+$B$~(line~\ref{algorithm:subsampling-majority-output}). Otherwise,  no
+information is 
 revealed~(line~\ref{algorithm:subsampling-majority-mute}).
@@
       \State{\textbf{if} $\mathit{count}_R \geq \mathit{count}_B$
-             \textbf{then} \Return { $R$ } \textbf{else} \Return {$B$} } 
+             \textbf{then} \Return { $1$ } \textbf{else} \Return {$0$} } 
       \label{algorithm:subsampling-majority-output}
@@
 $(\mathit{count}_R, \mathit{count}_B)$. There are only $m + 1$
-outcomes from $(m, 0), (m - 1, 1), \ldots, (0, m)$. Each outcome is
-represetned by a state in
+outcomes: $(m, 0), (m - 1, 1), \ldots, (0, m)$. Each outcome is
+represented by a state in
 Figure~\ref{figure:subsampling-majority-markov-chain}. From each
@@
 a finite set $\{ r_m, \ldots, r_M \}$ with the minimum $r_m$ and
-maximum $r_M$. For each $r \in \{ r_m, \ldots,
-r_M \}$, it is represented by a state in
-Figure~\ref{figure:subsampling-majority-markov-chain}. For instance,
+maximum $r_M$. 
+%For each $r \in \{ r_m, \ldots, r_M \}$, it is represented by a state
+%in Figure~\ref{figure:subsampling-majority-markov-chain}. 
+For instance,
 both outcomes $(m, 0)$ and $(0, m)$ transit to the state $r_M = 1/(4q)
@@
   (1/\delta)/\epsilon - r} \frac{\epsilon}{2}e^{-\epsilon|x|} dx$.
-From each state $r \in \{ r_m, \ldots, r_M \}$, it thus goes to the state
+From each state $r \in \{ r_m, \ldots, r_M \}$, it hence goes to the state
 $\top$ with probability $\int^{\infty}_{\ln (1/\delta)/\epsilon - r}
@@
 probability $1 - \int^{\infty}_{\ln (1/\delta)/\epsilon - r}
-\frac{\epsilon}{2}e^{-\epsilon|x|} dx$. Finally, the Markov chain
-moves to the state $R$ if $\mathit{count}_R \geq \mathit{count}_B$;
-otherwise, it moves to $B$.
+\frac{\epsilon}{2}e^{-\epsilon|x|} dx$ in
+Figure~\ref{figure:subsampling-majority-markov-chain}. Finally, the
+Markov chain 
+moves to the state $1$ if $\mathit{count}_R \geq \mathit{count}_B$;
+otherwise, it moves to $0$. Two dataset states are neighbors if they
+differ at most one member. For example, $rrb$ is a neighbor
+of $rrr$ and $rbb$ but not $bbb$.
 
@@
 
-      \node[node] (outR) at (5,  1.25) { $R$ };
-      \node[node] (outB) at (5, .25) { $B$ };
+      \node[node] (outR) at (5,  1.25) { $1$ };
+      \node[node] (outB) at (5, .25) { $0$ };
 

### 2018-06-07 syntax.tex
polish and spell check

@@
 and  \emph{until} operator
-($\until$) operator enclosed by two path formulae.
+($\until$) enclosed by two path formulae.
 We moreover define $F \phi \equiv true \until \phi$ and $G\phi \equiv \neg F (\neg\phi)$.

### 2018-06-08 above-threshold.tex
polish

@@
 the queries whose results are above the threshold; queries below the
-threshold only disclose minimal information and hence 
-irrelevant. Since query results above the threshold convey useful
-information, let us assume the mechanism will halt on the first such
+threshold only disclose minimal information and hence is
+irrelevant. Let us assume the mechanism will halt on the first such
 query result for simplicity. 
@@
 Assume that we have a threshold $t \in \{ 0, 1, \ldots, 5 \}$
-and queries $\{ f_i : f_i $ with sensitivity $\Delta
+and queries $\{ f_i : \Delta
 (f_i) = 1 \}$. In order to protect privacy, our mechanism applies the
@@
 threshold $t'$. For each query $f_i$, the truncated $\frac{1}{2}$-geometric
-mechanism is applied to its result $r_i$. If the perturbed result
+mechanism is applied to its result $r_i = f_i(x)$. If the perturbed result
 $r'_i$ is not less
-than the perturbed threshold $t'$, the mechanism halts with the input
+than the perturbed threshold $t'$, the mechanism halts with the output
 $\top$. Otherwise, it outputs $\bot$ and continues to the next query
@@
 
-      \node at (0, -1.65) { $\vdots$ };
+%      \node at (0, -1.65) { $\vdots$ };
 
@@
 
-%      (tt'2rr'0) edge [double] node [above] { $0|0, 1|0$ } (tt'2rr0)
-%      (tt'2rr'0) edge [double] node [above] { $0|1, 1|1, 2|1$ } (tt'2rr1)
-%      (tt'2rr'0) edge [double] node [above] { $1|2, 2|2$ } (tt'2rr2)
+      (tt'2rr'0) edge [double] node [above] 
+                 { \hide{$0|0, 1|0$} } (tt'2rr0)
+      (tt'2rr'0) edge [double] node [above] 
+                 { \hide{$0|1, 1|1, 2|1$} } (tt'2rr1)
+      (tt'2rr'0) edge [double] node [above] 
+                 { \hide{$1|2, 2|2$} } (tt'2rr2)
 
@@
 than the perturbed threshold. The model halts with the output $\top$
-by entering the state of a self loop. 
+by entering the state with a self loop. 
 Otherwise, it moves to one of $t'_1r_0$, $t'_1r_1$, or $t'_1r_2$
@@
 Now, the non-deterministic choices in the two computation cannot be 
-independent. Recall that the sensitivity of each query $f_i$ is
+independent. Recall that the sensitivity of each query is
 $1$. If the top computation moves to the state, say, $t'_1r_0$, it
 means the next query result on the dataset is $0$. Subsequently, the
-bottom computation can only move to $\underline{t}'_i\underline{r}_0$
-or $\underline{t}'_i\underline{r}_1$ depending on the perturbed
+bottom computation can only move to $\underline{t}'_j\underline{r}_0$
+or $\underline{t}'_j\underline{r}_1$ depending on its perturbed
 threshold. This is where actions are useful. Define the actions
@@
 their respective perturbed thresholds.
-Actions make sure the two computation on neighbors
+Actions make sure the two computation of neighbors
 is modeled properly. It remains to define the neighborhood

### 2018-06-08 dpCTL.tex
finalize
revise
polish

@@
 The logic $\dpCTLstar$ is designed to specify differentially private
-mechanisms. We introduce the differentially private path quantifer
-$\dpriv{\epsilon}{\delta}$ in $\dpCTLstar$. For any path formula
+mechanisms. We introduce the differentially private path quantifier
+$\dpriv{\epsilon}{\delta}$ and neighborhood relations for neighbors in
+$\dpCTLstar$. For any path formula 
 $\phi$, a state $s$ in a Markov chain $K$ satisfies
-$\dpriv{\epsilon}{\delta} \phi$ if the probability of having $\phi$-paths
-from $s$ is close to the probabilities of
-having $\phi$-paths from its neighbors.
+$\dpriv{\epsilon}{\delta} \phi$ if the probability of having paths 
+satisfying $\phi$ from $s$ is close to the probabilities of
+having paths satisfying $\phi$ from its neighbors.
 %We give semantics for both Markov chains and Markov decision processes.

### 2018-06-08 examples.tex
fix table row names
polish

@@
 propositions are used to designate intended interpretation on states
-(such as inputs or outputs). We demonstrates these ideas in examples.
+(such as inputs or outputs). We demonstrate these ideas in examples.
 
@@
 Consider the survey question: have you been diagnosed
-with the disease $X$? In order to protect surveyees' privacy, each
+with the disease $X$? In order to protect privacy, each surveyee
 answers the question as follows. The surveyee first flips a
@@
 \in \mathcal{X}$. Similarly, $\Pr[M (x) = 0] \leq e^{\ln 3} \Pr[M (x')
-= 0]$. The survey mechanism is $(\ln 3, 0)$-differentially private.
+= 0]$. The survey mechanism is hence $(\ln 3, 0)$-differentially private.
 The random noise boosts the probability of answering
@@
         \cline{2-4}
-        \multirow{2}*{\rotatebox{90}{state}} 
+        \multirow{2}*{\rotatebox{90}{$\mathit{input}$}} 
         &
@@
 respectively; the states $s$ and $t$ denote answers to the survey
-question and hence $\mathit{out}_Y \in L (s)$ and $\mathit{out}_N
+question and hence $\mathit{out}_1 \in L (s)$ and $\mathit{out}_0
 \in L (t)$.
@@
 The \emph{$\alpha$-geometric mechanism}
-outputs $f(x) + Y$ where $Y$ is a random variable with the geometric
+outputs $f(x) + Y$ on a dataset $x$ where $Y$ is a random variable with the geometric
 distribution~\cite{GRS:09:UUPM,GRS:12:UUPM} :
@@
 \Pr[Y = y] = \frac{1 - \alpha}{1 + \alpha}\alpha^{|y|}
-\textmd{ for } y \in \bbfZ.
+\textmd{ for } y \in \bbfZ
 \]
@@
       \cline{2-8}
-      \multirow{6}*{\rotatebox{90}{input}}
+      \multirow{6}*{\rotatebox{90}{$\mathit{input}$}}
       & 0 & {2}/{3}  & {1}/{6} & {1}/{12} & {1}/{24}  & {1}/{48} & {1}/{48}\\

### 2018-06-08 introduction.tex
polish

@@
 behaviors. A reflexive and symmetric user-defined binary relation
-over states is required to formalize similarity. We moreover add
+over states is required to formalize similar datasets. We moreover add
 the path quantifier $\dpriv{\epsilon}{\delta}$ for specifying similar
 behaviors. Informally, a state satisfies $\dpriv{\epsilon}{\delta}
-\phi$ if its probability of having path $\phi$ is close to
-those of similar states. Consider, for instance, a binary data analysis
+\phi$ if its probability of having path satisfying $\phi$ is close to
+those of similar states. Consider, for instance, a data analysis
 mechanism computing the likelihood ($\mathit{high}$ or $\mathit{low}$)
-of an epidemic. Then a state satisfying
+of an epidemic. A state satisfying
 $\dpriv{\epsilon}{\delta} (\F \mathit{high}) \wedge
 \dpriv{\epsilon}{\delta} (\F \mathit{low})$ denotes similar states
-have the same outcomes with similar probabilities.
+have similar probabilities on every outcomes.
 %\todo{not clear here}\lz{here it should be $\X$?}
@@
 \noindent
-\paragraph{Related Work.}
+\emph{Related Work.}
 An early attempt on formal verification of differential privacy
@@
 \noindent
-\paragraph{Contributions.} Our main contributions are threefold. 
+\emph{Contributions.} Our main contributions are threefold. 
 \begin{enumerate}
-\item We introduce the logic $\dpCTLstar$ for reasoning about properties in
-differential privacy. The logic is able to to express classical
+\item We introduce the logic $\dpCTLstar$ for reasoning about 
+differential privacy. The logic is able to to express subtle and generalized
 differentially private properties;

### 2018-06-08 main.tex
polish

@@
 \noindent
-\paragraph{Organization of the paper.}
+\emph{Organization of the paper.}
 Preliminaries are given in Section~\ref{section:preliminaries}.
-In Section~\ref{section:examples} we discuss how differentially private mechanisms are modeled as Markov chains.
+In Section~\ref{section:examples} we discuss how offline differentially private mechanisms are modeled as Markov chains.
 The logic $\dpCTLstar$ and its syntax are presented in Section~\ref{section:dpCTL}. The semantics over Markov chains and its model checking algorithm are given in Section~\ref{section:semantics}.
 Section~\ref{section:specifying-properties} discusses differential privacy properties using $\dpCTLstar$.
-More examples of differentially private mechanisms as Markov decision processes are given in Section~\ref{section:mdp}.
+More examples of online differentially private mechanisms as Markov decision processes are given in Section~\ref{section:mdp}.
 The semantics over Markov decision processes and undecidability of model checking is given in Section~\ref{section:semantics-mdp}.
@@
 \label{section:conclusions}
-We have introduced $\dpCTLstar$ to reason about properties in differential privacy, and have investigated its model checking problems.
+We have introduced $\dpCTLstar$ to reason about properties in differential privacy, and investigated its model checking problems.
 For Markov chains, the model checking problem has the same complexity as for $\PCTLstar$.
 The general MDP model checking problem however is undecidable.
-We have discussed some decidable special cases and an efficient yet sufficient condition to check differentially private subformulae.
+We have discussed some decidable special cases and a sufficient yet efficient condition to check differentially private subformulae.
 An interesting future work is to identify more

### 2018-06-08 preliminaries.tex
polish

@@
 We denote the data universe by $\mathcal{X}$; $x \in \mathcal{X}^n$ is
-a dataset with $n$ rows from the data universe. Two datasets $x$ and
+a \emph{dataset} with $n$ \emph{rows} from the data universe. Two datasets $x$ and
 $x'$ are \emph{neighbors} (denoted by $d(x, x') \leq 1$) if they are
@@
 oblivious mechanism depend on the query result $f (x)$.
-The order of rows in a dataset, for instance, is
+The order of rows, for instance, is
 irrelevant to oblivious mechanisms. Let $x, x'$
-be datasets and $\tilde{r} \in \tilde{R}$. We say the probability of
+be datasets and $\tilde{r} \in \tilde{R}$. The probability of
 the mechanism $M_f$ outputting $\tilde{r}$ on $x$ is
@@
 Informally, a differentially private mechanism has probabilistically
-similar behaviors on neighboring datasets. It will have similar
+similar behaviors on neighbors. It will have similar
 output distributions when any row is replaced by another in a given
@@
 %presence of any individual. 
-Individual privacy is thus preserved by differentially private mechanisms.
+Since the output distribution does not change significantly with the
+absence of any row in a dataset,
+individual privacy is thus preserved by differentially private mechanisms.
 
@@
 \sigma' \in S^+$ of the same length.
-Intuitively, its decision depends only on the length of the history.
+Intuitively, decisions of a query scheduler depend only on the length of the history.
 A path $\pi =

### 2018-06-08 semantics-mdp.tex
polish

@@
   e^{\epsilon} \cdot \myPr{s}{M_{\scheduler{S}}}{\neighbor{S}}{\phi}
-  + \delta
+  + \delta.
 \end{eqnarray*}
@@
 of having paths satisfying
-$\phi$ from neighbors. Recall that a history-dependent scheduler takes
-actions by previous states. Such a definition would allow
+$\phi$ from neighbors. Recall that a history-dependent scheduler chooses
+actions according to previous states. Such a definition would allow
 schedulers to take different actions from different states. 
@@
 (S, \Act, \wp, L)$, a neighborhood relation $\neighbor{S}$, $s \in
-S$, and a path formula $\phi$, consider the problem of deciding $M,
+S$, and a path formula $\phi$, consider the problem of checking $M,
 \neighbor{S}, s \models 
@@
 Proposition~\ref{proposition:sufficient-condition-mdp-model-checking} 
-check whether the mechanism is differentially private using the
+to check whether the mechanism is differentially private using the
 classical $\PCTL$ model checking algorithm for MDPs. 
@@
 is a sufficient condition. It only gives an upper bound of privacy
-parameters. Automatically computing tighter bounds may be possible.
+parameters. Tighter bounds may be computed by more sophisticated
+sufficient conditions.

### 2018-06-08 semantics.tex
polish

@@
 we say $s$ and $t$ are \emph{neighbors} or $t$ is a \emph{neighbor} of
-$s$. Define the satisfaction relation $K, \neighbor{S}, s
+$s$. 
+For any Markov chain $K$, neighborhood
+relation $N$ on $S$, $s \in S$, and a path formula $\phi$,
+define
+\[
+\myPr{s}{K}{N}{\phi} =
+\Pr [ \{ \pi : K, N, \pi \models \phi \textmd{ with } \pi_0 = s \} ].
+\]
+That is, $\myPr{s}{K}{N}{\phi}$ denotes the probability of paths
+satisfying $\phi$ from $s$ on $K$ with $N$. 
+Define the satisfaction relation $K, \neighbor{S}, s
 \models \Phi$ as follows.
@@
   & \textmd{ if } &
-  \Pr [\{ \pi : K, \neighbor{S}, \pi \models \phi \textmd{ with }
-                    \pi_0 = s\}] \in J \\
+                    \myPr{s}{K}{\neighbor{S}}{\phi} \in J\\
   K, \neighbor{S}, s \models \dpriv{\epsilon}{\delta} \phi
@@
   \textmd{for every } t \textmd{ with }  s \neighbor{S} t,
-      p(s) \leq e^{\epsilon} p(t) + \delta \textmd{ and }
-      p(t) \leq e^{\epsilon} p(s)  +  \\
-  & &  \delta \textmd{ where } p(x) = \Pr [\{
-      \pi : K, \neighbor{S}, \pi \models \phi \textmd{ with }
-      \pi_0 = x \}]
+      \myPr{s}{K}{\neighbor{S}}{\phi} \leq e^{\epsilon}
+      \myPr{t}{K}{\neighbor{S}}{\phi} + \delta
+      \textmd{ and }\\
+  & &  \myPr{t}{K}{\neighbor{S}}{\phi} \leq e^{\epsilon}
+       \myPr{s}{K}{\neighbor{S}}{\phi} + \delta
 \end{eqnarray*}
@@
 Moreover, the relation $K, \neighbor{S}, \pi \models \phi$ is defined as
-for the standard linear temporal logic formulae~\cite{ManneP92}. We only recall the semantics for $\X$ and $\until$:
+in the standard linear temporal logic formulae~\cite{ManneP92}. We only recall the semantics for the temporal operators $\X$ and $\until$:
 \begin{eqnarray*}
@@
 }
-\end{eqnarray*}
-
-For any Markov chain $K$, neighborhood
-relation $N$ on $S$, $s \in S$, and a path formula $\phi$,
-define
-\[
-\myPr{s}{K}{N}{\phi} =
-\Pr [ \{ \pi : K, N, \pi \models \phi \textmd{ with } \pi_0 = s \} ].
-\]
-That is, $\myPr{s}{K}{N}{\phi}$ denotes the probability of paths
-satisfying $\phi$ from $s$ on $K$ with $N$. The semantics of the
-differentially private operator hence becomes
-\begin{eqnarray*}
-  K, \neighbor{S}, s \models \dpriv{\epsilon}{\delta} \phi
-  & \textmd{ if } &
-  \textmd{for every } t \textmd{ with }  s \neighbor{S} t,
-      \myPr{s}{K}{\neighbor{S}}{\phi} \leq e^{\epsilon}
-      \myPr{t}{K}{\neighbor{S}}{\phi} + \delta
-      \textmd{ and }\\
-  & &  \myPr{t}{K}{\neighbor{S}}{\phi} \leq e^{\epsilon}
-       \myPr{s}{K}{\neighbor{S}}{\phi} + \delta
 \end{eqnarray*}
@@
 \begin{proposition}
-  Let $K = (S, \wp, L)$ be a Markov chain and $\neighbor{S}$ a
-  neighborhood relation on $S$.
+  Let $K = (S, \wp, L)$ be a Markov chain, $\neighbor{S}$ a
+  neighborhood relation on $S$, and $\phi$ a $\dpCTL$ path formula.
   $\{ s : K, \neighbor{S}, s \models \dpriv{\epsilon}{\delta} \phi \}$ is
-  computable within time polynomial in $|S|$.
+  computable within time polynomial in $|S|$ and $|\phi|$.
 \end{proposition}

### 2018-06-08 specifications.tex
polish

@@
 with $u N v$, we have $\myPr{u}{K}{N}{\X \mathit{out}_1}\le
-3\myPr{v}{K}{N}{\X \mathit{out}_1}$ for the probability of satisfying
+3\myPr{v}{K}{N}{\X \mathit{out}_1}$ for the probabilities of satisfying
 $\X \mathit{out}_1$ from $u$ and $v$.
@@
 $\psi$ for $k = 0, \ldots, 5$, then the $\frac{1}{2}$-geometric mechanism is
-$(\ln 2, 0)$-differentially private.
+$(\ln 2, 0)$-differentially private. For the subsampling majority
+mechanism (Section~\ref{subsection:subsampling}), consider the formula
+$\psi = \dpriv{\epsilon, \delta} (\F 0) \wedge \dpriv{\epsilon,
+  \delta} (\F 1)$. If a state satisfies $\psi$, its probability of
+outputting is $(\epsilon, \delta)$-close to those of its
+neighbor for every outcomes. The subsampling majority
+mechanism is $(\epsilon, \delta)$-differentially private.
 
@@
     \centering
-\resizebox{.70\columnwidth}{!}{
+\resizebox{.65\columnwidth}{!}{
     \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
@@
 We verify that this formula is satisfied for states $+$ and $-$.
-Moreover, the bound $\epsilon=\ln 9$ is tight, since the probability of satisfying $\X (\mathit{out}_1\wedge \X \mathit{out}_1)$ from states $+$ and $-$ are $\frac{9}{16}$ and $\frac{1}{16}$, respectively.
-Finally, the formula $\wedge_{a_1,a_2}\dpriv{\ln 9}{0} (\X (a_1\wedge \X a_2))$ specifies differential privacy for the model, where $a_1,a_2$ range over all atomic propositions $\{\mathit{out}_1, \mathit{out}_0\}$.
+Moreover, the bound $\epsilon=\ln 9$ is tight since the probability of satisfying $\X (\mathit{out}_1\wedge \X \mathit{out}_1)$ from states $+$ and $-$ are $\frac{9}{16}$ and $\frac{1}{16}$ respectively.
+Finally, the formula $\wedge_{a_1,a_2}\dpriv{\ln 9}{0} (\X (a_1\wedge \X a_2))$ specifies differential privacy for the model, where $a_1,a_2$ range over atomic propositions $\{\mathit{out}_1, \mathit{out}_0\}$.
 
@@
 This is a nested $\dpCTL$ formula, where the inner state formula $\dpriv{\ln 3}{0}(\X \mathit{out}_1)$ specifies the one-step differential privacy.
-Observe the inner formula is satisfied by all neighbors.
-Thus, the outer formula also has no privacy loss.
+Observe the inner formula is satisfied by all states.
+The outer formula has no privacy loss.
 \end{itemize}
@@
 states $+$ and $-$ since $\myPr{u}{K}{N}{\mathit{out}_1} =
-\frac{3}{4}$, $\frac{1}{4}$ for $u = +, -$ respectively (Figure~\ref{figure:2-dp}). A careful check ensures that one cannot decrease  $\delta_1=.5$ without increasing $\epsilon$.
+\frac{3}{4}$, $\frac{1}{4}$ for $u = +, -$ respectively (Figure~\ref{figure:2-dp}). A careful check infers that one cannot decrease  $\delta_1=.5$ without increasing $\epsilon$.
 Now consider the formula $\dpriv{\epsilon_2}{\delta_2} (\X

### 2018-06-08 subsampling.tex
polish

@@
 line~\ref{algorithm:subsampling-majority-noise},
-$\mathit{Lap}(p)$ denotes a continuous random
+$\mathit{Lap}(p)$ denotes the continuous random
 variable with the probability density
-function $f(x) = \frac{1}{2p}e^{-|x|/p}$ of Laplace distribution.
+function $f(x) = \frac{1}{2p}e^{-|x|/p}$ of the Laplace distribution.
 If the perturbed difference is sufficiently large,
@@
 Figure~\ref{figure:subsampling-majority-markov-chain}. From each
-dataset, the probability distribution of all outcomes gives the
-transition probability in the figure. Next, observe that  $|
+dataset, the probability distribution on all outcomes gives the
+transition probability. Next, observe that  $|
 \mathit{count}_R - \mathit{count}_B |$ can have only finitely many
@@
 probability $1 - \int^{\infty}_{\ln (1/\delta)/\epsilon - r}
-\frac{\epsilon}{2}e^{-\epsilon|x|} dx$ in
-Figure~\ref{figure:subsampling-majority-markov-chain}. Finally, the
-Markov chain 
-moves to the state $1$ if $\mathit{count}_R \geq \mathit{count}_B$;
+\frac{\epsilon}{2}e^{-\epsilon|x|} dx$. Finally, the
+Markov chain moves from the state $\top$ to $1$ if
+$\mathit{count}_R \geq \mathit{count}_B$; 
 otherwise, it moves to $0$. Two dataset states are neighbors if they

### 2018-06-08 syntax.tex
polish

@@
  by a path formula. A
-\emph{path} formula $\phi$ is simply a linear temporal logic formula, with temporal operator \emph{next}  ($\X$)
+\emph{path} formula $\phi$ is simply a linear temporal logic formula, with temporal operator \emph{next}  ($\X$) followed by a path formula, 
 and  \emph{until} operator
 ($\until$) enclosed by two path formulae.
-We moreover define $F \phi \equiv true \until \phi$ and $G\phi \equiv \neg F (\neg\phi)$.
+We define $F \phi \equiv true \until \phi$ and $G\phi \equiv \neg F (\neg\phi)$ as usual.
 

### 2018-06-12 above-threshold.tex
explain actions and add them to the figure.

@@
 
-      (t'1r'0) edge [double] node [above]
-              { \hide{$0|0, 0|1$} } (t'1r0)
-      (t'1r'0) edge [double] node [above] 
-              { \hide{$1|0, 1|1, 1|2$} } (t'1r1)
-      (t'1r'0) edge [double] node [below] 
-              { \hide{$2|1, 2|2$} } (t'1r2)
+      (t'1r'0) edge [double] node [right]
+              { \rotatebox{270}{$0\underline{0}, 0\underline{1}$} } (t'1r0)
+      (t'1r'0) edge [double] node [right=-8] 
+              { \rotatebox{315}{$1\underline{0}, 
+                                 1\underline{1}, 1\underline{2}$} } (t'1r1)
+      (t'1r'0) edge [double] node [right] 
+              { \rotatebox{334}{$2\underline{1}, 2\underline{2}$} } (t'1r2)
 
@@
 
-      (tt'2rr'0) edge [double] node [above] 
-                 { \hide{$0|0, 1|0$} } (tt'2rr0)
+      (tt'2rr'0) edge [double] node [right] 
+                 { \rotatebox{270}{$0\underline{0}, 1\underline{0}$} } (tt'2rr0)
       (tt'2rr'0) edge [double] node [above] 
@@
 
-      (tt'2rr'1) edge [double] node [above]
-                { \hide{$0|0, 1|0$} } (tt'2rr0)
+      (tt'2rr'1) edge [double] node [right=15,below=-11]
+                { \rotatebox{315}{$0\underline{0}, 1\underline{0}$} } (tt'2rr0)
       (tt'2rr'1) edge [double] node [above]
@@
 $\underline{t}'_2\underline{r}_0$ is associated with the actions
-$0\underline{0}$ and $1\underline{0}$ (but not $2\underline{0}$). On
-the action $0\underline{0}$, the top computation moves to $t'_ir_0$ and
-the bottom computation moves to $\underline{t}'_j\underline{r}_0$ by
-their respective perturbed thresholds.
+$0\underline{0}$ and $1\underline{0}$ (but not $2\underline{0}$). 
+Assume the perturbed thresholds of the top and bottom computation are
+$i$ and $j$ respectively. On the action $0\underline{0}$, the top
+computation moves to $t'_ir_0$ and 
+the bottom computation moves to $\underline{t}'_j\underline{r}_0$ 
 Actions make sure the two computation of neighbors
-is modeled properly. It remains to define the neighborhood
+is modeled properly. Now consider the action sequence
+$-,-,0\underline{1},-,2\underline{2},-,2\underline{1}$ from the states
+$t_0r_1$ and $\underline{t}_0\underline{r}_0$ (``$-$'' represents the
+purely probabilistic action). Together with the first query results, 
+it denotes four consecutive query results $1$, $0$, $2$, $2$ on the
+top computation, and $0$, $1$, $2$, $1$ on the bottom computation.
+Each action sequence models two sequences of query results: one on the
+top, the other on the bottom computation. Moreover, the difference of
+the corresponding query results on the two computation is at most one
+by the definition of the action set. Any sequence of adaptive query
+results is hence formalized by an action sequence in our model.
+
+It remains to define the neighborhood
 relation. Recall the sensitivity is $1$. Consider the neighborhood

### 2018-06-12 specifications.tex
minor
minor change

@@
 mechanism (Section~\ref{subsection:subsampling}), consider the formula
-$\psi = \dpriv{\epsilon, \delta} (\F 0) \wedge \dpriv{\epsilon,
-  \delta} (\F 1)$. If a state satisfies $\psi$, its probability of
+$\psi = \dpriv{\epsilon}{\delta} (\F 0) \wedge \dpriv{\epsilon}
+  {\delta} (\F 1)$. If a state satisfies $\psi$, its probability of
 outputting is $(\epsilon, \delta)$-close to those of its

### 2018-06-13 introduction.tex
polish

@@
 interact with its environment, it is formalized as a Markov
-chain. Otherwise, its interactions are formalized by actions in
+chain. Otherwise, its interactions are formalized by non-deterministic 
+actions in
 Markov decision processes. Our formalization effectively assumes that
-actions are controlled by adversaries. It is thus necessary to
-consider all privacy attacks from adversaries in order to establish
-differential privacy.
+actions are controlled by adversaries. It thus
+considers all privacy attacks from adversaries in order to establish
+differential privacy as required.
 %\todo{What is a state?}

### 2018-06-14 semantics.tex
add Depeng and Lijun's fix

@@
 having paths satisfying $\phi$ from $s$ is $(\epsilon, \delta)$-close
-those from every neighbor of $s$. Informally, it is
+to those from every neighbor of $s$. Informally, it is
 probabilistically similar to observe paths satisfying $\phi$ from $s$

### 2018-06-14 specifications.tex
add Depeng and Lijun's fix

--- 2018-06-08 specifications.tex
@@
 mechanism (Section~\ref{subsection:subsampling}), consider the formula
-$\psi = \dpriv{\epsilon, \delta} (\F 0) \wedge \dpriv{\epsilon,
-  \delta} (\F 1)$. If a state satisfies $\psi$, its probability of
+$\psi = \dpriv{\epsilon}{\delta} (\F 0) \wedge
+\dpriv{\epsilon}{\delta} (\F 1)$. If a state satisfies $\psi$, its
+probability of 
 outputting is $(\epsilon, \delta)$-close to those of its
@@
 compositional theorem, one has $\epsilon_2=2\epsilon_1=0$ and
-$\delta_2=2\delta_1=1$. However, we can check easily that one gets
+$\delta_2=2\delta_1=1$. However, we can get
 better privacy parameter $(0,.5)$ using the model checking algorithm
-because $\myPr{u}{K}{N}{\mathit{out}_1} = \frac{9}{16}$,
-$\frac{1}{16}$ for $u = +, -$ respectively.
-In general, compositional theorems for differential privacy only give
+since $\myPr{u}{K}{N}{\X (\mathit{out}_1 \wedge \X\mathit{out}_1)}
+= \frac{9}{16}$, $\frac{1}{16}$ for $u = +, -$ respectively.
+Generally, compositional theorems in differential privacy only give
 asymptotic upper bounds. Privacy parameters $\epsilon$ and $\delta$

### 2018-07-27 rebuttal-aplas.txt
add reviews and rebuttals

First of all, we would like to thank all reviewers' efforts and
comments.

The main contributions of this work are the formalization of privacy
protection mechanisms as Markov chains or Markov Decision Processes,
and its automated analyses. We choose to formalize interactive
mechanisms by non-determinism in Markov Decision Processes and
illustrate our formalization by the above threshold mechanism
(Algorithm 3) for clarity. In the above threshold mechanism, the
incoming queries f_i's are selected dynamically. Later queries may
depend on the results of previous queries. Hence the interaction is
implicit in the choices of incoming queries. It would be great if we
could give our formalization of more explicit interactions such as
those in streaming data. Due to space limit, we had to present a
simpler mechanism with implicit interactions. We also agree a thorough
investigation on the expressiveness of our formalization will be very
useful. It will definitely be an interesting future work. Our
(un)decidability results are for dpCTL*. As long as a mechanism can be
modeled by Markov chains, dpCTL* properties can be verified
automatically. Model checking dpCTL* over Markov Decision Processes on
the other hand is undecidable. Hence interactive mechanisms in general
cannot be verified with our current formalization.

We are certainly aware of various language-based techniques for
verifying differential privacy. Just like classical
programs/properties, we see both techniques are complementary.
Classical model checking aims to reduce human efforts by focusing on
abstract models with automated verification. It has helped programmers
verify simpler programs and properties reasonably successfully over
decades of research efforts. Our work has the same position. It does
not aim to verify all possible differentially private programs.
Instead, it hopefully will solve simpler problems so that programmers
can have more time for language-based techniques.

### 2018-07-28 rebuttal-aplas.txt
add Depeng's part

@@
 
-We are certainly aware of various language-based techniques for
-verifying differential privacy. Just like classical
+We are aware of various language-based techniques for verifying
+differential privacy (of course). Just like classical 
 programs/properties, we see both techniques are complementary.
@@
 can have more time for language-based techniques.
+
+F\phi (finally \phi) and G\phi (globally \phi) are two common temporal
+properties used in PCTL*. They are syntactic sugars (page 8).
+The time complexity in Proposition 1 is polynomial in
+|S|, which includes |S|^2. Actually the computations in PCTL* alone
+would take more than |S|^2. Since differential privacy requires the
+indistinguishable probabilities on all neighbors, thus we need to
+compute the probabilities and compare between neighbors. The
+relational assertion just tells us which pairs should be neighbors, without
+guaranteeing the probabilities. The other two examples actually show
+us explicit models and computations, though, which are much easier and
+obvious. There exist incoming edges for t1’r2’, such as from t1’r0, t1’r1
+and so on. We explain these transitions in section 7.1 and leave out
+some edges for clarity.

### 2018-07-28 above-threshold.tex

@@
 computation moves to $t'_ir_0$ and 
-the bottom computation moves to $\underline{t}'_j\underline{r}_0$ 
+the bottom computation moves to $\underline{t}'_j\underline{r}_0$.
 Actions make sure the two computation of neighbors

### 2018-08-21 main.tex
add authers and institutes

@@
 \usepackage{times}
+\usepackage[symbol]{footmisc}
+\renewcommand{\thefootnote}{\fnsymbol{footnote}}
 
@@
 \newif\ifpublic
-%\publictrue
+\publictrue
 
@@
 \author{
-Depeng Liu\inst{1}
+Depeng Liu\inst{1,2}
 \and
-Bow-Yaw Wang\inst{2}
+Bow-Yaw Wang\inst{3}
 \and
-Lijun Zhang\inst{1}}
+Lijun Zhang\inst{1,2}}
 \institute{
+State Key Laboratory of Computer Science, Institute of Software,
 Chinese Academy of Sciences
+\footnotemark[1]
 \and
-Academia Sinica
+University of Chinese Academy of Sciences
+\footnotemark[1]
+\and
+Institute of Information Science, Academia Sinica
+\footnotemark[2]
 }
@@
 \maketitle
+
+\footnotetext[1]{Partially supported by \todo{please fill in}}
+\footnotetext[2]{Partially supported by the Academia
+Sinica Thematic Project: Socially Accountable Privacy Framework for
+Secondary Data Usage.}
 

### 2018-09-03 introduction.tex
fix for camera ready
Remove a redundant word

@@
 with minimal efforts. More interaction
-between model checking~\cite{ManneP92,BK:08:PMC} and privacy analysis hopefully will follow.
+between model checking~\cite{CGPMC,BK:08:PMC} and privacy analysis hopefully will follow.
 
@@
 \item We introduce the logic $\dpCTLstar$ for reasoning about
-differential privacy. The logic is able to to express subtle and generalized
+differential privacy. The logic is able to express subtle and generalized
 differentially private properties;
